### 2013-08-13 chi2014/abstract.tex
latex paper structure

In this paper we describe the formatting requirements for
SIGCHI Conference Proceedings, and this sample file
offers recommendations on writing for the worldwide
SIGCHI readership. Please review this document even if
you have submitted to SIGCHI conferences before, some
format details have changed relative to previous years.

### 2013-08-13 chi2014/introduction.tex
latex paper structure

This format is to be used for submissions that are
published in the conference proceedings.  We wish to give
this volume a consistent, high-quality appearance. We
therefore ask that authors follow some simple
guidelines. In essence, you should format your paper
exactly like this document. The easiest way to do this is
simply to download a template from the conference web
site, and replace the content with your own material.

### 2013-08-13 bootstrap.sh
notebook export dependency

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 apt-get update
-apt-get install -y python-pip python-dev python-requests build-essential sqlite3 libfreetype6-dev libpng-dev python-scipy python-matplotlib python-tornado python-zmq python-nltk python-pandas
+apt-get install -y python-pip python-dev python-requests build-essential sqlite3 libfreetype6-dev libpng-dev python-scipy python-matplotlib python-tornado python-zmq python-nltk python-pandas pandoc
 pip install --upgrade ipython
 pip install turkflow numpy pyyaml
 

### 2013-08-14 bootstrap.sh
update to make tex files and more stats

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 apt-get update
-apt-get install -y python-pip python-dev python-requests build-essential sqlite3 libfreetype6-dev libpng-dev python-scipy python-matplotlib python-tornado python-zmq python-nltk python-pandas pandoc
+apt-get install -y python-pip python-dev python-requests build-essential sqlite3 libfreetype6-dev libpng-dev python-scipy python-matplotlib python-tornado python-zmq python-nltk python-pandas pandoc python-pygments python-sphinx
 pip install --upgrade ipython
 pip install turkflow numpy pyyaml
 

### 2013-08-14 src/originality/process_hits.py

@@ -366,7 +366,7 @@ for qs in question_sets:
     for cluster_num in subset:
         cluster_sizes[cluster_num] += 1
 
-    N1 = reduce(operator.add, [1 for key in subset if cluster_sizes[key] == 1])
+    N1 = len(key for key in subset if cluster_sizes[key] == 1)
     p0 = float(N1) / len(cluster_samples)
 
     print "\tsubset p0 size", i, "/10:", p0
@@ -374,8 +374,9 @@ for qs in question_sets:
     # if i == 9:
     #   print sorted(subset)
     #   print sorted(rest)
+    #   print sorted(cid for cid in rest if not cid in subset)
 
-    actual_p0 = float(sum(1 for cid in rest if not cid in subset)) / len(rest)
+    actual_p0 = float(len(cid for cid in rest if not cid in subset)) / len(rest)
 
     print "\trest p0", actual_p0
 

### 2013-08-14 src/main_experiment.py
pilot14 tested successfully

@@ -63,7 +63,8 @@ def get_questions_for_responses(n, append=False):
     return question
 
 def get_response_rewards(expid):
-    return [(10, 0.50, "%s_infinis" % expid, "infinis", True),
+    return [(75, 2.65, "%s_seventy_fives" % expid, "seventy_fives", False),
+            (100, 3.50, "%s_hundreds" % expid, "hundreds", False),
             ]
 
 def post_jobs(administrator_URL, responses_rewards, duration,

### 2013-09-01 src/common.py
added fallback questions, finally'

@@ -27,10 +27,13 @@ env = Environment(autoescape=guess_autoescape,
 ###
 
 class RandomStorm(TurkHITType):
-    def __init__(self, num_responses, max_assignments, reward, admin_url, admin_id, duration, append_ideas = False):
+    def __init__(self, num_responses, max_assignments, reward,
+            admin_url, admin_id, backup_problems, duration,
+            append_ideas = False):
         self.num_responses = 10 if num_responses is None else num_responses
         self.admin_url = admin_url
         self.admin_id = admin_id
+        self.backup_problems = backup_problems
 
         if append_ideas:
             self.append_ideas = True

### 2013-09-01 src/main_experiment.py

@@ -45,11 +45,11 @@ def gen_jobs_random(questions, num_assignments):
 def get_questions_for_responses(n, append=False):
     n = "as many as possible" if append else "%i" % (n)
 
-    charity_question = """<p>The Electronic Frontier Foundation (EFF) is a nonprofit whose goal is to protect individual rights with respect to digital and online technologies. For example, the EFF has initiated a lawsuit against the US government to limit the degree to which the US surveils its citizens via secret NSA programs. If you are unfamiliar with the EFF and its goals, read about it on its website (<a href="https://www.eff.org" target="_new">https://www.eff.org</a>) or via other online sources (such as Wikipedia).</p>
-      <p>Brainstorm %s <em>new</em> ways the EFF can raise funds and simultaneously increase awareness. Your ideas <em>must be different from their current methods</em>, which include donation pages, merchandise, web badges and banners, affiliate programs with Amazon and eBay, and donating things such as airmiles, cars, or stocks. See the full list of their current methods here: <a href="https://www.eff.org/helpout" target="_new">https://www.eff.org/helpout</a>. Be as specific as possible in your responses.</p>"""
+    charity_question = "<p>The Electronic Frontier Foundation (EFF) is a nonprofit whose goal is to protect individual rights with respect to digital and online technologies. For example, the EFF has initiated a lawsuit against the US government to limit the degree to which the US surveils its citizens via secret NSA programs. If you are unfamiliar with the EFF and its goals, read about it on its website (<a href=https://www.eff.org' target='_new'>https://www.eff.org</a>) or via other online sources (such as Wikipedia).</p>\
+      <p>Brainstorm %s <em>new</em> ways the EFF can raise funds and simultaneously increase awareness. Your ideas <em>must be different from their current methods</em>, which include donation pages, merchandise, web badges and banners, affiliate programs with Amazon and eBay, and donating things such as airmiles, cars, or stocks. See the full list of their current methods here: <a href='https://www.eff.org/helpout' target='_new'>https://www.eff.org/helpout</a>. Be as specific as possible in your responses.</p>"
 
-    turk_question = """<p>Mechanical Turk currently lacks a dedicated mobile app for performing HITs on smartphones (iPhone, Androids, etc.) or tablets (e.g., the iPad).</p>
-      <p>Brainstorm %s features for a mobile app to Mechanical Turk that would improve the worker's experience when performing HITs on mobile devices. Be as specific as possible in your responses.</p>"""
[…]

### 2013-09-01 src/main_experiment.py
pilot16 ready to go

@@ -35,6 +35,13 @@ def gen_jobs_for_question(question, num):
 
     return [{"question": question, "job_type": job_type} for i in range(num)]
 
+
+def gen_jobs_for_all_questions_long(questions_assignments):
+    jobs = []
+    for question, num_assignments in questions_assignments:
+      jobs.append(gen_jobs_for_question(question, num_assignments))
+    return list(itertools.chain.from_iterable(jobs))
+
 def gen_jobs_for_all_questions(questions, num_assignments_per_question):
     jobs = [gen_jobs_for_question(q, num_assignments_per_question) for q in questions]
     return list(itertools.chain.from_iterable(jobs))
@@ -62,13 +69,36 @@ def get_questions_for_responses(n, append=False):
 
     return question
 
+### This part is what changes for the experiment, mainly
+
+def get_num_participants(num_requested):
+  if num_requested == 5:
+    return [5, 5, 5, 5]
+  elif num_requested == 10:
+    return [5, 5, 5, 5]
+  elif num_requested == 20:
+    return [3, 0, 0, 0]
+  elif num_requested == 50:
+    return [6, 4, 5, 6]
+  elif num_requested == 75:
+    return [6, 5, 5, 5]
+  elif num_requested == 100:
+    return [7, 4, 5, 6]
+
 def get_response_rewards(expid):
-    return [(75, 2.65, "%s_seventy_fives" % expid, "seventy_fives", False),
+    return [
+            (5, 0.18, "%s_fives" % expid, "fives", False),
+            (10, 0.35, "%s_tens" % expid, "tens", False),
+            (20, 0.70, "%s_twenties" % expid, "twenties", False),
+            (50, 1.75, "%s_fifties" % expid, "fifties", False),
+            (75, 2.65, "%s_seventy_fives" % expid, "seventy_fives", False),
             (100, 3.50, "%s_hundreds" % expid, "hundreds", False),
             ]
 
+###
+
 def post_jobs(administrator_URL, responses_rewards, duration,
-              num_assignments_per_condition, tc, exp, random_type = False):
+              tc, exp, random_type = False):
 
     keys = []
 
@@ -76,10 +106,12 @@ def post_jobs(administrator_URL, responses_rewards, duration,
         questions = get_questions_for_responses(responses, append)
 
         jobs = None
-        if random_type:
-            jobs = gen_jobs_random(questions, num_assignments_per_condition)
-        else:
-            jobs = gen_jobs_for_all_questions(questions, num_assignments_per_condition)
[…]

### 2013-09-02 chi2014/method.tex
first pass of method section

We began the process of soliciting responses to brainstorming tasks in a crowd marketplace. Throughout the experimental process, particular aspects of the experimental design remained fixed.

The independent variables of interest were \emph{brainstorming problem} and \emph{number of ideas requested}.

Participants were recruited from Amazon's \emph{Mechanical Turk} (CITE), and were restricted to residents of the United States for a baseline expectation of english language comprehension and cultural familiarity. Mechanical Turk is an online marketplace in which members receive financial reward for completing \emph{Human Intelligence Tasks}, or HITs.

HITs were placed on the marketplace for each \emph{number of ideas requested} condition, with proportional rewards. Participants could accept at most one HIT in each of these conditions.

Upon accepting a HIT, participants are randomly assigned into a \emph{brainstorming problem} condition. We ensured that participants completing multiple HITs are not exposed to the same brainstorming task twice.

\subsubsection{Design Concerns}
Workers choose which HITs to complete, thus self-selection bias is a reasonable concern. This bias is also present in real-world HIT choice behaviour.

Upon accepting a HIT, participants were asked to give consent and informed that they could leave the study at any point without financial consequences. In regular online task markplaces, common practice would restrict workers from submitting a HIT as complete unless all responses were given. This is a threat to external validity.

\subsection{Task}

The brainstorming task is a form in a standard web browser. Participants are presented with a brief overview of the tenets of brainstorming, are presented a problem, and must enter some number of ideas to resolve the problem within an 18 hour period.

At the beginning of the form is a brief introduction to brainstorming, with a paraphrase of Osborne's four rules of brainstorming (CITE). These rules were manipulated to make sense within the constraints nominal brainstorming over the web medium. The rules, as displayed to the participants:

\begin{enumerate}
\item There are no bad ideas. Don't criticise your choices.
\item Wild ideas and building off of old ideas are okay.
\item Quantity of ideas is prioritized.
\item Combinations of ideas count as new ideas.
\end{enumerate}

The brainstorming task is below this, followed by a series of text entry inputs numbered through to the total number of ideas requested. Figure X (FIG) is an example of a typical task. We place a larger free text area at the bottom of the list where participants could enter any additional ideas.

\subsection{Pilots and Question Selection}

With this basic design in hand, we ran several pilots and experiments. In early pilots, we used classic problems from psychology literature on brainstorming, including the "thumb problem" and "broom problem" (CITE). Early results from brainstormers were unsatisfying and wildly divergent. We identified three key traits problems needed in order to evaluate them for creativity:

\begin{enumerate}
\item They must require creativity to resolve. Obvious answers do not satisfy the problem.
\item They must be problems that participants on Mechanical Turk would have the expertise to solve.
\item They must have an associated success metric.
\end{enumerate}

The primary researcher and two additional researchers familiar with crowd marketplace brainstorming tasks brainstormed a large variety of potential problems. These problems were iterativelly culled and refined over a series of further pilots until the above goals were felt to be achieved. This resulted in the following four questions, as presented to the participants:

\begin{enumerate}
\item \textbf{Charity}

The Electronic Frontier Foundation (EFF) is a nonprofit whose goal is to protect individual rights with respect to digital and online technologies. For example, the EFF has initiated a lawsuit against the US government to limit the degree to which the US surveils its citizens via secret NSA programs. If you are unfamiliar with the EFF and its goals, read about it on its website (https://www.eff.org) or via other online sources (such as Wikipedia).

Brainstorm N \emph{new} ways the EFF can raise funds and simultaneously increase awareness. Your ideas \emph{must be different from their current methods}, which include donation pages, merchandise, web badges and banners, affiliate programs with Amazon and eBay, and donating things such as airmiles, cars, or stocks. See the full list of their current methods here: https://www.eff.org/helpout. Be as specific as possible in your responses."

\item \textbf{Mechanical Turk}

"Mechanical Turk currently lacks a dedicated mobile app for performing HITs on smartphones (iPhone, Androids, etc.) or tablets (e.g., the iPad).

Brainstorm N features for a mobile app to Mechanical Turk that would improve the worker's experience when performing HITs on mobile devices. Be as specific as possible in your responses.

\item \textbf{MP3}

Many people have old iPods or MP3 players that they no longer use. Please brainstorm N uses for old iPods/MP3 players. Assume that the devices' batteries no longer work, though they can be powered via external power sources. Also be aware that devices may \emph{not} have displays. Be as specific as possible in your descriptions.

\item \textbf{Forgot Name}

Imagine you are in a social setting and you have forgotten the name of somebody you know. Brainstorm N ways you could learn their name without directly asking them. Be as specific as possible in your descriptions.

\end{enumerate}

Here do I talk about the unlimited condition?

\subsection{Experiment}

With the above problem set and the previously outlined task design (SEC), we collected responses. We chose 6 \emph{number of ideas requested} conditions that covered the spectrum (and beyond) of quantity of creativity requested from a single participant. Those conditions were: 5 ideas (with a corresponding reward of \$0.18 USD), 10 ideas (\$0.35), 20 (\$0.70), 50 (\$1.75), 75 (\$2.65), and 100 (\$3.50). To reiterate, participants chose a HIT to generate some number of ideas, and then were randomly assigned a problem.

\subsubsection{Data Collection}

Does it make sense to have a section here on what data we collected?

\subsection{Coding}

One researcher coded the data for category in a hierarchical scheme. For each question, ideas recieved from the participants were organized in a tree structure. A single node in the tree represents a particular solution to the problem.

To formalize the clustering method, we introduce some key terms:

\subsubsection{Similarity}
Similarity is a subjective measure of the equality of two nodes a, b on the basis of their key problem solution. a and b have high similarity if they solve the problem in the same way, and low similarity if their solutions have no common themes.

\subsubsection{Coverage}
A node a has high coverage of a node b if the problem solution in a is equal to or an abstraction of the problem solution in b. Coverage is not commutative.

\subsubsection{Symmetrical Coverage}
Two nodes a, b have symmetrical coverage if coverage(a, b) = coverage(b, a).

In the resulting tree, all parent nodes have high coverage over their child nodes, but all child nodes have low coverage over their parent nodes. All nodes have high affinity with their parent and their children. The clustering algorithm is outlined in figure (FIG).

\begin{figure*}[h]
\begin{verbatim}
for each idea:
  idea_node = new cluster with idea as member
  current_node = root
  do:
    best_match = max_affinity(idea_node, current_node.children)

    if best_match.affinity is low or current_node has no children:
      insert idea_node under current_node
      exit do
    else:
      if symmetrical_coverage(idea_node, best_match) and high:
        merge idea_node, best_match
        exit do
      else symmetrical_coverage(idea_node, best_match) and low:
        new_parent = new child node of current_node, summarizing cause of affinity
        insert best_match, idea_node under new_parent
        exit do
      else if coverage(idea_node, best_match) > coverage(best_match, idea_node):
        replace best_match with idea_node in tree
        current_node = idea_node
        idea_node = best_match
      else:
        current_node = best_match
\end{verbatim}
\caption{Manual clustering algorithm}
\label{fig:cluseringalg}
\end{figure*}

### 2013-09-03 src/clusterer/clustering_app.py
added clustering app to version control

@@ -1,0 +1,157 @@
+# -*- coding: utf-8 -*-
+
+# Form implementation generated from reading ui file 'clustering_app.ui'
+#
+# Created: Fri Aug 23 09:59:55 2013
+#      by: pyside-uic 0.2.14 running on PySide 1.2.0
+#
+# WARNING! All changes made in this file will be lost!
+
+from PySide import QtCore, QtGui
+
+class Ui_MainWindow(object):
+    def setupUi(self, MainWindow):
+        MainWindow.setObjectName("MainWindow")
+        MainWindow.resize(840, 755)
+        self.centralwidget = QtGui.QWidget(MainWindow)
+        self.centralwidget.setObjectName("centralwidget")
+        self.verticalLayout = QtGui.QVBoxLayout(self.centralwidget)
+        self.verticalLayout.setObjectName("verticalLayout")
+        self.verticalLayout_3 = QtGui.QVBoxLayout()
+        self.verticalLayout_3.setObjectName("verticalLayout_3")
+        self.horizontalLayout_2 = QtGui.QHBoxLayout()
+        self.horizontalLayout_2.setSizeConstraint(QtGui.QLayout.SetDefaultConstraint)
+        self.horizontalLayout_2.setObjectName("horizontalLayout_2")
+        self.label = QtGui.QLabel(self.centralwidget)
+        self.label.setObjectName("label")
+        self.horizontalLayout_2.addWidget(self.label)
+        self.combo_box_data_set = QtGui.QComboBox(self.centralwidget)
+        self.combo_box_data_set.setObjectName("combo_box_data_set")
+        self.horizontalLayout_2.addWidget(self.combo_box_data_set)
+        self.horizontalLayout_2.setStretch(0, 1)
+        self.horizontalLayout_2.setStretch(1, 1)
+        self.verticalLayout_3.addLayout(self.horizontalLayout_2)
+        self.horizontalLayout_3 = QtGui.QHBoxLayout()
+        self.horizontalLayout_3.setObjectName("horizontalLayout_3")
+        self.verticalLayout_4 = QtGui.QVBoxLayout()
+        self.verticalLayout_4.setObjectName("verticalLayout_4")
+        self.list_ideas = QtGui.QListView(self.centralwidget)
+        self.list_ideas.setSelectionMode(QtGui.QAbstractItemView.ExtendedSelection)
+        self.list_ideas.setObjectName("list_ideas")
+        self.verticalLayout_4.addWidget(self.list_ideas)
[…]

### 2013-09-03 src/clusterer/feature_extraction.py

@@ -1,0 +1,413 @@
+#!/usr/bin/python
+
+# Set of routines for transforming strings into higher-level codes
+# and for extracting features
+
+import popen2
+import re
+import nltk
+from nltk.corpus import wordnet as wn
+import operator
+import math
+
+# Random utility function
+# Assumes two sets of features passed in
+def bag_of_words_cos_similarity(l, r):
+  if len(l) == 0 or len(r) == 0:
+    return 0
+  common_set = l.intersection(r)
+  num_in_common = len(common_set)
+  denom = reduce(operator.mul, [math.sqrt(len(x)) for x in [l,r]], 1.0)
+  return num_in_common / denom
+
+def sanitize_raw_input(input_string):
+# Potentially do this, too:
+  #input_string = input_string.replace(" '", "'").replace("sn't", "s not").replace("n't", "not")
+  return input_string.encode('utf-8')
+
+def tokenize_no_punct(s):
+#    self.tokens = nltk.word_tokenize(self.answer)
+  s = re.sub(r'[^a-zA-Z0-9 ]', '', s)
+  return s.split()
+
+def tokenize_minimal_punct(s):
+  s = re.sub(r'[^a-zA-Z0-9,.!? ]', '', s)
+  return s.split()
+
+def tokenize_whitespace(s):
+  return s.split()
+
+def tokenize_lowercase_no_punct(s):
+  s = s.lower()
+  return tokenize_no_punct(s)
+
+# Returns a feature list that just includes the tokens
+def extract_identity(tokens):
+  return [[t] for t in tokens]
+
+def extract_spelling_suggestions(tokens):
+  return create_feature_list(tokens, spellcheck_word)
+
+def extract_best_spelling_suggestions(tokens):
+  return create_feature_list(tokens, get_best_spelling)
+
+def extract_number_codes(tokens):
+  return create_feature_list(tokens, code_number)
+
+def extract_punct_codes(tokens):
+  return create_feature_list(tokens, code_punctuation)
+
+def extract_has_punct(tokens):
+  return create_feature_list(tokens, code_has_punctuation)
+
+# Adds codes if words other than first word are capitalized
+def extract_later_caps_codes(tokens):
+  if len(tokens) > 1:
+    return create_feature_list(tokens[0], lambda x:[]) + create_feature_list(tokens[1:], code_caps)
+  return create_feature_list(tokens, lambda x:[])
+
[…]

### 2013-09-03 src/clusterer/import_data.py

@@ -1,0 +1,144 @@
+#!/usr/bin/python
+
+import csv
+import os
+import sqlite3 as db
+import re
+import nltk
+import feature_extraction as fe
+import naive_bayes as nb
+import operator
+import math
+import sys
+
+sys.setrecursionlimit(10000)
+
+db_name = 'clusters.db'
+clusters_dir = '/home/fil/Dropbox/crowbrain_share/data/idea-clusters-2013.08.19/'
+
+def create_tables(cursor):
+  sql = [
+    '''CREATE TABLE IF NOT EXISTS
+          ideas (id INTEGER PRIMARY KEY AUTOINCREMENT,
+                 cluster_num INTEGER NOT NULL,
+                 idea TEXT NOT NULL,
+                 idea_num INTEGER NOT NULL,
+                 worker_id TEXT NOT NULL,
+                 post_date TEXT NOT NULL,
+                 num_ideas_requested INTEGER NOT NULL,
+                 question_code TEXT NOT NULL)''',
+    '''CREATE TABLE IF NOT EXISTS
+          used_ideas(idea_id INTEGER REFERENCES ideas(id))''',
+    '''CREATE TABLE IF NOT EXISTS
+          similarity_scores (idea1 INTEGER NOT NULL REFERENCES ideas(id),
+                             idea2 INTEGER NOT NULL REFERENCES ideas(id),
+                             score DOUBLE)''',
+    '''CREATE TABLE IF NOT EXISTS
+          clusters (question_code TEXT NOT NULL PRIMARY KEY,
+                    cluster_text TEXT NOT NULL)''',
+    '''CREATE INDEX IF NOT EXISTS
+          ideas_idea_index ON ideas(idea)''',
+    '''CREATE INDEX IF NOT EXISTS
+          used_ideas_index ON used_ideas(idea_id)''',
+    '''CREATE INDEX IF NOT EXISTS
+          similarity_score_index_1 ON similarity_scores(idea1)''',
+    '''CREATE INDEX IF NOT EXISTS
+          similarity_score_index_2 ON similarity_scores(idea2)''',
+    '''CREATE INDEX IF NOT EXISTS
+          similarity_score_index_3 ON similarity_scores(idea1, idea2)''',
+  ]
+  for s in sql:
+    cursor.execute(s)
+
+def import_file(fname, cursor):
+  with open(fname) as fin:
+    print "Importing", fname
+    reader = csv.reader(fin)
+    reader.next() # Read header
+    for row in reader:
[…]

### 2013-09-03 src/clusterer/naive_bayes.py

@@ -1,0 +1,211 @@
+#!/usr/bin/python
+
+import operator
+import re
+import pprint
+import math
+
+# Dependencies: nltk, WordNet for nltk, and hunspell (installed locally, only if spell checking)
+
+TAG_COUNTS = 0
+TAG_TO_FEATURE_COUNTS = 1
+TAG_PROB = 2
+TAG_TO_FEATURE_PROB = 3
+STALE_PROBABILITIES_FLAG = 4
+def create_model():
+# First two dictionaries are the count dictionaries, second two are the probability dictionaries
+  return [{}, {}, {}, {}, False]
+
+# instance_features are the extracted features for an instance. instance_labels is a list of strings associated with this instance and its features
+def update_model(model, instance_features, instance_labels):
+  tag_counts = model[TAG_COUNTS]
+  tag_to_feature_counts = model[TAG_TO_FEATURE_COUNTS]
+  update_counts_from_feature_set(instance_features, instance_labels, tag_counts, tag_to_feature_counts)
+  model[STALE_PROBABILITIES_FLAG] = True
+
+# Removes the data represented by this instance's features and labels from the model
+def remove_from_model(model, instance_features, instance_labels):
+  tag_counts = model[TAG_COUNTS]
+  tag_to_feature_counts = model[TAG_TO_FEATURE_COUNTS]
+  remove_counts_from_feature_set(instance_features, instance_labels, tag_counts, tag_to_feature_counts)
+  model[STALE_PROBABILITIES_FLAG] = True
+  
+# features are the extracted feature set
+# Returns list of (tag, log(prob)) tuples, ordered most to least likely
+def label_features(model, features):
+  if model[STALE_PROBABILITIES_FLAG]:
+    update_model_probabilities_from_counts(model)
+  return get_tag_predictions_from_feature_set(features, model[TAG_PROB], model[TAG_TO_FEATURE_PROB])
+
+def print_model_info(model):
+  print "NB Model:"
+  print "Num different labels:", len(model[TAG_COUNTS])
+  print "Labels:", model[TAG_COUNTS].keys()
+  print "Total number of label data points:", reduce(operator.add, model[TAG_COUNTS].values(), 0)
+
+def pretty_print_model(model):
+  import pprint
+  print "Tag counts:"
+  pprint.pprint(model[TAG_COUNTS])
+  print
+  print "Tag probabilities:"
+  pprint.pprint(model[TAG_PROB])
+  print
+  print "Token counts:"
[…]

### 2013-09-03 src/clusterer/qt_app.py

@@ -1,0 +1,553 @@
+#!/usr/bin/python
+import os
+import sys
+import json
+import re
+import string
+from PySide import QtCore, QtGui
+import sqlite3 as db
+import clustering_app
+import csv
+from collections import deque, defaultdict
+
+db_file_name = 'clusters.db'
+
+def get_line_indent(l):
+  depth_count = 0
+  l = l.strip()
+  front = l
+  while front and front[0] == '-':
+    depth_count = depth_count + 1
+    front = front[1:]
+  return depth_count
+
+class IdeaListModel(QtCore.QAbstractListModel):
+  def __init__(self, parent=None):
+    super(IdeaListModel, self).__init__(parent)
+    self.conn = db.connect(db_file_name)
+    self.cur_question_code = ""
+    self.cur_ideas = [] # Each entry: (id, idea)
+    self.last_label = None # hack
+    self._init_from_db()
+  def _init_from_db(self):
+    self.cur_question_code = self.get_question_codes()[0]
+    self.read_in_ideas()
+  def read_in_ideas(self):
+    self.cur_ideas = []
+    cursor = self.conn.cursor()
+    cursor.execute("SELECT id, idea FROM ideas WHERE question_code = ? AND id NOT IN (SELECT idea_id FROM used_ideas) ORDER BY id ASC", (self.cur_question_code,))
+    for idea_id, idea in cursor.fetchall():
+      self.cur_ideas.append((idea_id, idea))
+    cursor.close()
+
+  def resolve_right(self):
+    # This is so awful, please never look at this awful code again
+    text = self.get_cluster_text()
+    old_count = len(set(self.extract_idea_ids_from_text(text)))
+
+    new_text = ""
+
+    present_ids = []
+    not_found_answers = []
+
+    id_stacks = defaultdict(list)
+
+    # find real IDs
+    cursor = self.conn.cursor()    
+    for l in text.splitlines():
+      id_loc = l.find('(_')
+      if id_loc > 0:
+        old_id = l[id_loc+2:-1]
+        l_p = l[0:id_loc - 2]
+        l_p = l_p.strip('-')
+        
+        if l_p not in id_stacks:
+          cursor.execute("SELECT id FROM ideas WHERE question_code = ? AND idea = ?", (self.cur_question_code, l_p,))
+          for (idea_id,) in cursor.fetchall():
+            id_stacks[l_p].append(idea_id)
+
+        # check if idea is somehow not in the new list
+        if l_p not in id_stacks:
+          not_found_answers.append(l)
[…]

### 2013-09-03 src/clusterer/import_data.py
IdeaTreeModel with import from text, save to SQL. All untested, of course

@@ -20,7 +20,6 @@ def create_tables(cursor):
   sql = [
     '''CREATE TABLE IF NOT EXISTS
           ideas (id INTEGER PRIMARY KEY AUTOINCREMENT,
-                 cluster_num INTEGER NOT NULL,
                  idea TEXT NOT NULL,
                  idea_num INTEGER NOT NULL,
                  worker_id TEXT NOT NULL,
@@ -34,8 +33,16 @@ def create_tables(cursor):
                              idea2 INTEGER NOT NULL REFERENCES ideas(id),
                              score DOUBLE)''',
     '''CREATE TABLE IF NOT EXISTS
-          clusters (question_code TEXT NOT NULL PRIMARY KEY,
-                    cluster_text TEXT NOT NULL)''',
+          clusters (id INTEGER PRIMARY KEY AUTOINCREMENT,
+                    question_code TEXT NOT NULL,
+                    label TEXT NOT NULL)''',
+    '''CREATE TABLE IF NOT EXISTS
+          cluster_heirarchy (child INTEGER NOT NULL REFERENCES clusters(id),
+                             parent INTEGER NOT NULL REFERENCES clusters(id))''',
+    '''CREATE TABLE IF NOT EXISTS
+          idea_clusters (idea_id INTEGER NOT NULL REFERENCES ideas(id),
+                         cluster_id INTEGER NOT NULL REFERENCES clusters(id),
+                         UNIQUE(idea_id) ON CONFLICT REPLACE)''',
     '''CREATE INDEX IF NOT EXISTS
           ideas_idea_index ON ideas(idea)''',
     '''CREATE INDEX IF NOT EXISTS

### 2013-09-03 src/clusterer/qt_app.py

@@ -10,7 +10,7 @@ import clustering_app
 import csv
 from collections import deque, defaultdict
 
-db_file_name = 'clusters.db'
+db_file_name = '~/Dropbox/clusterer/clusters.db'
 
 def get_line_indent(l):
   depth_count = 0
@@ -20,6 +20,173 @@ def get_line_indent(l):
     depth_count = depth_count + 1
     front = front[1:]
   return depth_count
+
+def extract_idea_ids_from_text(s):
+  matches = []
+  for match in re.findall('\(_id:[0-9]+\)', s):
+    matches.append(int(re.findall('[0-9]+', match)[0]))
+  return matches
+
+class IdeaTreeNode(object):
+  def __init__(self, ideas, parent, label = None):
+    assert not isinstance(ideas, basestring)
+    self.label = label
+    self.ideas = ideas # list of (text, id) tuples
+    self.parent = parent
+    self.children = []
+
+  def merge(self, other_node):
+    self.ideas.extend(other_node.ideas)
+
+  def append_child(self, child_node):
+    self.children.append(child_node)
+
+  def child_count(self):
+    return len(self.children)
+
+  def child(self, row):
+    return self.children[row]
+
+  def label(self):
+    if self.label:
+      return self.label
+    else if len(self.ideas) > 0:
+      return self.ideas[0][0][:30]
+    else:
+      return "NO LABEL, NO IDEAS"
+
+  def tooltip(self):
+    return '\n'.join(self.ideas)
+
+  def row(self):
+    if self.parent:
+      return self.parent.children.index(self)
+    return 0
+
+  def parent(self):
+    return self.parent
+
+class IdeaTreeModel(QtCore.QAbstractItemModel):
+  def __init__(self, question_code, parent=None):
+    super(IdeaTreeModel, self).__init__(parent)
+    self.root = IdeaTreeNode([], None, question_code)
+    self.conn = db.connect(db_file_name)
+    self.question_code = question_code
+
+  def columnCount(self, parent = None):
+    return 1
+
+  def data(self, index, role):
+    item = index.internalPointer()
+    if role == QtCore.Qt.DisplayRole:
+      return item.label()
+    else if role == QtCore.Qt.ToolTipRole:
+      return item.tooltip()
+    return None
+
+  def index(self, row, column, parent):
+    p = parent.internalPointer if parent.isValid() else self.root
+    c = p.child(row)
+    return self.createIndex(row, column, c)
+
+  def parent(self, index):
[…]

### 2013-09-03 src/clusterer/qt_app.py
redone display, save, load code is idea-complete sans testing

@@ -97,6 +97,9 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
 
     return self.createIndex(p.row(), 0, p)
 
+  def rowCount(self, parent=QtCore.QModelIndex()):
+    p = parent.internalPointer() if parent.isValid() else self.root
+    return p.child_count()
 
   def get_next_depth(self, dq):
     for l in dq:
@@ -179,12 +182,81 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
 
       return clus_id
 
-
     def save(self):
       cursor = self.conn.cursor()
       cursor.execute("DELETE FROM clusters WHERE question_code=?")
 
-      self._save_node(self.root)
+      self._save_node(self.root, cursor)
+
+      cursor.close()
+
+    def load(self):
+      cursor = self.conn.cursor()
+
+      cluster_dict = dict()
+
+      # Load clusters
+      cursor.execute("""SELECT id, label FROM clusters
+                        WHERE question_code = ?""", (self.question_code,))
+      for i, l in cursor.fetchall():
+        node = IdeaTreeNode([], None, l)
+        cluster_dict[i] = node
+
+      # Organize into tree
+      cursor.execute("""SELECT child, parent FROM cluster_hierarchy
+                        WHERE child IN (SELECT id FROM clusters
+                                        WHERE question_code = ?)""",
+                     (self.question_code,))
+      for c, p in cursor.fetchall():
+        child = cluster_dict[c]
+        parent = cluster_dict[p]
+
+        child.parent = parent
+        parent.append_child(child)
+
+      # Load ideas into clusters
+      cursor.execute("""SELECT cluster_id, id, idea
+                        FROM ideas INNER JOIN idea_clusters
+                        ON ideas.id = idea_clusters.idea_id
+                        WHERE ideas.question_code = ?""",
+                     (self.question_code,))
+      for cid, iid, idea in cursor.fetchall():
+        cluster_dict[cid].ideas.append((iid, idea))
+
+      cursor.close()
+
+    def export_clusters(self, filename):
+
+      cursor = self.conn.cursor()
+      with open("%s_%s_clusters.csv" % (filename, question_code), 'w') as cfout:
+        cwriter = csv.writer(cfout)
+        cwriter.writerow(['question_code', 'cluster', 'cluster_parent', 'cluster_label'])
+
+        cursor.execute("""SELECT child, parent, label
[…]

### 2013-09-03 bootstrap.sh
can import from text successfully

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 apt-get update
-apt-get install -y python-pip python-dev python-requests build-essential sqlite3 libfreetype6-dev libpng-dev python-scipy python-matplotlib python-tornado python-zmq python-nltk python-pandas pandoc python-pygments python-sphinx
+apt-get install -y python-pip python-dev python-requests build-essential sqlite3 libfreetype6-dev libpng-dev python-scipy python-matplotlib python-tornado python-zmq python-nltk python-pandas pandoc python-pygments python-sphinx pyside-tools
 pip install --upgrade ipython
 pip install turkflow numpy pyyaml
 

### 2013-09-03 src/clusterer/clustering_app.py

@@ -2,7 +2,7 @@
 
 # Form implementation generated from reading ui file 'clustering_app.ui'
 #
-# Created: Fri Aug 23 09:59:55 2013
+# Created: Tue Sep  3 19:14:11 2013
 #      by: pyside-uic 0.2.14 running on PySide 1.2.0
 #
 # WARNING! All changes made in this file will be lost!
@@ -47,26 +47,19 @@ class Ui_MainWindow(object):
         self.button_move_down = QtGui.QPushButton(self.centralwidget)
         self.button_move_down.setObjectName("button_move_down")
         self.horizontalLayout_5.addWidget(self.button_move_down)
-        self.button_move_down_split = QtGui.QPushButton(self.centralwidget)
-        self.button_move_down_split.setObjectName("button_move_down_split")
-        self.horizontalLayout_5.addWidget(self.button_move_down_split)
         self.verticalLayout_4.addLayout(self.horizontalLayout_5)
         self.horizontalLayout_8 = QtGui.QHBoxLayout()
         self.horizontalLayout_8.setObjectName("horizontalLayout_8")
         self.button_resolve_lost = QtGui.QPushButton(self.centralwidget)
         self.button_resolve_lost.setObjectName("button_resolve_lost")
         self.horizontalLayout_8.addWidget(self.button_resolve_lost)
-        self.button_resolve_right = QtGui.QPushButton(self.centralwidget)
-        self.button_resolve_right.setObjectName("button_resolve_right")
-        self.horizontalLayout_8.addWidget(self.button_resolve_right)
         self.verticalLayout_4.addLayout(self.horizontalLayout_8)
         self.horizontalLayout_3.addLayout(self.verticalLayout_4)
         self.verticalLayout_5 = QtGui.QVBoxLayout()
         self.verticalLayout_5.setObjectName("verticalLayout_5")
-        self.text_edit_clusters = QtGui.QPlainTextEdit(self.centralwidget)
-        self.text_edit_clusters.setLineWrapMode(QtGui.QPlainTextEdit.NoWrap)
-        self.text_edit_clusters.setObjectName("text_edit_clusters")
-        self.verticalLayout_5.addWidget(self.text_edit_clusters)
+        self.tree_main = QtGui.QTreeView(self.centralwidget)
+        self.tree_main.setObjectName("tree_main")
+        self.verticalLayout_5.addWidget(self.tree_main)
[…]

### 2013-09-03 src/clusterer/import_data.py

@@ -13,7 +13,7 @@ import sys
 
 sys.setrecursionlimit(10000)
 
-db_name = 'clusters.db'
+db_name = '/home/fil/Dropbox/clusters.db'
 clusters_dir = '/home/fil/Dropbox/crowbrain_share/data/idea-clusters-2013.08.19/'
 
 def create_tables(cursor):
@@ -22,6 +22,7 @@ def create_tables(cursor):
           ideas (id INTEGER PRIMARY KEY AUTOINCREMENT,
                  idea TEXT NOT NULL,
                  idea_num INTEGER NOT NULL,
+                 auto_cluster_num INTEGER NOT NULL,
                  worker_id TEXT NOT NULL,
                  post_date TEXT NOT NULL,
                  num_ideas_requested INTEGER NOT NULL,
@@ -63,7 +64,7 @@ def import_file(fname, cursor):
     reader = csv.reader(fin)
     reader.next() # Read header
     for row in reader:
-      cursor.execute("INSERT INTO ideas(cluster_num, idea, idea_num, worker_id, post_date, num_ideas_requested, question_code) VALUES(?, ?, ?, ?, ?, ?, ?)", row)
+      cursor.execute("INSERT INTO ideas(auto_cluster_num, idea, idea_num, worker_id, post_date, num_ideas_requested, question_code) VALUES(?, ?, ?, ?, ?, ?, ?)", row)
 
 def get_stems(cursor):
   cursor.execute('SELECT idea FROM ideas')

### 2013-09-03 src/clusterer/qt_app.py

@@ -7,10 +7,11 @@ import string
 from PySide import QtCore, QtGui
 import sqlite3 as db
 import clustering_app
+import import_cluster_text
 import csv
 from collections import deque, defaultdict
 
-db_file_name = '~/Dropbox/clusterer/clusters.db'
+db_file_name = '/home/fil/Dropbox/clusters.db'
 
 def get_line_indent(l):
   depth_count = 0
@@ -29,8 +30,8 @@ def extract_idea_ids_from_text(s):
 
 class IdeaTreeNode(object):
   def __init__(self, ideas, parent, label = None):
-    assert not isinstance(ideas, basestring)
-    self.label = label
+    assert not isinstance(ideas, str)
+    self._label = label
     self.ideas = ideas # list of (text, id) tuples
     self.parent = parent
     self.children = []
@@ -48,22 +49,22 @@ class IdeaTreeNode(object):
     return self.children[row]
 
   def label(self):
-    if self.label:
-      return self.label
-    else if len(self.ideas) > 0:
+    if self._label:
+      return self._label
+    elif len(self.ideas) > 0:
       return self.ideas[0][0][:30]
     else:
       return "NO LABEL, NO IDEAS"
 
   def tooltip(self):
-    return '\n'.join(self.ideas)
+    return '\n'.join(i[0] for i in self.ideas)
 
   def row(self):
     if self.parent:
       return self.parent.children.index(self)
     return 0
 
-  def parent(self):
+  def get_parent(self):
     return self.parent
 
 class IdeaTreeModel(QtCore.QAbstractItemModel):
@@ -80,18 +81,18 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
     item = index.internalPointer()
     if role == QtCore.Qt.DisplayRole:
       return item.label()
-    else if role == QtCore.Qt.ToolTipRole:
+    elif role == QtCore.Qt.ToolTipRole:
       return item.tooltip()
     return None
 
   def index(self, row, column, parent):
-    p = parent.internalPointer if parent.isValid() else self.root
+    p = parent.internalPointer() if parent.isValid() else self.root
     c = p.child(row)
     return self.createIndex(row, column, c)
 
   def parent(self, index):
-    c = index.InternalPointer()
-    p = c.parent()
+    c = index.internalPointer()
+    p = c.get_parent()
     if p == self.root:
       return QtCore.QModelIndex()
 
@@ -108,7 +109,7 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
[…]

### 2013-09-04 src/clusterer/import_data.py
can save maybe? seems to work

@@ -38,7 +38,7 @@ def create_tables(cursor):
                     question_code TEXT NOT NULL,
                     label TEXT NOT NULL)''',
     '''CREATE TABLE IF NOT EXISTS
-          cluster_heirarchy (child INTEGER NOT NULL REFERENCES clusters(id),
+          cluster_hierarchy (child INTEGER NOT NULL REFERENCES clusters(id),
                              parent INTEGER NOT NULL REFERENCES clusters(id))''',
     '''CREATE TABLE IF NOT EXISTS
           idea_clusters (idea_id INTEGER NOT NULL REFERENCES ideas(id),

### 2013-09-04 src/clusterer/qt_app.py

@@ -166,104 +166,107 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
 
     self.endResetModel()
 
-    def _save_node(self, node, cursor):
-      # Create cluster
-      cursor.execute("""INSERT INTO clusters(question_code, label)
-                        VALUES (?, ?)""", (self.question_code, node.label()))
-      clus_id = cursor.lastrowid
+  def _save_node(self, node, cursor):
+    # Create cluster
+    cursor.execute("""INSERT INTO clusters(question_code, label)
+                      VALUES (?, ?)""", (self.question_code, node.label()))
+    clus_id = cursor.lastrowid
 
-      # Map ideas
-      for i, text in node.ideas:
-        cursor.execute("""INSERT INTO idea_clusters(idea_id, cluster_id)
-                          VALUES (?, ?)""", (i, clus_id))
+    # Map ideas
+    for i, text in node.ideas:
+      cursor.execute("""INSERT INTO idea_clusters(idea_id, cluster_id)
+                        VALUES (?, ?)""", (i, clus_id))
 
-      # Create children
-      child_ids = [self._save_node(c, cursor) for c in node.children]
-      
-      # Create parent-child relationships
-      for child_id in child_ids:
-        cursor.execute("""INSERT INTO cluster_hierarchy(child, parent)
-                          VALUES(?, ?)""", (child_id, clus_id))
+    # Create children
+    child_ids = [self._save_node(c, cursor) for c in node.children]
+    
+    # Create parent-child relationships
+    for child_id in child_ids:
+      cursor.execute("""INSERT INTO cluster_hierarchy(child, parent)
+                        VALUES(?, ?)""", (child_id, clus_id))
 
-      return clus_id
+    return clus_id
 
-    def save(self):
-      cursor = self.conn.cursor()
-      cursor.execute("DELETE FROM clusters WHERE question_code=?")
+  def save(self):
+    cursor = self.conn.cursor()
+    cursor.execute("DELETE FROM clusters WHERE question_code=?",
+                   (self.question_code,))
 
-      self._save_node(self.root, cursor)
+    self._save_node(self.root, cursor)
 
-      cursor.close()
+    self.conn.commit()
+    cursor.close()
+    self.conn.commit()
 
-    def load(self):
-      self.beginResetModel()
-      cursor = self.conn.cursor()
+  def load(self):
+    self.beginResetModel()
+    cursor = self.conn.cursor()
 
[…]

### 2013-09-04 src/clusterer/import_data.py
can save and load, maybe sped up with indices

@@ -49,6 +49,14 @@ def create_tables(cursor):
     '''CREATE INDEX IF NOT EXISTS
           used_ideas_index ON used_ideas(idea_id)''',
     '''CREATE INDEX IF NOT EXISTS
+          clusters_cluster_index ON clusters(id)''',
+    '''CREATE INDEX IF NOT EXISTS
+          cluster_h_cluster_index ON cluster_hierarchy(child)''',
+    '''CREATE INDEX IF NOT EXISTS
+          idea_clusters_idea_index ON idea_clusters(idea_id)''',
+    '''CREATE INDEX IF NOT EXISTS
+          idea_clusters_cluster_index ON idea_clusters(cluster_id)''',
+    '''CREATE INDEX IF NOT EXISTS
           similarity_score_index_1 ON similarity_scores(idea1)''',
     '''CREATE INDEX IF NOT EXISTS
           similarity_score_index_2 ON similarity_scores(idea2)''',

### 2013-09-04 src/clusterer/qt_app.py

@@ -74,6 +74,8 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
     self.conn = db.connect(db_file_name)
     self.question_code = question_code
 
+    self.load()
+
   def columnCount(self, parent = None):
     return 1
 
@@ -166,6 +168,8 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
 
     self.endResetModel()
 
+    idea_model.make_ideas_used(used_ids)
+
   def _save_node(self, node, cursor):
     # Create cluster
     cursor.execute("""INSERT INTO clusters(question_code, label)
@@ -173,7 +177,7 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
     clus_id = cursor.lastrowid
 
     # Map ideas
-    for i, text in node.ideas:
+    for text, i in node.ideas:
       cursor.execute("""INSERT INTO idea_clusters(idea_id, cluster_id)
                         VALUES (?, ?)""", (i, clus_id))
 
@@ -202,6 +206,7 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
     self.beginResetModel()
     cursor = self.conn.cursor()
 
+    root_cluster = None
     cluster_dict = dict()
 
     # Load clusters
@@ -210,6 +215,11 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
     for i, l in cursor.fetchall():
       node = IdeaTreeNode([], None, l)
       cluster_dict[i] = node
+      if root_cluster is None:
+        root_cluster = node
+
+    if root_cluster is None:
+      return
 
     # Organize into tree
     cursor.execute("""SELECT child, parent FROM cluster_hierarchy
@@ -230,9 +240,12 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
                       WHERE ideas.question_code = ?""",
                    (self.question_code,))
     for cid, iid, idea in cursor.fetchall():
-      cluster_dict[cid].ideas.append((iid, idea))
+      cluster_dict[cid].ideas.append((idea, iid))
 
     cursor.close()
+
+    self.root = root_cluster
+
     self.endResetModel()
 
   def export_clusters(self, filename):
@@ -354,18 +367,17 @@ class IdeaListModel(QtCore.QAbstractListModel):
     idea = cursor.fetchone()[0]
     cursor.close()
     return idea
-  # Returns list of strings to display
-  def make_ideas_used(self, indexes):
+
[…]

### 2013-09-04 src/clusterer/clustering_app.py
can import all from text

@@ -2,7 +2,7 @@
 
 # Form implementation generated from reading ui file 'clustering_app.ui'
 #
-# Created: Tue Sep  3 19:14:11 2013
+# Created: Wed Sep  4 10:07:49 2013
 #      by: pyside-uic 0.2.14 running on PySide 1.2.0
 #
 # WARNING! All changes made in this file will be lost!

### 2013-09-04 src/clusterer/qt_app.py

@@ -125,11 +125,17 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
 
     idea, ids = idea_model.get_ids_for_text(l)
     new_ids = [i for i in ids if not i in used_ids]
+    if len(new_ids) == 0:
+        print("Couldn't find idea:", l)
+        return None
     new_id = new_ids[0]
     used_ids.append(new_id)
     return IdeaTreeNode([(idea, new_id)], current_node, idea)
 
   def import_from_text(self, text, idea_model):
+    if len(text) == 0:
+        return
+
     self.beginResetModel()
     used_ids = []
 
@@ -156,6 +162,8 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
           current_node.merge(next_node)
 
         elif depth_count > current_depth:
+          if depth_count - 1 != current_depth:
+            print(line)
           assert depth_count - 1 == current_depth
           current_node.append_child(next_node)
           current_node = next_node

### 2013-09-04 src/clusterer/qt_app.py
first pass at implementing idea addition algorithm

@@ -42,6 +42,10 @@ class IdeaTreeNode(object):
   def append_child(self, child_node):
     self.children.append(child_node)
 
+  def remove_child(self, child_node):
+    if child_node in self.children:
+      del self.children[self.children.index(child_node)]
+
   def child_count(self):
     return len(self.children)
 
@@ -66,6 +70,14 @@ class IdeaTreeNode(object):
 
   def get_parent(self):
     return self.parent
+
+  def get_ids(self):
+    ids = [i[1] for i in self.ideas]
+
+    for c in self.children:
+      ids += c.get_ids()
+
+    return ids
 
 class IdeaTreeModel(QtCore.QAbstractItemModel):
   def __init__(self, question_code, parent=None):
@@ -248,11 +260,27 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
                       WHERE ideas.question_code = ?""",
                    (self.question_code,))
     for cid, iid, idea in cursor.fetchall():
-      cluster_dict[cid].ideas.append((idea, iid))
+      if cid in cluster_dict:
+        cluster_dict[cid].ideas.append((idea, iid))
 
     cursor.close()
 
     self.root = root_cluster
+
+    self.endResetModel()
+
+  def get_ids(self, index):
+    if index.isValid():
+      return index.internalPointer().get_ids()
+    else:
+      return []
+
+  def remove_node(self, index):
+    self.beginResetModel()
+    # This is probably a memory leak, I might be keeping a reference around
+    if index.isValid():
+      node = index.internalPointer()
+      node.parent.remove_child(node)
 
     self.endResetModel()
 
@@ -309,11 +337,10 @@ class IdeaListModel(QtCore.QAbstractListModel):
     for idea_id, idea in cursor.fetchall():
       self.cur_ideas.append((idea_id, idea))
     cursor.close()
+  
+  def get_ideas_for_indices(self, indices):
+    return [self.cur_ideas[i] for i in indices]
 
-  def resolve_right(self):
-    # This is so awful, please never look at this awful code again
-    text = self.get_cluster_text()
-    
   def get_ids_for_text(self, text):
     l = text
     id_loc = l.find('(_')
@@ -332,12 +359,9 @@ class IdeaListModel(QtCore.QAbstractListModel):
     else:
       return None
 
-  def resolve(self):
+  def resolve(self, idea_tree_model):
     # fix lost ideas
-    text = self.get_cluster_text()
[…]

### 2013-09-04 src/clusterer/clustering_app.py
can select similarity cluster. on to coverage

@@ -2,7 +2,7 @@
 
 # Form implementation generated from reading ui file 'clustering_app.ui'
 #
-# Created: Wed Sep  4 10:07:49 2013
+# Created: Wed Sep  4 12:29:07 2013
 #      by: pyside-uic 0.2.14 running on PySide 1.2.0
 #
 # WARNING! All changes made in this file will be lost!

### 2013-09-04 src/clusterer/qt_app.py

@@ -8,6 +8,7 @@ from PySide import QtCore, QtGui
 import sqlite3 as db
 import clustering_app
 import import_cluster_text
+import similarity_selector
 import csv
 from collections import deque, defaultdict
 
@@ -79,10 +80,18 @@ class IdeaTreeNode(object):
 
     return ids
 
+  def get_all_ideas(self):
+    ret = self.ideas
+
+    for c in self.children:
+      ret += c.get_all_ideas()
+
+    return ret
+
 class IdeaTreeModel(QtCore.QAbstractItemModel):
-  def __init__(self, question_code, parent=None):
+  def __init__(self, question_code, parent=None, root = None):
     super(IdeaTreeModel, self).__init__(parent)
-    self.root = IdeaTreeNode([], None, question_code)
+    self.root = IdeaTreeNode([], None, question_code) if not root else root
     self.conn = db.connect(db_file_name)
     self.question_code = question_code
 
@@ -341,6 +350,23 @@ class IdeaListModel(QtCore.QAbstractListModel):
   def get_ideas_for_indices(self, indices):
     return [self.cur_ideas[i] for i in indices]
 
+  def mean_similarity(self, group1_ids, group2_ids):
+    cursor = self.conn.cursor()
+    tups = [(x, y) for x in group1_ids for y in group2_ids]
+
+    scores = []
+
+    for id1, id2 in tups:
+      cursor.execute("""SELECT score FROM similarity_scores
+                        WHERE idea1 = ? AND idea2 = ?""", (id1, id2))
+      res = cursor.fetchone()
+      if res:
+        scores.append(res[0])
+      else:
+        print("Missing similarity score", id1, id2)
+
+    return sum(scores) / len(scores)
+
   def get_ids_for_text(self, text):
     l = text
     id_loc = l.find('(_')
@@ -553,13 +579,60 @@ class AppWindow(QtGui.QMainWindow):
     self.idea_tree_models[qc].import_from_text(text, self.idea_model)
 
   def coverage_prompt(self, node1, node2):
-    print "TODO"
+    print("TODO")
 
-  def affinity_prompt(self, node, compare_nodes):
-    print "TODO"
+  def similarity_prompt(self, node, compare_root):
+    scores = []
+
+    nis = node.get_all_ideas()
+
+    for c in compare_root.children:
+      cis = c.get_all_ideas()
+      if len(cis) == 0:
+        print(c.label())
+      s = self.idea_model.mean_similarity([i[1] for i in nis],
[…]

### 2013-09-04 src/clusterer/clustering_app.py
all gialogs implemented, but not sure if it works

@@ -2,7 +2,7 @@
 
 # Form implementation generated from reading ui file 'clustering_app.ui'
 #
-# Created: Wed Sep  4 12:29:07 2013
+# Created: Wed Sep  4 13:28:07 2013
 #      by: pyside-uic 0.2.14 running on PySide 1.2.0
 #
 # WARNING! All changes made in this file will be lost!

### 2013-09-04 src/clusterer/qt_app.py

@@ -9,6 +9,8 @@ import sqlite3 as db
 import clustering_app
 import import_cluster_text
 import similarity_selector
+import coverage
+import title
 import csv
 from collections import deque, defaultdict
 
@@ -95,7 +97,8 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
     self.conn = db.connect(db_file_name)
     self.question_code = question_code
 
-    self.load()
+    if root is None:
+      self.load()
 
   def columnCount(self, parent = None):
     return 1
@@ -579,7 +582,28 @@ class AppWindow(QtGui.QMainWindow):
     self.idea_tree_models[qc].import_from_text(text, self.idea_model)
 
   def coverage_prompt(self, node1, node2):
-    print("TODO")
+    dialog = QtGui.QDialog()
+    dialog.ui = coverage.Ui_Dialog()
+    dialog.ui.setupUi(dialog)
+
+    dialog.ui.lbl_idea1.setText(node1.label())
+    dialog.ui.lbl_idea2.setText(node2.label())
+
+    dialog.ui.btn_both_high.clicked.connect(lambda x=0: dialog.done(x))
+    dialog.ui.btn_both_low.clicked.connect(lambda x=1: dialog.done(x))
+    dialog.ui.btn_12.clicked.connect(lambda x=2: dialog.done(x))
+    dialog.ui.btn_21.clicked.connect(lambda x=3: dialog.done(x))
+
+    ret = dialog.exec()
+
+    if ret == 0:
+      return True, True
+    elif ret == 1:
+      return False, False
+    elif ret == 2:
+      return True, False
+    elif ret == 3:
+      return False, True
 
   def similarity_prompt(self, node, compare_root):
     scores = []
@@ -631,8 +655,22 @@ class AppWindow(QtGui.QMainWindow):
 
     return None
 
-  def cluster_label_prompt(self, ideas):
-    print("TODO")
+  def cluster_label_prompt(self, node):
+    dialog = QtGui.QDialog()
+    dialog.ui = title.Ui_Dialog()
+    dialog.ui.setupUi(dialog)
+
+    dialog.ui.btn_ok.clicked.connect(lambda x=0: dialog.done(x))
+    dialog.ui.btn_none.clicked.connect(lambda x=1: dialog.done(x))
+
+    itm = IdeaTreeModel(self.idea_model.cur_question_code, root = node)
+    dialog.ui.treeView.setModel(itm)
+
[…]

### 2013-09-04 src/clusterer/qt_app.py
adding an idea works in one, perfect case

@@ -44,9 +44,11 @@ class IdeaTreeNode(object):
 
   def append_child(self, child_node):
     self.children.append(child_node)
+    child_node.parent = self
 
   def remove_child(self, child_node):
     if child_node in self.children:
+      child_node.parent = None
       del self.children[self.children.index(child_node)]
 
   def child_count(self):
@@ -201,6 +203,10 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
     self.endResetModel()
 
     idea_model.make_ideas_used(used_ids)
+
+  def bad_fake_reset(self):
+    self.beginResetModel()
+    self.endResetModel()
 
   def _save_node(self, node, cursor):
     # Create cluster
@@ -663,7 +669,13 @@ class AppWindow(QtGui.QMainWindow):
     dialog.ui.btn_ok.clicked.connect(lambda x=0: dialog.done(x))
     dialog.ui.btn_none.clicked.connect(lambda x=1: dialog.done(x))
 
-    itm = IdeaTreeModel(self.idea_model.cur_question_code, root = node)
+    # Add a fake parent if there is none
+    if node.parent is None:
+      fake_parent = IdeaTreeNode([], None)
+      fake_parent.append_child(node)
+      node.parent = fake_parent
+
+    itm = IdeaTreeModel(self.idea_model.cur_question_code, root = node.parent)
     dialog.ui.treeView.setModel(itm)
 
     ret = dialog.exec()
@@ -679,8 +691,6 @@ class AppWindow(QtGui.QMainWindow):
     indexes = [i.row() for i in self.ui.list_ideas.selectedIndexes()]
     ideas = self.idea_model.get_ideas_for_indices(indexes)
 
-    #self.idea_model.make_ideas_used([i[0] for i in ideas])
-
     # I stored it backwards from Mike
     ideas_flipped = [(i[1], i[0]) for i in ideas]
 
@@ -688,13 +698,11 @@ class AppWindow(QtGui.QMainWindow):
     self.cluster_label_prompt(new_node)
     current_node = itm.root
 
-    return
-
     while True:
       best_match = self.similarity_prompt(new_node, current_node)
 
       if best_match is None:
-        # current_node.append_child(new_node)
+        current_node.append_child(new_node)
         break
 
       else:
@@ -707,7 +715,7 @@ class AppWindow(QtGui.QMainWindow):
           # Create new node under the current node
           # TODO: move this into the tree class
[…]

### 2013-09-04 src/clusterer/qt_app.py
some niceties

@@ -42,9 +42,13 @@ class IdeaTreeNode(object):
   def merge(self, other_node):
     self.ideas.extend(other_node.ideas)
 
+  def sort_children(self):
+    self.children = sorted(self.children, key=lambda x: x.label().lower())
+
   def append_child(self, child_node):
     self.children.append(child_node)
     child_node.parent = self
+    self.sort_children()
 
   def remove_child(self, child_node):
     if child_node in self.children:
@@ -61,7 +65,7 @@ class IdeaTreeNode(object):
     if self._label:
       return self._label
     elif len(self.ideas) > 0:
-      return self.ideas[0][0][:30]
+      return self.ideas[0][0]
     else:
       return "NO LABEL, NO IDEAS"
 
@@ -280,6 +284,12 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
     for cid, iid, idea in cursor.fetchall():
       if cid in cluster_dict:
         cluster_dict[cid].ideas.append((idea, iid))
+
+    for key in cluster_dict:
+      # Don't load useless clusters (artifacts from import)
+      c = cluster_dict[key]
+      if len(c.get_all_ideas()) == 0:
+        c.parent.remove_child(c)
 
     cursor.close()
 
@@ -594,6 +604,8 @@ class AppWindow(QtGui.QMainWindow):
 
     dialog.ui.lbl_idea1.setText(node1.label())
     dialog.ui.lbl_idea2.setText(node2.label())
+    dialog.ui.lbl_idea1.setToolTip(node1.tooltip())
+    dialog.ui.lbl_idea2.setToolTip(node2.tooltip())
 
     dialog.ui.btn_both_high.clicked.connect(lambda x=0: dialog.done(x))
     dialog.ui.btn_both_low.clicked.connect(lambda x=1: dialog.done(x))
@@ -619,7 +631,7 @@ class AppWindow(QtGui.QMainWindow):
     for c in compare_root.children:
       cis = c.get_all_ideas()
       if len(cis) == 0:
-        print(c.label())
+        print("No idea under", c.label())
       s = self.idea_model.mean_similarity([i[1] for i in nis],
                                           [i[1] for i in cis])
       scores.append((c, s))
@@ -677,6 +689,7 @@ class AppWindow(QtGui.QMainWindow):
 
     itm = IdeaTreeModel(self.idea_model.cur_question_code, root = node.parent)
[…]

### 2013-09-05 src/clusterer/clustering_app.py
clustering app small updates

@@ -2,7 +2,7 @@
 
 # Form implementation generated from reading ui file 'clustering_app.ui'
 #
-# Created: Wed Sep  4 13:28:07 2013
+# Created: Thu Sep  5 17:39:06 2013
 #      by: pyside-uic 0.2.14 running on PySide 1.2.0
 #
 # WARNING! All changes made in this file will be lost!

### 2013-09-05 src/clusterer/qt_app.py

@@ -39,8 +39,13 @@ class IdeaTreeNode(object):
     self.parent = parent
     self.children = []
 
+  def add_ideas(self, ideas):
+    present_ids = [iid for idea, iid in self.ideas]
+    ideas = [(idea, iid) for (idea, iid) in ideas if not iid in present_ids]
+    self.ideas.extend(ideas)
+
   def merge(self, other_node):
-    self.ideas.extend(other_node.ideas)
+    self.add_ideas(other_node.ideas)
 
   def sort_children(self):
     self.children = sorted(self.children, key=lambda x: x.label().lower())
@@ -58,19 +63,35 @@ class IdeaTreeNode(object):
   def child_count(self):
     return len(self.children)
 
+  def all_child_nodes(self):
+    ret = self.children.copy()
+
+    for c in self.children:
+      ret += c.all_child_nodes()
+
+    return ret
+
   def child(self, row):
+    if len(self.children) <= row:
+      return None
     return self.children[row]
 
   def label(self):
     if self._label:
       return self._label
-    elif len(self.ideas) > 0:
-      return self.ideas[0][0]
+    elif len(self.get_all_ideas()) > 0:
+      return self.get_all_ideas()[0][0]
     else:
       return "NO LABEL, NO IDEAS"
 
+  def long_label(self):
+    if self.parent is not None:
+      return self.parent.long_label() + "/" + self.label()
+    else:
+      return self.label()
+
   def tooltip(self):
-    return '\n'.join(i[0] for i in self.ideas)
+    return '\n'.join(i[0] for i in self.get_all_ideas())
 
   def row(self):
     if self.parent:
@@ -89,7 +110,7 @@ class IdeaTreeNode(object):
     return ids
 
   def get_all_ideas(self):
-    ret = self.ideas
+    ret = self.ideas.copy()
 
     for c in self.children:
       ret += c.get_all_ideas()
@@ -120,6 +141,8 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
   def index(self, row, column, parent):
     p = parent.internalPointer() if parent.isValid() else self.root
     c = p.child(row)
+    if c is None:
+      return None
     return self.createIndex(row, column, c)
 
   def parent(self, index):
@@ -283,13 +306,14 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
                    (self.question_code,))
     for cid, iid, idea in cursor.fetchall():
[…]

### 2013-09-06 src/clusterer/import_data.py
use new clusters for import

@@ -14,7 +14,7 @@ import sys
 sys.setrecursionlimit(10000)
 
 db_name = '/home/fil/Dropbox/clusters.db'
-clusters_dir = '/home/fil/Dropbox/crowbrain_share/data/idea-clusters-2013.08.19/'
+clusters_dir = '/home/fil/Dropbox/crowbrain_share/data/idea-clusters-2013.09.05/'
 
 def create_tables(cursor):
   sql = [

### 2013-09-06 src/clusterer/qt_app.py
better tooltips

@@ -86,12 +86,22 @@ class IdeaTreeNode(object):
 
   def long_label(self):
     if self.parent is not None:
-      return self.parent.long_label() + "/" + self.label()
+      l = self.parent.label()
+      if l == 'forgot_name' or l == 'iPod' or l == 'charity' or l == 'turk':
+        return self.label()
+      else:
+        return self.parent.long_label() + "/" + self.label()
     else:
       return self.label()
 
   def tooltip(self):
-    return '\n'.join(i[0] for i in self.get_all_ideas())
+    ret = '\n'.join(i[0] for i in self.ideas)
+
+    for c in self.children:
+      tt = c.tooltip()
+      ret = ret + '\n' + '\n'.join(['\t' + t for t in tt.split('\n')]) + '\n'
+
+    return ret
 
   def row(self):
     if self.parent:
@@ -746,19 +756,26 @@ class AppWindow(QtGui.QMainWindow):
         break
 
       else:
+        bm_parent = best_match.parent
+        if bm_parent is None:
+          assert(False)
+
         # each is true if high, false if low
         cov_n_b, cov_b_n = self.coverage_prompt(new_node, best_match)
         if cov_n_b == cov_b_n and cov_n_b:
           best_match.merge(new_node)
           new_node = best_match
+
+          if best_match._label is None:
+            self.cluster_label_prompt(best_match)
           break
         elif cov_n_b == cov_b_n:
           # Create new node under the current node
           # TODO: move this into the tree class
           new_parent = IdeaTreeNode([], current_node, None)
-          current_node.append_child(new_parent)
+          bm_parent.append_child(new_parent)
 
-          current_node.remove_child(best_match)
+          bm_parent.remove_child(best_match)
 
           new_parent.append_child(best_match)
           new_parent.append_child(new_node)
@@ -774,10 +791,9 @@ class AppWindow(QtGui.QMainWindow):
           if new_node._label is None:
             self.cluster_label_prompt(new_node)
 
-          current_node.remove_child(best_match)
-          current_node.append_child(new_node)
-          current_node = new_node
-          current_node.append_child(best_match)
+          bm_parent.remove_child(best_match)
+          bm_parent.append_child(new_node)
+          new_node.append_child(best_match)
 
           self.cluster_label_prompt(best_match)
           break

### 2013-09-06 src/clusterer/import_data.py
should have fixed the saving of fake labels, but it's not a big deal either way

@@ -36,7 +36,7 @@ def create_tables(cursor):
     '''CREATE TABLE IF NOT EXISTS
           clusters (id INTEGER PRIMARY KEY AUTOINCREMENT,
                     question_code TEXT NOT NULL,
-                    label TEXT NOT NULL)''',
+                    label TEXT)''',
     '''CREATE TABLE IF NOT EXISTS
           cluster_hierarchy (child INTEGER NOT NULL REFERENCES clusters(id),
                              parent INTEGER NOT NULL REFERENCES clusters(id))''',

### 2013-09-06 src/clusterer/qt_app.py

@@ -34,7 +34,10 @@ def extract_idea_ids_from_text(s):
 class IdeaTreeNode(object):
   def __init__(self, ideas, parent, label = None):
     assert not isinstance(ideas, str)
-    self._label = label
+    if label is not None and len(label) > 0:
+      self._label = label
+    else:
+      self._label = None
     self.ideas = ideas # list of (text, id) tuples
     self.parent = parent
     self.children = []
@@ -248,7 +251,7 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
   def _save_node(self, node, cursor):
     # Create cluster
     cursor.execute("""INSERT INTO clusters(question_code, label)
-                      VALUES (?, ?)""", (self.question_code, node.label()))
+                      VALUES (?, ?)""", (self.question_code, node._label))
     clus_id = cursor.lastrowid
 
     # Map ideas

### 2013-09-06 src/clusterer/clustering_app.py
ignore some files

@@ -2,7 +2,7 @@
 
 # Form implementation generated from reading ui file 'clustering_app.ui'
 #
-# Created: Thu Sep  5 17:39:06 2013
+# Created: Fri Sep  6 09:31:24 2013
 #      by: pyside-uic 0.2.14 running on PySide 1.2.0
 #
 # WARNING! All changes made in this file will be lost!

### 2013-09-06 src/clusterer/clustering_app.py
remove clustering app python file from tracking

@@ -1,128 +1,0 @@
-# -*- coding: utf-8 -*-
-
-# Form implementation generated from reading ui file 'clustering_app.ui'
-#
-# Created: Fri Sep  6 09:31:24 2013
-#      by: pyside-uic 0.2.14 running on PySide 1.2.0
-#
-# WARNING! All changes made in this file will be lost!
-
-from PySide import QtCore, QtGui
-
-class Ui_MainWindow(object):
-    def setupUi(self, MainWindow):
-        MainWindow.setObjectName("MainWindow")
-        MainWindow.resize(840, 755)
-        self.centralwidget = QtGui.QWidget(MainWindow)
-        self.centralwidget.setObjectName("centralwidget")
-        self.verticalLayout = QtGui.QVBoxLayout(self.centralwidget)
-        self.verticalLayout.setObjectName("verticalLayout")
-        self.verticalLayout_3 = QtGui.QVBoxLayout()
-        self.verticalLayout_3.setObjectName("verticalLayout_3")
-        self.horizontalLayout_2 = QtGui.QHBoxLayout()
-        self.horizontalLayout_2.setSizeConstraint(QtGui.QLayout.SetDefaultConstraint)
-        self.horizontalLayout_2.setObjectName("horizontalLayout_2")
-        self.label = QtGui.QLabel(self.centralwidget)
-        self.label.setObjectName("label")
-        self.horizontalLayout_2.addWidget(self.label)
-        self.combo_box_data_set = QtGui.QComboBox(self.centralwidget)
-        self.combo_box_data_set.setObjectName("combo_box_data_set")
-        self.horizontalLayout_2.addWidget(self.combo_box_data_set)
-        self.horizontalLayout_2.setStretch(0, 1)
-        self.horizontalLayout_2.setStretch(1, 1)
-        self.verticalLayout_3.addLayout(self.horizontalLayout_2)
-        self.horizontalLayout_3 = QtGui.QHBoxLayout()
-        self.horizontalLayout_3.setObjectName("horizontalLayout_3")
-        self.verticalLayout_4 = QtGui.QVBoxLayout()
-        self.verticalLayout_4.setObjectName("verticalLayout_4")
-        self.list_ideas = QtGui.QListView(self.centralwidget)
-        self.list_ideas.setSelectionMode(QtGui.QAbstractItemView.ExtendedSelection)
-        self.list_ideas.setObjectName("list_ideas")
-        self.verticalLayout_4.addWidget(self.list_ideas)
[…]

### 2013-09-10 src/main_experiment.py
pilot17

@@ -62,10 +62,15 @@ def get_questions_for_responses(n, append=False):
 
     forgot_question = """<p>Imagine you are in a social setting and you have forgotten the name of somebody you know. Brainstorm %s ways you could learn their name without directly asking them. Be as specific as possible in your descriptions.</p>"""
 
-    question = [charity_question % (n),
-                turk_question % (n),
+    question = [
+                #charity_question % (n),
+                #turk_question % (n),
                 mp3_question %(n),
-                forgot_question %(n)]
+                mp3_question %(n),
+                mp3_question %(n),
+                mp3_question %(n),
+                #forgot_question %(n)
+               ]
 
     return question
 
@@ -73,26 +78,18 @@ def get_questions_for_responses(n, append=False):
 
 def get_num_participants(num_requested):
   if num_requested == 5:
-    return [5, 5, 5, 5]
+    return [0, 0, 39, 0]
   elif num_requested == 10:
-    return [5, 5, 5, 5]
-  elif num_requested == 20:
-    return [3, 0, 0, 0]
-  elif num_requested == 50:
-    return [6, 4, 5, 6]
-  elif num_requested == 75:
-    return [6, 5, 5, 5]
-  elif num_requested == 100:
-    return [7, 4, 5, 6]
+    return [0, 0, 13, 0]
 
 def get_response_rewards(expid):
     return [
             (5, 0.18, "%s_fives" % expid, "fives", False),
             (10, 0.35, "%s_tens" % expid, "tens", False),
-            (20, 0.70, "%s_twenties" % expid, "twenties", False),
-            (50, 1.75, "%s_fifties" % expid, "fifties", False),
-            (75, 2.65, "%s_seventy_fives" % expid, "seventy_fives", False),
-            (100, 3.50, "%s_hundreds" % expid, "hundreds", False),
+            #(20, 0.70, "%s_twenties" % expid, "twenties", False),
+            #(50, 1.75, "%s_fifties" % expid, "fifties", False),
+            #(75, 2.65, "%s_seventy_fives" % expid, "seventy_fives", False),
+            #(100, 3.50, "%s_hundreds" % expid, "hundreds", False),
             ]
 
 ###

### 2013-09-10 exps/make_heatmap.py
refactor runs to completion. Much nicer

@@ -1,0 +1,118 @@
+#!/usr/bin/python
+
+import json
+
+html_template = '''
+<html>
+  <title>Brainstorming Heatmaps</title>
+  <script src="https://ajax.googleapis.com/ajax/libs/jquery/2.0.3/jquery.min.js"></script>
+  <script type="text/javascript">
+    var table_data = %s;
+    function make_table(table_data) {
+      var return_str = '<table id="data_table">';
+      var headers = table_data["headers"];
+      var checked_headers = [];
+      return_str += "<tr>";
+      for (var i = 0; i < headers.length; i++) {
+        var header_name = headers[i];
+        console.log(header_name + " Checked: " + $("#"+header_name).prop('checked'));
+        if ($("#" + header_name).prop('checked')) {
+          console.log("Writing " + header_name);
+          return_str += "<td>" + header_name + "</td>";
+          checked_headers.push(header_name);
+        }
+      }
+      return_str += "</tr>";
+      for (var i = 0; i < table_data["rows"].length; i++) {
+        return_str += "<tr>";
+        var row = table_data["rows"][i];
+        for (var j = 0; j < checked_headers.length; j++) {
+          var table_cell_color = row[checked_headers[j]];
+          console.log("Color: " + table_cell_color);
+          return_str += '<td bgcolor="' + table_cell_color + '">&nbsp;</td>';
+        }
+        return_str += "</tr>";
+      }
+      return_str += "</table>";
+      return return_str;
+    }
+    function make_checkboxes(table_data) {
+      var return_str = "<form>";
+      var headers = table_data['headers'];
+      for (var i = 0; i < headers.length; i++) {
+        var header_name = headers[i];
+        return_str += '<input type="checkbox" id="' + header_name + '" name="' + header_name + '" value="' + header_name + '" checked="1">' + header_name + '<br>';
+      }
+      return_str += "</form>";
+      return return_str;
+    }
+    function add_listeners(table_data) {
+      var headers = table_data['headers'];
+      for (var i = 0; i < headers.length; i++) {
+        var header_name = headers[i];
+        $("#" + header_name).change(function() {
+          $("#data_table").replaceWith(make_table(table_data));
+        });
+      }
+    }
+    $(document).ready(function() {
+    $("#checkboxes").append(make_checkboxes(table_data));
[…]

### 2013-09-10 exps/make_heatmap.py
sweet heatmaps

@@ -7,38 +7,43 @@ html_template = '''
   <title>Brainstorming Heatmaps</title>
   <script src="https://ajax.googleapis.com/ajax/libs/jquery/2.0.3/jquery.min.js"></script>
   <script type="text/javascript">
-    var table_data = %s;
-    function make_table(table_data) {
+    var table_datas = %s;
+
+    function make_tables(table_datas) {
+      var return_str = '<table><tr>';
+      for (var table_key in table_datas['tables']) {
+        table_data = table_datas['tables'][table_key];
+        return_str += "<td style='vertical-align:top; border-right:1px solid black; padding-left:5px; padding-right:5px;'><div id='data_table_" + table_key + "'>";
+        return_str += make_table(table_data, table_datas['headers']);
+        return_str += "</div></td>";
+      }
+      return return_str
+    }
+
+    function make_table(table_data, headers) {
       var return_str = '<table id="data_table">';
-      var headers = table_data["headers"];
-      var checked_headers = [];
       return_str += "<tr>";
       for (var i = 0; i < headers.length; i++) {
         var header_name = headers[i];
-        console.log(header_name + " Checked: " + $("#"+header_name).prop('checked'));
-        if ($("#" + header_name).prop('checked')) {
-          console.log("Writing " + header_name);
-          return_str += "<td>" + header_name + "</td>";
-          checked_headers.push(header_name);
-        }
+        return_str += "<td class='" + header_name + "'>" + header_name + "</td>";
       }
       return_str += "</tr>";
-      for (var i = 0; i < table_data["rows"].length; i++) {
+      for (var i = 0; i < table_data.length; i++) {
         return_str += "<tr>";
-        var row = table_data["rows"][i];
-        for (var j = 0; j < checked_headers.length; j++) {
-          var table_cell_color = row[checked_headers[j]];
-          console.log("Color: " + table_cell_color);
-          return_str += '<td bgcolor="' + table_cell_color + '">&nbsp;</td>';
+        var row = table_data[i];
+        for (var j = 0; j < headers.length; j++) {
+          var table_cell_color = row[headers[j]];
+          //console.log("Color: " + table_cell_color);
+          return_str += '<td bgcolor="' + table_cell_color + '" class="' + headers[j] + '">&nbsp;</td>';
         }
         return_str += "</tr>";
       }
       return_str += "</table>";
       return return_str;
     }
-    function make_checkboxes(table_data) {
[…]

### 2013-09-10 exps/make_heatmap.py
heatmap distances

@@ -1,10 +1,12 @@
 #!/usr/bin/python
 
-import json
+import json, math
 
 html_template = '''
 <html>
   <title>Brainstorming Heatmaps</title>
+  <style type='text/css'>
+  </style>
   <script src="https://ajax.googleapis.com/ajax/libs/jquery/2.0.3/jquery.min.js"></script>
   <script type="text/javascript">
     var table_datas = %s;
@@ -98,8 +100,11 @@ def process_data(data_lists):
           if val > 1.0:
             print "Error: Value is >1", val
             val = 1.0
-          if val < 0:
+          elif val < 0:
             print "Error: Value is <0", val
+            val = 0
+          if math.isnan(val):
+            #print "Error, value is None"
             val = 0
           as_int = int((1.0-val)*255)
           data_dict[key] = "#%02x%02x%02x" % (as_int, as_int, as_int)

### 2013-09-11 src/main_experiment.py
setup should be complete

@@ -78,15 +78,17 @@ def get_questions_for_responses(n, append=False):
 
 def get_num_participants(num_requested):
   if num_requested == 5:
-    return [0, 0, 39, 0]
+    return [0, 0, 50 0]
   elif num_requested == 10:
+    return [0, 0, 26, 0]
+  elif num_requested == 20:
     return [0, 0, 13, 0]
 
 def get_response_rewards(expid):
     return [
             (5, 0.18, "%s_fives" % expid, "fives", False),
             (10, 0.35, "%s_tens" % expid, "tens", False),
-            #(20, 0.70, "%s_twenties" % expid, "twenties", False),
+            (20, 0.70, "%s_twenties" % expid, "twenties", False),
             #(50, 1.75, "%s_fifties" % expid, "fifties", False),
             #(75, 2.65, "%s_seventy_fives" % expid, "seventy_fives", False),
             #(100, 3.50, "%s_hundreds" % expid, "hundreds", False),

### 2013-09-11 src/main_experiment.py
progress

@@ -78,7 +78,7 @@ def get_questions_for_responses(n, append=False):
 
 def get_num_participants(num_requested):
   if num_requested == 5:
-    return [0, 0, 50 0]
+    return [0, 0, 50, 0]
   elif num_requested == 10:
     return [0, 0, 26, 0]
   elif num_requested == 20:

### 2013-09-13 chi2014/method.tex
post-meeting tweaks
more meeting touches
Some more method

@@
 
-Many people have old iPods or MP3 players that they no longer use.
-Please brainstorm N uses for old iPods/MP3 players.
+Many people have old MP3 players or MP3 players that they no longer use.
+Please brainstorm N uses for old MP3 players/MP3 players.
 Assume that the devices' batteries no longer work, though they can be powered via external power sources.
@@
 
-Here do I talk about the unlimited condition?
+Classic brainstorming procedure asks for as many ideas as possible in a period of time. We instead asked participants for a specific number of responses: 5, 10, 20, 50, 75 or 100.
+
+This decision was motivated by our early experience asking respondants to simply come up with "as many ideas as possible" in a generous time frame.
+Over N respondents, we received a mean of N ideas (std N), despite giving a financial reward N times that of any of our explicitly numbered conditions.
+This unlimited condition provided a great deal of leeway for participants to "game the system", and receive the full reward with very little work, so we dropped the condition.
 
@@
 
-Does it make sense to have a section here on what data we collected?
+HITs were augmented with JavaScript to collect data in addition to brainstorming responses.
+For each response, we collected the time of the first activation and last de-activation of the form element.
+We collected the time the HIT was accepted and the time it was submitted.
+We took a has of the Mechanical Turk worker ID as a unique identifier of our participants.
+
+\subsection{Terminology}
+
+In examining the data, we found responses advocating a specific problem-solving strategy would nonetheless vary in generality.
+We introduce terminology here that will allow us to better describe and formalize these differences.
+
+An \emph{instance} is a single response by a single participant to a brainstorming question. For example, Figure FIG gives four \emph{instances} provided in the \emph{MP3 player} brainstorming task.
+
+\begin{figure}[!h]
+    \begin{enumerate}
+        \item "Use as a mirror"
+        \item "Mirror"
+        \item "Use it as a mirror for makeup"
+        \item "Tape it to a telephone pole and use as a reflector"
+    \end{enumerate}
+\end{figure}
+
+An instance is associated with exactly one \emph{idea}.
+An idea may have 0 or more instances, all of which describe an identical solution to a problem.
+In Figure FIG, "Use as a mirror" and "Mirror" belong to the same \emph{idea}.
+However, "Use it as a mirror for makeup" does not belong to the same idea, as it adds additional information to the problem solution; it is more specific.
+
+Despite this, it is clear the ideas have some commonality.
+It is desireable to encode this relationship.
+To that end, we organize ideas in \emph{category trees}.
+In a category tree, ideas act as nodes.
+Each idea may have 0 or 1 parent idea (only the root has 0) and 0 or more child ideas.
+An example category tree for the instances in Figure FIG is given in Figure FIG.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{sample_category_tree}
+\end{figure}
+
+If an idea \emph{A} is a parent of idea \emph{B}, A is a generalization of B: all ideas in A are are suitably general to encompass all ideas in B. This difference is relative to the ideas in question.
+
+A \emph{question forest} is simply the collection of all category trees generated in response to a particular question. Different trees in a question forest have no special relationship in meaning.
 
@@
 
-One researcher coded the data for category in a hierarchical scheme.
-For each question, ideas recieved from the participants were organized in a tree structure.
-A single node in the tree represents a particular solution to the problem.
+One researcher coded the data to produce the idea forests described above.
+The clustering algorithm is described in Figure FIG.
+Instances are added one at a time to the existing idea forest.
+In brief summary, the root idea that most closely matches the strategy in the instance is selected, and then that root idea and the new instance are compared in generality.
+This process may be repeated depending on this generality relationship, until instance is either placed in an existing idea or a new idea node is created.
 
-To formalize the clustering method, we introduce some key terms:
+The clustering algorithm relies on the human intelligence of the researcher for three key decisions:
 
 \subsubsection{Similarity}
-Similarity is a subjective measure of the equality of two nodes a, b on the basis of their key problem solution.
-a and b have high similarity if they solve the problem in the same way, and low similarity if their solutions have no common themes.
+Two ndodes a and b have high similarity if they solve the problem in the same way, and low similarity if their solutions have no common themes.
+For example, in Figure FIG, "Mirror" and "Tape it to a telephone pole and use as a reflector" have high similarity, as they both solve the problem of using a broken MP3 player by using it to reflect light.
 
 \subsubsection{Coverage}
-A node a has high coverage of a node b if the problem solution in a is equal to or an abstraction of the problem solution in b. Coverage is not commutative.
+A node a has high coverage of a node b if the problem solution in a is equal to or an abstraction of the problem solution in b.
+Coverage is not commutative.
+To return to the example, "Mirror" has high coverage of "Tape it to a telephone pole and use as a reflector", as the latter is a specific use for a mirror.
+On the other hand, the latter idea has low coverage of the former.
 
-\subsubsection{Symmetrical Coverage}
-Two nodes a, b have symmetrical coverage if coverage(a, b) = coverage(b, a).
-
-In the resulting tree, all parent nodes have high coverage over their child nodes, but all child nodes have low coverage over their parent nodes.
-All nodes have high affinity with their parent and their children.
-The clustering algorithm is outlined in figure (FIG).
+\subsubsection{Artificial Ideas}
+Occasionally, two nodes have high similarity but no coverage.
+For example, "Use it as a mirror for makeup" and "Tape it to a telephone pole and use as a reflector" use the MP3 player in the same way, but neither is a generalization of the other.
+In this case, the algorithm appeals to the human coder to provide a lable for a new idea that is a generalization of both.
 
@@
 \begin{verbatim}
-for each idea:
-  idea_node = new cluster with idea as member
-  current_node = root
+for each instance:
+  idea_node = new node including instance
+  current_node = root of forest
   do:
-    best_match = max_affinity(idea_node, current_node.children)
+    best_match = max_similarity(idea_node, current_node.children)
 
-    if best_match.affinity is low or current_node has no children:
+    if best_match.similarity is low or current_node has no children:
       insert idea_node under current_node
@@
     else:
-      if symmetrical_coverage(idea_node, best_match) and high:
+      if coverage(idea_node, best_match) == coverage(best_match, idea_node) == high:
         merge idea_node, best_match
         exit do
-      else symmetrical_coverage(idea_node, best_match) and low:
-        new_parent = new child node of current_node, summarizing cause of affinity
+      else if coverage(idea_node, best_match) == coverage(best_match, idea_node) == low:
+        new_parent = new artifical idea node
         insert best_match, idea_node under new_parent
+        insert new_parent under current_node
         exit do
@@
 \end{figure*}
+
+After completion of the coding algorithm, we were left with an idea forest as described in section SEC.

### 2013-09-13 chi2014/method.tex
typo

@@
 
-If an idea \emph{A} is a parent of idea \emph{B}, A is a generalization of B: all ideas in A are are suitably general to encompass all ideas in B. This difference is relative to the ideas in question.
+If an idea \emph{A} is a parent of idea \emph{B}, A is a generalization of B: all instances in A are are suitably general to encompass all instances in B.
+This difference is relative to the ideas in question.
 

### 2013-09-13 src/main_experiment.py
whocares

@@ -78,7 +78,7 @@ def get_questions_for_responses(n, append=False):
 
 def get_num_participants(num_requested):
   if num_requested == 5:
-    return [0, 0, 50, 0]
+    return [0, 0, 40, 0]
   elif num_requested == 10:
     return [0, 0, 26, 0]
   elif num_requested == 20:
@@ -87,8 +87,8 @@ def get_num_participants(num_requested):
 def get_response_rewards(expid):
     return [
             (5, 0.18, "%s_fives" % expid, "fives", False),
-            (10, 0.35, "%s_tens" % expid, "tens", False),
-            (20, 0.70, "%s_twenties" % expid, "twenties", False),
+            #(10, 0.35, "%s_tens" % expid, "tens", False),
+            #(20, 0.70, "%s_twenties" % expid, "twenties", False),
             #(50, 1.75, "%s_fifties" % expid, "fifties", False),
             #(75, 2.65, "%s_seventy_fives" % expid, "seventy_fives", False),
             #(100, 3.50, "%s_hundreds" % expid, "hundreds", False),

### 2013-09-14 src/classify/process_hits.py
all the output and clustering of pilot18

@@ -18,6 +18,8 @@ base_dirs = ['/home/fil/Dropbox/crowbrain_share/experiments/pilot14-2013.08.16',
              '/home/fil/Dropbox/crowbrain_share/experiments/pilot13-2013-07-18',
              '/home/fil/Dropbox/crowbrain_share/experiments/pilot16-2013.09.02',
              '/home/fil/Dropbox/crowbrain_share/experiments/pilot17-2013.09.10',
+             '/home/fil/Dropbox/crowbrain_share/experiments/pilot18-2013.09.12',
+             '/home/fil/Dropbox/crowbrain_share/experiments/pilot18_2-2013.09.13',
              '/home/fil/Dropbox/crowbrain_share/experiments/pilot12-2013.07.16',
              '/home/fil/Dropbox/crowbrain_share/experiments/pilot11-2013.07.15',]
 #base_dir = '/home/mterry/Dropbox/crowbrain_share/experiments/pilot13-2013-07-18'

### 2013-09-14 src/clusterer/import_data.py

@@ -14,7 +14,7 @@ import sys
 sys.setrecursionlimit(10000)
 
 db_name = '/home/fil/Dropbox/clusters.db'
-clusters_dir = '/home/fil/Dropbox/crowbrain_share/data/idea-clusters-2013.09.05/'
+clusters_dir = '/home/fil/Dropbox/crowbrain_share/data/idea-clusters-2013.09.13/'
 
 def create_tables(cursor):
   sql = [

### 2013-09-15 chi2014/introduction.tex
more writing

@@
-This format is to be used for submissions that are
-published in the conference proceedings.  We wish to give
-this volume a consistent, high-quality appearance. We
-therefore ask that authors follow some simple
-guidelines. In essence, you should format your paper
-exactly like this document. The easiest way to do this is
-simply to download a template from the conference web
-site, and replace the content with your own material.
+Whether in research or industry, productivity can be hamstrung by the need to generate creative or novel ideas.
+Given the ubiquity of idea generation tasks, it is desireable to provide generalizable automated methods of solving them.
+Automated computational methods to solve specific problems are unlikely to appear without significant advances in natural language processing and artificial intelligence.
+As a result, we are left with human-centric solutions, such as brainstorming.
+
+Brainstorming is a tool proposed by Osborne [CITE] that has achieved widespread popularity and similarly widespread research and controversy.
+Brainstorming methods have been applied in many scenarios, from spatially co-located participants to electronic brainstorming over computer networks [CITE].
+The introduction of crowd marketplaces such as Amazon's Mechanical Turk [CITE] and the corresponding paradigm of massively parallel Human Intelligence Tasks provide both a natural venue to solicit ideas and a new context in which to establish models of brainstorming activity.
+
+This setting introduces unique challenges.
+For example, crowd marketplaces are driven by financial transactions, while participants in traditional brainstorming settings are required and motivated to generate ideas for a given duration.
+Guidelines and points of comparison are needed.
+
+In the course of their research in Human-Computer Interactions, two of the authors went looking for these sets of guidelines to apply them to a crowd idea generation task.
+While previous work has utilized brainstorming techniques in crowd marketplaces [CITE], we were unable to find any generalizable models or guidelines for brainstorming in the context of the crowd.
+Models such as Nijstaf and Stroebe's SIAM [CITE] make and test predictions regarding in-person brainstorming, but we did not find similar baselines for electronic or crowd brainstorming.
+
+This paper and corresponding data set provide a baseline for the idea generation process in crowd marketplaces.
+We establish a corpus of brainstorming data in a restricted design space with two axes: idea generation problem, and number of responses requested of a single crowd worker.
+
+We provide a method for clustering responses to idea generation problems and provide a detailed description of the resulting data set, which has been made publically available.
+We demonstrate the saturation of ideas received over time, and show that the rate of saturation, and size of the saturated idea set, originality of ideas and generality of ideas vary between conditions.
+
+Furthermore, we examine predictions made by Nijstaf and Stroebe's model of in-person brainstorming and test them in the context of crowd brainstorming.
+We find that the tested predictions hold: that ideas generated in sequence tend to belong to the same category, and that the time spent to generate when changing categories is signficantly greater than when generating ideas within the same category.
+
+Finally, we briefly outline qualitative findings in the data, and recommend future work motivated by these findings.

### 2013-09-15 chi2014/method.tex
more writing

@@
-We began the process of soliciting responses to brainstorming tasks in a crowd marketplace. Throughout the experimental process, particular aspects of the experimental design remained fixed.
+\section{Method}
+
+In order to develop an adequate model of baseline brainstorming performance in crowd marketplaces, it was necessary to establish a corpus of data.
+Over the course of four months we repeatedly collected responses to brainstorming tasks.
+Throughout the experimental process, particular aspects of the experimental design remained fixed.
 
@@
 
-Upon accepting a HIT, participants are randomly assigned into a \emph{brainstorming problem} condition.
+Upon accepting a HIT, participants were randomly assigned into a \emph{brainstorming problem} condition.
 We ensured that participants completing multiple HITs are not exposed to the same brainstorming task twice.
@@
 
-\subsubsection{Data Collection}
+\subsubsection{Data Collected}
+
+In total, 341 HITs were complteed by 280 distinct participants.
+A breakdown of responses across conditions is given in table TAB.
+In cases when a single participant answered the same question multiple times, all HITs after the first were dropped for subsequent analysis.
+
+\begin{tabular}[h!]{r | l l l l l l }
+& 5 & 10 & 20 & 50 & 75 & 100 \\ \hline \hline
+HITs (MP3) & 59 & 49 & 23 & 10 & 10 & 10 \\
+HITs unique (MP3) & 57 & 39 & 21 & 9 & 10 & 10 \\ \hline
+HITs (Turk) & 11 & 10 & 11 & 10 & 9 & 9 \\
+HITs unique (Turk) & 11 & 9 & 11 & 9 & 9 & 7 \\ \hline
+HITs (Forgot) & 10 & 10 & 10 & 12 & 11 & 11 \\
+HITs unique (Forgot) & 10 & 8 & 9 & 12 & 10 & 10 \\ \hline
+HITs (Charity) & 9 & 10 & 10 & 8 & 10 & 9 \\
+HITs unique (Charity) & 9 & 9 & 10 & 7 & 9 & 9 \\ \hline
+HITs (total) \\
+HITs unique (total) \\ \hline
+\end{tabular}
+
+There is significantly more data for the MP3 condition than for any of the others.
+Despite at least 7 unique HITs for each of the 24 problem x number conditions, we found that the limited number of total responses for the 5, 10 and 20 conditions precluded significant analysis.
+To that end, we selected the MP3 player problem to further expand the data set to 400 (TODO: maybe not true) responses per number condition.
 
@@
 
-In examining the data, we found responses advocating a specific problem-solving strategy would nonetheless vary in generality.
-We introduce terminology here that will allow us to better describe and formalize these differences.
+As we began to examine the data, it quickly began apparent that there are nuances in discussing responses to brainstorming tasks.
+In the interest of facilitating this discussion, we introduce terms here that will be used for the remainder of the paper.
 
@@
 
-One researcher coded the data to produce the idea forests described above.
+One researcher coded responses the MP3 problem to produce an idea forest as described above. The MP3 problem was chosen because it was subjectively easiest to distinguish between and categorize ideas.
+
 The clustering algorithm is described in Figure FIG.
@@
 
-The clustering algorithm relies on the human intelligence of the researcher for three key decisions:
-
-\subsubsection{Similarity}
-Two ndodes a and b have high similarity if they solve the problem in the same way, and low similarity if their solutions have no common themes.
-For example, in Figure FIG, "Mirror" and "Tape it to a telephone pole and use as a reflector" have high similarity, as they both solve the problem of using a broken MP3 player by using it to reflect light.
-
-\subsubsection{Coverage}
-A node a has high coverage of a node b if the problem solution in a is equal to or an abstraction of the problem solution in b.
-Coverage is not commutative.
-To return to the example, "Mirror" has high coverage of "Tape it to a telephone pole and use as a reflector", as the latter is a specific use for a mirror.
-On the other hand, the latter idea has low coverage of the former.
-
-\subsubsection{Artificial Ideas}
-Occasionally, two nodes have high similarity but no coverage.
-For example, "Use it as a mirror for makeup" and "Tape it to a telephone pole and use as a reflector" use the MP3 player in the same way, but neither is a generalization of the other.
-In this case, the algorithm appeals to the human coder to provide a lable for a new idea that is a generalization of both.
-
-\begin{figure*}[h]
+\begin{figure*}[h!]
+\small
 \begin{verbatim}
@@
 
+The clustering algorithm relies on the human intelligence of the researcher for three key decisions:
+
+\subsubsection{Similarity}
+Two ndodes a and b have high similarity if they solve the problem in the same way, and low similarity if their solutions have no common themes.
+For example, in Figure FIG, "Mirror" and "Tape it to a telephone pole and use as a reflector" have high similarity, as they both solve the problem of using a broken MP3 player by using it to reflect light.
+
+\subsubsection{Coverage}
+A node a has high coverage of a node b if the problem solution in a is equal to or an abstraction of the problem solution in b.
+Coverage is not commutative.
+To return to the example, "Mirror" has high coverage of "Tape it to a telephone pole and use as a reflector", as the latter is a specific use for a mirror.
+On the other hand, the latter idea has low coverage of the former.
+
+\subsubsection{Artificial Ideas}
+Occasionally, two nodes have high similarity but no coverage.
+For example, "Use it as a mirror for makeup" and "Tape it to a telephone pole and use as a reflector" use the MP3 player in the same way, but neither is a generalization of the other.
+In this case, the algorithm appeals to the human coder to provide a lable for a new idea that is a generalization of both.
+
 After completion of the coding algorithm, we were left with an idea forest as described in section SEC.

### 2013-09-15 chi2014/related.tex
more writing

\section{Related Work}

% TODO Note: I plaigarized the wording on these definitions from wikipedia at this point, should pull from original Osborne at some point
Brainstorming was originally proposed in XXXX by Osborne [CITE]. Brainstorming makes several recommendations for idea generation in groups, the core of which are a set of four rules: \emph{focus on quantity}, \emph{withhold criticism}, \emph{welcome unusual ideas}, and \emph{combine and improve ideas}.

Since then, brainstorming as proposed by Osborne has been extensively evaluated. Very quickly, Taylor et. al established that \emph{nominal groups}, or individuals brainstorming in isolated whose ideas are later combined, performed better than real groups in terms of number of ideas generated [CITE]. Furthermore, Diehl and Stroebe found extremely high correlation (XX) between the number of ideas generated and the quality of ideas, a finding that has been verified many times since (Another citation here would be nice). In Bouchard and Hare [CITE], nominal brainstorming groups were able to generate quantities of ideas linear in the number of members in the group.

Electronic brainstorming is a variant in which individuals brainstorm at separate computer terminals, with the exchange of ideas performed over the network. Participants can enter ideas while simultaneously viewing the ideas of all other ideas, anonymously presented on-screen. Gallupe [CITE] found that electronic brainstorming reduced the inhibiting social effects of \emph{production blocking} (the inability for multiple people to speak at once), and \emph{evaluation apprehension} (the fear of judgment) in comparison to in-person groups. However, Pinsonneault et al identify four productivity impediments introduced by group electronic brainstorming, and show that there is little evidence to support the proposition that electronic brainstorming out-performs nominal brainstorming [CITE].

As crowd marketplaces do not tend to include built-in support for real-time communication between workers, nominal brainstorming groups provide the most reasonable assumption for providing a baseline assessment of performance. It is fortuitous, then, that prior work suggests that nominal groups are likely equivalent at worst to groups with the ability to communicate.

This choice is supported by other users of crowd marketplaces for idea generation. Yu and Nickerson use many individuals in an evolutionary algorithm to generate creative chair designs [CITE]. This technique is compared to single participants generating single ideas, but no baseline is established for idea generators individually exhausting a design space.

Something about another example. Maybe the Little iterative vs parallel paper? It includes creative tasks.

% TODO: part of this is Pao-exact, assuming it's cool because she's going to be an author on the paper
While there are no models of nominal brainstorming in the context of a crowd marketplace, several researchers have modelled the idea generation process for individuals and groups. We emphasize the former for its applicability to nominal groups. Nijstad and Stroebe define the \emph{search for ideas in associative memory} (SIAM) model. SIAM assumes two stages of idea generation: knowledge activation based on search of working memory, and idea production within a single activated image. This model verifies three predictions for individual brainstorming behaviour:

\begin{enumerate}
\item an idea is more likely to be followed by an idea in the same category than would be expected by chance
\item new ideas in the same category as the previous idea are generated faster than ideas between categories
\item increased clustering (generation of ideas within-category) is positively correlated with overall high production
\end{enumerate}

Should we have another model here? I haven't read any yet.

In summary, while brainstorming has been examined in the traditional and electronic settings, its application in crowd marketplaces has received limited attention.

### 2013-09-15 chi2014/summary.tex
more writing

\section{Summary of Data}

The resulting forest was composed of 3007 instances, 1213 ideas, and 335 category trees. The number of instances, ideas and categories present in each of the number conditions is given in table TAG. In addition, table TAG shows the number of "non-singleton" categories, namely, those categories containing at least two instances. Intuitively, non-singleton categories represent those categories in which some exploration actually occured. These represent categories in which we can potentially identify the idea production phase of the SIAM model [].

\begin{tabular}[h!]{r | l l l l l l }
& 5 & 10 & 20 & 50 & 75 & 100 \\ \hline \hline
HITs & 57 & 47 & 23 & 10 & 10 & 10 \\
instances & 293 & 471 & 453 & 500 & 634 & 855 \\
ideas & 171 & 249 & 278 & 341 & 443 & 177 \\
category trees & 76 & 93 & 100 & 123 & 172 & 177 \\
non-singleton trees & 28 & 41 & 41 & 50 & 49 & 61 \\
\end{tabular}

Figure FIG shows the distribution of instances in non-singleton categories. The majority of the "mass" of instances are in trees with less than 70 instances.

% TODO: this figure needs better labeling for the second part
\begin{figure}[!h]
    \centering
    \includegraphics[width=0.9\columnwidth]{instances_in_subtrees}
\end{figure}

### 2013-09-15 chi2014/method.tex
pre breaking things

@@
 After completion of the coding algorithm, we were left with an idea forest as described in section SEC.
+
+\subsection{Measures}
+
+With the described tree structure, we were able to begin deriving measures of interest with respect to instances, ideas, and category trees.
+In particular, we used the category forest structure to assign originality and generality scores to ideas and categories.
+Furthermore, we use the category tree to identify remixing of ideas within the brainstorming run of a participant.
+
+\subsubsection{Originality}
+
+As our measure of originality, we use \emph{o-score}, introduced by Jansson and Smith [CITE].
+An idea's o-score is $1 - p(idea)$, where $p(idea) = (number of instances of that idea)/(number of instances total)$.
+We similarly calculate o-score for categories.
+Throughout this paper, we will occasionally refer to the \emph{category o-score} of an idea.
+This simply refers to the o-score of the category tree to which that idea belongs.
+
+\subsubsection{Generality}
+
+Similar ideas are presented differently between different instances.
+The most major distinction is generality.
+We use \emph{idea height}, the height of an idea in it's corresponding category tree to quantify this notion of generality.
+An illustration of how idea height is calculated is given in figure FIG.
+Idea height is a relative scale that is comparable within-tree but not between-trees.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{idea_height}
+\end{figure}

### 2013-09-16 chi2014/introduction.tex
citations and some figure markup

@@
 
-Brainstorming is a tool proposed by Osborne [CITE] that has achieved widespread popularity and similarly widespread research and controversy.
-Brainstorming methods have been applied in many scenarios, from spatially co-located participants to electronic brainstorming over computer networks [CITE].
-The introduction of crowd marketplaces such as Amazon's Mechanical Turk [CITE] and the corresponding paradigm of massively parallel Human Intelligence Tasks provide both a natural venue to solicit ideas and a new context in which to establish models of brainstorming activity.
+Brainstorming is a tool proposed by Osborne \cite{osborn_applied_1957} that has achieved widespread popularity and similarly widespread research and controversy.
+Brainstorming methods have been applied in many scenarios, from spatially and temporally co-located participants to electronic brainstorming over computer networks \cite{pinsonneault_electronic_1999}.
+The introduction of crowd marketplaces such as Amazon's Mechanical Turk \cite{_amazon_????} and the corresponding paradigm of massively parallel Human Intelligence Tasks provide both a natural venue to solicit ideas and a new context in which to establish models of brainstorming activity.
 
@@
 While previous work has utilized brainstorming techniques in crowd marketplaces [CITE], we were unable to find any generalizable models or guidelines for brainstorming in the context of the crowd.
-Models such as Nijstaf and Stroebe's SIAM [CITE] make and test predictions regarding in-person brainstorming, but we did not find similar baselines for electronic or crowd brainstorming.
+Models such as Nijstaf and Stroebe's SIAM \cite{nijstad_how_2006} make and test predictions regarding in-person brainstorming, but we did not find similar baselines for electronic or crowd brainstorming.
 
@@
 
-Furthermore, we examine predictions made by Nijstaf and Stroebe's model of in-person brainstorming and test them in the context of crowd brainstorming.
+Furthermore, we examine predictions made by Nijstad and Stroebe's model of in-person brainstorming and test them in the context of crowd brainstorming.
 We find that the tested predictions hold: that ideas generated in sequence tend to belong to the same category, and that the time spent to generate when changing categories is signficantly greater than when generating ideas within the same category.

### 2013-09-16 chi2014/method.tex
citations and some figure markup

@@
 
-Participants were recruited from Amazon's \emph{Mechanical Turk} (CITE), and were restricted to residents of the United States for a baseline expectation of english language comprehension and cultural familiarity.
+Participants were recruited from Amazon's \emph{Mechanical Turk} \cite{_amazon_????}, and were restricted to residents of the United States for a baseline expectation of english language comprehension and cultural familiarity.
 Mechanical Turk is an online marketplace in which members receive financial reward for completing \emph{Human Intelligence Tasks}, or HITs.
@@
 
-At the beginning of the form is a brief introduction to brainstorming, with a paraphrase of Osborne's four rules of brainstorming (CITE).
+At the beginning of the form is a brief introduction to brainstorming, with a paraphrase of Osborne's four rules of brainstorming \cite{osborn_applied_1957}.
 These rules were manipulated to make sense within the constraints nominal brainstorming over the web medium.
@@
 With this basic design in hand, we ran several pilots and experiments.
-In early pilots, we used classic problems from psychology literature on brainstorming, including the "thumb problem" and "broom problem" (CITE).
+In early pilots, we used classic problems from psychology literature on brainstorming, including the "thumb problem" and "broom problem" \cite{isaksen_review_1997}.
 Early results from brainstormers were unsatisfying and wildly divergent.
@@
 
-An \emph{instance} is a single response by a single participant to a brainstorming question. For example, Figure FIG gives four \emph{instances} provided in the \emph{MP3 player} brainstorming task.
+An \emph{instance} is a single response by a single participant to a brainstorming question.
+For example, figure~\ref{fig:sample_instances} gives four \emph{instances} provided in the \emph{MP3 player} brainstorming task.
 
@@
     \end{enumerate}
+    \label{sample_instances}
 \end{figure}
@@
 
-The clustering algorithm is described in Figure FIG.
+The clustering algorithm is described in Figure~\ref{fig:cluseringalg}.
 Instances are added one at a time to the existing idea forest.
@@
 
-As our measure of originality, we use \emph{o-score}, introduced by Jansson and Smith [CITE].
+As our measure of originality, we use \emph{o-score}, introduced by Jansson and Smith \cite{jansson_design_1991}.
 An idea's o-score is $1 - p(idea)$, where $p(idea) = (number of instances of that idea)/(number of instances total)$.
@@
     \includegraphics[width=0.9\columnwidth]{idea_height}
+    \label{idea_height}
 \end{figure}

### 2013-09-16 chi2014/related.tex
citations and some figure markup

@@
 % TODO Note: I plaigarized the wording on these definitions from wikipedia at this point, should pull from original Osborne at some point
-Brainstorming was originally proposed in XXXX by Osborne [CITE].
+Brainstorming was originally proposed in XXXX by Osborne \cite{osborn_applied_1957}.
 Brainstorming makes several recommendations for idea generation in groups, the core of which are a set of four rules: \emph{focus on quantity}, \emph{withhold criticism}, \emph{welcome unusual ideas}, and \emph{combine and improve ideas}.
@@
 Very quickly, Taylor et.
-al established that \emph{nominal groups}, or individuals brainstorming in isolated whose ideas are later combined, performed better than real groups in terms of number of ideas generated [CITE].
-Furthermore, Diehl and Stroebe found extremely high correlation (XX) between the number of ideas generated and the quality of ideas, a finding that has been verified many times since (Another citation here would be nice).
-In Bouchard and Hare [CITE], nominal brainstorming groups were able to generate quantities of ideas linear in the number of members in the group.
+al established that \emph{nominal groups}, or individuals brainstorming in isolated whose ideas are later combined, performed better than real groups in terms of number of ideas generated \cite{taylor_does_1958}.
+Furthermore, Diehl and Stroebe found that groups which generated good ideas also generated many ideas \cite{diehl_productivity_1987}, a finding that has been verified many times since (Another citation here would be nice).
+In Bouchard and Hare \cite{bouchard_jr_size_1970}, nominal brainstorming groups were able to generate quantities of ideas linear in the number of members in the group.
 
@@
 Participants can enter ideas while simultaneously viewing the ideas of all other ideas, anonymously presented on-screen.
-Gallupe [CITE] found that electronic brainstorming reduced the inhibiting social effects of \emph{production blocking} (the inability for multiple people to speak at once), and \emph{evaluation apprehension} (the fear of judgment) in comparison to in-person groups.
-However, Pinsonneault et al identify four productivity impediments introduced by group electronic brainstorming, and show that there is little evidence to support the proposition that electronic brainstorming out-performs nominal brainstorming [CITE].
+Gallupe \cite{gallupe_electronic_1992} found that electronic brainstorming reduced the inhibiting social effects of \emph{production blocking} (the inability for multiple people to speak at once), and \emph{evaluation apprehension} (the fear of judgment) in comparison to in-person groups.
+However, Pinsonneault et al identify four productivity impediments introduced by group electronic brainstorming, and show that there is little evidence to support the proposition that electronic brainstorming out-performs nominal brainstorming \cite{pinsonneault_electronic_1999}.
 
@@
 This choice is supported by other users of crowd marketplaces for idea generation.
-Yu and Nickerson use many individuals in an evolutionary algorithm to generate creative chair designs [CITE].
+Yu and Nickerson use many individuals in an evolutionary algorithm to generate creative chair designs \cite{yu_cooks_2011}.
 This technique is compared to single participants generating single ideas, but no baseline is established for idea generators individually exhausting a design space.
@@
 We emphasize the former for its applicability to nominal groups.
-Nijstad and Stroebe define the \emph{search for ideas in associative memory} (SIAM) model.
+Nijstad and Stroebe define the \emph{search for ideas in associative memory} (SIAM) model \cite{nijstad_how_2006}.
 SIAM assumes two stages of idea generation: knowledge activation based on search of working memory, and idea production within a single activated image.

### 2013-09-16 chi2014/goals.tex
goals section first pass

\section{Goals and Hypotheses}

Our primary motivation is to develop baseline for brainstorming performance in crowd marketplaces. Brainstorming performance is often measured in terms of creativity. However, there any many interpretations of creativity [CITE a couple citations]. Instead we are interested in performance on variables that are commonly \emph{components} of a creativity score: quantity, originality, and surprise factor.

As our primary goal, we examine these outcomes with respect to the number of answer requested from workers. This is the simplest of conditions that can vary when choosing a design for a crowd brainstorming problem. Furthermore, we detail the phenomenon of \emph{riffing} (known in the idea generation literature as \emph{clustering}, a term we use for its classification meaning later in the paper).

In addition to the high-level goal of modelling quantity, originality and surprise factor in brainstorming results, we began with several explicit hypotheses. \textbf{Hypothesis 1:} the rate of new ideas will reduce towards 0 exponentially as a function of the number of responses gathered. Less formally, by continually soliciting solutions, we will eventually reach a saturation point at which the majority of new solutions are derivative of previously seen solutions.

\textbf{Hypothesis 2:} there is a set general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.

In addition, trends identified by prior work in spatial and temporal co-located brainstorming may or may not be exist in crowd brainstorming. We selected the trends most testable with our data set to replicate their findings. Nijstad and Stroebe's SIAM model \cite{nijstad_how_2006} makes three non-subjective predictions, two of which retain their applicability in this setting:

\begin{itemize}
\item \textbf{Hypothesis 3:} An idea from one semantic category should more often be followed by an idea from the same category than expected by random chance
\item \textbf{Hypothesis 4:} idea generation when changing semantic categories takes longer than when repetiting categories
\end{itemize}

Parnes et al \cite{parnes_effects_1961} showed that more high quality ideas were generated in the latter half of brainstorming run than the former half. In keeping of our philosophy regarding creativity and quality measurements, we restrict the content of this test to originality. \textbf{Hypothesis 5:} Ideas generated in the latter half of a brainstorming session are of higher originality than ideas generated in the former half.

### 2013-09-16 chi2014/conclusions.tex
missing text files

\section{Conclusion}

We set out to establish a baseline for brainstorming performance in crowd marketplaces, a new setting for idea generation in with spatial and temporal co-presence are not guaranteed. We did this by establishing a corpus of brainstorming responses, publically available at URL. Furthermore, we introduce the \emph{idea forest}, an encoding of brainstorming responses that captures differences in generality we discovered, an introduced an algorithm for producing idea forests.

We found that originality increased as brainstorming participants were asked to generate more ideas individually. Generality, however, decreased. In both cases, we found a "sweet spot" for this change at the 20 idea mark; participants' ideas reach their peak creativity and minimum generality after their 20th contribution.

Furthermore, we verified that the findings and predictions of prior work in which participants were spatially and temporally copresent. We found that, in line with the predictions of Jigstad and Stroebe's SIAM model \cite{nijstad_how_2006}, ideas are more likely to follow an idea of the same category than would be randomly expected, and that the time spent to generate an idea within-category is lesser than that to generate ideas between categories. Furthermore, we recreated the finding of Parnes et al \cite{parnes_effects_1961} that responses in the latter part of a participant's idea generation process are more creative than those earlier.

We identified three qualitative phenomena in of brainstorming results. First, that brainstorming results are subject to the popular consciousness of the time they are gathered. Second, that workers are more likely to take HITs with higher absolute payout, even if relative payout is identical.

The most salient recommendation we can make is easily summarized: if you use the crowd for brainstorming and desire originality, it is most productive to ask few participants for more than 20 ideas each.

### 2013-09-16 chi2014/future.tex
missing text files

\section{Future Work}

\subsection{Interventions}

\begin{itemize}
\item multi-phase brainstorming: introducing common ideas first to force workers more quickly into the originality zone
\end{itemize}

\subsection{Predicting Turker performance}

\subsection{Automatic Idea Forest Generation}

### 2013-09-16 chi2014/qualitative.tex
missing text files

\section{Qualitative findings}

I'm not going into any of these in detail because I'm not sure if they should go in.

\subsection{The Miley Cyrus affect}

\begin{itemize}
\item last batch of HITs had several references to Miley Cyrus
\item previously, this pop culture reference never came up
\item suggests the pool of available ideas is time-sensitive
\end{itemize}

\subsection{Reward vs Relative Reward in Completion Time}

\begin{itemize}
\item HITs with more overall reward perform faster, despite same payout per idea
\item design implications
\end{itemize}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{batch_time}
    \caption{Time to complete batch of HITs}
\end{figure}

\subsection{Personas}

\begin{itemize}
\item Comment on the effect of single participants - the zoomed in 75'er effect.
\end{itemize}

### 2013-09-16 chi2014/results.tex
missing text files

\section{Results}

\subsection{Between-Conditions Comparisons}

\subsubsection{Saturation of ideas}

One of our hypotheses (HYP) was that as the number of instances gathered increased, the rate of new ideas, new categories and new singleton categories would converge towards 0.

Figure FIG shows the rate of new idea, category, and non-singleton category generation  over time for each number requested conditions. These plots have been smoothed by randomly shuffling the order in which ideas were receieved 500 times and taking the mean cumulative count at each point in time, and plotting with a window size of 10. In an ideal scenario where each instance provided a new unique idea and category, the lines would horizontal across the top of the figure.

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{rate_new_idea_over_time}
    \includegraphics[width=0.9\columnwidth]{rate_new_category_over_time}
    \includegraphics[width=0.9\columnwidth]{rate_new_ns_category_over_time}
    \caption{Rate of idea generation}
\end{figure}

Visually, the rate of idea, category and singleton category generation seems to trend towards 0. We tested this hypothesis...

Additionally Figure FIG indicates that there is some effect of number of ideas requested on the rate of generation. Generally, conditions with more requested responses generated ideas and categories at a higher rate. We tested this...

\subsubsection{Originality}

We were also interested in the effect of the number requested condition on originality. Originality, as measured by o-scores at both the category and idea level, is compared cross condition in figure FIG. Some stat tests here...

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{subtree_oscore}
    \includegraphics[width=0.9\columnwidth]{idea_oscore}
    \caption{Error bars are standard error}
\end{figure}

The more ideas requested, the more original the ideas produced. This provides support for the findings of Parnes et al \cite{parnes_effects_1961}, namely that more original ideas are found in extended idea generation events. This finding will be examined in the next section.

\subsubsection{Generality}

As discussed above, we use height in category tree as a measure of an ideas generality. Generality (and its reverse, sepcificity) are both potentially desireable outcomes of a brainstorming task.

The idea-height across number conditions is given in figure FIG.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{height_in_subtree}
    \includegraphics[width=0.9\columnwidth]{height_in_subtree_5}
    \caption{Generality of ideas}
\end{figure}

The top panel in Figure FIG suggests that ideas tend to be less general in the conditions where we request more ideas. However, the bottom panel, which shows generality only for the first 5 responses, has no such effect. This suggests that only answers later on in the larger request batches become less general, an effect that we further examine in the following section.

In this previous section, we showed that the rate of idea convergence, the originality of ideas, and the generality of ideas varied between the number requested conditions. These relationships are useful to understand when deciding what crowd brainstorming strategy to follow. What this birds eye view excludes is an understanding of \emph{how} these differences occur. To that end, we shift our attention to the context of a single brainstorming run.

\subsection{Brainstorming Runs}

The large, experiment-scale trends we describe above are composed of ideas generated in many individual brainstorming runs.

\subsubsection{Originality}

In the previous section, we showed that originality at both the idea and category tree level rose as the number of responses solicited increased. Parnes et al fround that ideas generated in the latter half of a brainstorming session were rated more original in colocated, cotemporal idea generation groups \cite{parnes_effects_1961}. We hypothesize that we should see a difference in idea and category oscore as a function of position in the brainstorming run, with later ideas performing better. Figure FIG presents the mean oscroe ratings on this scale.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{run_idea_oscore}
    \includegraphics[width=0.9\columnwidth]{run_category_oscore}
    \caption{Error bars are standard error}
\end{figure}

There appears to be a clear, uniform increase in originality on both measures through the 20th idea, at which point originality peters out, or, in the case of category o-score, becomes highly variant. We tested the hypothesis that ideas were less creative after the 20th...

Thus, we confirm Parnes et al's prior result: later ideas are more original than those early in the run. Furthermore, we identify 20 ideas as a key inflection point past which participants have realiably exhausted their supply of readily available ideas and begin generating their most original ideas.

\subsubsection{Generality}

Similarly, generatility was found to decrease as the number of ideas requested increased. Figure FIG shows generality over the course of a brainstorming run.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{run_height_in_subtree}
    \caption{Error bars are standard error}
\end{figure}

Unsurprisingly, generality performs inversely to originality - the most general ideas are the least original. Also like originality, participants seem to exhaust their supply of general ideas by around the 20th idea, at which point they produce ideas at a fairly uniform generality. We tested this...

\subsubsection{SIAM hypotheses}

\cite{nijstad_how_2006}

### 2013-09-16 chi2014/goals.tex
some results discussion

@@
 
-\textbf{Hypothesis 2:} there is a set general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.
+In addition, we expect this saturation to occur at a different rate between the conditions. \textbf{Hypothesis 2:} the rate of new ideas will reduce to 0 in some conditions more quickly than others.
+
+\textbf{Hypothesis 3:} there is a set general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.
 
@@
 \begin{itemize}
-\item \textbf{Hypothesis 3:} An idea from one semantic category should more often be followed by an idea from the same category than expected by random chance
-\item \textbf{Hypothesis 4:} idea generation when changing semantic categories takes longer than when repetiting categories
+\item \textbf{Hypothesis 4:} An idea from one semantic category should more often be followed by an idea from the same category than expected by random chance
+\item \textbf{Hypothesis 5:} idea generation when changing semantic categories takes longer than when repetiting categories
 \end{itemize}
@@
 In keeping of our philosophy regarding creativity and quality measurements, we restrict the content of this test to originality.
-\textbf{Hypothesis 5:} Ideas generated in the latter half of a brainstorming session are of higher originality than ideas generated in the former half.
+\textbf{Hypothesis 6:} Ideas generated in the latter half of a brainstorming session are of higher originality than ideas generated in the former half.

### 2013-09-16 chi2014/method.tex
some results discussion

--- 2013-09-15 chi2014/method.tex
@@
 
-In order to develop an adequate model of baseline brainstorming performance in crowd marketplaces, it was necessary to establish a corpus of data.
+In order to develop models and test hypotheses, it was necessary to establish a significant corpus of brainstorming data data.
 Over the course of four months we repeatedly collected responses to brainstorming tasks.
-Throughout the experimental process, particular aspects of the experimental design remained fixed.
 
-The independent variables of interest were \emph{brainstorming problem} and \emph{number of ideas requested}.
-
-Participants were recruited from Amazon's \emph{Mechanical Turk} (CITE), and were restricted to residents of the United States for a baseline expectation of english language comprehension and cultural familiarity.
+Participants were recruited from Amazon's \emph{Mechanical Turk} \cite{_amazon_????}, and were restricted to residents of the United States for a baseline expectation of english language comprehension and cultural familiarity.
 Mechanical Turk is an online marketplace in which members receive financial reward for completing \emph{Human Intelligence Tasks}, or HITs.
 
-HITs were placed on the marketplace for each \emph{number of ideas requested} condition, with proportional rewards. Participants could accept at most one HIT in each of these conditions.
+HITs were placed on the marketplace asking for a variable \emph{number of ideas}, with proportional rewards.
+These conditions were : 5 ideas (with a corresponding reward of \$0.18 USD), 10 ideas (\$0.35), 20 (\$0.70), 50 (\$1.75), 75 (\$2.65), and 100 (\$3.50).
 
-Upon accepting a HIT, participants were randomly assigned into a \emph{brainstorming problem} condition.
-We ensured that participants completing multiple HITs are not exposed to the same brainstorming task twice.
+This served as the independent variable in all data gathering.
+It must be stated that under normal brainstorming conditions, participants are asked to generate as many ideas as possible within a fixed time frame, instead of a fixed number.
+We excluded such a condition for two reasons.
+First, it is practice in crowd creativity tasks to solicit a fixed number of ideas (e.g.
+Yu and Nickerson \cite{yu_cooks_2011}).
+Secondly, we piloted asking respondants for "as many ideas as possible" in a generous time frame.
+Over N respondents, we received a mean of N ideas (std N), despite giving a financial of \$\$\$\$.
+This suggested a propensity to "game the system", a behaviour we wanted to exclude.
 
-\subsubsection{Design Concerns}
-Workers choose which HITs to complete, thus self-selection bias is a reasonable concern. This bias is also present in real-world HIT choice behaviour.
+While the brainstorming question was manipulated in pilot studies, it was fixed for the data reported on in this paper:
+
+\begin{itemize}
+\item \textbf{MP3}
+
+Many people have old MP3 players or MP3 players that they no longer use.
+Please brainstorm N uses for old MP3 players/MP3 players.
+Assume that the devices' batteries no longer work, though they can be powered via external power sources.
+Also be aware that devices may \emph{not} have displays.
+Be as specific as possible in your descriptions.
+\end{itemize}
+
+This problem was selected for three qualities:
+
+\begin{enumerate}
+\item It requires creativity to resolve. Obvious answers do not satisfy the problem.
+\item Mechanical Turk participants have the expertise to solve the problem.
+\item It has an associated success metric.
+\end{enumerate}
+
+Workers choose which \emph{number of ideas} condition to participate in, thus self-selection bias is a reasonable concern. However, this bias is also present in real-world HIT choice behaviour.
 
 Upon accepting a HIT, participants were asked to give consent and informed that they could leave the study at any point without financial consequences.
-In regular online task markplaces, common practice would restrict workers from submitting a HIT as complete unless all responses were given.
-This is a threat to external validity.
+Under normal conditions, the participants would be expected to complete all responses to receive compensation.
+
+Should we be dropping these runs, or at least testing they don't have an affect?
 
@@
 
-The brainstorming task is below this, followed by a series of text entry inputs numbered through to the total number of ideas requested.
+The brainstorming problem is below this, followed by a series of text entry inputs numbered through to the total number of ideas requested.
 Figure X (FIG) is an example of a typical task.
@@
 
-\subsection{Pilots and Question Selection}
-
-With this basic design in hand, we ran several pilots and experiments.
-In early pilots, we used classic problems from psychology literature on brainstorming, including the "thumb problem" and "broom problem" (CITE).
-Early results from brainstormers were unsatisfying and wildly divergent.
-We identified three key traits problems needed in order to evaluate them for creativity:
-
-\begin{enumerate}
-\item They must require creativity to resolve. Obvious answers do not satisfy the problem.
-\item They must be problems that participants on Mechanical Turk would have the expertise to solve.
-\item They must have an associated success metric.
-\end{enumerate}
-
-The primary researcher and two additional researchers familiar with crowd marketplace brainstorming tasks brainstormed a large variety of potential problems.
-These problems were iterativelly culled and refined over a series of further pilots until the above goals were felt to be achieved.
-This resulted in the following four questions, as presented to the participants:
-
-\begin{enumerate}
-\item \textbf{Charity}
-
-The Electronic Frontier Foundation (EFF) is a nonprofit whose goal is to protect individual rights with respect to digital and online technologies.
-For example, the EFF has initiated a lawsuit against the US government to limit the degree to which the US surveils its citizens via secret NSA programs.
-If you are unfamiliar with the EFF and its goals, read about it on its website (https://www.eff.org) or via other online sources (such as Wikipedia).
-
-Brainstorm N \emph{new} ways the EFF can raise funds and simultaneously increase awareness.
-Your ideas \emph{must be different from their current methods}, which include donation pages, merchandise, web badges and banners, affiliate programs with Amazon and eBay, and donating things such as airmiles, cars, or stocks.
-See the full list of their current methods here: https://www.eff.org/helpout.
-Be as specific as possible in your responses."
-
-\item \textbf{Mechanical Turk}
-
-"Mechanical Turk currently lacks a dedicated mobile app for performing HITs on smartphones (iPhone, Androids, etc.) or tablets (e.g., the iPad).
-
-Brainstorm N features for a mobile app to Mechanical Turk that would improve the worker's experience when performing HITs on mobile devices. Be as specific as possible in your responses.
-
-\item \textbf{MP3}
-
-Many people have old MP3 players or MP3 players that they no longer use.
-Please brainstorm N uses for old MP3 players/MP3 players.
-Assume that the devices' batteries no longer work, though they can be powered via external power sources.
-Also be aware that devices may \emph{not} have displays.
-Be as specific as possible in your descriptions.
-
-\item \textbf{Forgot Name}
-
-Imagine you are in a social setting and you have forgotten the name of somebody you know.
-Brainstorm N ways you could learn their name without directly asking them.
-Be as specific as possible in your descriptions.
-
-\end{enumerate}
-
-Classic brainstorming procedure asks for as many ideas as possible in a period of time. We instead asked participants for a specific number of responses: 5, 10, 20, 50, 75 or 100.
-
-This decision was motivated by our early experience asking respondants to simply come up with "as many ideas as possible" in a generous time frame.
-Over N respondents, we received a mean of N ideas (std N), despite giving a financial reward N times that of any of our explicitly numbered conditions.
-This unlimited condition provided a great deal of leeway for participants to "game the system", and receive the full reward with very little work, so we dropped the condition.
-
-\subsection{Experiment}
-
-With the above problem set and the previously outlined task design (SEC), we collected responses.
-We chose 6 \emph{number of ideas requested} conditions that covered the spectrum (and beyond) of quantity of creativity requested from a single participant.
-Those conditions were: 5 ideas (with a corresponding reward of \$0.18 USD), 10 ideas (\$0.35), 20 (\$0.70), 50 (\$1.75), 75 (\$2.65), and 100 (\$3.50).
-To reiterate, participants chose a HIT to generate some number of ideas, and then were randomly assigned a problem.
-
 \subsubsection{Data Collected}
 
-In total, 341 HITs were complteed by 280 distinct participants.
+In total, 161 HITs were complteed by 146 distinct participants, for a total of 3007 responses.
 A breakdown of responses across conditions is given in table TAB.
-In cases when a single participant answered the same question multiple times, all HITs after the first were dropped for subsequent analysis.
+In cases when a single participant participated in multiple conditions, all HITs after the first were dropped for subsequent analysis.
 
@@
 & 5 & 10 & 20 & 50 & 75 & 100 \\ \hline \hline
-HITs (MP3) & 59 & 49 & 23 & 10 & 10 & 10 \\
-HITs unique (MP3) & 57 & 39 & 21 & 9 & 10 & 10 \\ \hline
-HITs (Turk) & 11 & 10 & 11 & 10 & 9 & 9 \\
-HITs unique (Turk) & 11 & 9 & 11 & 9 & 9 & 7 \\ \hline
-HITs (Forgot) & 10 & 10 & 10 & 12 & 11 & 11 \\
-HITs unique (Forgot) & 10 & 8 & 9 & 12 & 10 & 10 \\ \hline
-HITs (Charity) & 9 & 10 & 10 & 8 & 10 & 9 \\
-HITs unique (Charity) & 9 & 9 & 10 & 7 & 9 & 9 \\ \hline
-HITs (total) \\
-HITs unique (total) \\ \hline
+HITs completed & 59 & 49 & 23 & 10 & 10 & 10 \\
+HITs completed (no repeats) & 57 & 39 & 21 & 9 & 10 & 10 \\
+solutions (no repeats) & 283 & 372 & 413 & 450 & 635 & 855 \\
 \end{tabular}
 
-There is significantly more data for the MP3 condition than for any of the others.
-Despite at least 7 unique HITs for each of the 24 problem x number conditions, we found that the limited number of total responses for the 5, 10 and 20 conditions precluded significant analysis.
-To that end, we selected the MP3 player problem to further expand the data set to 400 (TODO: maybe not true) responses per number condition.
-
 HITs were augmented with JavaScript to collect data in addition to brainstorming responses.
-For each response, we collected the time of the first activation and last de-activation of the form element.
+For each response, we collected the time of the first activation and last de-activation of the corresponding form element.
 We collected the time the HIT was accepted and the time it was submitted.
-We took a has of the Mechanical Turk worker ID as a unique identifier of our participants.
+To uniquely identify participants, we computed a hash of their Mechanical Turk worker ID.
+A hash was used to protect privacy after data release.
 
@@
 
-As we began to examine the data, it quickly began apparent that there are nuances in discussing responses to brainstorming tasks.
-In the interest of facilitating this discussion, we introduce terms here that will be used for the remainder of the paper.
+To facilitate discussion, we introduce terms here that we will use throughout the rest of the paper.
 
-An \emph{instance} is a single response by a single participant to a brainstorming question. For example, Figure FIG gives four \emph{instances} provided in the \emph{MP3 player} brainstorming task.
+An \emph{instance} is a single solution by a single participant to a brainstorming question.
+Alternatively, an instance may be thought of as the text entered into a single field in the task form.
+For example, figure~\ref{fig:sample_instances} gives four \emph{instances} from the data set.
 
@@
     \end{enumerate}
+    \label{sample_instances}
 \end{figure}
 
-An instance is associated with exactly one \emph{idea}.
-An idea may have 0 or more instances, all of which describe an identical solution to a problem.
+A \emph{brainstorming run}, or \emph{run}, is the temporally-ordered series of instances that were received in a single HIT by a single user.
+As stated above, after dropping repeat data, all brainstorming runs were completed by distinct participants.
+
+An \emph{idea} is a unique solution to the brainstorming problem.
+An instance is associated with exactly one \emph{idea}, and all instances in the same idea should be paraphrases of one another.
 In Figure FIG, "Use as a mirror" and "Mirror" belong to the same \emph{idea}.
-However, "Use it as a mirror for makeup" does not belong to the same idea, as it adds additional information to the problem solution; it is more specific.
+However, "Use it as a mirror for makeup" does not belong to the same idea, as it contains additional information.
 
@@
 It is desireable to encode this relationship.
-To that end, we organize ideas in \emph{category trees}.
+To that end, we propose a hierarchical clustering of ideas in \emph{category trees}.
 In a category tree, ideas act as nodes.
@@
 If an idea \emph{A} is a parent of idea \emph{B}, A is a generalization of B: all instances in A are are suitably general to encompass all instances in B.
-This difference is relative to the ideas in question.
+This hierarchical clustering allows us to classify ideas as simultaneously different in solution stragy but different in solution details.
 
-A \emph{question forest} is simply the collection of all category trees generated in response to a particular question. Different trees in a question forest have no special relationship in meaning.
+We further distinguish category trees into \emph{singleton} and \emph{non-singleton} category trees.
+A singleton category tree is composed of only one idea node, the root.
+A non-singleton category tree has at least two idea nodes.
+We introduce non-singleton category trees as correspondents to the images in Nisjstad and Stroebe's SIAM model \cite{nijstad_how_2006}.
+On SIAM, the idea generation process is represented by a series of images activations within which idea production occurs.
+Non-singleton trees by definition represent strategies that have been produced in at least two distinct ways.
 
-\subsection{Coding}
+The \emph{idea forest} is the collection of all category trees. Different trees in the forest have no particular relationship.
 
-One researcher coded responses the MP3 problem to produce an idea forest as described above. The MP3 problem was chosen because it was subjectively easiest to distinguish between and categorize ideas.
+\subsection{Clustering}
 
-The clustering algorithm is described in Figure FIG.
+In addition to the notion of an idea forest, we also propose a clustering algorithm driven by human intelligence to create the data structure.
+The algorithm is described in Figure~\ref{fig:cluseringalg}.
 Instances are added one at a time to the existing idea forest.
@@
 \subsubsection{Similarity}
-Two ndodes a and b have high similarity if they solve the problem in the same way, and low similarity if their solutions have no common themes.
+Two nodes a and b have high similarity if they solve the problem in the same way, and low similarity if their solutions have no common themes.
 For example, in Figure FIG, "Mirror" and "Tape it to a telephone pole and use as a reflector" have high similarity, as they both solve the problem of using a broken MP3 player by using it to reflect light.
@@
 
-After completion of the coding algorithm, we were left with an idea forest as described in section SEC.
+The primary author followed the algorithm to cluster the 3007 response instances into ideas and trees.

### 2013-09-16 chi2014/results.tex
some results discussion

@@
 
-\subsection{Between-Conditions Comparisons}
+We can use the uniqueness measure to understand the rate at which new ideas arrive.
+As we gather more and more ideas in the course of an experiment, we expect that we will see fewer ideas that are not already in our idea pool.
+Figure FIG shows the cumulative count of ideas as a function of the number of instances gathered in the experiment.
+To smooth the plot, the order in which instances were received was shuffled 100 times, and the mean of the cumulative unique idea count was taken at each step.
 
-\subsubsection{Saturation of ideas}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{ideas_over_time}
+    \caption{Cumulative idea count}
+\end{figure}
 
-One of our hypotheses (HYP) was that as the number of instances gathered increased, the rate of new ideas, new categories and new singleton categories would converge towards 0.
+The horizontal line in Figure FIG represent the idea case, in which each instance represents a different idea (i.e.
+every answer we receive is a new, unique idea).
+As expected, the rate at which ideas are gained appears to taper off.
 
-Figure FIG shows the rate of new idea, category, and non-singleton category generation  over time for each number requested conditions.
-These plots have been smoothed by randomly shuffling the order in which ideas were receieved 500 times and taking the mean cumulative count at each point in time, and plotting with a window size of 10.
-In an ideal scenario where each instance provided a new unique idea and category, the lines would horizontal across the top of the figure.
+Figure FIG more closely examines the rate, the primary property of interest.
+The panels, from top to bottom, are the rate at which new ideas, categories, and non-singleton categories are generated, respectively.
+Similar to Figure FIG, these are based on shuffling the order of instances 10000 times and taking the mean of the derivative at each point.
+In this case, the ideal (a 1:1:1 instance:idea:category ratio) is represented by the horizontal bar across the top of the chart.
 
-\begin{figure}
+\begin{figure}[h!]
     \centering
-    \includegraphics[width=0.9\columnwidth]{rate_new_idea_over_time}
-    \includegraphics[width=0.9\columnwidth]{rate_new_category_over_time}
-    \includegraphics[width=0.9\columnwidth]{rate_new_ns_category_over_time}
+    \includegraphics[width=0.7\columnwidth]{rate_new_idea_over_time}
+    \includegraphics[width=0.7\columnwidth]{rate_new_category_over_time}
+    \includegraphics[width=0.7\columnwidth]{rate_new_ns_category_over_time}
     \caption{Rate of idea generation}
 \end{figure}
+
+The number of ideas, categories and non-singleton each drop off exponentially over time.
+It is unsurprising that the number of new categories should drop off faster than the number of new ideas, as we expect that there are fewer categories of ideas than there are individual ideas themselves.
+
+The third panel, which shows non-singleton categories over time, tells a distinctly different story from the others.
+While in the idea and category plots, the height-ordering of rates of generation between categories remains generally identical, there is a point of intersection and reversal in the non-singleton category plot at around the 30 instances point.
+Before this point, the lower conditions actually generate new categories at a \emph{faster} rate.
+
+In lower conditions, brainstormers will quickly offer up big, popular, non-singleton categories.
+Beyond the inflection point, the category pool is saturated with these low-hanging fruit, and only brainstormers tasked with generating more ideas will find the smaller category.
+This is a more nuanced view of the tree node/instance quartiles in Section SEC; condition 5 brainstormers cover the spectrum of category sizes while condition 100 brainstormers pull from the smaller category trees in the forest.
+
+\subsection{Hypothesis 1}
 
@@
 
-\subsubsection{Originality}
+\subsection{Originality}
 
@@
 This finding will be examined in the next section.
-
-\subsubsection{Generality}
-
-As discussed above, we use height in category tree as a measure of an ideas generality. Generality (and its reverse, sepcificity) are both potentially desireable outcomes of a brainstorming task.
-
-The idea-height across number conditions is given in figure FIG.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.9\columnwidth]{height_in_subtree}
-    \includegraphics[width=0.9\columnwidth]{height_in_subtree_5}
-    \caption{Generality of ideas}
-\end{figure}
-
-The top panel in Figure FIG suggests that ideas tend to be less general in the conditions where we request more ideas.
-However, the bottom panel, which shows generality only for the first 5 responses, has no such effect.
-This suggests that only answers later on in the larger request batches become less general, an effect that we further examine in the following section.
-
-In this previous section, we showed that the rate of idea convergence, the originality of ideas, and the generality of ideas varied between the number requested conditions.
-These relationships are useful to understand when deciding what crowd brainstorming strategy to follow.
-What this birds eye view excludes is an understanding of \emph{how} these differences occur.
-To that end, we shift our attention to the context of a single brainstorming run.
 

### 2013-09-16 chi2014/summary.tex
some results discussion

@@
-\section{Summary of Data}
+\section{Summary of Idea Forest}
 
-The resulting forest was composed of 3007 instances, 1213 ideas, and 335 category trees.
-The number of instances, ideas and categories present in each of the number conditions is given in table TAG.
-In addition, table TAG shows the number of "non-singleton" categories, namely, those categories containing at least two instances.
-Intuitively, non-singleton categories represent those categories in which some exploration actually occured.
-These represent categories in which we can potentially identify the idea production phase of the SIAM model [].
+The resulting forest was composed of 3007 instances, 1213 ideas, 335 category trees, and 129 non-singleton category trees.
+These metrics are decomposed by number condition is given in table TAB.
+A bird's eye view of the complete idea forest is in Figure FIG.
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{idea_forest}
+    \caption{Idea Forest}
+\end{figure}
+
+\begin{table}
 	\begin{tabular}[h!]{r | l l l l l l }
-& 5 & 10 & 20 & 50 & 75 & 100 \\ \hline \hline
+	\textbf{number condition} & 5 & 10 & 20 & 50 & 75 & 100 \\ \hline \hline
 	HITs & 57 & 47 & 23 & 10 & 10 & 10 \\
@@
 	\end{tabular}
+\end{table}
 
-Figure FIG shows the distribution of instances in non-singleton categories. The majority of the "mass" of instances are in trees with less than 70 instances.
+The height of category trees follows a roughly Poisson distribution, with the maximum tree height observed at 5.
+Summary statistics across condition are in table TAB.
+The third quartile drops as the number condition increases.
+While the upper conditions still discover categories trees with heights as great as 5, they generate many more "short" trees.
+This suggests that the ideas generated in the upper conditions are less common.
 
-% TODO: this figure needs better labeling for the second part
-\begin{figure}[!h]
-    \centering
-    \includegraphics[width=0.9\columnwidth]{instances_in_subtrees}
-\end{figure}
+\begin{table}
+	\begin{tabular}[h!]{r | l l l l l l l}
+	\textbf{number condition} & 5 & 10 & 20 & 50 & 75 & 100 & all \\ \hline \hline
+	median & 2 & 2 & 2 & 2 & 2 & 2 & 2\\
+	first quartile & 2 & 1  & 1 & 1 & 1 & 1 & 1\\
+	third quartile & 2 & 3 &3 &3 &2 &2 & 2\\
+	\end{tabular}
+	\caption{Category tree height}
+\end{table}
+
+The number of nodes and number of instances are also roughly poisson distributed (Table TAB).
+In keeping with the findings for height, trees found in the upper number conditions are smaller than trees found in the lower conditions.
+The precipitous drop in median tree size suggests that in the largest condition, participants are finding rare ideas with only a few variants and less than 14 instances.
+
+\begin{table}
+\begin{tabular}[h!]{r | l l l l l l l}
+	\textbf{number condition} & 5 & 10 & 20 & 50 & 75 & 100 & all \\ \hline \hline
+	number of nodes& \\ \hline
+	median &3&4&4&3&2&2&2 \\
+	first quartile &2&1&1&1&1&1&1 \\
+	third quartile &15&12&12&10&5&5&5 \\
+	number of instances& \\ \hline
+	median &14&12&12&9&4&4&4 \\
+	first quartile &4&3&4&2&1&1&1 \\
+	third quartile &48&31&29&21&14&13&13 \\
+	\end{tabular}
+	\caption{Number of ideas and instances in trees}
+\end{table}
+
+To better understand what this distribution of tree sizes means...

### 2013-09-17 chi2014/goals.tex
so close on results :D

@@
 This is the simplest of conditions that can vary when choosing a design for a crowd brainstorming problem.
-Furthermore, we detail the phenomenon of \emph{riffing} (known in the idea generation literature as \emph{clustering}, a term we use for its classification meaning later in the paper).
+Furthermore, we detail the phenomenon of \emph{riffing} (known in the idea generation literature as \emph{clustering}, a term we must unfortunately reserve for it's classification-related meaning).
 

### 2013-09-17 chi2014/measures.tex
so close on results :D

\section{Measures}

\subsubsection{Originality}

As our measure of originality, we use \emph{o-score}, introduced by Jansson and Smith \cite{jansson_design_1991}. An idea's o-score is $1 - p(idea)$, where $p(idea) = (number of instances of that idea)/(number of instances total)$. The o-score for category trees is similarly calculated. We will occasionally refer to the \emph{category o-score} of an idea. This simply refers to the o-score of the category tree to which that idea belongs.

\subsubsection{Uniqueness}

To examine the quantity of idea produced, we must first have some method of comparing two ideas to see if they are unique. We use three such comparisons. The first, \emph{idea equality}, equates two instances if they belong to the same idea node. The second, \emph{category equality}, equates two instances if their idea nodes are in the same tree. The former requires exact equality in meaning, while the latter requires only that the same strategy is employed.

The final measure, \emph{look-back equality}, is non-symmetrical and has meaning only in the context of a single brainstorming run. An instance \emph{a} is look-back equal to an \emph{earlier} instance \emph{b} if \emph{b} is in either \emph{a}'s idea node or any of \emph{a}'s parents, siblings, or children. This definition specifically identifies when a later instance in a run can be considered to have been influenced by earlier instance.

### 2013-09-17 chi2014/results.tex
so close on results :D

@@
 
-\subsection{Hypothesis 1}
+\subsubsection{Hypothesis 1}
 
@@
 Generally, conditions with more requested responses generated ideas and categories at a higher rate.
-We tested this...
+
+Test here...
+
+\subsubsection{Hypothesis 2}
+
+Test here...
 
@@
 
-We were also interested in the effect of the number requested condition on originality.
-Originality, as measured by o-scores at both the category and idea level, is compared cross condition in figure FIG.
-Some stat tests here...
+We measure differences in originality between conditions by examining the idea o-score and category o-score.
+The distributions of these scores are summarized in Figure FIG.
+The left panel compares idea o-score and the right compares category o-score.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{oscores_conditions}
+\end{figure}
+
+The more ideas requested, the more original the ideas produced.
+As suggested by the instance quartile range in section SEC, higher number conditions produce idea in trees with fewer instances, thus the high category o-score.
+We also see that these conditions produce \emph{ideas} with fewer instances.
+
+As the originality rises, however, so does the number of outliers.
+To understand this phenomenon, we need to more closely observe the distribution of originality scores between not just conditions, but ordinal position in a brianstorming run.
+
+Figure FIG provides this visualization.
+The upper panel gives the mean idea o-score as a function of ordinal position in all 100 condition brainstorming runs.
+The o-score for each ordinal position is the mean of all idea o-scores for that ordinal position in a run.
+Following this, the plot is further smoothed by making each point the mean of a sliding window of size 10 around its ordinal position.
+The bottom panel is a similar plot for category-oscore.
+To aid in interpretation, the first, second and third quartiles are also represented on the plot.
+The error bars are standard error.
 
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{subtree_oscore}
-    \includegraphics[width=0.9\columnwidth]{idea_oscore}
-    \caption{Error bars are standard error}
+    \includegraphics[width=0.9\columnwidth]{idea_oscore_order_100}
+    \caption{Idea o-score as a function of order (100 condition)}
+    \includegraphics[width=0.9\columnwidth]{cat_oscore_order_100}
+    \caption{Category o-score as a function of order (100 condition)}
 \end{figure}
 
-The more ideas requested, the more original the ideas produced.
-This provides support for the findings of Parnes et al \cite{parnes_effects_1961}, namely that more original ideas are found in extended idea generation events.
-This finding will be examined in the next section.
-
-\subsection{Brainstorming Runs}
-
-The large, experiment-scale trends we describe above are composed of ideas generated in many individual brainstorming runs.
-
-\subsubsection{Originality}
-
-In the previous section, we showed that originality at both the idea and category tree level rose as the number of responses solicited increased.
-Parnes et al fround that ideas generated in the latter half of a brainstorming session were rated more original in colocated, cotemporal idea generation groups \cite{parnes_effects_1961}.
-We hypothesize that we should see a difference in idea and category oscore as a function of position in the brainstorming run, with later ideas performing better.
-Figure FIG presents the mean oscroe ratings on this scale.
+The only part of the originality score that falls outside the first and third quartiles is the beginning of the run.
+Participants in the higher numbered conditions generate more original ideas overall, but not until they have exceeded some threshold of common ideas.
+Figure FIG replicates figure FIG but across all conditions, and the same pattern of originality growth until an originality peak - around 20 ideas - is reached.
+Note that because of the additional HITs, there are far more examples of early-order ideas in the 5, 10 and 20 conditions than in the 50, 75, or 100, so this pattern is present without any dominating affect by the upper conditions.
 
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{run_idea_oscore}
-    \includegraphics[width=0.9\columnwidth]{run_category_oscore}
-    \caption{Error bars are standard error}
+    \includegraphics[width=0.9\columnwidth]{idea_oscore_order}
+    \caption{Idea o-score as a function of order}
+    \includegraphics[width=0.9\columnwidth]{cat_oscore_order}
+    \caption{Category o-score as a function of order}
 \end{figure}
 
-There appears to be a clear, uniform increase in originality on both measures through the 20th idea, at which point originality peters out, or, in the case of category o-score, becomes highly variant.
-We tested the hypothesis that ideas were less creative after the 20th...
+A similar, although less severe, pattern is observed for category o-score.
+There is a gradual increase in category originality until the 20 instance threshold is reached, at which point variance of the scores increases rapidly and no clear pattern is observable.
+This is explained by the relationship between common ideas and common categories.
+We expect that unoriginal ideas belong to unoriginal categories, as a high instance count for an idea increases the instance count for its category tree.
+However, we cannot make the inverse assumption, that a high originality idea belongs to a high originality category.
+Once the common idea are exhausted, then, there is no relationship between idea originality and category originality.
 
-Thus, we confirm Parnes et al's prior result: later ideas are more original than those early in the run.
-Furthermore, we identify 20 ideas as a key inflection point past which participants have realiably exhausted their supply of readily available ideas and begin generating their most original ideas.
+\subsubsection{Hypothesis 6}
 
-\subsubsection{Generality}
+This threshold at 20 ideas brings us to \textbf{Hypothesis 6:} Ideas generated in the latter half of a brainstorming session are of higher originality than ideas generated in the former half.
+With our empirical evidence, we can instead test that ideas \emph{more than 20 instances into a brainstorming session} are more original than those in the first 20 instances, regardless of condition.
 
-Similarly, generatility was found to decrease as the number of ideas requested increased. Figure FIG shows generality over the course of a brainstorming run.
+Test here...
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.9\columnwidth]{run_height_in_subtree}
-    \caption{Error bars are standard error}
-\end{figure}
-
-Unsurprisingly, generality performs inversely to originality - the most general ideas are the least original.
-Also like originality, participants seem to exhaust their supply of general ideas by around the 20th idea, at which point they produce ideas at a fairly uniform generality.
-We tested this...
+\subsection{Riffing}
 

### 2013-09-17 chi2014/summary.tex
first draft

@@
 \begin{table}
-	\begin{tabular}[h!]{r | l l l l l l }
-	\textbf{number condition} & 5 & 10 & 20 & 50 & 75 & 100 \\ \hline \hline
-	HITs & 57 & 47 & 23 & 10 & 10 & 10 \\
-	instances & 293 & 471 & 453 & 500 & 634 & 855 \\
-	ideas & 171 & 249 & 278 & 341 & 443 & 177 \\
-	category trees & 76 & 93 & 100 & 123 & 172 & 177 \\
-	non-singleton trees & 28 & 41 & 41 & 50 & 49 & 61 \\
+	\begin{tabular}[h!]{r | l l l l l l l}
+	\textbf{number condition} & 5 & 10 & 20 & 50 & 75 & 100 & all \\ \hline \hline
+	HITs & 57 & 47 & 23 & 10 & 10 & 10 & 146\\
+	instances & 293 & 471 & 453 & 500 & 634 & 855 & 3007\\
+	ideas & 171 & 249 & 278 & 341 & 443 & 177 &1212\\
+	category trees & 76 & 93 & 100 & 123 & 172 & 177 &321\\
+	non-singleton trees & 28 & 41 & 41 & 50 & 49 & 61 &92\\
 	\end{tabular}
@@
 	\textbf{number condition} & 5 & 10 & 20 & 50 & 75 & 100 & all \\ \hline \hline
-	median & 2 & 2 & 2 & 2 & 2 & 2 & 2\\
+	median & 2 & 2 & 2 & 2 & 2 & 2 & 1\\
 	first quartile & 2 & 1  & 1 & 1 & 1 & 1 & 1\\
@@
 	number of nodes& \\ \hline
-	median &3&4&4&3&2&2&2 \\
+	median &3&4&4&3&2&2&1 \\
 	first quartile &2&1&1&1&1&1&1 \\
-	third quartile &15&12&12&10&5&5&5 \\
+	third quartile &15&12&12&10&5&5&3 \\
 	number of instances& \\ \hline
-	median &14&12&12&9&4&4&4 \\
+	median &14&12&12&9&4&4&2 \\
 	first quartile &4&3&4&2&1&1&1 \\
-	third quartile &48&31&29&21&14&13&13 \\
+	third quartile &48&31&29&21&14&13&6 \\
 	\end{tabular}

### 2013-09-17 src/modelling/output_cumulative_csv.py
modelling code

@@ -1,0 +1,12 @@
+#!/usr/bin/python
+
+import json
+
+with open('cumulative_ideas_categories.json') as fin:
+  rows = json.loads(fin.read())
+
+with open('cumulative_data.csv', 'w') as fout:
+  header = ['tuple_id', 'id', 'condition', 'response_num', 'num_cum_ideas', 'num_cum_categories']
+  fout.write('\t'.join(header) + '\n')
+  for row in rows:
+    fout.write('\t'.join([str(x) for x in row]) + '\n')

### 2013-09-17 src/modelling/stan_log_reg_model.r

@@ -1,0 +1,234 @@
+require(rstan)
+
+kruschke_script_dir = '/home/mterry/Documents/probability_work/kruschke_scripts'
+do_plot_post = function(data) {
+  cur_dir = getwd()
+  setwd(kruschke_script_dir)
+  source('plotPost.R')
+  plotPost(data)
+  setwd(cur_dir)
+}
+  
+do_stan_cumulative_analysis_log_model = function(df, measure) {
+model_string = "
+data {
+  int N;
+  real y[N];
+  int x[N];
+  int num_cond;
+  int cond_group[N];
+}
+
+parameters {
+  real<lower=0> b_cond[num_cond];
+  real<lower=0> x_scale;
+  real x_trans;
+  real y_trans;
+  real<lower=2> log_base;
+  real<lower=0> sigma;
+}
+
+transformed parameters {
+}
+
+model {
+  real mu[N];
+  for (i in 1:N) {
+    mu[i] <- y_trans + b_cond[cond_group[i]] * log(x_scale * x[i] + x_trans) / log(log_base);
+     y[i] ~ normal(mu[i], sigma);
+  }
+}
+";
+  num_conditions = max(df$condition_num)
+  data_list = list(
+    N = nrow(df),
+    y = df[,measure],
+    x = df$response_num+1,
+    cond_group = df$condition_num,
+    num_cond = num_conditions
+  );
+
+  n_chains = 3;
+  num_saved_steps = 20000;
+  n_iter = ceiling(num_saved_steps / n_chains)
+  print(paste("Fitting log model for", measure))
+  fit = stan(model_code = model_string, data = data_list, iter = n_iter, chains = n_chains)
+
+  print(paste("Log model for", measure))
+  print(fit)
+  par(mfrow=c(1,1))
+  plot(fit)
+
+  list_of_arrays = extract(fit, permuted=TRUE)
+  b0_list = list_of_arrays$b0
+  b_cond_list = list_of_arrays$b_cond
+  x_scale_list = list_of_arrays$x_scale
+  x_trans_list = list_of_arrays$x_trans
+  sigma_list = list_of_arrays$sigma
+
+  par(mfrow=c(2,2))
+  hist(b0_list, main=paste("b0", measure))
+  hist(x_scale_list, main=paste("x_scale", measure))
+  hist(x_trans_list, main=paste("x_trans", measure))
+  hist(sigma_list, main=paste("sigma", measure))
+  b_cond = array(dim=c(num_conditions))
[…]

### 2013-09-17 chi2014/results.tex

@@
 
-\subsection{Riffing}
+\subsection{Riffing: We need to change the name of this}
 
-\subsubsection{SIAM hypotheses}
+At the level of the individual brainstorming run, we examine a new phenomenon related to uniqueness above.
+Within the SIAM model, \emph{riffing} (or \emph{clustering} as it is known in the work) occurs when a single participant activates a single image (analogous to a non-singleton category) and produces multiple ideas from that image under a fail state is reached and a transition is undergone to a new image.
+Unsurprisingly, we observed riffing activity in our data.
+An example of a single participant's riffed responses is given in Figure FIG.
+The second column is dark if the instance is a \emph{look-back equal} to any idea earlier in the run.
+The first column is dark if the instance is a \emph{source} instance, meaning it is the first in a chain of look-back equal instances.
+Riff chains can either be consecutive (as in the example), or separated with instances in between.
+The third column simply provides the text for that instance.
 
-\cite{nijstad_how_2006}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\columnwidth]{10_riff}
+    \caption{Example of riffing in a size 10 run}
+\end{figure}
+
+The participant hits on the categories of using the device's hard drive for storage, recycling the device or its materials, and selling the device.
+within these categories, they generate 2-3 ideas each.
+This is a fairly typical example of riffing.
+In larger conditions, it is common for participants to reach further back and return to an idea multiple time.
+Table TAB provides a summary of riffing between conditions.
+
+\begin{table}
+\begin{tabular}[h!]{r | l l l l l l l}
+    \hline \hline \textbf{condition} & 5 & 10 & 20  \\ \hline \hline
+    \% riffs & 13.65 & 19.28 & 25.17  \\
+    source instances & 10.58 & 12.92 & 15.01 \\
+    riffs per source & 1.29 & 1.49 & 1.68 \\
+    median length of consecutive chain & 2 & 2 & 2 \\
+    median distance to previous in chain & 1 & 2 & 2 \\
+    median distance to first in chain & 2 & 3 & 5 \\ \hline \hline
+    \textbf{condition} & 50 & 75 & 100 \\ \hline \hline
+    \% riffs & 39.6 & 32.81 & 42.81 \\
+    source instances & 19 & 17.19 & 18.13\\
+    riffs per source & 2.08 & 1.91 & 2.36\\
+    median length of consecutive chain & 2 & 2 & 2\\
+    median distance to previous in chain & 7 & 10 & 5\\
+    median distance to first in chain & 16.5 & 18 & 26.5 \\
+    \end{tabular}
+    \caption{Riffing statistics by condition}
+\end{table}
+
+The higher the number of instances requested, the more participants riff on their old ideas.
+Similarly, upper conditions use more of their ideas as sources for riffing, and riff more times on each of those ideas.
+The primary surprise is that they do not riff more consecutively.
+The median length of a consecutive riff chain remains low across conditions.
+This leaves us with the image of a participant in the upper conditions searching through their old ideas for something to build upon one instance at a time.
+
+Another outlier of note is the relatively short median distance to a previous idea in a chain for the 100 condition.
+Participants in the 100 condition may feel more pressure to remix, and return to earlier ideas more quickly than those in the 50 or 75 conditions.
+
+\subsubsection{Hypothesis 4}
+
+Riffing activity is analogous to idea production within an image as described by Nijstad and Stroebe's SIAM model \cite{nijstad_how_2006}.
+Thus, we should expect to confirm their hypothesis, \textbf{Hypothesis 4}, that an idea from one semantic category should more often by followed by an idea from the same category than expected by random chance.
+
+Test here...
+
+\subsection{Completion Time}
+
+As another aspect of how participants complete brainstorming runs, we are interested in the time spent on each response.
+The first, second and third quartiles for each condition are given in boxplots in figure FIG.
+The whiskers extend to 1.5 times the inner quartile range.
+Surprisingly, we see many instances taking up to 28018382ms (7.78 hours) to complete.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{time_spent_condition}
+    \caption{Time spent per instance}
+\end{figure}
+
+Either our participants were exceedingly concerned with crafting their responses, or there is another influence at work.
+Figure FIG gives the mean time spent on an instance by order, separated by condition.
+Most of the highly variant response times take place early in the brainstorming run.
+This suggests that at least one participant accepted a brainstorming task, took an initial stab at instances, and then left the task for up to 8 hours before continuing.
+Beyond this, we might expect to see time spent per instance go up as participants have to dig deeper to find responses, but no such effect is consistently observed.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{time_spent_order}
+    \caption{Time spent per instance}
+\end{figure}
+
+\subsubsection{Hypotheses 5}
+
+We must also consider time spent in light of the hypotheses made by Nijstad and Stroede \cite{nijstad_how_2006}.
+The second, \textbf{Hypothesis 5}, states that idea generation when changing semantic categories should take longer than idea generation within categories.
+
+Test here...

### 2013-09-17 chi2014/summary.tex
fixed an error where repeat workers were counted towards cluster size

@@
 	ideas & 171 & 249 & 278 & 341 & 443 & 177 &1212\\
-	category trees & 76 & 93 & 100 & 123 & 172 & 177 &321\\
-	non-singleton trees & 28 & 41 & 41 & 50 & 49 & 61 &92\\
+	category trees & 72 & 79 & 93 & 114 & 172 & 177 &321\\
+	non-singleton trees & 28 & 34 & 40 & 48 & 49 & 61 &92\\
 	\end{tabular}
@@
 	\textbf{number condition} & 5 & 10 & 20 & 50 & 75 & 100 & all \\ \hline \hline
-	median & 2 & 2 & 2 & 2 & 2 & 2 & 1\\
-	first quartile & 2 & 1  & 1 & 1 & 1 & 1 & 1\\
-	third quartile & 2 & 3 &3 &3 &2 &2 & 2\\
+	median & 2 & 2 & 2 & 2 & 2 & 2 & 2\\
+	first quartile & 2 & 2  & 1 & 1 & 1 & 1 & 1\\
+	third quartile & 3 & 3 &3 &3 &2 &2 & 2\\
 	\end{tabular}
@@
 	number of nodes& \\ \hline
-	median &3&4&4&3&2&2&1 \\
-	first quartile &2&1&1&1&1&1&1 \\
-	third quartile &15&12&12&10&5&5&3 \\
+    median &5&5&4&3&2&2&2 \\
+	first quartile &2&2&1&1&1&1&1 \\
+	third quartile &17&17&14&10&5&5&5 \\
 	number of instances& \\ \hline
-	median &14&12&12&9&4&4&2 \\
-	first quartile &4&3&4&2&1&1&1 \\
-	third quartile &48&31&29&21&14&13&6 \\
+	median &14&14&12&10&4&4&4 \\
+    first quartile &4&4&4&2&1&1&1 \\
+	third quartile &47&45&30&23&13&13&13 \\
 	\end{tabular}

### 2013-09-17 src/clusterer/qt_app.py
some hypothesis testing

@@ -2,6 +2,7 @@
 import os
 import sys
 import json
+import pickle
 import re
 import string
 from PySide import QtCore, QtGui
@@ -371,6 +372,11 @@ class IdeaTreeModel(QtCore.QAbstractItemModel):
     with open("%s_%s_clusters.txt" % (filename, self.question_code), 'w') as cfout:
       lines = self.root.as_text()
       cfout.write('\n'.join(lines))
+
+    pf = "%s_%s_IdeaTreeNode.pickle" % (filename, self.question_code)
+    print(pf)
+    with open(pf, 'wb') as cfout:
+        pickle.dump(self.root, cfout)
 
 
   def export_clusters(self, filename):

### 2013-09-17 chi2014/introduction.tex
New intro, no abstract, new paper title

--- 2013-08-13 chi2014/introduction.tex
@@
-This format is to be used for submissions that are
-published in the conference proceedings.  We wish to give
-this volume a consistent, high-quality appearance. We
-therefore ask that authors follow some simple
-guidelines. In essence, you should format your paper
-exactly like this document. The easiest way to do this is
-simply to download a template from the conference web
-site, and replace the content with your own material.
+%Whether in research or industry, productivity can be hamstrung by the need to generate creative or novel ideas.
+Given the ubiquity of idea generation tasks, it is desireable to provide generalizable automated methods of solving them.
+Automated computational methods to solve specific problems are unlikely to appear without significant advances in natural language processing and artificial intelligence.
+As a result, we are left with human-centric solutions, such as brainstorming.
+
+Microtask marketplaces such as Amazon's Mechanical Turk have proven to be useful platforms for outsourcing a wide range of tasks, from proofreading documents \cite{soylent} to estimating the caloric content of pictures of food \cite{platemate}.
+These platforms make possible human computation systems \cite{quinn_2011}, hybrid computational systems that tap humans' abilities to solve problems that currently cannot be solved by machines alone.
+
+A particularly compelling use of microtask marketplaces is to tap into the collective creativity of crowds.
+This enables one to quickly sample a wide range of ideas from a diverse population, which is useful for tasks such as brainstorming or design.
+Aaron Koblin's art piece ``The Sheep Market'' \cite{koblin_2009} was an early demonstration of this potential.
+In this project, he asked 10,000 individuals to draw a ``sheep facing left,'' yielding a wide range of unique results.
+More recently, there has been interest in systems that support crowd-based design.
+For example, Yu and Nickerson developed a system that combines genetic algorithms with design input from crowd workers to explore a design space \cite{yu_cooks_2011}.
+
+In this paper, we focus on crowd-based brainstorming.
+While brainstorming has been an implicit component of many creative endeavors in microtask marketplaces, we are unaware of any research that specifically describes and models brainstorming in this environment.
+Furthermore, while there is a rich literature describing brainstorming processes of individuals and groups (e.g., \cite{osborn_applied_1957, pinsonneault_electronic_1999, nijstad_how_2006} OTHERS?), microtask marketplaces are qualitatively different from previously studied environments, calling into question how much of this past research maps onto this environment.
+For example, brainstorming participants in a microtask marketplace are paid for their participation, and are often eager to complete a task as fast as possible to move on to the next paying task [REF?].
+Given these and other differences, there are many open questions about how brainstorming unfolds in this context.
+
+%Brainstorming is a tool proposed by Osborne \cite{osborn_applied_1957} that has achieved widespread popularity and similarly widespread research and controversy.
+Brainstorming methods have been applied in many scenarios, from spatially and temporally co-located participants to electronic brainstorming over computer networks \cite{pinsonneault_electronic_1999}.
+The introduction of crowd marketplaces such as Amazon's Mechanical Turk \cite{_amazon_????} and the corresponding paradigm of massively parallel Human Intelligence Tasks provide both a natural venue to solicit ideas and a new context in which to establish models of brainstorming activity.
+
+%This setting introduces unique challenges.
+For example, crowd marketplaces are driven by financial transactions, while participants in traditional brainstorming settings are required and motivated to generate ideas for a given duration.
+Guidelines and points of comparison are needed.
+
+% In the course of their research in Human-Computer Interactions, two of the authors went looking for these sets of guidelines to apply them to a crowd idea generation task.
+While previous work has utilized brainstorming techniques in crowd marketplaces [CITE], we were unable to find any generalizable models or guidelines for brainstorming in the context of the crowd.
+Models such as Nijstaf and Stroebe's SIAM \cite{nijstad_how_2006} make and test predictions regarding in-person brainstorming, but we did not find similar baselines for electronic or crowd brainstorming.
+
+This paper presents results from a series of experiments intended to model brainstorming on microtask marketplaces (specifically, Amazon's Mechanical Turk).
+In our experiments, we asked individuals to brainstorm ways to use an MP3 player whose batteries were dead.
+We varied the number of ideas requested (5, 10, 20, 50, 75, 100, or unlimited), paying a fixed amount per idea (3.5 cents).
+Across all conditions, 154 workers submitted responses to 170 HITs (Human Intelligence Tasks), yielding 3292 ideas.
+From these ideas, we present a series of models characterizing features of crowd brainstorming, such as the rate of novel idea production, the rate at which novel {\em categories\/} of ideas are produced (where a category represents a set of conceptually similar ideas that are distinct from other categories), and the relative novelty of ideas, all as functions of number of responses requested.
+
+Our primary findings and contributions are as follows:
+\begin{itemize}
+\item The rates at which novel ideas and categories of ideas are produced follow logarithmic growth models, as a function of the total number of responses received, from all participants.
+In other words (and as one would expect), there are diminishing returns from asking more people to submit ideas.
+However, the actual rates differ per condition, as described below.
+\item The rates for generating novel ideas are linked to the number of responses requested: The more responses requested from individuals, the higher the rate of novel idea generation, in general.
+\item People are likely to come up with the same overall set of ideas early in their brainstorming process, but then diverge from this set of common ideas as they generate more ideas.
+Our data suggest that the first 20 ideas are the most unoriginal (i.e., the most common).
+This finding complements the previous finding, and further suggests that there is little value in requesting fewer than 20 ideas from workers.
+\item Individuals' patterns of idea generation mimic those identified by Nijstaf and Stroebe's SIAM (search for ideas in associative memory) model \cite{nijstad_how_2006}: Workers produce a set of ideas from working memory, then switch strategies to generate ideas that are permutations of other ideas, a process we dub {\em riffing\/}.
+In our data, we see the prevalence of riffing steadily increase over the first 20 responses, after which its prevalence plateaus.
+\item Requesting ``as many ideas as possible'' for a fixed price is inadvisable, as it leads to a relatively low number of ideas produced (in our experiments, fewer than 20)
+\end{itemize}
+
+Our results provide baseline models useful for two audiences: 1) those who wish to leverage crowds for brainstorming, and 2) researchers interested in developing techniques to optimize the crowd brainstorming processes.
+For the former, our models provide clear guidance and recommendations for various parameters of brainstorming in this environment (such as the number of questions to request from workers).
+For the latter group, our models provide a baseline for comparing novel techniques intended to increase the quality of brainstorming output by crowds.
+
+%We provide a method for clustering responses to idea generation problems and provide a detailed description of the resulting data set, which has been made publically available.
+We demonstrate the saturation of ideas received over time, and show that the rate of saturation, and size of the saturated idea set, originality of ideas and generality of ideas vary between conditions.
+
+%Furthermore, we examine predictions made by Nijstad and Stroebe's model of in-person brainstorming and test them in the context of crowd brainstorming.
+We find that the tested predictions hold: that ideas generated in sequence tend to belong to the same category, and that the time spent to generate when changing categories is signficantly greater than when generating ideas within the same category.
+
+%Finally, we briefly outline qualitative findings in the data, and recommend future work motivated by these findings.
+
+In the rest of this paper, we review prior work, then describe our experimental design, results from our experiments, and implications for practitioners and researchers.

### 2013-09-17 chi2014/introduction.tex
Minor edits to intro

@@
 In this paper, we focus on crowd-based brainstorming.
-While brainstorming has been an implicit component of many creative endeavors in microtask marketplaces, we are unaware of any research that specifically describes and models brainstorming in this environment.
+While brainstorming has been an implicit component of many creative endeavors in microtask marketplaces, we are unaware of any research that specifically describes and models the brainstorming process in this environment.
 Furthermore, while there is a rich literature describing brainstorming processes of individuals and groups (e.g., \cite{osborn_applied_1957, pinsonneault_electronic_1999, nijstad_how_2006} OTHERS?), microtask marketplaces are qualitatively different from previously studied environments, calling into question how much of this past research maps onto this environment.

### 2013-09-17 chi2014/introduction.tex
Minor edits to intro

@@
 In our data, we see the prevalence of riffing steadily increase over the first 20 responses, after which its prevalence plateaus.
-\item Requesting ``as many ideas as possible'' for a fixed price is inadvisable, as it leads to a relatively low number of ideas produced (in our experiments, fewer than 20)
+\item Requesting ``as many ideas as possible'' for a fixed price (fixed for the entire HIT, rather than per idea) is inadvisable, as it leads to a relatively low number of ideas produced (in our experiments, fewer than 20)
 \end{itemize}

### 2013-09-17 src/modelling/hyp6.r
Hypothesis 6 testing, with shifting break point between left and right

@@ -1,0 +1,120 @@
+require(rstan)
+
+kruschke_script_dir = '/home/fil/code/kruschke/samples'
+cur_dir = getwd()
+setwd(kruschke_script_dir)
+source('HDIofMCMC.R')
+setwd(cur_dir)
+
+model_string = "
+data {
+  int N;
+  real y[N];
+}
+
+parameters {
+  real<lower=0> mu;
+  real<lower=0> sigma;
+}
+
+model {
+  for (i in 1:N) {
+    y[i] ~ normal(mu, sigma);
+  }
+}
+";
+  
+add.error.bars <- function(X,ydown, yup,w,col=1){
+    arrows(X, ydown, X, yup, code=3,angle=90,length=w,col=col);
+}
+
+fit_model = function(df, condition, measure) {
+  sdf = subset(df, num_requested == condition)
+  
+  max_order = max(sdf$order)
+  print(max_order)
+
+  xs = seq(1, max_order)
+  lys = c()
+  rys = c()
+  l_bot = c()
+  r_bot = c()
+  l_top = c()
+  r_top = c()
+
+  pass = 1
+  last_pass = 0
+
+  for (i in 1:max_order) {
+    ldf = subset(sdf, order < i)
+    rdf = subset(sdf, order >= i)
+
+    l_data_list = list(
+      N = nrow(ldf),
+      y = ldf[,measure]
+    );
+
+    n_chains = 3;
+    num_saved_steps = 20000;
+    n_iter = ceiling(num_saved_steps / n_chains)
+    print(paste("Fitting for", measure, condition,i, "left"))
+    lfit = stan(model_code = model_string, data = l_data_list, iter = n_iter, chains = n_chains)
+
+    r_data_list = list(
+      N = nrow(rdf),
+      y = rdf[,measure]
+    );
+
+    print(paste("Fitting for", measure, condition,i, "right"))
+    rfit = stan(model_code = model_string, data = r_data_list, iter = n_iter, chains = n_chains)
+
+    llist = extract(lfit, permuted=TRUE)
+    rlist = extract(rfit, permuted=TRUE)
+
+    lmu_list = llist$mu
+    rmu_list = rlist$mu
+
+    left_hdi = HDIofMCMC(lmu_list)
+    right_hdi = HDIofMCMC(rmu_list)
+    
+    lys = c(lys, mean(lmu_list))
+    rys = c(rys, mean(rmu_list))
+
[…]

### 2013-09-17 chi2014/introduction.tex

@@
 In other words (and as one would expect), there are diminishing returns from asking more people to submit ideas.
-However, the actual rates differ per condition, as described below.
+However, the actual rates differ per condition, as described below
 \item The rates for generating novel ideas are linked to the number of responses requested: The more responses requested from individuals, the higher the rate of novel idea generation, in general.
+[TODO: There is likely past work that has a similar finding; we should relate this finding to it.
+Taylor Berry Block as related?
+Diehl as related?]
 \item People are likely to come up with the same overall set of ideas early in their brainstorming process, but then diverge from this set of common ideas as they generate more ideas.
 Our data suggest that the first 20 ideas are the most unoriginal (i.e., the most common).
-This finding complements the previous finding, and further suggests that there is little value in requesting fewer than 20 ideas from workers.
+This finding complements the previous finding, and further suggests that there is little value in requesting fewer than 20 ideas from workers
 \item Individuals' patterns of idea generation mimic those identified by Nijstaf and Stroebe's SIAM (search for ideas in associative memory) model \cite{nijstad_how_2006}: Workers produce a set of ideas from working memory, then switch strategies to generate ideas that are permutations of other ideas, a process we dub {\em riffing\/}.
-In our data, we see the prevalence of riffing steadily increase over the first 20 responses, after which its prevalence plateaus.
+In our data, we see the prevalence of riffing steadily increase over the first 20 responses, after which its prevalence plateaus
 \item Requesting ``as many ideas as possible'' for a fixed price (fixed for the entire HIT, rather than per idea) is inadvisable, as it leads to a relatively low number of ideas produced (in our experiments, fewer than 20)
+\item The time spent producing an idea matches the times reported in previous studies \cite{nijstad_how_2006}
 \end{itemize}
+
+%Overall, across a number of metrics and perspectives (e.g., where novel ideas show up in the brainstorming process, time spent producing ideas), crowd brainstorming shares a number of similarities with brainstorming in more traditional environments.
 

### 2013-09-17 chi2014/related.tex

--- 2013-09-15 chi2014/related.tex
@@
 \section{Related Work}
+In this section, we review related work in the areas of brainstorming and crowdsourcing.
 
 % TODO Note: I plaigarized the wording on these definitions from wikipedia at this point, should pull from original Osborne at some point
-Brainstorming was originally proposed in XXXX by Osborne [CITE].
-Brainstorming makes several recommendations for idea generation in groups, the core of which are a set of four rules: \emph{focus on quantity}, \emph{withhold criticism}, \emph{welcome unusual ideas}, and \emph{combine and improve ideas}.
+In 1957, Osborne formalized the brainstorming process, providing a set of recommendations for idea generation by groups \cite{osborn_applied_1957}: \emph{focus on quantity}, \emph{withhold criticism}, \emph{welcome unusual ideas}, and \emph{combine and improve ideas}.
+These recommendations form the basis for the instructions we provide to workers in our experiments.
 
-Since then, brainstorming as proposed by Osborne has been extensively evaluated.
-Very quickly, Taylor et.
-al established that \emph{nominal groups}, or individuals brainstorming in isolated whose ideas are later combined, performed better than real groups in terms of number of ideas generated [CITE].
-Furthermore, Diehl and Stroebe found extremely high correlation (XX) between the number of ideas generated and the quality of ideas, a finding that has been verified many times since (Another citation here would be nice).
-In Bouchard and Hare [CITE], nominal brainstorming groups were able to generate quantities of ideas linear in the number of members in the group.
+Following Osborne's work, Taylor et al.\ established that brainstorming by \emph{nominal groups}, or individuals working in isolation from one another, yields better results than collocated groups in terms of number of ideas generated \cite{taylor_does_1958}.
+These effects can be partially explained by the absence of counterproductive social forces, such as the fear of judgment.
+Bouchard and Hare \cite{bouchard_jr_size_1970}, later found that nominal brainstorming groups are able to generate quantities of ideas linear in the number of members in the group.
 
-Electronic brainstorming is a variant in which individuals brainstorm at separate computer terminals, with the exchange of ideas performed over the network.
-Participants can enter ideas while simultaneously viewing the ideas of all other ideas, anonymously presented on-screen.
-Gallupe [CITE] found that electronic brainstorming reduced the inhibiting social effects of \emph{production blocking} (the inability for multiple people to speak at once), and \emph{evaluation apprehension} (the fear of judgment) in comparison to in-person groups.
-However, Pinsonneault et al identify four productivity impediments introduced by group electronic brainstorming, and show that there is little evidence to support the proposition that electronic brainstorming out-performs nominal brainstorming [CITE].
+Electronic brainstorming is a variant in which individuals brainstorm at separate computer terminals, with the exchange of ideas performed over the network \cite{gallupe_electronic_1992}.
+Participants can enter ideas while simultaneously viewing the ideas of all other ideas, which are anonymously presented on-screen.
+As with nominal brainstorming, Gallupe \cite{gallupe_electronic_1992} observed that electronic brainstorming can reduce counterproductive social effects, such as production blocking (the inability for multiple people to speak at once) and fear of judgment, compared to physically collocated groups.
+However, Pinsonneault et al.\ \cite{pinsonneault_electronic_1999} identify productivity impediments introduced by group electronic brainstorming, such as being distracted by other ideas appearing, or individuals limiting their idea generation to avoid replicating the ideas of others.
+Given these drawbacks, their results suggest there is little evidence to support the proposition that electronic brainstorming outperforms nominal brainstorming.
 
-As crowd marketplaces do not tend to include built-in support for real-time communication between workers, nominal brainstorming groups provide the most reasonable assumption for providing a baseline assessment of performance.
-It is fortuitous, then, that prior work suggests that nominal groups are likely equivalent at worst to groups with the ability to communicate.
+Finally, Diehl and Stroebe found that groups that generate good ideas also generate \emph{many} ideas \cite{diehl_productivity_1987}, a finding that has been verified many times since [TODO: Need more backing refs here].
 
-This choice is supported by other users of crowd marketplaces for idea generation.
-Yu and Nickerson use many individuals in an evolutionary algorithm to generate creative chair designs [CITE].
-This technique is compared to single participants generating single ideas, but no baseline is established for idea generators individually exhausting a design space.
-
-Something about another example. Maybe the Little iterative vs parallel paper? It includes creative tasks.
-
-% TODO: part of this is Pao-exact, assuming it's cool because she's going to be an author on the paper
-While there are no models of nominal brainstorming in the context of a crowd marketplace, several researchers have modelled the idea generation process for individuals and groups.
-We emphasize the former for its applicability to nominal groups.
-Nijstad and Stroebe define the \emph{search for ideas in associative memory} (SIAM) model.
-SIAM assumes two stages of idea generation: knowledge activation based on search of working memory, and idea production within a single activated image.
-This model verifies three predictions for individual brainstorming behaviour:
+In terms of individual brainstorming processes, Nijstad and Stroebe provide a model for individual idea generation dubbed SIAM, or the search for ideas in associative memory.
+\cite{nijstad_how_2006}.
+SIAM defines two stages of idea generation: the generation of ideas based working memory, and idea production within a single activated ``image'' (where an image can be thought of as a category of ideas).
+This model predicts the following:
 
@@
 
-Should we have another model here? I haven't read any yet.
+Later, we will present evidence in support of this model from the data we collected.
+
+In the realm of crowd-based brainstorming, there are a number of ways this process could be realized.
+The most basic configuration mirrors that of nominal brainstorming, where individuals brainstorm ideas in isolation of one another, and the resulting ideas are passed on to the requestor (the individual posting the brainstorming task).
+More sophisticated brainstorming environments are possible, such as those that provide awareness of ideas that others have generated, or those that rely on the crowd to organize or rank the ideas produced by others.
+In this paper, we focus on the most basic setup, where individuals brainstorm in isolation of one another.
+
+While we are unaware of specific studies of brainstorming processes in crowd marketplaces, there is a range of related research examining more creative processes in this environment.
+For example, Yu and Nickerson created a system in which individuals participate in an evolutionary algorithm to generate creative chair designs \cite{yu_cooks_2011}.
+Compared to ideas produced by single individuals, they found the ideas produced as a result of their genetic algorithm were more creative.
+However, individuals produced only one idea each, compared to the many ideas typically produced by individuals in a brainstorming session.
+Thus, while their approach demonstrates the creative potential of the crowd, their results say little about brainstorming processes by crowds.
+
+TODO: More crowdsourcing related work here.
+Maybe the Little iterative vs parallel paper?
+It includes creative tasks.
+Perhaps describe some of the unique characteristics of microtask marketplaces, like people's desire to perform a task and move on...
 

### 2013-09-18 chi2014/method.tex
so close on results :D
replaces examples

@@
     \begin{enumerate}
-        \item "Use as a mirror"
-        \item "Mirror"
-        \item "Use it as a mirror for makeup"
-        \item "Tape it to a telephone pole and use as a reflector"
+        \item "Storage Container"
+        \item "Small Storage Box"
+        \item "Coin Storage"
+        \item "Travel Jewelry Case"
     \end{enumerate}
@@
 An instance is associated with exactly one \emph{idea}, and all instances in the same idea should be paraphrases of one another.
-In Figure FIG, "Use as a mirror" and "Mirror" belong to the same \emph{idea}.
-However, "Use it as a mirror for makeup" does not belong to the same idea, as it contains additional information.
+In Figure FIG, "Storage Container" and "Small Storage Box" belong to the same \emph{idea}.
+However, "Coin Storage" does not belong to the same idea, as it contains additional information.
 
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{sample_category_tree}
+    \includegraphics[width=0.9\columnwidth]{example_instances}
 \end{figure}
@@
 
-\begin{figure*}[h!]
+\begin{figure*}[ht]
 \small
@@
 Two nodes a and b have high similarity if they solve the problem in the same way, and low similarity if their solutions have no common themes.
-For example, in Figure FIG, "Mirror" and "Tape it to a telephone pole and use as a reflector" have high similarity, as they both solve the problem of using a broken MP3 player by using it to reflect light.
+For example, in Figure FIG, "Small Storage Box" and "Travel Jewelry Case" have high similarity, as they both solve the problem of using a broken MP3 player by using it to reflect light.
 
@@
 Coverage is not commutative.
-To return to the example, "Mirror" has high coverage of "Tape it to a telephone pole and use as a reflector", as the latter is a specific use for a mirror.
+To return to the example, "Small Storage Box" has high coverage of "Travel Jewelry Case", as the latter is a just a specific use for the former.
 On the other hand, the latter idea has low coverage of the former.
@@
 Occasionally, two nodes have high similarity but no coverage.
-For example, "Use it as a mirror for makeup" and "Tape it to a telephone pole and use as a reflector" use the MP3 player in the same way, but neither is a generalization of the other.
+For example, "Coin Storage" and "Travel Jewelry Case" use the MP3 player in the same way, but neither is a generalization of the other.
 In this case, the algorithm appeals to the human coder to provide a lable for a new idea that is a generalization of both.

### 2013-09-18 chi2014/goals.tex
Quick pass through the goals section

--- 2013-09-16 chi2014/goals.tex
@@
 
-Our primary motivation is to develop baseline for brainstorming performance in crowd marketplaces.
+To characterize brainstorming in microtask marketplaces, we are interested in modeling the following quantities:
+\begin{itemize}
+\item the rate of unique idea production
+\item the rate of production of unique \emph{categories} of ideas
+\item the relative originality of ideas
+\item individual practices in idea generation
+\end{itemize}
+
+We will examine these rates as a function of the number of ideas requested from workers.
+
+%Our primary motivation is to develop baseline for brainstorming performance in crowd marketplaces.
 Brainstorming performance is often measured in terms of creativity.
@@
 
-As our primary goal, we examine these outcomes with respect to the number of answer requested from workers.
+%As our primary goal, we examine these outcomes with respect to the number of answer requested from workers.
 This is the simplest of conditions that can vary when choosing a design for a crowd brainstorming problem.
-Furthermore, we detail the phenomenon of \emph{riffing} (known in the idea generation literature as \emph{clustering}, a term we use for its classification meaning later in the paper).
+Furthermore, we detail the phenomenon of \emph{riffing} (known in the idea generation literature as \emph{clustering}, a term we must unfortunately reserve for it's classification-related meaning).
 
-In addition to the high-level goal of modelling quantity, originality and surprise factor in brainstorming results, we began with several explicit hypotheses.
-\textbf{Hypothesis 1:} the rate of new ideas will reduce towards 0 exponentially as a function of the number of responses gathered.
-Less formally, by continually soliciting solutions, we will eventually reach a saturation point at which the majority of new solutions are derivative of previously seen solutions.
+In modeling these quantities, we have several hypotheses related to results received in the aggregate (across all participants), as well as individuals' brainstorming processes.
 
-\textbf{Hypothesis 2:} there is a set general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.
+\textbf{Hypothesis 1:} The rate of new ideas will diminish exponentially toward zero as a function of the number of responses gathered.
+Less formally, by continually soliciting solutions, we will eventually reach a saturation point at which any idea a worker submits is likely to be the same or practically identical to previously seen solutions.
+Related to this, we expect the number of \emph{categories} of ideas to diminish exponentially as well.
 
-In addition, trends identified by prior work in spatial and temporal co-located brainstorming may or may not be exist in crowd brainstorming.
-We selected the trends most testable with our data set to replicate their findings.
-Nijstad and Stroebe's SIAM model \cite{nijstad_how_2006} makes three non-subjective predictions, two of which retain their applicability in this setting:
+\textbf{Hypothesis 2:} The rate of new ideas will tend toward zero faster in some conditions than others.
+In particular, we expect conditions requesting a smaller number of ideas will tend toward zero faster, for reasons hypothesized next.
 
-\begin{itemize}
-\item \textbf{Hypothesis 3:} An idea from one semantic category should more often be followed by an idea from the same category than expected by random chance
-\item \textbf{Hypothesis 4:} idea generation when changing semantic categories takes longer than when repetiting categories
-\end{itemize}
+\textbf{Hypothesis 3:} We anticipate there is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.
+In other words, we expect there to be a relatively fixed set of ideas that ``everyone'' thinks of first.
 
-Parnes et al \cite{parnes_effects_1961} showed that more high quality ideas were generated in the latter half of brainstorming run than the former half.
-In keeping of our philosophy regarding creativity and quality measurements, we restrict the content of this test to originality.
-\textbf{Hypothesis 5:} Ideas generated in the latter half of a brainstorming session are of higher originality than ideas generated in the former half.
+Parnes et al \cite{parnes_effects_1961} showed that more high quality ideas are generated in the latter half of a brainstorming run than the former half.
+Because of the difficulty in reliably assessing the quality of an idea for a brainstorming exercise, we instead consider the o-score of an idea to indirectly measure this concept.
+
+\textbf{Hypothesis 4:} Ideas generated in the second half of a brainstorming session have higher o-scores than ideas generated in the first half.
+
+With respect to individual brainstorming processes, we have several hypotheses derived from Nijstad and Stroebe's work and their SIAM model \cite{nijstad_how_2006}:
+
+\textbf{Hypothesis 5:} An idea from one semantic category of ideas should more often be followed by an idea from the same category than expected by random chance.
+
+\textbf{Hypothesis 6:} The time it takes to generate an idea should be longer when changing semantic categories than when generating ideas within a category.

### 2013-09-18 chi2014/results.tex
hypothesis 5 results

@@
 
-We must also consider time spent in light of the hypotheses made by Nijstad and Stroede \cite{nijstad_how_2006}.
-The second, \textbf{Hypothesis 5}, states that idea generation when changing semantic categories should take longer than idea generation within categories.
+The Nijstad and Stroede \cite{nijstad_how_2006} model also provides \textbf{Hypothesis 5}: \emph{idea generation when changing semantic categories should take longer than idea generation within categories}.
+An examination of the distribution of data for time spent per instance suggested a log-normal distribution, with the frequency of occurances quickly rising until a peak was reached, followed by a gradual drop-off.
+As with previous tests, we used the Stan package for Bayesian inference [CITE].
+We fit, two models: $t_w = \text{lognormal}(\mu_w, \sigma_w)$ for within-category idea generation and $t_b = \text{lognormal}(\mu_b, \sigma_b)$ for between-category idea generation, where the $t$ parameter in each model represents the time spent generating an instance.
 
-Test here...
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{hyp5_comparison}
+    \caption{Time spent to generate an instance}
+\end{figure}
+
+The resulting fit models are shown in Figure FIG, overlapping the log-space histograms for observed values.
+The mean for the within-category condition was 9.5 in log space (13.4 seconds, 95\% HDI 9.4-9.6), while the mean for the between-categories condition was 10.2 (27.0 seconds, 95\% HDI 10.2-10.3).
+The variances were 1.1 (3 seconds) and 1.2 (3.3 seconds) respectively.
+These difference were both sigificant.
+Within-category instance generation took on average 12.6 seconds less than between-categories.
+This is consistent with the findings of Nijstad and Stoebe, who reported differences of 6-12 seconds.

### 2013-09-18 chi2014/results.tex
hypothesis 4 stats

@@
 Riffing activity is analogous to idea production within an image as described by Nijstad and Stroebe's SIAM model \cite{nijstad_how_2006}.
-Thus, we should expect to confirm their hypothesis, \textbf{Hypothesis 4}, that an idea from one semantic category should more often by followed by an idea from the same category than expected by random chance.
+Thus, we should expect to confirm their hypothesis, \textbf{Hypothesis 4}, that \emph{an idea from one semantic category should more often by followed by an idea from the same category than expected by random chance}.
 
-Test here...
+To test this, we modelled the probability of any category being followed by the same category with a Bernoulli distribution.
+Each instance not the first in its run was a trial.
+If the instance was in the same category tree as the previous instance, the trial was a pass.
+Otherwise, it was a failure.
+The Bernoulli $\theta$ was 0.156 (95\% HDI 0.143 - 0.170).
+
+The most common category tree in our data set, \emph{music player}, represents only 5.65\% of instances.
+This is well below the 0.143 lower-bound of HDI.
+Thus, we accept the hypothesis that an instance follows a previous instance in the same category with greater probability than is explained by random chance.
 

### 2013-09-18 chi2014/results.tex
hypothesis 3 test

@@
 Once the common idea are exhausted, then, there is no relationship between idea originality and category originality.
+
+\subsubsection{Hypothesis 3}
+
+This interpretation supports \textbf{Hypothesis 3:} There is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.
+We test this interpretation by constructing a set of the top 66 (5\%) most common ideas.
+If our intuition is correct, these ideas should be more likely to appear in the first 5 instances of a run than they are to appear in any of the following instances.
+We designate two Bernoulli variables.
+The first, defined by parameter $\theta_f$, represents the probability that an instance is in the common set given it is in the first five instances of a run.
+The second, defined by parameter $\theta_l$, is the probability that an instance is in the common pool given it is one of the \emph{last} (>5) instances in a run..
+
+We fit these parameters using data from all runs with a total number of instances greater than 10.
+We chose this so that no run could contribute evidence to $\theta_f$ without also contributing to $\theta_l$.
+The resulting parameters were $\theta_f = 0.564$ (95\% HDI 0.517-0.611) and $\theta_l = 0.323$ (95\% HDI 0.303-0.342).
+This difference is significant, supporing the hypothesis that common ideas are over-represented in the early part of a brainstorming run, regardless of conditon.
 

### 2013-09-18 chi2014/abstract.tex
abstract, introduction pass

@@
-In this paper we describe the formatting requirements for
-SIGCHI Conference Proceedings, and this sample file
-offers recommendations on writing for the worldwide
-SIGCHI readership. Please review this document even if
-you have submitted to SIGCHI conferences before, some
-format details have changed relative to previous years.
+Microtask marketplaces enable hybrid computational systems that use crowds to solve problems of creativity, such as brainstorming.
+We present a first characterization of brainstorming in this context, providing models for the rate of idea generation, the rate of category generation, the originality of ideas and the novelty of ideas.
+We demonstrate that the rates of idea and category generation can be closely fit by a logarithmic model, that these rates roughly increase if fewer participants are asked for more responses, and that this distinction is the result of a \emph{burn in period} of common ideas at the beginning of a brainstorming task.
+Furthermore, we explore the phenomenon of \emph{riffing} in brainstorming tasks, and replicate two related findings from prior work.
+Our results provide recommendations for those leveraging the crowd for brainstorming, and comparable models of creative idea generation research community.

### 2013-09-18 chi2014/introduction.tex
abstract, introduction pass

@@
 
-Microtask marketplaces such as Amazon's Mechanical Turk have proven to be useful platforms for outsourcing a wide range of tasks, from proofreading documents \cite{soylent} to estimating the caloric content of pictures of food \cite{platemate}.
-These platforms make possible human computation systems \cite{quinn_2011}, hybrid computational systems that tap humans' abilities to solve problems that currently cannot be solved by machines alone.
+Microtask marketplaces such as Amazon's Mechanical Turk have proven to be useful platforms for outsourcing a wide range of tasks, from proofreading documents \cite{soylent} to estimating the caloric content of food \cite{platemate}.
+These platforms make possible human computation systems \cite{quinn_2011}, hybrid systems that tap humans' abilities to solve problems that currently cannot be solved by machines alone.
 
@@
 While brainstorming has been an implicit component of many creative endeavors in microtask marketplaces, we are unaware of any research that specifically describes and models the brainstorming process in this environment.
-Furthermore, while there is a rich literature describing brainstorming processes of individuals and groups (e.g., \cite{osborn_applied_1957, pinsonneault_electronic_1999, nijstad_how_2006} OTHERS?), microtask marketplaces are qualitatively different from previously studied environments, calling into question how much of this past research maps onto this environment.
-For example, brainstorming participants in a microtask marketplace are paid for their participation, and are often eager to complete a task as fast as possible to move on to the next paying task [REF?].
-Given these and other differences, there are many open questions about how brainstorming unfolds in this context.
+Furthermore, while there is a rich literature describing brainstorming processes of individuals and groups (e.g., \cite{osborn_applied_1957, pinsonneault_electronic_1999, nijstad_how_2006} OTHERS?), microtask marketplaces have qualitative differences.
+For example, brainstorming participants in a microtask marketplace are paid for their participation, and are often eager to complete a task as fast as possible to maximize earnings [REF?].
+Furthermore, participants in microtask marketplaces cannot be expected to be spatially or temporally co-located.
+There many open questions about how brainstorming unfolds in this context.
 
@@
 
-This paper presents results from a series of experiments intended to model brainstorming on microtask marketplaces (specifically, Amazon's Mechanical Turk).
+This paper presents the results of a series of experiments intended to model brainstorming on microtask marketplaces (specifically, Amazon's Mechanical Turk).
 In our experiments, we asked individuals to brainstorm ways to use an MP3 player whose batteries were dead.
@@
 \begin{itemize}
-\item The rates at which novel ideas and categories of ideas are produced follow logarithmic growth models, as a function of the total number of responses received, from all participants.
-In other words (and as one would expect), there are diminishing returns from asking more people to submit ideas.
-However, the actual rates differ per condition, as described below
-\item The rates for generating novel ideas are linked to the number of responses requested: The more responses requested from individuals, the higher the rate of novel idea generation, in general.
+\item The rates at which novel ideas and categories of ideas are produced follow logarithmic growth models, as a function of the total number of responses receieved across participants.
+Thus, the rate of novel ideas decreases exponentially.
+\item The rates for generating novel ideas changes with the number of responses requested per individual, even if the overall number of responses is held constant.
+In general, a higher rate of novel idea generation is achieved by asking fewer participants for more responses each.
 [TODO: There is likely past work that has a similar finding; we should relate this finding to it.
@@
 \item People are likely to come up with the same overall set of ideas early in their brainstorming process, but then diverge from this set of common ideas as they generate more ideas.
-Our data suggest that the first 20 ideas are the most unoriginal (i.e., the most common).
+The first 20 ideas are the least original (i.e., the most common).
 This finding complements the previous finding, and further suggests that there is little value in requesting fewer than 20 ideas from workers
@@
 In our data, we see the prevalence of riffing steadily increase over the first 20 responses, after which its prevalence plateaus
-\item Requesting ``as many ideas as possible'' for a fixed price (fixed for the entire HIT, rather than per idea) is inadvisable, as it leads to a relatively low number of ideas produced (in our experiments, fewer than 20)
+\item Requesting ``as many ideas as possible'' for a fixed price (fixed for the entire HIT, rather than per idea) is inadvisable, as it leads to a relatively low number of ideas produced (in our experiments, fewer than 15)
 \item The time spent producing an idea matches the times reported in previous studies \cite{nijstad_how_2006}

### 2013-09-18 chi2014/method.tex
Pass through the method section

@@
 \section{Method}
+In this section, we describe our process of recruiting and remunerating participants, the brainstorming question selected, and the user interface provided the worker.
+We then describe the data collected, and our method of organizing it into a hierarchical cluster.
 
-In order to develop models and test hypotheses, it was necessary to establish a significant corpus of brainstorming data data.
-Over the course of four months we repeatedly collected responses to brainstorming tasks.
+\subsection{Recruitment and Remuneration}
+To study brainstorming in a crowd environment, we published brainstorming HITs (human intelligence tasks) on Amazon's Mechanical Turk.
+We restricted our worker pool to residents of the United States to increase the chances of English language comprehension and a shared cultural background.
+No other screening was performed on workers.
 
-Participants were recruited from Amazon's \emph{Mechanical Turk} \cite{_amazon_????}, and were restricted to residents of the United States for a baseline expectation of english language comprehension and cultural familiarity.
-Mechanical Turk is an online marketplace in which members receive financial reward for completing \emph{Human Intelligence Tasks}, or HITs.
+HITs were placed on the marketplace asking for 5, 10, 20, 50, 75, 100, or an unlimited number of ideas for a brainstorming problem.
+We paid a fixed amount per HIT, computed at a rate of approximately 3.5 cents per idea requested, yielding reward amounts of 18 cents (5), 35 cents (10), 70 cents (20), \$1.75 (50), \$2.65 (75), and \$3.50 (100).
+For the unlimited condition, we offered \$1.75, pricing it the same as the 50 ideas condition.
+HITs were advertised with the number of responses requested.
+Workers had 18 hours to submit the HIT once accepted.
 
-HITs were placed on the marketplace asking for a variable \emph{number of ideas}, with proportional rewards.
-These conditions were : 5 ideas (with a corresponding reward of \$0.18 USD), 10 ideas (\$0.35), 20 (\$0.70), 50 (\$1.75), 75 (\$2.65), and 100 (\$3.50).
+\subsection{Brainstorming Question}
+In conducting this research, we tested several brainstorming questions, including classic brainstorming questions (e.g., ``provide alternative uses for a mop''), a request to brainstorm factors that influence weight gain and weight loss, and more goal-oriented questions, such as one requesting people to brainstorm ways to raise funds for a charity.
+Ultimately, we found the goal-oriented brainstorming questions to be the most compelling since they reflect what we believe would be real-world uses of brainstorming on Mechanical Turk.
 
-This served as the independent variable in all data gathering.
+The final brainstorming question we chose for this work is the following:
+
+TODO: Indent, italicize this:
+Many people have old MP3 players or MP3 players that they no longer use.
+Please brainstorm N uses for old MP3 players/MP3 players.
+Assume that the devices' batteries no longer work, though they can be powered via external power sources.
+Also be aware that devices may \emph{not} have displays.
+Be as specific as possible in your descriptions.
+
+%This served as the independent variable in all data gathering.
 It must be stated that under normal brainstorming conditions, participants are asked to generate as many ideas as possible within a fixed time frame, instead of a fixed number.
@@
 
-While the brainstorming question was manipulated in pilot studies, it was fixed for the data reported on in this paper:
+We chose to focus on this problem because a) we expected workers to be familiar with problem area (old MP3 players), b) it represents a real-world problem for which innovative solutions are welcome (namely, what to do with old electronics that are still partially functional), and c) we anticipated there to be a wide range of creative solutions possible.
+All data presented in this paper refer only to responses received to this particular question.
 
-\begin{itemize}
-\item \textbf{MP3}
+Initially, workers were randomly paired with a question from a pool of questions.
+However, we eventually dropped all other questions, so workers would only receive the MP3 player question.
+When testing multiple questions, we attempted to prevent workers from answering the same question twice (for example, if they chose the HIT requesting 10 responses, then the HIT requesting 20 responses) .
+However, we were not always successful in preventing this from happening.
+Thus, we only use a worker's first set of responses if they answered the question multiple times.
 
-Many people have old MP3 players or MP3 players that they no longer use.
-Please brainstorm N uses for old MP3 players/MP3 players.
-Assume that the devices' batteries no longer work, though they can be powered via external power sources.
-Also be aware that devices may \emph{not} have displays.
-Be as specific as possible in your descriptions.
-\end{itemize}
+%Because workers could choose which \emph{number of ideas} condition to participate in, self-selection bias is a reasonable concern.
+However, this bias is also present in real-world HIT choice behaviour.
 
-This problem was selected for three qualities:
+\subsection{Worker's Brainstorming Interface}
 
-\begin{enumerate}
-\item It requires creativity to resolve. Obvious answers do not satisfy the problem.
-\item Mechanical Turk participants have the expertise to solve the problem.
-\item It has an associated success metric.
-\end{enumerate}
+Upon accepting a HIT, participants were asked to give consent for participating in the study, and informed that they could leave the study at any point without financial consequences.
+Under more realistic, real-world conditions, workers would be expected to complete all responses to receive compensation.
+As we will report later, the incidence of workers abusing this policy was relatively low.
 
-Workers choose which \emph{number of ideas} condition to participate in, thus self-selection bias is a reasonable concern. However, this bias is also present in real-world HIT choice behaviour.
+The brainstorming task was presented in a standard web-based form that included an introduction, the specific brainstorming problem, a series of fields for entering the individual responses, and a larger text area where workers could enter additional ideas (TODO: Figure X).
 
-Upon accepting a HIT, participants were asked to give consent and informed that they could leave the study at any point without financial consequences.
-Under normal conditions, the participants would be expected to complete all responses to receive compensation.
-
-Should we be dropping these runs, or at least testing they don't have an affect?
-
-\subsection{Task}
-
-The brainstorming task is a form in a standard web browser.
-Participants are presented with a brief overview of the tenets of brainstorming, are presented a problem, and must enter some number of ideas to resolve the problem within an 18 hour period.
-
-At the beginning of the form is a brief introduction to brainstorming, with a paraphrase of Osborne's four rules of brainstorming \cite{osborn_applied_1957}.
-These rules were manipulated to make sense within the constraints nominal brainstorming over the web medium.
-The rules, as displayed to the participants:
+The introduction Osborne's four rules of brainstorming \cite{osborn_applied_1957}, adapted to this context:
 
@@
 
-The brainstorming problem is below this, followed by a series of text entry inputs numbered through to the total number of ideas requested.
-Figure X (FIG) is an example of a typical task.
-We place a larger free text area at the bottom of the list where participants could enter any additional ideas.
+In addition to the visible input fields, HITs were instrumented to collect timing data.
+Specifically, we recorded the time of the first activation and last de-activation of each form element, the time the HIT was accepted, and the time it was submitted.
 
-\subsubsection{Data Collected}
+\subsection{Data Collected}
 
-In total, 161 HITs were complteed by 146 distinct participants, for a total of 3007 responses.
-A breakdown of responses across conditions is given in table TAB.
-In cases when a single participant participated in multiple conditions, all HITs after the first were dropped for subsequent analysis.
+Initially, we sought 10 HITs for each condition.
+However, we found that collecting only 10 HITs for the lower response conditions (5, 10, 20 responses) limited our ability to make meaningful comparisons between responses received from the larger response conditions (50, 75, 100).
+Accordingly, we sought to receive a minimum of 400 ideas for each condition.
+We found the 5 and 10 conditions to be particularly unpopular, and were not able to fully reach the desired number of ideas in these conditions.
+
+In total, 154 workers submitted 170 HITs, for a total of 3292 responses.
+As we received data in the unlimited condition, it became obvious that this was a poor strategy to employ -- the number of responses received was less than 15 in all cases.
+Accordingly, we remove that data from our analysis.
+
+If a worker completed multiple HITs (i.e., generated responses for multiple conditions), there is the risk their responses and idea generation processes evolved over time, making it difficult to compare their responses in the second HIT with responses from others who are performing the same HIT without any previous exposure to the problem.
+Accordingly, we retain only the first HIT submitted by the same worker.
+
+After removing these data, we are left with 146 distinct participants for a total of 146 HITs, yielding 3007 responses. A breakdown of responses across conditions is given in table TODO: TAB.
 
@@
 
-HITs were augmented with JavaScript to collect data in addition to brainstorming responses.
-For each response, we collected the time of the first activation and last de-activation of the corresponding form element.
-We collected the time the HIT was accepted and the time it was submitted.
-To uniquely identify participants, we computed a hash of their Mechanical Turk worker ID.
-A hash was used to protect privacy after data release.
+%To uniquely identify participants, we computed a hash of their Mechanical Turk worker ID. A hash was used to protect privacy after data release.
 
-\subsection{Terminology}
+\subsection{Hierarchical Clustering}
 
-To facilitate discussion, we introduce terms here that we will use throughout the rest of the paper.
+To analyze the data, we organized it into sets hierarchical clusters, where the clusters encode similarity relationships between ideas.
+Specifically, parent nodes are more generalized versions of child nodes.
+This clustering enables a wide range of data analysis.
+We introduce specific terminology, then describe the clusters and the process used to create them.
 
+\subsubsection{Clustering: Terminology}
 An \emph{instance} is a single solution by a single participant to a brainstorming question.
-Alternatively, an instance may be thought of as the text entered into a single field in the task form.
-For example, figure~\ref{fig:sample_instances} gives four \emph{instances} from the data set.
+Most concretely, it is the text entered into a single field in the web form.
+For example, figure~\ref{fig:sample_instances} provides an example of four \emph{instances} from workers.
 
@@
 
-A \emph{brainstorming run}, or \emph{run}, is the temporally-ordered series of instances that were received in a single HIT by a single user.
-As stated above, after dropping repeat data, all brainstorming runs were completed by distinct participants.
+A \emph{brainstorming run}, or \emph{run}, is the temporally-ordered series of instances that were received in a single HIT by a single worker.
 
 An \emph{idea} is a unique solution to the brainstorming problem.
-An instance is associated with exactly one \emph{idea}, and all instances in the same idea should be paraphrases of one another.
-In Figure FIG, "Storage Container" and "Small Storage Box" belong to the same \emph{idea}.
-However, "Coin Storage" does not belong to the same idea, as it contains additional information.
+An instance is associated with exactly one \emph{idea}, though multiple instances can be associated with the same idea (i.e., when multiple workers come up with the same idea).
+Multiple instances of the same idea do not need to be identically phrased to refer to the same idea.
+For example, in Figure FIG, we consider ``Storage Container'' and ``Small Storage Box'' as two instances of the same \emph{idea}.
+However, we do not consider ``Coin Storage'' to be the same idea, since it provides a more specific use case than the previous two.
 
-Despite this, it is clear the ideas have some commonality.
-It is desireable to encode this relationship.
-To that end, we propose a hierarchical clustering of ideas in \emph{category trees}.
+Despite the differences in specificity between ``Coin Storage'' and the more generic containers, it is useful to encode the fact that these ideas are related.
+We use hierarchical clustering to encode that relationship.
+We call a hierarchical cluster of ideas a \emph{category trees}.
+
 In a category tree, ideas act as nodes.
-Each idea may have 0 or 1 parent idea (only the root has 0) and 0 or more child ideas.
+Each idea may optionally have a single parent idea, and zero or more child ideas.
+If an idea \emph{A} is a parent of idea \emph{B}, A is a generalization of B.
 An example category tree for the instances in Figure FIG is given in Figure FIG.
@@
 
-If an idea \emph{A} is a parent of idea \emph{B}, A is a generalization of B: all instances in A are are suitably general to encompass all instances in B.
-This hierarchical clustering allows us to classify ideas as simultaneously different in solution stragy but different in solution details.
+Category trees can either be \emph{singletons} or \emph{non-singletons}.
+A singleton category tree is composed of only one \emph{idea node}, the root.
+A non-singleton category tree has \emph{at least two idea nodes}.
+Non-singleton category trees correspond to the ``images'' described in Nisjstad and Stroebe's SIAM model \cite{nijstad_how_2006}.
 
-We further distinguish category trees into \emph{singleton} and \emph{non-singleton} category trees.
-A singleton category tree is composed of only one idea node, the root.
-A non-singleton category tree has at least two idea nodes.
-We introduce non-singleton category trees as correspondents to the images in Nisjstad and Stroebe's SIAM model \cite{nijstad_how_2006}.
-On SIAM, the idea generation process is represented by a series of images activations within which idea production occurs.
-Non-singleton trees by definition represent strategies that have been produced in at least two distinct ways.
+Finally, the \emph{idea forest} is the collection of all category trees.
 
-The \emph{idea forest} is the collection of all category trees. Different trees in the forest have no particular relationship.
+\subsubsection{Clustering: Process}
+To produce the idea forest, the primary author added instances to the idea forest one at a time using the following algorithm.
 
-\subsection{Clustering}
+If an instance described an idea already represented as a node in the forest, the instance was added to that node. Otherwise, a new node was produced.
 
-In addition to the notion of an idea forest, we also propose a clustering algorithm driven by human intelligence to create the data structure.
-The algorithm is described in Figure~\ref{fig:cluseringalg}.
-Instances are added one at a time to the existing idea forest.
-In brief summary, the root idea that most closely matches the strategy in the instance is selected, and then that root idea and the new instance are compared in generality.
+If a node was a generalization of a second node, the second node was introduced as a child of the first.
+In some cases, two ideas were clearly related, though no generalization of the two existed in the data set.
+In these cases, we introduced a node that is a generalization of the two.
+For example, if we received only ``Coin Storage'' and ``Travel Jewelry Case'' as ideas, we would create a node labeled ``Storage Case'' to capture the fact that the two are related concepts.
+These ``artificial'' nodes were marked to distinguish them from true generalizations found in the data.
+
+Clustering was performed using a custom-built tool that assisted in finding instances that were similar to a given instance.
+Briefly, the tool represented each instance using a bag of words model which included synonyms for words.
+The words were stemmed, then individual instances were ranking according to a cosine similarity metric.
+
+The primary author clustered the 3007 response instances into ideas and category trees.
+A researcher not involved in the project randomly sampled 40 category trees to assess whether ideas within the tree should be there, and whether parent-child relationships correctly encoded generalizations of concepts.
+Results indicate that 8\% of the data would be X by the independent researcher, and 0.7\% of the data would be Y by the independent researcher.
+
+%In brief summary, the root idea that most closely matches the strategy in the instance is selected, and then that root idea and the new instance are compared in generality.
 This process may be repeated depending on this generality relationship, until instance is either placed in an existing idea or a new idea node is created.
 
-\begin{figure*}[ht]
-\small
-\begin{verbatim}
-for each instance:
-  idea_node = new node including instance
-  current_node = root of forest
-  do:
-    best_match = max_similarity(idea_node, current_node.children)
+% \begin{figure*}[ht]
+% \small
+% \begin{verbatim}
+% for each instance:
+%   idea_node = new node including instance
+%   current_node = root of forest
+%   do:
+%     best_match = max_similarity(idea_node, current_node.children)
 
-    if best_match.similarity is low or current_node has no children:
-      insert idea_node under current_node
-      exit do
-    else:
-      if coverage(idea_node, best_match) == coverage(best_match, idea_node) == high:
-        merge idea_node, best_match
-        exit do
-      else if coverage(idea_node, best_match) == coverage(best_match, idea_node) == low:
-        new_parent = new artifical idea node
-        insert best_match, idea_node under new_parent
-        insert new_parent under current_node
-        exit do
-      else if coverage(idea_node, best_match) > coverage(best_match, idea_node):
-        replace best_match with idea_node in tree
-        current_node = idea_node
-        idea_node = best_match
-      else:
-        current_node = best_match
-\end{verbatim}
-\caption{Manual clustering algorithm}
-\label{fig:cluseringalg}
-\end{figure*}
-
-The clustering algorithm relies on the human intelligence of the researcher for three key decisions:
-
-\subsubsection{Similarity}
-Two nodes a and b have high similarity if they solve the problem in the same way, and low similarity if their solutions have no common themes.
-For example, in Figure FIG, "Small Storage Box" and "Travel Jewelry Case" have high similarity, as they both solve the problem of using a broken MP3 player by using it to reflect light.
-
-\subsubsection{Coverage}
-A node a has high coverage of a node b if the problem solution in a is equal to or an abstraction of the problem solution in b.
-Coverage is not commutative.
-To return to the example, "Small Storage Box" has high coverage of "Travel Jewelry Case", as the latter is a just a specific use for the former.
-On the other hand, the latter idea has low coverage of the former.
-
-\subsubsection{Artificial Ideas}
-Occasionally, two nodes have high similarity but no coverage.
-For example, "Coin Storage" and "Travel Jewelry Case" use the MP3 player in the same way, but neither is a generalization of the other.
-In this case, the algorithm appeals to the human coder to provide a lable for a new idea that is a generalization of both.
-
-The primary author followed the algorithm to cluster the 3007 response instances into ideas and trees.
+%     if best_match.similarity is low or current_node has no children:
+%       insert idea_node under current_node
+%       exit do
+%     else:
+%       if coverage(idea_node, best_match) == coverage(best_match, idea_node) == high:
+%         merge idea_node, best_match
+%         exit do
+%       else if coverage(idea_node, best_match) == coverage(best_match, idea_node) == low:
+%         new_parent = new artifical idea node
+%         insert best_match, idea_node under new_parent
+%         insert new_parent under current_node
+%         exit do
+%       else if coverage(idea_node, best_match) > coverage(best_match, idea_node):
+%         replace best_match with idea_node in tree
+%         current_node = idea_node
+%         idea_node = best_match
+%       else:
+%         current_node = best_match
+% \end{verbatim}
+% \caption{Manual clustering algorithm}
+% \label{fig:cluseringalg}
+% \end{figure*}

### 2013-09-18 chi2014/results.tex
Unfinished pass through results.tex

@@
 
-We can use the uniqueness measure to understand the rate at which new ideas arrive.
-As we gather more and more ideas in the course of an experiment, we expect that we will see fewer ideas that are not already in our idea pool.
-Figure FIG shows the cumulative count of ideas as a function of the number of instances gathered in the experiment.
-To smooth the plot, the order in which instances were received was shuffled 100 times, and the mean of the cumulative unique idea count was taken at each step.
+\subsection{Idea Forest}
+
+Hierarchical clustering yielded an idea forest composed of 3007 instances, 1213 unique ideas, 335 category trees, and 129 non-singleton category trees.
+In general, category trees were shallow, with a maximum tree height of 5.
+The number of nodes within a category tree ranged from 1 to X, with the distribution of values roughly following a Poisson distribution.
+The number of instances per node (denoting equivalent response instances) also roughly follows a Poisson distribution, with number of instances ranging from X to Y.
+
+\begin{table}
+	\begin{tabular}[h!]{r | l l l l l l l}
+	\textbf{number condition} & 5 & 10 & 20 & 50 & 75 & 100 & all \\ \hline \hline
+	HITs & 57 & 47 & 23 & 10 & 10 & 10 & 146\\
+	instances & 293 & 471 & 453 & 500 & 634 & 855 & 3007\\
+	ideas & 171 & 249 & 278 & 341 & 443 & 177 &1212\\
+	category trees & 72 & 79 & 93 & 114 & 172 & 177 &321\\
+	non-singleton trees & 28 & 34 & 40 & 48 & 49 & 61 &92\\
+	\end{tabular}
+\end{table}
+
+% are also roughly poisson distributed (Table TAB).
+In keeping with the findings for height, trees found in the upper number conditions are smaller than trees found in the lower conditions.
+The precipitous drop in median tree size suggests that in the largest condition, participants are finding rare ideas with only a few variants and less than 14 instances.
+
+% \begin{table}
+% \begin{tabular}[h!]{r | l l l l l l l}
+% 	\textbf{number condition} & 5 & 10 & 20 & 50 & 75 & 100 & all \\ \hline \hline
+% 	number of nodes& \\ \hline
+%     median &5&5&4&3&2&2&2 \\
+% 	first quartile &2&2&1&1&1&1&1 \\
+% 	third quartile &17&17&14&10&5&5&5 \\
+% 	number of instances& \\ \hline
+% 	median &14&14&12&10&4&4&4 \\
+%     first quartile &4&4&4&2&1&1&1 \\
+% 	third quartile &47&45&30&23&13&13&13 \\
+% 	\end{tabular}
+% 	\caption{Number of ideas and instances in trees}
+% \end{table}
+
+% \begin{figure}[h!]
+%     \centering
+%     \includegraphics[width=0.9\columnwidth]{idea_forest}
+%     \caption{Idea Forest}
+% \end{figure}
+
+% The height of category trees follows a roughly Poisson distribution, with the maximum tree height observed at 5.
+Summary statistics across condition are in table TAB.
+The third quartile drops as the number condition increases.
+While the upper conditions still discover categories trees with heights as great as 5, they generate many more "short" trees.
+This suggests that the ideas generated in the upper conditions are less common.
+
+% \begin{table}
+% 	\begin{tabular}[h!]{r | l l l l l l l}
+% 	\textbf{number condition} & 5 & 10 & 20 & 50 & 75 & 100 & all \\ \hline \hline
+% 	median & 2 & 2 & 2 & 2 & 2 & 2 & 2\\
+% 	first quartile & 2 & 2  & 1 & 1 & 1 & 1 & 1\\
+% 	third quartile & 3 & 3 &3 &3 &2 &2 & 2\\
+% 	\end{tabular}
+% 	\caption{Category tree height}
+% \end{table}
+
+\subsection{Rates of Idea and Category Production}
+The idea forest allows us to model the rate at which new ideas and new categories of ideas are produced, testing hypotheses 1 and 2 (i.e., rates of new ideas and categories of ideas will diminish over time, and will do so differentially per condition).
+
+Figure FIG shows the cumulative count of new ideas as a function of the number of instances gathered in the experiment. For reference, a perfect ``1:1'' idea generation line is also plotted.
+
+We model this quantity using the following logarithmic model:
+
+TODO: num_of_new_ideas = b0 + b_cond * log(x_scale * x + x_trans)
+
+In our model, we assume b0, x_scale, and x_trans do not vary between conditions, while b_cond (essentially, a scaling factor) varies on a per-condition basis.
+We fit the parameters to the data using a Bayesian approach [REF Kruschke], using the STAN modeling software to perform the actual fit [REF STAN].
+The resulting posterior distributions for the parameters allow us to determine whether the conditions are significantly different from one another by comparing whether the posterior distributions' highest density intervals (HDIs) overlap or not.
+In our analyses, we use an HDI of 95\%.
+
+Analysis of the b_cond HDIs reveals that there are no significant differences between the 5 and 10 conditions, with all other conditions significantly different from one another except for the 50 and 100 conditions.
+
+Taking the derivative of the models allows us to examine the same data, but as rates at which new ideas are produced.
+As expected, rates of new ideas quickly taper off as a function of the number of ideas received.
 
@@
 \end{figure}
-
-The horizontal line in Figure FIG represent the idea case, in which each instance represents a different idea (i.e.
-every answer we receive is a new, unique idea).
-As expected, the rate at which ideas are gained appears to taper off.
-
-Figure FIG more closely examines the rate, the primary property of interest.
-The panels, from top to bottom, are the rate at which new ideas, categories, and non-singleton categories are generated, respectively.
-Similar to Figure FIG, these are based on shuffling the order of instances 10000 times and taking the mean of the derivative at each point.
-In this case, the ideal (a 1:1:1 instance:idea:category ratio) is represented by the horizontal bar across the top of the chart.
 
@@
 
-The number of ideas, categories and non-singleton each drop off exponentially over time.
-It is unsurprising that the number of new categories should drop off faster than the number of new ideas, as we expect that there are fewer categories of ideas than there are individual ideas themselves.
+We can perform a similar analysis for the number of new idea categories that arrive over time.
+As would be surmised from the previous graphs, we observe a similar pattern, which can also be modeled using a logarithmic model.
+As would also be expected, the number of new idea categories produced increases more slowly than the number of new ideas.
+
+We see a very similar story with respect to differences in rates between conditions.
+The 5 and 10 conditions are virtually identical, while all other conditions are significantly different from one another.
+
+Collectively, these results largely affirm hypotheses 1 and 2: The rate of new idea and new idea categories production quickly tapers off and tends toward zero over time.
+Furthermore, with a few exceptions, rates differ between conditions.
 
@@
 
-\subsubsection{Hypothesis 1}
+The final measure, \emph{look-back equality}, is non-symmetrical and has meaning only in the context of a single brainstorming run.
+An instance \emph{a} is look-back equal to an \emph{earlier} instance \emph{b} if \emph{b} is in either \emph{a}'s idea node or any of \emph{a}'s parents, siblings, or children.
+This definition specifically identifies when a later instance in a run can be considered to have been influenced by earlier instance.
 
-Visually, the rate of idea, category and singleton category generation seems to trend towards 0. We tested this hypothesis...
+\subsubsection{Originality}
 
-Additionally Figure FIG indicates that there is some effect of number of ideas requested on the rate of generation.
-Generally, conditions with more requested responses generated ideas and categories at a higher rate.
+As our measure of originality, we use \emph{o-score}, introduced by Jansson and Smith \cite{jansson_design_1991}.
+An idea's o-score is $1 - p(idea)$, where $p(idea) = (number of instances of that idea)/(number of instances total)$.
+The o-score for category trees is similarly calculated.
+We will occasionally refer to the \emph{category o-score} of an idea.
+This simply refers to the o-score of the category tree to which that idea belongs.
 
-Test here...
-
-\subsubsection{Hypothesis 2}
-
-Test here...
+\subsubsection{Uniqueness}
 

### 2013-09-18 chi2014/method.tex

@@
 
-The brainstorming task was presented in a standard web-based form that included an introduction, the specific brainstorming problem, a series of fields for entering the individual responses, and a larger text area where workers could enter additional ideas (TODO: Figure X).
+The brainstorming task was presented in a standard web-based form that included an introduction, the specific brainstorming problem, a series of fields for entering the individual responses, and a larger text area where workers could enter additional ideas (Figure~\ref{fig:sample_task}).
 
-The introduction Osborne's four rules of brainstorming \cite{osborn_applied_1957}, adapted to this context:
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{sample_task}
+    \caption{Sample task on Mechanical Turk}
+    \label{fig:sample_task}
+\end{figure}
+
+The introduction listed Osborne's four rules of brainstorming \cite{osborn_applied_1957}, adapted to this context:
 
@@
 If a worker completed multiple HITs (i.e., generated responses for multiple conditions), there is the risk their responses and idea generation processes evolved over time, making it difficult to compare their responses in the second HIT with responses from others who are performing the same HIT without any previous exposure to the problem.
-Accordingly, we retain only the first HIT submitted by the same worker.
+Accordingly, we retain only the first HIT submitted by each worker.
 
-After removing these data, we are left with 146 distinct participants for a total of 146 HITs, yielding 3007 responses. A breakdown of responses across conditions is given in table TODO: TAB.
+After removing these data, we are left with 146 distinct participants for a total of 146 HITs, yielding 3007 responses.
+A breakdown of responses across conditions is given in Table~\ref{tab:data_collected}.
 
-\begin{tabular}[h!]{r | l l l l l l }
-& 5 & 10 & 20 & 50 & 75 & 100 \\ \hline \hline
-HITs completed & 59 & 49 & 23 & 10 & 10 & 10 \\
-HITs completed (no repeats) & 57 & 39 & 21 & 9 & 10 & 10 \\
-solutions (no repeats) & 283 & 372 & 413 & 450 & 635 & 855 \\
+\begin{table}[h!]
+\begin{tabular}{r | l l l l l l l }
+    & 5 & 10 & 20 & 50 & 75 & 100 & all \\ \hline
+    participants & 57 & 39 & 21 & 9 & 10 & 10 & 146 \\
+    solutions & 283 & 372 & 413 & 450 & 635 & 855 & 3007
 \end{tabular}
+\caption{Data after removing unlimited condition and repeats}
+\label{tab:data_collected}
+\end{table}
 
@@
 Most concretely, it is the text entered into a single field in the web form.
-For example, figure~\ref{fig:sample_instances} provides an example of four \emph{instances} from workers.
+For example, four \emph{instances} from the data set:
 
-\begin{figure}[!h]
 \begin{enumerate}
@@
 \end{enumerate}
-    \label{sample_instances}
-\end{figure}
 
@@
 If an idea \emph{A} is a parent of idea \emph{B}, A is a generalization of B.
-An example category tree for the instances in Figure FIG is given in Figure FIG.
+An example category tree for the instances above is given in Figure~\ref{fig:example_instances}.
 
@@
     \includegraphics[width=0.9\columnwidth]{example_instances}
+    \caption{Example category tree}
+    \label{fig:example_instances}
 \end{figure}

### 2013-09-18 chi2014/results.tex
Working through the results section.

@@
 We model this quantity using the following logarithmic model:
-
-TODO: num_of_new_ideas = b0 + b_cond * log(x_scale * x + x_trans)
-
-In our model, we assume b0, x_scale, and x_trans do not vary between conditions, while b_cond (essentially, a scaling factor) varies on a per-condition basis.
+\[n_{ideas} = b_0 + b_{cond} * \log (x_{scale} * x + x_{offset})\]
+where $n_{ideas}$ represents the number of new ideas, $b_0$, $x_{scale}$, and $x_{offset}$ are parameters shared between conditions that serve to position the curve, and $b_{cond}$ acts as a per-condition scaling factor.
 We fit the parameters to the data using a Bayesian approach [REF Kruschke], using the STAN modeling software to perform the actual fit [REF STAN].
-The resulting posterior distributions for the parameters allow us to determine whether the conditions are significantly different from one another by comparing whether the posterior distributions' highest density intervals (HDIs) overlap or not.
+In modeling the parameters, we use uninformative priors.
+The resulting posterior distributions allow us to determine whether the conditions are significantly different from one another by determining whether the highest density intervals (HDIs) of $b_{cond}$'s posterior distributions overlap or not.
 In our analyses, we use an HDI of 95\%.
 
-Analysis of the b_cond HDIs reveals that there are no significant differences between the 5 and 10 conditions, with all other conditions significantly different from one another except for the 50 and 100 conditions.
+Analysis of the HDIs of $b_{cond}$ reveals that there are no significant differences between the 5 and 10 conditions, with all other conditions significantly different from one another except for the 50 and 100 conditions.
+[TODO: Table of fitted values and their HDIs.]
 
@@
 The 5 and 10 conditions are virtually identical, while all other conditions are significantly different from one another.
+[TODO: Table of values and HDIs]
 
 Collectively, these results largely affirm hypotheses 1 and 2: The rate of new idea and new idea categories production quickly tapers off and tends toward zero over time.
-Furthermore, with a few exceptions, rates differ between conditions.
+Furthermore, with a few exceptions, rates differ between conditions -- new idea and idea category production is generally linked to the number of ideas requested.
 
-The third panel, which shows non-singleton categories over time, tells a distinctly different story from the others.
+Given the observed differences in production of new ideas and idea categories, it is worthwhile to better understand \emph{why} these differences exist.
+
+%The third panel, which shows non-singleton categories over time, tells a distinctly different story from the others.
 While in the idea and category plots, the height-ordering of rates of generation between categories remains generally identical, there is a point of intersection and reversal in the non-singleton category plot at around the 30 instances point.
@@
 
-In lower conditions, brainstormers will quickly offer up big, popular, non-singleton categories.
-Beyond the inflection point, the category pool is saturated with these low-hanging fruit, and only brainstormers tasked with generating more ideas will find the smaller category.
-This is a more nuanced view of the tree node/instance quartiles in Section SEC; condition 5 brainstormers cover the spectrum of category sizes while condition 100 brainstormers pull from the smaller category trees in the forest.
-
-The final measure, \emph{look-back equality}, is non-symmetrical and has meaning only in the context of a single brainstorming run.
-An instance \emph{a} is look-back equal to an \emph{earlier} instance \emph{b} if \emph{b} is in either \emph{a}'s idea node or any of \emph{a}'s parents, siblings, or children.
-This definition specifically identifies when a later instance in a run can be considered to have been influenced by earlier instance.
-
-\subsubsection{Originality}
+Early in the brainstorming process, individuals typically produce the same sets of ideas from the same sets of idea categories.
+It is only after the produce about 20 ideas that they start producing less generic ideas.
+This phenomenon can be observed by examining the o-scores of ideas and idea categories, as well as the size of the category trees.
+[TODO] [Mike stopped here for the night]
 
@@
 
-\subsubsection{Uniqueness}
-
-\subsection{Originality}
-
 We measure differences in originality between conditions by examining the idea o-score and category o-score.
@@
 The left panel compares idea o-score and the right compares category o-score.
+
+%.
+Beyond the inflection point, the category pool is saturated with these low-hanging fruit, and only brainstormers tasked with generating more ideas will find the smaller category.
+This is a more nuanced view of the tree node/instance quartiles in Section SEC; condition 5 brainstormers cover the spectrum of category sizes while condition 100 brainstormers pull from the smaller category trees in the forest.
+
+% The final measure, \emph{look-back equality}, is non-symmetrical and has meaning only in the context of a single brainstorming run.
+An instance \emph{a} is look-back equal to an \emph{earlier} instance \emph{b} if \emph{b} is in either \emph{a}'s idea node or any of \emph{a}'s parents, siblings, or children.
+This definition specifically identifies when a later instance in a run can be considered to have been influenced by earlier instance.
 

### 2013-09-18 chi2014/introduction.tex
Putting some references for introduction

@@
 Furthermore, while there is a rich literature describing brainstorming processes of individuals and groups (e.g., \cite{osborn_applied_1957, pinsonneault_electronic_1999, nijstad_how_2006} OTHERS?), microtask marketplaces have qualitative differences.
-For example, brainstorming participants in a microtask marketplace are paid for their participation, and are often eager to complete a task as fast as possible to maximize earnings [REF?].
+For example, crowd workers exhibit high variance of effort in completing a task \cite{soylent}.
+Brainstorming participants in a microtask marketplace are paid for their participation and some might be more eager to complete a task as fast as possible to maximize earnings \cite{kittur2008crowdsourcing}.
 Furthermore, participants in microtask marketplaces cannot be expected to be spatially or temporally co-located.
@@
 In general, a higher rate of novel idea generation is achieved by asking fewer participants for more responses each.
-[TODO: There is likely past work that has a similar finding; we should relate this finding to it.
+This result corresponds to previous finding that making participants to extend more effort generating ideas--in this case by requiring them to generate a lot of ideas--will lead to an increasing proportion of good ideas\cite{parnes_effects_1961}.[TODO: There is likely past work that has a similar finding; we should relate this finding to it.
 Taylor Berry Block as related?
-Diehl as related?]
+Diehl as related?][Pao's note: can't find more directly related stuff mainly because most of the previous lit focus on time limit task rather than number of ideas.
+This is actually a good argument for the paper since time limit might not work in this marketplace.]
 \item People are likely to come up with the same overall set of ideas early in their brainstorming process, but then diverge from this set of common ideas as they generate more ideas.

### 2013-09-18 chi2014/related.tex
Putting some references for introduction

@@
 % TODO Note: I plaigarized the wording on these definitions from wikipedia at this point, should pull from original Osborne at some point
-In 1957, Osborne formalized the brainstorming process, providing a set of recommendations for idea generation by groups \cite{osborn_applied_1957}: \emph{focus on quantity}, \emph{withhold criticism}, \emph{welcome unusual ideas}, and \emph{combine and improve ideas}.
+In the 1950s, Osborne formalized the brainstorming process, providing a set of recommendations for idea generation by groups \cite{osborn_applied_1957}: \emph{focus on quantity}, \emph{withhold criticism}, \emph{welcome unusual ideas}, and \emph{combine and improve ideas}.
 These recommendations form the basis for the instructions we provide to workers in our experiments.

### 2013-09-18 chi2014/related.tex

@@
 More sophisticated brainstorming environments are possible, such as those that provide awareness of ideas that others have generated, or those that rely on the crowd to organize or rank the ideas produced by others.
+Examples for these environments are platforms for generating social innovation ideas or product concepts like OpenIDEO, Quirky and Threadless that employ structured process where participants take on different responsibility from generating ideas to evaluating others' ideas.
 In this paper, we focus on the most basic setup, where individuals brainstorm in isolation of one another.
 
-While we are unaware of specific studies of brainstorming processes in crowd marketplaces, there is a range of related research examining more creative processes in this environment.
+While we are unaware of specific studies of brainstorming processes in crowd marketplaces, there is a range of related research examining more creative processes in this environment \cite{lewis2011affective, kittur2011crowdforge, Zhang:2012:HCT:2207676.2207708}.
 For example, Yu and Nickerson created a system in which individuals participate in an evolutionary algorithm to generate creative chair designs \cite{yu_cooks_2011}.
@@
 
-TODO: More crowdsourcing related work here.
+Little et el.
+compared the iterative and parallel human computation processes on a brainstorming task for names for fabricated companies \cite{little2010exploring}.
+In the iterative condition, participants saw all names suggested by previous participants.
+Participants in the parallel condition worked in isolation without seeing other participants' ideas.
+Subjective ratings of the quality of ideas reveal that while the ideas from the iterative condition were rated higher on average, the best ideas from the parallel condition have higher scores than the best ideas from the iterative conditions.
+The parallel process is more likely to generate names with high ratings, advocating the case for nominal brainstorming over traditional electronic brainstorming.
+However, the study did not mention categorization of generated ideas, an important measure on flexibility of idea generation \cite{lewis2011affective, nijstad_how_2006, finke1992creative, shah2003metrics}.
+Each participant also generated only five ideas, leaving the number of ideas requested per participant an open dimension to be explored.
+
+%The nature of microtask marketplace also re.. .. identified high variance of turk eff
+
+%In this paper, we focus on crowd-based brainstorming.
+While brainstorming has been an implicit component of many creative endeavors in microtask marketplaces, we are unaware of any research that specifically describes and models the brainstorming process in this environment.
+Furthermore, while there is a rich literature describing brainstorming processes of individuals and groups (e.g., \cite{osborn_applied_1957, pinsonneault_electronic_1999, nijstad_how_2006} OTHERS?), microtask marketplaces have qualitative differences.
+For example, crowd workers exhibit high variance of effort in completing a task \cite{soylent}.
+Brainstorming participants in a microtask marketplace are paid for their participation and some might be more eager to complete a task as fast as possible to maximize earnings \cite{kittur2008crowdsourcing}.
+Furthermore, participants in microtask marketplaces cannot be expected to be spatially or temporally co-located.
+There many open questions about how brainstorming unfolds in this context.
+
+%TODO: More crowdsourcing related work here.
 Maybe the Little iterative vs parallel paper?

### 2013-09-18 chi2014/related.tex
Related work: nature of crowd market

@@
 
-%The nature of microtask marketplace also re.. .. identified high variance of turk eff
+The nature of microtask marketplace also calls for an alternative structure for brainstorming tasks in this setting.
+Researchers have found that 30\% or more submission on Mechanical Turk may be low quality \cite{kittur2008crowdsourcing}.
+For example, submissions to a brainstorming task for a company name are sometimes grammartically awkward or offensive \cite{little2010exploring}.
+These low quality responses were not found in walk-in brainstorming studies.
+Bernstein et al.
+identified high variance of effort Amazon Mechanical Turk workers put on task \cite{soylent}.
+They characterized two worker personas at opposite ends of the effort spectrum: the {\em Eager Beaver\/} who goes beyond task requirement and the {\em Lazy Turker\/} who does as little work as necessary to get paid.
+A brainstorming task that asks participants to generate as many ideas as they can under a set time limit might not be practical.
+A {\em Lazy Turker\/} can simply fill in a few ideas and wait for the time to run out instead of exerting full effort.
+This requirement does not map well with most previous work on brainstorming where the task are time-limited.
 

### 2013-09-18 chi2014/method.tex

@@
 However, we eventually dropped all other questions, so workers would only receive the MP3 player question.
-When testing multiple questions, we attempted to prevent workers from answering the same question twice (for example, if they chose the HIT requesting 10 responses, then the HIT requesting 20 responses) .
+%When testing multiple questions, we attempted to prevent workers from answering the same question twice (for example, if they chose the HIT requesting 10 responses, then the HIT requesting 20 responses) .
 However, we were not always successful in preventing this from happening.
@@
 Accordingly, we sought to receive a minimum of 400 ideas for each condition.
-We found the 5 and 10 conditions to be particularly unpopular, and were not able to fully reach the desired number of ideas in these conditions.
+We found the 5 and 10 conditions to be particularly unpopular, and were not able to reach this minimum.
 
@@
 This clustering enables a wide range of data analysis.
-We introduce specific terminology, then describe the clusters and the process used to create them.
+We introduce specific terminology, motivate the technique, then describe the clusters and the process used to create them.
 
@@
 
+\subsubsection{Clustering: Motivation}
+
+Initially, a coding scheme was employed in which a master list of unique ideas was formed, and each response was assigned a categorical code from this list.
+It proved difficult to determine and adhere to a constant granularity of distinction.
+For example, it is desireable to encode some equality between "use [MP3 players] to build a chair, and "use [MP3 players] to build a table".
+Information loss results from either assigning both the same code or each different codes.
+An interval specificity scale is also difficult to employ; intuitively, we cannot compare the distance between "audiobooks" and "audio books for the blind" with the distance between the latter and "[audio] instructions for the blind on a variety of things".
+
+A category tree allows the encoding of generalization relationships such that useful metrics can be extracted without any assumptions about distance.
+For example, The number of descendants of an idea gives a measure of how broadly that node was explored. 
+The prevalance of similar responses can be approximated at multiple levels; the number of instances under a tree represents the popularity of a category, while the same for an individual node provides a weighted score of ideas that are exactly equivalance.
+An idea can be considered more closely related to its adjacent ideas than it is to all other ideas in the tree.
+Using this common structure to operationalize originality and uniqueness ensures consistency in analysis.
+Furthermore, the category forest relies primarily on generalization decisions for construction, which are more clearly defined than equivalence.
+
 \subsubsection{Clustering: Process}
@@
 A researcher not involved in the project randomly sampled 40 category trees to assess whether ideas within the tree should be there, and whether parent-child relationships correctly encoded generalizations of concepts.
-Results indicate that 8\% of the data would be X by the independent researcher, and 0.7\% of the data would be Y by the independent researcher.
+10\% (HDI 6-14) of instances were marked as specific than their associated idea.
+0.7\% (HDI 1.3-2.8\%) of instances were marked as not generalized by their parents.
+No significant effect was found between conditions for probability of either error.
 

### 2013-09-18 chi2014/method.tex
Added a very brief section on data analysis

@@
 
+\subsection{Data Analysis}
+Throughout this paper, we will fit models to data and give confidence metrics for the parameters of those models.
+Our analysis techniques are based on Kruschke's \emph{Doing Bayesian Data Analysis: A Tutorial in R and Bugs} \cite{kruschke_doing_2010}.
+We express our confidence in parameters in terms of the 95\% HDI (highest density interval), or the interval which contains 95\% of the mass of our posterior belief distribution.
+
+To fit the models, we use the Bayesian inference software Stan \cite{stan-software:2013} via the package RStan.
+
 %In brief summary, the root idea that most closely matches the strategy in the instance is selected, and then that root idea and the new instance are compared in generality.

### 2013-09-18 chi2014/introduction.tex
Add more references

@@
 While brainstorming has been an implicit component of many creative endeavors in microtask marketplaces, we are unaware of any research that specifically describes and models the brainstorming process in this environment.
-Furthermore, while there is a rich literature describing brainstorming processes of individuals and groups (e.g., \cite{osborn_applied_1957, pinsonneault_electronic_1999, nijstad_how_2006} OTHERS?), microtask marketplaces have qualitative differences.
+Furthermore, while there is a rich literature describing brainstorming processes of individuals and groups (e.g., \cite{osborn_applied_1957, pinsonneault_electronic_1999, nijstad_how_2006,brown1998modeling}), microtask marketplaces have qualitative differences.
 For example, crowd workers exhibit high variance of effort in completing a task \cite{soylent}.
@@
 In general, a higher rate of novel idea generation is achieved by asking fewer participants for more responses each.
-This result corresponds to previous finding that making participants to extend more effort generating ideas--in this case by requiring them to generate a lot of ideas--will lead to an increasing proportion of good ideas\cite{parnes_effects_1961}.[TODO: There is likely past work that has a similar finding; we should relate this finding to it.
+This result corresponds to previous finding that making participants to extend more effort generating ideas--in this case by requiring them to generate a lot of ideas--will lead to an increasing proportion of good ideas\cite{parnes_effects_1961}.%[TODO: There is likely past work that has a similar finding; we should relate this finding to it.
 Taylor Berry Block as related?

### 2013-09-18 chi2014/related.tex
Related work: nature of crowd market
Add more references

@@
 
-Finally, Diehl and Stroebe found that groups that generate good ideas also generate \emph{many} ideas \cite{diehl_productivity_1987}, a finding that has been verified many times since [TODO: Need more backing refs here].
+Finally, Diehl and Stroebe found that groups that generate good ideas also generate \emph{many} ideas \cite{diehl_productivity_1987}, a finding that has been verified many times \cite{briggs1997quality, parnes1959effects, parnes_effects_1961, shah2003metrics, cross1996creativity}.
 
@@
 A {\em Lazy Turker\/} can simply fill in a few ideas and wait for the time to run out instead of exerting full effort.
-This requirement does not map well with most previous work on brainstorming where the task are time-limited.
-
-%In this paper, we focus on crowd-based brainstorming.
-While brainstorming has been an implicit component of many creative endeavors in microtask marketplaces, we are unaware of any research that specifically describes and models the brainstorming process in this environment.
-Furthermore, while there is a rich literature describing brainstorming processes of individuals and groups (e.g., \cite{osborn_applied_1957, pinsonneault_electronic_1999, nijstad_how_2006} OTHERS?), microtask marketplaces have qualitative differences.
-For example, crowd workers exhibit high variance of effort in completing a task \cite{soylent}.
-Brainstorming participants in a microtask marketplace are paid for their participation and some might be more eager to complete a task as fast as possible to maximize earnings \cite{kittur2008crowdsourcing}.
-Furthermore, participants in microtask marketplaces cannot be expected to be spatially or temporally co-located.
-There many open questions about how brainstorming unfolds in this context.
+The motivational structure of microstask marketplace does not map well with most previous work on brainstorming where the sessions are time-limited.
 

### 2013-09-18 chi2014/goals.tex
Goals: brainstorming metrices

@@
 \section{Goals and Hypotheses}
+
+Researchers have used different metrices to assess the brainstorming process and its outputs.
+Typically, the measures of concern are the novelty of generated ideas, the practicality/appropriateness of ideas, the variety of ideas (flexibility), and quantity of ideas generated.
+\cite{finke1992creative, shah2003metrics}.
+The most common approach to measure novelty and appropriateness is to ask domain experts to subjectively make scale-rated assessment of each \cite{lewis2011affective, amabile_1983}.
+Depending on the types of artifacts and researchers' interest, these two measures are sometimes combined into one measure of quality \cite{little2010exploring}.
+Instead of subjective ratings, some researchers opt for a measure that can be derived objectively.
+For example, Jansson et al.
+used `o' scores, derived from the frequency an idea occur, to evaluate products novelty \cite{jansson_design_1991}.
+Flexibility is measured by the number of explored ``categories'' in a design space or by subjective rating from domain experts.
+\cite{lewis2011affective, marsh1996examples}.
+Category or feature-based quantities also yield other metrics such as within-catgories fluency and  the length of train-of-thought when participants focus on a particular category \cite{nijstad_how_2006}.
+
+% Researchers have long debated how to best assess creativity in a way that consistently captures the complex dimensions of creativity [35].
+The most commonly adopted approach [1, 5, 17] and the one we employ is to ask domain experts to subjectively make scale-rated assessments of creativity.
+Typically, domain experts rate originality, feasibility, elaboration, and flexibility and count the number of ideas generated [3].
 

### 2013-09-18 chi2014/related.tex

@@
 
-Following Osborne's work, Taylor et al.\ established that brainstorming by \emph{nominal groups}, or individuals working in isolation from one another, yields better results than collocated groups in terms of number of ideas generated \cite{taylor_does_1958}.
-These effects can be partially explained by the absence of counterproductive social forces, such as the fear of judgment.
+Following Osborne's work, Taylor et al.\ established that brainstorming by \emph{nominal groups}, or individuals working in isolation from one another whose results are later combined, yields better results than collocated groups in terms of number of ideas generated \cite{taylor_does_1958}.
+These effects can be partially explained by the absence of counterproductive social forces, such as production blocking (the inability for multiple people to speak at once) and the fear of judgment.
 Bouchard and Hare \cite{bouchard_jr_size_1970}, later found that nominal brainstorming groups are able to generate quantities of ideas linear in the number of members in the group.
@@
 Participants can enter ideas while simultaneously viewing the ideas of all other ideas, which are anonymously presented on-screen.
-As with nominal brainstorming, Gallupe \cite{gallupe_electronic_1992} observed that electronic brainstorming can reduce counterproductive social effects, such as production blocking (the inability for multiple people to speak at once) and fear of judgment, compared to physically collocated groups.
-However, Pinsonneault et al.\ \cite{pinsonneault_electronic_1999} identify productivity impediments introduced by group electronic brainstorming, such as being distracted by other ideas appearing, or individuals limiting their idea generation to avoid replicating the ideas of others.
+As with nominal brainstorming, Gallupe \cite{gallupe_electronic_1992} observed that electronic brainstorming can reduce counterproductive social effects.
+However, Pinsonneault et al.\ \cite{pinsonneault_electronic_1999} identify productivity impediments introduced by group electronic brainstorming, such as being distracted by other ideas appearing, or individuals witholding ideas they feel are derivative of others.
 Given these drawbacks, their results suggest there is little evidence to support the proposition that electronic brainstorming outperforms nominal brainstorming.
@@
 SIAM defines two stages of idea generation: the generation of ideas based working memory, and idea production within a single activated ``image'' (where an image can be thought of as a category of ideas).
-This model predicts the following:
+This model makes predictions applicable to all brainstorming contexts, namely that an idea is more likely to be followed by an idea in the same category than would be expected by chance, and new ideas in the same category as the previous idea are generated faster than ideas between categories.
+Nijstad and Stroebe verify these in the context of nominal brainstorming.
 
-\begin{enumerate}
-\item an idea is more likely to be followed by an idea in the same category than would be expected by chance
-\item new ideas in the same category as the previous idea are generated faster than ideas between categories
-\item increased clustering (generation of ideas within-category) is positively correlated with overall high production
-\end{enumerate}
+The simplist imaginable instantiation of brainstorming in a microtask marketplace mirrors that of nominal brainstorming.
+More sophisticated brainstorming environments are possible, such as those that provide awareness of ideas that others have generated (electronic brainstorming), or those that rely on the crowd to organize or rank the ideas produced by others, as in evolutionary techniques \cite{yu_cooks_2011}.
+%Examples for these environments are platforms for generating social innovation ideas or product concepts like OpenIDEO, Quirky and Threadless that employ structured process where participants take on different responsibility from generating ideas to evaluating others' ideas.
+In this paper, we examine nominal brainstorming to establish a baseline.
 
-Later, we will present evidence in support of this model from the data we collected.
-
-In the realm of crowd-based brainstorming, there are a number of ways this process could be realized.
-The most basic configuration mirrors that of nominal brainstorming, where individuals brainstorm ideas in isolation of one another, and the resulting ideas are passed on to the requestor (the individual posting the brainstorming task).
-More sophisticated brainstorming environments are possible, such as those that provide awareness of ideas that others have generated, or those that rely on the crowd to organize or rank the ideas produced by others.
-Examples for these environments are platforms for generating social innovation ideas or product concepts like OpenIDEO, Quirky and Threadless that employ structured process where participants take on different responsibility from generating ideas to evaluating others' ideas.
-In this paper, we focus on the most basic setup, where individuals brainstorm in isolation of one another.
-
-While we are unaware of specific studies of brainstorming processes in crowd marketplaces, there is a range of related research examining more creative processes in this environment \cite{lewis2011affective, kittur2011crowdforge, Zhang:2012:HCT:2207676.2207708}.
-For example, Yu and Nickerson created a system in which individuals participate in an evolutionary algorithm to generate creative chair designs \cite{yu_cooks_2011}.
-Compared to ideas produced by single individuals, they found the ideas produced as a result of their genetic algorithm were more creative.
-However, individuals produced only one idea each, compared to the many ideas typically produced by individuals in a brainstorming session.
-Thus, while their approach demonstrates the creative potential of the crowd, their results say little about brainstorming processes by crowds.
+While we are unaware of specific studies of brainstorming in microtask marketplaces, there is a range of related research examining other creative processes \cite{lewis2011affective, kittur2011crowdforge, Zhang:2012:HCT:2207676.2207708}.
+For example, Yu and Nickerson created an evolutionary crowd algorithm to generate creative chair designs \cite{yu_cooks_2011}.
+Their work and others demonstrates the creative potential of the crowd, but does not specifically address brainstorming or analogous processes in which participants generate many ideas.
 

### 2013-09-18 chi2014/results.tex

@@
 The number of nodes within a category tree ranged from 1 to X, with the distribution of values roughly following a Poisson distribution.
-The number of instances per node (denoting equivalent response instances) also roughly follows a Poisson distribution, with number of instances ranging from X to Y.
+The number of instances per node (i.e., the number of equivalent idea instances) also roughly follows a Poisson distribution, with number of instances ranging from X to Y.
 
@@
 \subsection{Rates of Idea and Category Production}
-The idea forest allows us to model the rate at which new ideas and new categories of ideas are produced, testing hypotheses 1 and 2 (i.e., rates of new ideas and categories of ideas will diminish over time, and will do so differentially per condition).
+The idea forest allows us to model the rate at which new ideas and new categories of ideas are produced, testing hypotheses 1 and 2 (i.e., the notion that rates of new ideas and categories of ideas will diminish over time, and will do so as a function condition).
 
@@
 where $n_{ideas}$ represents the number of new ideas, $b_0$, $x_{scale}$, and $x_{offset}$ are parameters shared between conditions that serve to position the curve, and $b_{cond}$ acts as a per-condition scaling factor.
+This model does not explicitly model variability due to different workers, a point we return to later.
+
 We fit the parameters to the data using a Bayesian approach [REF Kruschke], using the STAN modeling software to perform the actual fit [REF STAN].
@@
 
-Analysis of the HDIs of $b_{cond}$ reveals that there are no significant differences between the 5 and 10 conditions, with all other conditions significantly different from one another except for the 50 and 100 conditions.
-[TODO: Table of fitted values and their HDIs.]
+Analysis of the HDIs of $b_{cond}$ reveals no significant differences between the 5 and 10 conditions, with all other conditions significantly different from one another except for the 50 and 100 conditions.
+[TODO: Table of fitted values and their HDIs.] Notable in our data is the 75 condition, which rises above the 100 condition in this plot.
+Further analysis reveals that a single worker is responsible for that additional rise in number of new ideas.
+Removing that worker from the data reveals that the 75 condition is not different from either the 50 or 100 conditions.
 
@@
 As would be surmised from the previous graphs, we observe a similar pattern, which can also be modeled using a logarithmic model.
-As would also be expected, the number of new idea categories produced increases more slowly than the number of new ideas.
+In the fitted model, we see that the number of new idea categories produced increases more slowly than the number of new ideas.
+This finding suggests that although individuals may produce new, unique ideas over time, these new ideas are within the \emph{same} categories.
+In other words, they are ``riffing'' -- coming up with permutations on existing ideas.
 
-We see a very similar story with respect to differences in rates between conditions.
-The 5 and 10 conditions are virtually identical, while all other conditions are significantly different from one another.
-[TODO: Table of values and HDIs]
+Collectively, these results largely affirm hypotheses 1 and 2: The rates of new idea and new idea categories production quickly taper off and tend toward zero over time.
+Furthermore, we see differences in rates between conditions.
+In our data, we see three clear groups of rates: the 5 and 10 conditions, the 20 condition, and the 50, 75, and 100 condition.
 
-Collectively, these results largely affirm hypotheses 1 and 2: The rate of new idea and new idea categories production quickly tapers off and tends toward zero over time.
-Furthermore, with a few exceptions, rates differ between conditions -- new idea and idea category production is generally linked to the number of ideas requested.
-
-Given the observed differences in production of new ideas and idea categories, it is worthwhile to better understand \emph{why} these differences exist.
+%Given the observed differences in production of new ideas and idea categories, it is worthwhile to better understand \emph{why} these differences exist.
 
@@
 
-Early in the brainstorming process, individuals typically produce the same sets of ideas from the same sets of idea categories.
-It is only after the produce about 20 ideas that they start producing less generic ideas.
-This phenomenon can be observed by examining the o-scores of ideas and idea categories, as well as the size of the category trees.
-[TODO] [Mike stopped here for the night]
+\subsection{Characteristics of an Idea Within a Run}
+We now turn our attention to individual runs, and characterize the ideas produced throughout a run.
 
-As our measure of originality, we use \emph{o-score}, introduced by Jansson and Smith \cite{jansson_design_1991}.
+\subsubsection{Most Popular Ideas}
+Our results suggest that individuals typically produce ideas from a common set idea categories early in the brainstorming process, then diverge and produce more original ideas.
+This phenomenon can be observed by examining the o-scores of ideas and idea categories in Figures X and Y, which show the distribution of o-scores for individual ideas and idea categories, as a function of response number, across all conditions.
+
+%As our measure of originality, we use \emph{o-score}, introduced by Jansson and Smith \cite{jansson_design_1991}.
 An idea's o-score is $1 - p(idea)$, where $p(idea) = (number of instances of that idea)/(number of instances total)$.
@@
 
-We measure differences in originality between conditions by examining the idea o-score and category o-score.
+% We measure differences in originality between conditions by examining the idea o-score and category o-score.
 The distributions of these scores are summarized in Figure FIG.
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\columnwidth]{oscores_conditions}
-\end{figure}
+% \begin{figure}[h!]
+%     \centering
+%     \includegraphics[width=0.9\columnwidth]{oscores_conditions}
+% \end{figure}
 
-The more ideas requested, the more original the ideas produced.
+% The more ideas requested, the more original the ideas produced.
 As suggested by the instance quartile range in section SEC, higher number conditions produce idea in trees with fewer instances, thus the high category o-score.
@@
 
-As the originality rises, however, so does the number of outliers.
+% As the originality rises, however, so does the number of outliers.
 To understand this phenomenon, we need to more closely observe the distribution of originality scores between not just conditions, but ordinal position in a brianstorming run.
 
-Figure FIG provides this visualization.
+% Figure FIG provides this visualization.
 The upper panel gives the mean idea o-score as a function of ordinal position in all 100 condition brainstorming runs.
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.9\columnwidth]{idea_oscore_order_100}
-    \caption{Idea o-score as a function of order (100 condition)}
-    \includegraphics[width=0.9\columnwidth]{cat_oscore_order_100}
-    \caption{Category o-score as a function of order (100 condition)}
-\end{figure}
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.9\columnwidth]{idea_oscore_order_100}
+%     \caption{Idea o-score as a function of order (100 condition)}
+%     \includegraphics[width=0.9\columnwidth]{cat_oscore_order_100}
+%     \caption{Category o-score as a function of order (100 condition)}
+% \end{figure}
 
-The only part of the originality score that falls outside the first and third quartiles is the beginning of the run.
+% The only part of the originality score that falls outside the first and third quartiles is the beginning of the run.
 Participants in the higher numbered conditions generate more original ideas overall, but not until they have exceeded some threshold of common ideas.
@@
     \includegraphics[width=0.9\columnwidth]{idea_oscore_order}
-    \caption{Idea o-score as a function of order}
+    \caption{Idea o-score as a function of response number, all conditions}
     \includegraphics[width=0.9\columnwidth]{cat_oscore_order}
-    \caption{Category o-score as a function of order}
+    \caption{Category o-score as a function of response number, all conditions}
 \end{figure}
 
-A similar, although less severe, pattern is observed for category o-score.
-There is a gradual increase in category originality until the 20 instance threshold is reached, at which point variance of the scores increases rapidly and no clear pattern is observable.
-This is explained by the relationship between common ideas and common categories.
+In both instances, we observe a steady increase in o-scores for the first 20-30 responses.
+After these first 20-30 responses, the o-scores essentially plateau, with no clear discernible pattern for either ideas or idea categories.
+
+These graphs lend support for hypothesis 3, the notion that there is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.
+
+To more explicitly confirm this hypothesis, we extracted the top 5\% most common ideas (66 in total).
+If our intuition is correct, these ideas should be more likely to appear in the first five instances of a run than they are to appear in any of the following instances.
+We introduce two Bernoulli random variables to model this.
+The first, defined by parameter $\theta_f$ represents the probability that a response instance is in the top 5\% most common ideas, given it is found in the first five instances of a run ($P(\theta_f|top)$).
+The second, defined by parameter $\theta_l$, is the probability that an instance is in the common pool given it is one of the \emph{last} (>5) instances in a run ($P(\theta_l|top)$).
+
+We fit these parameters using data from all runs that have more than 10 responses.
+We chose this constraint so that no run could contribute evidence to $\theta_f$ without also contributing to $\theta_l$.
+The resulting parameters were $\theta_f = 0.564$ (95\% HDI 0.517-0.611) and $\theta_l = 0.323$ (95\% HDI 0.303-0.342).
+This difference is significant, supporing the hypothesis that common ideas are over-represented in the early part of a brainstorming run.
+
+%This is explained by the relationship between common ideas and common categories.
 We expect that unoriginal ideas belong to unoriginal categories, as a high instance count for an idea increases the instance count for its category tree.
@@
 
-\subsubsection{Hypothesis 3}
-
-This interpretation supports \textbf{Hypothesis 3:} There is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.
-We test this interpretation by constructing a set of the top 66 (5\%) most common ideas.
-If our intuition is correct, these ideas should be more likely to appear in the first 5 instances of a run than they are to appear in any of the following instances.
-We designate two Bernoulli variables.
-The first, defined by parameter $\theta_f$, represents the probability that an instance is in the common set given it is in the first five instances of a run.
-The second, defined by parameter $\theta_l$, is the probability that an instance is in the common pool given it is one of the \emph{last} (>5) instances in a run..
-
-We fit these parameters using data from all runs with a total number of instances greater than 10.
-We chose this so that no run could contribute evidence to $\theta_f$ without also contributing to $\theta_l$.
-The resulting parameters were $\theta_f = 0.564$ (95\% HDI 0.517-0.611) and $\theta_l = 0.323$ (95\% HDI 0.303-0.342).
-This difference is significant, supporing the hypothesis that common ideas are over-represented in the early part of a brainstorming run, regardless of conditon.
-
-\subsubsection{Hypothesis 6}
-
-This threshold at 20 ideas brings us to \textbf{Hypothesis 6:} Ideas generated in the latter half of a brainstorming session are of higher originality than ideas generated in the former half.
-With our empirical evidence, we can instead test that ideas \emph{more than 20 instances into a brainstorming session} are more original than those in the first 20 instances, regardless of condition.
+\subsubsection{Originality of Ideas}
+Complementing the view of what ideas are generated when, we now consider idea and idea category o-scores as a function of response number.
+This analysis allows us to examine hypothesis 4, or our expectation that o-scores increase as a function of response number.
 
@@
 
-\subsection{Riffing: We need to change the name of this}
+\subsection{Idea Generation Patterns}
+Given the clear differences in characteristics of ideas as a function of response number, we now consider idea generation patterns observable from the data.
+In this section, we examine these patterns in the context of hypotheses 5 and 6.
+Specifically, these hypotheses draw upon the SIAM model and predict that individuals will generate ideas from an idea category until they exhaust that category, at which point they switch to another category (hypothesis 5).
+Furthermore, SIAM predicts that this category switching will be observable as taking more time to generate an idea in a new category than in the existing category (hypothesis 6).
 
-At the level of the individual brainstorming run, we examine a new phenomenon related to uniqueness above.
-Within the SIAM model, \emph{riffing} (or \emph{clustering} as it is known in the work) occurs when a single participant activates a single image (analogous to a non-singleton category) and produces multiple ideas from that image under a fail state is reached and a transition is undergone to a new image.
-Unsurprisingly, we observed riffing activity in our data.
-An example of a single participant's riffed responses is given in Figure FIG.
-The second column is dark if the instance is a \emph{look-back equal} to any idea earlier in the run.
-The first column is dark if the instance is a \emph{source} instance, meaning it is the first in a chain of look-back equal instances.
-Riff chains can either be consecutive (as in the example), or separated with instances in between.
+To determine category switching, we examine whether an idea was assigned to the same category tree as the previous, or whether it was assigned to a different category tree.
+We also examine the timing data collected from the web app to examine the time to generate an idea as a function of whether it is in the same category as the previous idea.
+
+Our data show clear evidence in support of both hypotheses. As an example, below is an example of a worker riffing on a single idea.
+
+%An example of a single participant's riffed responses is given in Figure FIG.
+The second column is dark if the response instance is a permutation of any idea earlier in the run.
+The first column is dark if the instance is a \emph{source} instance, meaning it is the source for later permutations.
+Permutation chains can either be consecutive (as in this example), or separated with instances in between.
 The third column simply provides the text for that instance.
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.5\columnwidth]{10_riff}
-    \caption{Example of riffing in a size 10 run}
-\end{figure}
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.5\columnwidth]{10_riff}
+%     \caption{Example of riffing in a size 10 run}
+% \end{figure}
 
-The participant hits on the categories of using the device's hard drive for storage, recycling the device or its materials, and selling the device.
+%The participant hits on the categories of using the device's hard drive for storage, recycling the device or its materials, and selling the device.
 within these categories, they generate 2-3 ideas each.
@@
 
-The higher the number of instances requested, the more participants riff on their old ideas.
-Similarly, upper conditions use more of their ideas as sources for riffing, and riff more times on each of those ideas.
-The primary surprise is that they do not riff more consecutively.
-The median length of a consecutive riff chain remains low across conditions.
-This leaves us with the image of a participant in the upper conditions searching through their old ideas for something to build upon one instance at a time.
+Examining our data, we find that the higher the number of instances requested, the more likely participants are to use old ideas as inspiration for new ideas (i.e., ``riff'' on previous ideas).
+However, these riff sequences are typically quite small in length, across conditions.
 
-Another outlier of note is the relatively short median distance to a previous idea in a chain for the 100 condition.
-Participants in the 100 condition may feel more pressure to remix, and return to earlier ideas more quickly than those in the 50 or 75 conditions.
-
-\subsubsection{Hypothesis 4}
-
-Riffing activity is analogous to idea production within an image as described by Nijstad and Stroebe's SIAM model \cite{nijstad_how_2006}.
-Thus, we should expect to confirm their hypothesis, \textbf{Hypothesis 4}, that \emph{an idea from one semantic category should more often by followed by an idea from the same category than expected by random chance}.
-
-To test this, we modelled the probability of any category being followed by the same category with a Bernoulli distribution.
-Each instance not the first in its run was a trial.
-If the instance was in the same category tree as the previous instance, the trial was a pass.
-Otherwise, it was a failure.
-The Bernoulli $\theta$ was 0.156 (95\% HDI 0.143 - 0.170).
-
-The most common category tree in our data set, \emph{music player}, represents only 5.65\% of instances.
-This is well below the 0.143 lower-bound of HDI.
+To explicitly test hypothesis 5, we model the probability of any idea category being followed by the same idea category with a Bernoulli random variable.
+Using a uniform beta prior, we find the probability of a response instance being in the same idea category as the previous to be 15.6\% (95\% HDI 14.3\% - 17.0\%).
+To establish a baseline to understand whether this probability is greater than chance, the most popular category tree in our data set is \emph{music player}, representing 5.65\% of all instances.
+Thus, the greatest probability of one instance being in the same category as the previous instance, by chance, is 5.65\%, well below the 14.3\% lower bound of the calculated HDI.
 Thus, we accept the hypothesis that an instance follows a previous instance in the same category with greater probability than is explained by random chance.
 
-\subsection{Completion Time}
+To test hypothesis 6, we now examine the time it takes to generate a response instance.
 
-As another aspect of how participants complete brainstorming runs, we are interested in the time spent on each response.
-The first, second and third quartiles for each condition are given in boxplots in figure FIG.
-The whiskers extend to 1.5 times the inner quartile range.
-Surprisingly, we see many instances taking up to 28018382ms (7.78 hours) to complete.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.9\columnwidth]{time_spent_condition}
-    \caption{Time spent per instance}
-\end{figure}
-
-Either our participants were exceedingly concerned with crafting their responses, or there is another influence at work.
-Figure FIG gives the mean time spent on an instance by order, separated by condition.
-Most of the highly variant response times take place early in the brainstorming run.
-This suggests that at least one participant accepted a brainstorming task, took an initial stab at instances, and then left the task for up to 8 hours before continuing.
-Beyond this, we might expect to see time spent per instance go up as participants have to dig deeper to find responses, but no such effect is consistently observed.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.9\columnwidth]{time_spent_order}
-    \caption{Time spent per instance}
-\end{figure}
-
-\subsubsection{Hypotheses 5}
-
-The Nijstad and Stroede \cite{nijstad_how_2006} model also provides \textbf{Hypothesis 5}: \emph{idea generation when changing semantic categories should take longer than idea generation within categories}.
-An examination of the distribution of data for time spent per instance suggested a log-normal distribution, with the frequency of occurances quickly rising until a peak was reached, followed by a gradual drop-off.
-As with previous tests, we used the Stan package for Bayesian inference [CITE].
-We fit, two models: $t_w = \text{lognormal}(\mu_w, \sigma_w)$ for within-category idea generation and $t_b = \text{lognormal}(\mu_b, \sigma_b)$ for between-category idea generation, where the $t$ parameter in each model represents the time spent generating an instance.
+An examination of the distribution of data for time spent per instance suggests a log-normal distribution to the data.
+Using STAN, we fit two models: $t_w = \text{lognormal}(\mu_w, \sigma_w)$ for within-category idea generation and $t_b = \text{lognormal}(\mu_b, \sigma_b)$ for between-category idea generation, where the $t$ parameter in each model represents the time spent generating an instance.
 
@@
 Within-category instance generation took on average 12.6 seconds less than between-categories.
-This is consistent with the findings of Nijstad and Stoebe, who reported differences of 6-12 seconds.
+This is consistent with the findings of Nijstad and Stoebe, who reported differences of 6-12 seconds between conditions.
+Thus, we find support for hypothesis 6 and also show the applicability of the SIAM model to brainstorming in microtask marketplaces.
+
+%Figure FIG displays the mean time spent on an instance by order, separated by condition.
+Most of the highly variant response times take place early in the brainstorming run.
+This suggests that at least one participant accepted a brainstorming task, took an initial stab at instances, and then left the task for up to 8 hours before continuing.
+Beyond this, we might expect to see time spent per instance go up as participants have to dig deeper to find responses, but no such effect is consistently observed.
+
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.9\columnwidth]{time_spent_condition}
+%     \caption{Time spent per instance}
+% \end{figure}
+
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.9\columnwidth]{time_spent_order}
+%     \caption{Time spent per instance}
+% \end{figure}
+
+%The Nijstad and Stroede \cite{nijstad_how_2006} model also provides \textbf{Hypothesis 5}: \emph{idea generation when changing semantic categories should take longer than idea generation within categories}.

### 2013-09-18 chi2014/method.tex
fix quotes in method section

@@
 \begin{enumerate}
-    \item "Storage Container"
-    \item "Small Storage Box"
-    \item "Coin Storage"
-    \item "Travel Jewelry Case"
+    \item ``Storage Container"
+    \item ``Small Storage Box"
+    \item ``Coin Storage"
+    \item ``Travel Jewelry Case"
 \end{enumerate}
@@
 It proved difficult to determine and adhere to a constant granularity of distinction.
-For example, it is desireable to encode some equality between "use [MP3 players] to build a chair, and "use [MP3 players] to build a table".
+For example, it is desireable to encode some equality between ``use [MP3 players] to build a chair, and "use [MP3 players] to build a table".
 Information loss results from either assigning both the same code or each different codes.
-An interval specificity scale is also difficult to employ; intuitively, we cannot compare the distance between "audiobooks" and "audio books for the blind" with the distance between the latter and "[audio] instructions for the blind on a variety of things".
+An interval specificity scale is also difficult to employ; intuitively, we cannot compare the distance between ``audiobooks" and ``audio books for the blind" with the distance between the latter and ``[audio] instructions for the blind on a variety of things".
 

### 2013-09-18 chi2014/method.tex
Method: blockquote

@@
 The final brainstorming question we chose for this work is the following:
-
-TODO: Indent, italicize this:
-Many people have old MP3 players or MP3 players that they no longer use.
+\begin{quote}
+{\em Many people have old MP3 players or MP3 players that they no longer use.
 Please brainstorm N uses for old MP3 players/MP3 players.
 Assume that the devices' batteries no longer work, though they can be powered via external power sources.
-Also be aware that devices may \emph{not} have displays.
-Be as specific as possible in your descriptions.
+Also be aware that devices may {\em not \/} have displays.
+Be as specific as possible in your descriptions.\/}
+\end{quote}
 

### 2013-09-18 chi2014/introduction.tex
Minor edits to Intro

--- 2013-09-18 chi2014/introduction.tex
@@
 Microtask marketplaces such as Amazon's Mechanical Turk have proven to be useful platforms for outsourcing a wide range of tasks, from proofreading documents \cite{soylent} to estimating the caloric content of food \cite{platemate}.
-These platforms make possible human computation systems \cite{quinn_2011}, hybrid systems that tap humans' abilities to solve problems that currently cannot be solved by machines alone.
+These platforms make possible human computation systems \cite{quinn_2011}, hybrid systems that leverage humans' abilities to solve problems that currently cannot be solved by machines alone.
 
@@
 While brainstorming has been an implicit component of many creative endeavors in microtask marketplaces, we are unaware of any research that specifically describes and models the brainstorming process in this environment.
-Furthermore, while there is a rich literature describing brainstorming processes of individuals and groups (e.g., \cite{osborn_applied_1957, pinsonneault_electronic_1999, nijstad_how_2006} OTHERS?), microtask marketplaces have qualitative differences.
-For example, brainstorming participants in a microtask marketplace are paid for their participation, and are often eager to complete a task as fast as possible to maximize earnings [REF?].
+Furthermore, while there is a rich literature describing brainstorming processes of individuals and groups (e.g., \cite{osborn_applied_1957, pinsonneault_electronic_1999, nijstad_how_2006,brown1998modeling}), there are important qualitative differences in microtask marketplaces.
+For example, crowd workers exhibit high variance of effort in completing a task \cite{soylent}.
+Brainstorming participants in a microtask marketplace are also paid for their participation, and some might be more eager than others to complete a task as fast as possible to maximize earnings \cite{kittur2008crowdsourcing}.
 Furthermore, participants in microtask marketplaces cannot be expected to be spatially or temporally co-located.
-There many open questions about how brainstorming unfolds in this context.
+Collectively, these differences make it unclear how brainstorming unfolds in this context.
 
@@
 
-This paper presents the results of a series of experiments intended to model brainstorming on microtask marketplaces (specifically, Amazon's Mechanical Turk).
+This paper presents the results of a series of experiments intended to model brainstorming in microtask marketplaces (specifically, Amazon's Mechanical Turk).
 In our experiments, we asked individuals to brainstorm ways to use an MP3 player whose batteries were dead.
@@
 Across all conditions, 154 workers submitted responses to 170 HITs (Human Intelligence Tasks), yielding 3292 ideas.
-From these ideas, we present a series of models characterizing features of crowd brainstorming, such as the rate of novel idea production, the rate at which novel {\em categories\/} of ideas are produced (where a category represents a set of conceptually similar ideas that are distinct from other categories), and the relative novelty of ideas, all as functions of number of responses requested.
+From these ideas, we present a series of models characterizing features of crowd brainstorming, such as the rate of novel idea production, the rate at which novel {\em categories\/} of ideas are produced (where a category represents a set of conceptually similar ideas that are distinct from other categories), qualities of the ideas (for example, whether they are common or not), and characteristics of the brainstorming process itself.
 
@@
 \begin{itemize}
-\item The rates at which novel ideas and categories of ideas are produced follow logarithmic growth models, as a function of the total number of responses receieved across participants.
-Thus, the rate of novel ideas decreases exponentially.
-\item The rates for generating novel ideas changes with the number of responses requested per individual, even if the overall number of responses is held constant.
+\item The rates at which novel ideas and categories of ideas are produced follow logarithmic growth models. Thus, the rates of novel ideas and categories of ideas decrease exponentially.
+%\item The rates for generating novel ideas changes with the number of responses requested per individual, even if the overall number of responses is held constant.
 In general, a higher rate of novel idea generation is achieved by asking fewer participants for more responses each.
-[TODO: There is likely past work that has a similar finding; we should relate this finding to it.
+This result corresponds to previous finding that making participants to extend more effort generating ideas--in this case by requiring them to generate a lot of ideas--will lead to an increasing proportion of good ideas\cite{parnes_effects_1961}.
+%[TODO: There is likely past work that has a similar finding; we should relate this finding to it.
 Taylor Berry Block as related?
-Diehl as related?]
+Diehl as related?][Pao's note: can't find more directly related stuff mainly because most of the previous lit focus on time limit task rather than number of ideas.
+This is actually a good argument for the paper since time limit might not work in this marketplace.]
 \item People are likely to come up with the same overall set of ideas early in their brainstorming process, but then diverge from this set of common ideas as they generate more ideas.
-The first 20 ideas are the least original (i.e., the most common).
-This finding complements the previous finding, and further suggests that there is little value in requesting fewer than 20 ideas from workers
+Given our results, we find there is little value in requesting fewer than 20 ideas from workers, since these early ideas are among the most common.
 \item Individuals' patterns of idea generation mimic those identified by Nijstaf and Stroebe's SIAM (search for ideas in associative memory) model \cite{nijstad_how_2006}: Workers produce a set of ideas from working memory, then switch strategies to generate ideas that are permutations of other ideas, a process we dub {\em riffing\/}.
-In our data, we see the prevalence of riffing steadily increase over the first 20 responses, after which its prevalence plateaus
-\item Requesting ``as many ideas as possible'' for a fixed price (fixed for the entire HIT, rather than per idea) is inadvisable, as it leads to a relatively low number of ideas produced (in our experiments, fewer than 15)
-\item The time spent producing an idea matches the times reported in previous studies \cite{nijstad_how_2006}
+In our data, we see the prevalence of riffing steadily increase across the first 20 responses, after which its prevalence plateaus.
+\item We also replicate Nijstaf and Stroebe's findings with respect to the time it takes to generate a new idea \cite{nijstaf_how_2006}.
+\item Requesting ``as many ideas as possible'' for a fixed price (fixed for the entire HIT, rather than per idea) is inadvisable, as it leads to a relatively low number of ideas produced (in our experiments, fewer than 15).
 \end{itemize}

### 2013-09-18 chi2014/related.tex
Tweaks to related work.

@@
 
-Following Osborne's work, Taylor et al.\ established that brainstorming by \emph{nominal groups}, or individuals working in isolation from one another whose results are later combined, yields better results than collocated groups in terms of number of ideas generated \cite{taylor_does_1958}.
+Following Osborne's work, Taylor et al.\ established that brainstorming by \emph{nominal groups}, or individuals working in isolation from one another, yields better results than collocated groups in terms of number of ideas generated \cite{taylor_does_1958}.
 These effects can be partially explained by the absence of counterproductive social forces, such as production blocking (the inability for multiple people to speak at once) and the fear of judgment.
-Bouchard and Hare \cite{bouchard_jr_size_1970}, later found that nominal brainstorming groups are able to generate quantities of ideas linear in the number of members in the group.
+Bouchard and Hare \cite{bouchard_jr_size_1970} later found that nominal brainstorming groups are able to generate quantities of ideas linear in the number of members in the group.
 
@@
 As with nominal brainstorming, Gallupe \cite{gallupe_electronic_1992} observed that electronic brainstorming can reduce counterproductive social effects.
-However, Pinsonneault et al.\ \cite{pinsonneault_electronic_1999} identify productivity impediments introduced by group electronic brainstorming, such as being distracted by other ideas appearing, or individuals witholding ideas they feel are derivative of others.
+However, Pinsonneault et al.\ \cite{pinsonneault_electronic_1999} identify productivity impediments introduced by group electronic brainstorming, such as being distracted by other ideas appearing, and individuals witholding ideas they feel are derivative of others.
 Given these drawbacks, their results suggest there is little evidence to support the proposition that electronic brainstorming outperforms nominal brainstorming.
@@
 
-In terms of individual brainstorming processes, Nijstad and Stroebe provide a model for individual idea generation dubbed SIAM, or the search for ideas in associative memory.
-\cite{nijstad_how_2006}.
-SIAM defines two stages of idea generation: the generation of ideas based working memory, and idea production within a single activated ``image'' (where an image can be thought of as a category of ideas).
-This model makes predictions applicable to all brainstorming contexts, namely that an idea is more likely to be followed by an idea in the same category than would be expected by chance, and new ideas in the same category as the previous idea are generated faster than ideas between categories.
-Nijstad and Stroebe verify these in the context of nominal brainstorming.
+Shifting our focus to an \emph{individual's} brainstorming processes, Nijstad and Stroebe provide a model for individual idea generation dubbed SIAM, or the ``search for ideas in associative memory'' \cite{nijstad_how_2006}.
+SIAM defines two stages of idea generation: the generation of ideas based on working memory, and idea production within a single activated ``image'' (where an image can be thought of as a category of ideas).
+This model makes predictions applicable to all brainstorming contexts, with one of the most important predictions being that an idea is more likely to be followed by an idea in the same category than would be expected by chance.
+Their model also predicts that new ideas in the same category as the previous idea are generated faster than ideas between categories.
+Nijstad and Stroebe verified these predictions in the context of nominal brainstorming.
 
@@
 %Examples for these environments are platforms for generating social innovation ideas or product concepts like OpenIDEO, Quirky and Threadless that employ structured process where participants take on different responsibility from generating ideas to evaluating others' ideas.
-In this paper, we examine nominal brainstorming to establish a baseline.
+In this paper, we constrain our focus to nominal brainstorming in this environment with the goal of establishing a set of baselines.
 
@@
 For example, Yu and Nickerson created an evolutionary crowd algorithm to generate creative chair designs \cite{yu_cooks_2011}.
-Their work and others demonstrates the creative potential of the crowd, but does not specifically address brainstorming or analogous processes in which participants generate many ideas.
+Their work demonstrates the creative potential of the crowd, but does not specifically address brainstorming or analogous processes in which participants generate many ideas.
 
@@
 Participants in the parallel condition worked in isolation without seeing other participants' ideas.
-Subjective ratings of the quality of ideas reveal that while the ideas from the iterative condition were rated higher on average, the best ideas from the parallel condition have higher scores than the best ideas from the iterative conditions.
-The parallel process is more likely to generate names with high ratings, advocating the case for nominal brainstorming over traditional electronic brainstorming.
+Subjective ratings of the quality of ideas reveal that while the ideas from the iterative condition were rated higher on average, the best ideas from the parallel condition were rated higher than the best ideas from the iterative conditions.
+Thus, the parallel process was found to be more likely to generate names with high ratings, advocating the case for nominal brainstorming over traditional electronic brainstorming.
 However, the study did not mention categorization of generated ideas, an important measure on flexibility of idea generation \cite{lewis2011affective, nijstad_how_2006, finke1992creative, shah2003metrics}.
@@
 
-The nature of microtask marketplace also calls for an alternative structure for brainstorming tasks in this setting.
+The nature of the microtask marketplace may also require alternative ways of structuring the brainstorming activity in this environment.
+% calls for an alternative structure for brainstorming tasks in this setting.
 Researchers have found that 30\% or more submission on Mechanical Turk may be low quality \cite{kittur2008crowdsourcing}.
@@
 These low quality responses were not found in walk-in brainstorming studies.
-Bernstein et al.
-identified high variance of effort Amazon Mechanical Turk workers put on task \cite{soylent}.
+For exapmle, Bernstein et al.
+found high variance of effort from Amazon Mechanical Turk workers \cite{soylent}.
 They characterized two worker personas at opposite ends of the effort spectrum: the {\em Eager Beaver\/} who goes beyond task requirement and the {\em Lazy Turker\/} who does as little work as necessary to get paid.
-A brainstorming task that asks participants to generate as many ideas as they can under a set time limit might not be practical.
-A {\em Lazy Turker\/} can simply fill in a few ideas and wait for the time to run out instead of exerting full effort.
-The motivational structure of microstask marketplace does not map well with most previous work on brainstorming where the sessions are time-limited.
+Given this finding, a brainstorming task that asks participants to generate as many ideas as they can in a fixed time period might not be cost-effective, since a {\em Lazy Turker\/} can simply fill in a few ideas and wait for the time to run out instead of exerting full effort.
+
+ %The motivational structure of microstask marketplace does not map well with most previous work on brainstorming where the sessions are time-limited.
 
@@
 
-In summary, while brainstorming has been examined in the traditional and electronic settings, its application in crowd marketplaces has received limited attention.
+In summary, while brainstorming has been examined in the traditional and electronic settings, its application in crowd marketplaces has received limited attention, leaving many open questions about how this process occurs in this environment.

### 2013-09-18 chi2014/abstract.tex
Edit abstract

@@
 Microtask marketplaces enable hybrid computational systems that use crowds to solve problems of creativity, such as brainstorming.
+We present a first characterization of brainstorming in this context, providing models for the rate of idea generation, the rate of category generation and the novelty of generated ideas.
+We demonstrate that the rates of idea and category generation can be closely fit by a logarithmic model, and identify a \emph{burn in period} of common ideas at the beginning of a brainstorming task when people came up with the same overall set of ideas before diverging from this common set.
+Our results suggest a cutoff of this period at around the 20th idea.
+Furthermore, we explore the phenomenon of \emph{riffing} in brainstorming tasks, and replicate two related findings from prior work.
+Our results provide recommendations for those leveraging the crowd for brainstorming, and comparable models of creative idea generation research community.
+
+%Microtask marketplaces enable hybrid computational systems that use crowds to solve problems of creativity, such as brainstorming.
 We present a first characterization of brainstorming in this context, providing models for the rate of idea generation, the rate of category generation, the originality of ideas and the novelty of ideas.

### 2013-09-18 chi2014/abstract.tex
Change novelty to originality

@@
 Microtask marketplaces enable hybrid computational systems that use crowds to solve problems of creativity, such as brainstorming.
-We present a first characterization of brainstorming in this context, providing models for the rate of idea generation, the rate of category generation and the novelty of generated ideas.
+We present a first characterization of brainstorming in this context, providing models for the rate of idea generation, the rate of category generation and the originality of generated ideas.
 We demonstrate that the rates of idea and category generation can be closely fit by a logarithmic model, and identify a \emph{burn in period} of common ideas at the beginning of a brainstorming task when people came up with the same overall set of ideas before diverging from this common set.

### 2013-09-18 chi2014/results.tex
Minor edits to Intro
hypothesis 4

@@
 
-Hierarchical clustering yielded an idea forest composed of 3007 instances, 1213 unique ideas, 335 category trees, and 129 non-singleton category trees.
+Hierarchical clustering yielded an idea forest composed of 3007 instances, 1213 unique ideas, 335 category trees, and 129 non-singleton category trees (Table~\ref{tab:idea_forest}).
 In general, category trees were shallow, with a maximum tree height of 5.
-The number of nodes within a category tree ranged from 1 to X, with the distribution of values roughly following a Poisson distribution.
-The number of instances per node (i.e., the number of equivalent idea instances) also roughly follows a Poisson distribution, with number of instances ranging from X to Y.
-
+The number of nodes within a category tree ranged from 1 to 66, with the distribution of values roughly following a Poisson distribution.
+The number of instances per node (i.e., the number of equivalent idea instances) also roughly follows a Poisson distribution, with number of instances ranging from 1 to 63.
 \begin{table}
 	\begin{tabular}[h!]{r | l l l l l l l}
-	\textbf{number condition} & 5 & 10 & 20 & 50 & 75 & 100 & all \\ \hline \hline
-	HITs & 57 & 47 & 23 & 10 & 10 & 10 & 146\\
-	instances & 293 & 471 & 453 & 500 & 634 & 855 & 3007\\
-	ideas & 171 & 249 & 278 & 341 & 443 & 177 &1212\\
-	category trees & 72 & 79 & 93 & 114 & 172 & 177 &321\\
-	non-singleton trees & 28 & 34 & 40 & 48 & 49 & 61 &92\\
+	\textbf{number condition} & 5 & 10 & 20 & 50 \\ \hline \hline
+	HITs & 57 & 47 & 23 & 10\\
+	instances & 293 & 471 & 453 & 500 \\
+	ideas & 171 & 249 & 278 & 341 \\
+	category trees & 72 & 79 & 93 & 114 \\
+	non-singleton trees & 28 & 34 & 40 & 48 \\ \hline \hline
+	\textbf{number condition} & 75 & 100 & all \\ \hline \hline
+    HITs & 10 & 10 & 146\\
+	instances & 634 & 855 & 3007\\
+	ideas & 443 & 177 &1212\\
+	category trees & 172 & 177 &321\\
+	non-singleton trees & 49 & 61 &92\\ 
 	\end{tabular}
+    \caption{Result counts between conditions}
+    \label{tab:idea_forest}
 \end{table}
@@
     \caption{Idea o-score as a function of response number, all conditions}
+    \label{fig:idea_oscore_order}
     \includegraphics[width=0.9\columnwidth]{cat_oscore_order}
     \caption{Category o-score as a function of response number, all conditions}
+    \label{fig:cat_oscore_order}
 \end{figure}
@@
 
-\subsubsection{Originality of Ideas}
-Complementing the view of what ideas are generated when, we now consider idea and idea category o-scores as a function of response number.
-This analysis allows us to examine hypothesis 4, or our expectation that o-scores increase as a function of response number.
+\subsubsection{Hypothesis 4}
+The steady increase in originality seen in the first 20 instances (Figure~\ref{fig:idea_oscore_order}, \ref{fig:cat_oscore_order}), provides evidence for \textbf{Hypothesis 4:} ideas generated in the second half a brainstorming session have higher o-scores than those in the first half.
+With the point of change seen at the 20 point, we use this as our division point, fitting two each for idea and category o-scores: instances in the first 20 of runs, and instances in the second 20 runs.
+In examining the data, it became clear that o-scores do not follow a remotely normal distribution.
+Instead, we fit a beta distributions to each data division.
+The results are in Table~\ref{tab:hyp4} and Figure~\ref{fig:idea_oscore_hyp4}.
 
-Test here...
+\begin{table}
+	\begin{tabular}[h!]{r | l l }
+	    & $\alpha$ median & $\alpha$ 95\% HDI \\ \hline \hline
+        idea o-score (\textless=20) & 135.86 & 122.39-149.72 \\
+        idea o-score (\textgreater20) & 259.84 & 240.47-278.81 \\ \hline \hline
+	    & $\beta$ median & $\beta$ 95\% HDI \\ \hline \hline
+        category o-score (\textless=20) & 1.16 & 1.08-1.25 \\
+        category o-score (\textgreater20) & 0.82 & 0.77-0.86 \\
+	\end{tabular}
+    \caption{Hypothesis 4 Beta models}
+    \label{tab:hyp4}
+\end{table}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{hyp4_ideas}
+    \caption{Beta models for idea o-score}
+    \label{fig:idea_oscore_hyp4}
+    \includegraphics[width=0.9\columnwidth]{hyp4_cats}
+    \caption{Beta models for category o-score}
+    \label{fig:cat_oscore_hyp4}
+\end{figure}
+
+The idea o-score models are significantly different in the $\alpha$ parameter, with mean o-scores of 0.995 and 0.997 for the $\leq20$ and $\textgreater20$ conditions respectively.
+The category o-scores are significantly different in both parameters, with mean o-scores of 0.975 and 0.982.
+
+Participants generated more original ideas after their first 20 instances than they did after.
+Not only does this support the hypothesis, it provides a clear guideline for maximizing originality in brainstorming tasks on microtask marketplaces: it is more cost-effective to solicit $\textgreater20$ ideas from a smaller pool of participants than to ask for less from more.
 

### 2013-09-18 chi2014/goals.tex
Goals: brainstorming metrices
Little wordplay

--- 2013-09-18 chi2014/goals.tex
@@
 
-To characterize brainstorming in microtask marketplaces, we are interested in modeling the following quantities:
-\begin{itemize}
-\item the rate of unique idea production
-\item the rate of production of unique \emph{categories} of ideas
-\item the relative originality of ideas
-\item individual practices in idea generation
-\end{itemize}
+% Researchers have long debated how to best assess creativity in a way that consistently captures the complex dimensions of creativity [35].
+The most commonly adopted approach [1, 5, 17] and the one we employ is to ask domain experts to subjectively make scale-rated assessments of creativity.
+Typically, domain experts rate originality, feasibility, elaboration, and flexibility and count the number of ideas generated [3].
 
-We will examine these rates as a function of the number of ideas requested from workers.
+To characterize brainstorming in microtask marketplaces, we are interested in modeling rates at which ideas and idea categories are generated, the originality of ideas as measured by o-scores, and individuals' idea generation processes.
 
@@
 
-In modeling these quantities, we have several hypotheses related to results received in the aggregate (across all participants), as well as individuals' brainstorming processes.
+In modeling these quantities, we wish to explore the following hypotheses:
 
-\textbf{Hypothesis 1:} The rate of new ideas will diminish exponentially toward zero as a function of the number of responses gathered.
-Less formally, by continually soliciting solutions, we will eventually reach a saturation point at which any idea a worker submits is likely to be the same or practically identical to previously seen solutions.
-Related to this, we expect the number of \emph{categories} of ideas to diminish exponentially as well.
+\textbf{Hypothesis 1:} The probability that an idea is novel decreases non-linearly as a function of the number of responses received.
+Similarly, we expect the rate of new idea categories to decrease non-linearly over time.
 
-\textbf{Hypothesis 2:} The rate of new ideas will tend toward zero faster in some conditions than others.
-In particular, we expect conditions requesting a smaller number of ideas will tend toward zero faster, for reasons hypothesized next.
+\textbf{Hypothesis 2:} There is a set of general, common ideas that make up the first several responses of every crowd brainstorming session.
+In other words, we expect there to be a relatively fixed set of ideas that most people think of first.
 
-\textbf{Hypothesis 3:} We anticipate there is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.
-In other words, we expect there to be a relatively fixed set of ideas that ``everyone'' thinks of first.
-
-Parnes et al \cite{parnes_effects_1961} showed that more high quality ideas are generated in the latter half of a brainstorming run than the former half.
+%Parnes et al \cite{parnes_effects_1961} showed that more high quality ideas are generated in the latter half of a brainstorming run than the former half.
 Because of the difficulty in reliably assessing the quality of an idea for a brainstorming exercise, we instead consider the o-score of an idea to indirectly measure this concept.
 
-\textbf{Hypothesis 4:} Ideas generated in the second half of a brainstorming session have higher o-scores than ideas generated in the first half.
+\textbf{Hypothesis 3:} O-scores generated in the second half of a brainstorming session will be higher than ideas generated in the first half.
+This hypothesis provides a complementary perspective to the former.
 
-With respect to individual brainstorming processes, we have several hypotheses derived from Nijstad and Stroebe's work and their SIAM model \cite{nijstad_how_2006}:
+With respect to individual brainstorming processes, we have a pair of hypotheses derived from Nijstad and Stroebe's work and their SIAM model \cite{nijstad_how_2006}:
 
-\textbf{Hypothesis 5:} An idea from one semantic category of ideas should more often be followed by an idea from the same category than expected by random chance.
+\textbf{Hypothesis 4:} An idea from one semantic category of ideas should more often be followed by an idea from the same category than expected by random chance.
 
-\textbf{Hypothesis 6:} The time it takes to generate an idea should be longer when changing semantic categories than when generating ideas within a category.
+\textbf{Hypothesis 5:} The time it takes to generate an idea should be longer when changing semantic categories than when generating ideas within a category.

### 2013-09-18 chi2014/related.tex

@@
 
-% TODO Note: I plaigarized the wording on these definitions from wikipedia at this point, should pull from original Osborne at some point
+\subsection{Traditional Brainstorming}
 In the 1950s, Osborne formalized the brainstorming process, providing a set of recommendations for idea generation by groups \cite{osborn_applied_1957}: \emph{focus on quantity}, \emph{withhold criticism}, \emph{welcome unusual ideas}, and \emph{combine and improve ideas}.
@@
 
+To assess the brainstorming process and its outputs, a number of measures have been proposed.
+Typically, the measures of concern are the novelty of generated ideas, the practicality/appropriateness of ideas, the variety of ideas (flexibility), and the quantity of ideas generated.
+\cite{finke1992creative, shah2003metrics}.
+
+The most common approach to measure novelty and appropriateness is to ask domain experts to subjectively make scale-rated assessment of each \cite{lewis2011affective, amabile_1983}.
+Depending on the types of artifacts and researchers' interest, these two measures are sometimes combined into one measure of quality \cite{little2010exploring}.
+
+Instead of subjective ratings, some researchers opt for a measure that can be derived objectively.
+For example, Jansson et al.
+used `o' scores, derived from the frequency with which an idea occurs, to evaluate products novelty \cite{jansson_design_1991}.
+We adopt this use of o-score in our later analyses.
+
+Flexibility is often measured by the number of explored ``categories'' in a design space, or by subjective rating from domain experts \cite{lewis2011affective, marsh1996examples}.
+Category or feature-based quantities also yield other metrics such as within-catgories fluency and the length of train-of-thought when participants focus on a particular category \cite{nijstad_how_2006}.
+Building in this past line of work, we consider both individual ideas and categories of ideas.
+
+\subsection{Crowdsourced Brainstorming}
 The simplist imaginable instantiation of brainstorming in a microtask marketplace mirrors that of nominal brainstorming.
@@
 
-In summary, while brainstorming has been examined in the traditional and electronic settings, its application in crowd marketplaces has received limited attention, leaving many open questions about how this process occurs in this environment.
+In summary, while brainstorming has been examined in traditional and electronic settings, its application in crowd marketplaces has received limited attention, leaving many open questions about how this process occurs in this environment.

### 2013-09-18 chi2014/results.tex
Fixed plots

--- 2013-09-18 chi2014/results.tex
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{ideas_over_time}
-    \caption{Cumulative idea count}
+    \includegraphics[width=0.9\columnwidth]{cumulative_ideas}
+    \caption{Cumulative number of novel ideas}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.7\columnwidth]{rate_new_idea_over_time}
-    \includegraphics[width=0.7\columnwidth]{rate_new_category_over_time}
-    \includegraphics[width=0.7\columnwidth]{rate_new_ns_category_over_time}
-    \caption{Rate of idea generation}
+    \includegraphics[width=0.9\columnwidth]{cumulative_categories}
+    \caption{Cumulative number of novel idea categories}
 \end{figure}
+
+% \begin{figure}[h!]
+%     \centering
+%     \includegraphics[width=0.7\columnwidth]{rate_new_idea_over_time}
+%     \includegraphics[width=0.7\columnwidth]{rate_new_category_over_time}
+%     \includegraphics[width=0.7\columnwidth]{rate_new_ns_category_over_time}
+%     \caption{Rate of idea generation}
+% \end{figure}
 
@@
 This analysis allows us to examine hypothesis 4, or our expectation that o-scores increase as a function of response number.
-
-Test here...
+We examine o-scores for both ideas and idea categories.
 

### 2013-09-18 chi2014/related.tex
typo

@@
 \subsection{Crowdsourced Brainstorming}
-The simplist imaginable instantiation of brainstorming in a microtask marketplace mirrors that of nominal brainstorming.
+The simplest imaginable instantiation of brainstorming in a microtask marketplace mirrors that of nominal brainstorming.
 More sophisticated brainstorming environments are possible, such as those that provide awareness of ideas that others have generated (electronic brainstorming), or those that rely on the crowd to organize or rank the ideas produced by others, as in evolutionary techniques \cite{yu_cooks_2011}.

### 2013-09-18 chi2014/results.tex
Some revision, fixing of lots of technical problems

--- 2013-09-18 chi2014/results.tex
@@
 
-Figure FIG shows the cumulative count of new ideas as a function of the number of instances gathered in the experiment. For reference, a perfect ``1:1'' idea generation line is also plotted.
+Figure~\ref{fig:cumulative_ideas} shows the cumulative count of new ideas as a function of the number of instances gathered in the experiment.
+For reference, a perfect ``1:1'' idea generation line is also plotted.
 
@@
 
-We fit the parameters to the data using a Bayesian approach [REF Kruschke], using the STAN modeling software to perform the actual fit [REF STAN].
+We fit the parameters to the data using a Bayesian approach \cite{kruschke_doing_201}, using the STAN modeling software to perform the actual fit \cite{stan-software:2013}.
 In modeling the parameters, we use uninformative priors.
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{ideas_over_time}
-    \caption{Cumulative idea count}
+    \includegraphics[width=0.9\columnwidth]{cumulative_ideas}
+    \caption{Cumulative number of novel ideas}
+    \label{fig:cumulative_ideas}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.7\columnwidth]{rate_new_idea_over_time}
-    \includegraphics[width=0.7\columnwidth]{rate_new_category_over_time}
-    \includegraphics[width=0.7\columnwidth]{rate_new_ns_category_over_time}
-    \caption{Rate of idea generation}
+    \includegraphics[width=0.9\columnwidth]{cumulative_categories}
+    \caption{Cumulative number of novel idea categories}
+    \label{fig:cumulative_categories}
 \end{figure}
+
+% \begin{figure}[h!]
+%     \centering
+%     \includegraphics[width=0.7\columnwidth]{rate_new_idea_over_time}
+%     \includegraphics[width=0.7\columnwidth]{rate_new_category_over_time}
+%     \includegraphics[width=0.7\columnwidth]{rate_new_ns_category_over_time}
+%     \caption{Rate of idea generation}
+% \end{figure}
 
@@
 Our results suggest that individuals typically produce ideas from a common set idea categories early in the brainstorming process, then diverge and produce more original ideas.
-This phenomenon can be observed by examining the o-scores of ideas and idea categories in Figures X and Y, which show the distribution of o-scores for individual ideas and idea categories, as a function of response number, across all conditions.
+This phenomenon can be observed by examining the o-scores of ideas and idea categories in Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order}, which show the distribution of o-scores for individual ideas and idea categories, as a function of response number, across all conditions.
 
@@
 The first, defined by parameter $\theta_f$ represents the probability that a response instance is in the top 5\% most common ideas, given it is found in the first five instances of a run ($P(\theta_f|top)$).
-The second, defined by parameter $\theta_l$, is the probability that an instance is in the common pool given it is one of the \emph{last} (>5) instances in a run ($P(\theta_l|top)$).
+The second, defined by parameter $\theta_l$, is the probability that an instance is in the common pool given it is one of the \emph{last} ($\textgreater5$) instances in a run ($P(\theta_l|\overline{top})$).
 
@@
 
-\subsubsection{Hypothesis 4}
-The steady increase in originality seen in the first 20 instances (Figure~\ref{fig:idea_oscore_order}, \ref{fig:cat_oscore_order}), provides evidence for \textbf{Hypothesis 4:} ideas generated in the second half a brainstorming session have higher o-scores than those in the first half.
-With the point of change seen at the 20 point, we use this as our division point, fitting two each for idea and category o-scores: instances in the first 20 of runs, and instances in the second 20 runs.
-In examining the data, it became clear that o-scores do not follow a remotely normal distribution.
-Instead, we fit a beta distributions to each data division.
-The results are in Table~\ref{tab:hyp4} and Figure~\ref{fig:idea_oscore_hyp4}.
+\subsubsection{Originality}
+
+The increase in o-score over time in Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order}) also supports hypothesis 4, suggest by Parnes et al \cite{parnes_effects_1961}, that ideas generated in the second half a brainstorming session have higher o-scores than those in the first half.
+We test a stronger guarantee implied by our data: that ideas after the first 20 have higher o-scores than the first 20.
+Although we expected the o-scores to be normally distributed in each condition, they were not.
+Instead, we fit Beta distributions to o-score for each of the $\leq20$, $\textgreater20$ conditions.
+Table~\ref{tab:hyp4} and Figure~\ref{fig:idea_oscore_hyp4} show the resulting parameters and models.
 
@@
 
-The idea o-score models are significantly different in the $\alpha$ parameter, with mean o-scores of 0.995 and 0.997 for the $\leq20$ and $\textgreater20$ conditions respectively.
-The category o-scores are significantly different in both parameters, with mean o-scores of 0.975 and 0.982.
+The $\alpha$ parameter is significantly different in both the idea and category o-score comparison.
+The mean idea o-score is 0.997 in the $\textgreater20$ versus 0.995 in the $\leq20$ condition.
+Similarly, the $\leq20$ condition is favourable in category o-score with a mean of 0.982 vs 0.975.
 
-Participants generated more original ideas after their first 20 instances than they did after.
-Not only does this support the hypothesis, it provides a clear guideline for maximizing originality in brainstorming tasks on microtask marketplaces: it is more cost-effective to solicit $\textgreater20$ ideas from a smaller pool of participants than to ask for less from more.
+This support the hypothesis, it provides a clear guideline for maximizing originality in brainstorming tasks on microtask marketplaces: it is more cost-effective to solicit $\textgreater20$ ideas from a smaller pool of participants than to ask for less from more.
 
@@
 In this section, we examine these patterns in the context of hypotheses 5 and 6.
-Specifically, these hypotheses draw upon the SIAM model and predict that individuals will generate ideas from an idea category until they exhaust that category, at which point they switch to another category (hypothesis 5).
-Furthermore, SIAM predicts that this category switching will be observable as taking more time to generate an idea in a new category than in the existing category (hypothesis 6).
+Specifically, these hypotheses draw upon the SIAM model's conceptualization of brainstorming as a process in ideas are generated within one category image at a time.
 
-To determine category switching, we examine whether an idea was assigned to the same category tree as the previous, or whether it was assigned to a different category tree.
-We also examine the timing data collected from the web app to examine the time to generate an idea as a function of whether it is in the same category as the previous idea.
-
-Our data show clear evidence in support of both hypotheses. As an example, below is an example of a worker riffing on a single idea.
+In our data, we identify this pattern of creating new instances based on old ones as ``riffing'' (Table~\ref{tab:riffing}).
+The higher the number of instances requested, the more likely participants were to use old ideas as inspiration for new ideas (i.e., ``riff'' on previous ideas).
+However, these riff sequences are typically quite small in length, across conditions.
 
@@
     \caption{Riffing statistics by condition}
+    \label{tab:riffing}
 \end{table}
 
-Examining our data, we find that the higher the number of instances requested, the more likely participants are to use old ideas as inspiration for new ideas (i.e., ``riff'' on previous ideas).
-However, these riff sequences are typically quite small in length, across conditions.
+As an example of riffing, below is an excerpt of a particular participant's brainstorming run.
+
+\begin{quote}
+    \begin{enumerate}
+        \item recycle
+        \item take apart and refurbish
+        \item scrap it for any precious metals inside (some have gold contacts)
+    \end{enumerate}
+\end{quote}
+
+The participant first touches on the general idea of recycling.
+Then, they become more specific by reducing the scope of recyling to specific parts.
+Finally, they limit the scope of parts to precious metals, particularly gold.
+At this point, the participant exhausts their working memory of the category an moves on.
+
+SIAM predict that individuals will generate ideas from an idea category until they exhaust that category, at which point they switch to another category (hypothesis 5).
+Furthermore, SIAM predicts that this category switching will be observable as taking more time to generate an idea in a new category than in the existing category (hypothesis 6).
+
+\subsubsection{Category changing}
+
+To determine category switching, we examine whether an idea was assigned to the same category tree as the previous, or whether it was assigned to a different category tree.
 
@@
 
-To test hypothesis 6, we now examine the time it takes to generate a response instance.
+\subsubsection{Idea generation time}
+
+To test hypothesis 6, we now examine the time it takes to generate a response instance. We calculate this as the time from the first activation of the form widget to the last.
 
 An examination of the distribution of data for time spent per instance suggests a log-normal distribution to the data.
-Using STAN, we fit two models: $t_w = \text{lognormal}(\mu_w, \sigma_w)$ for within-category idea generation and $t_b = \text{lognormal}(\mu_b, \sigma_b)$ for between-category idea generation, where the $t$ parameter in each model represents the time spent generating an instance.
+We fit two models: $t_w = \text{lognormal}(\mu_w, \sigma_w)$ for within-category idea generation and $t_b = \text{lognormal}(\mu_b, \sigma_b)$ for between-category idea generation, where the $t$ parameter in each model represents the time spent generating an instance.
 
@@
     \caption{Time spent to generate an instance}
+    \label{fig:hyp5_comparison}
 \end{figure}
 
-The resulting fit models are shown in Figure FIG, overlapping the log-space histograms for observed values.
+The resulting fit models are shown in Figure~\ref{fig:hyp5_comparison}, overlapping the log-space histograms for observed values.
 The mean for the within-category condition was 9.5 in log space (13.4 seconds, 95\% HDI 9.4-9.6), while the mean for the between-categories condition was 10.2 (27.0 seconds, 95\% HDI 10.2-10.3).

### 2013-09-18 chi2014/conclusions.tex
fix quotations

@@
 Generality, however, decreased.
-In both cases, we found a "sweet spot" for this change at the 20 idea mark; participants' ideas reach their peak creativity and minimum generality after their 20th contribution.
+In both cases, we found a ``sweet spot" for this change at the 20 idea mark; participants' ideas reach their peak creativity and minimum generality after their 20th contribution.
 

### 2013-09-18 chi2014/abstract.tex
Abstract edit

@@
 We present a first characterization of brainstorming in this context, providing models for the rate of idea generation, the rate of category generation and the originality of generated ideas.
-We demonstrate that the rates of idea and category generation can be closely fit by a logarithmic model, and identify a \emph{burn in period} of common ideas at the beginning of a brainstorming task when people came up with the same overall set of ideas before diverging from this common set.
-Our results suggest a cutoff of this period at around the 20th idea.
+We demonstrate that the rates of idea and category generation can be closely fit by a logarithmic model, and show that the first few responses of a brainstorming run are drawn from a pool of very common, obvious ideas.
+Our results suggest a cutoff for this strategy 20th idea.
 Furthermore, we explore the phenomenon of \emph{riffing} in brainstorming tasks, and replicate two related findings from prior work.

### 2013-09-18 chi2014/results.tex

@@
 
-We fit the parameters to the data using a Bayesian approach \cite{kruschke_doing_201}, using the STAN modeling software to perform the actual fit \cite{stan-software:2013}.
+We fit the parameters to the data using a Bayesian approach \cite{kruschke_doing_2010}, using the STAN modeling software to perform the actual fit \cite{stan-software:2013}.
 In modeling the parameters, we use uninformative priors.

### 2013-09-18 chi2014/results.tex
New results section

--- 2013-09-18 chi2014/results.tex
@@
 Hierarchical clustering yielded an idea forest composed of 3007 instances, 1213 unique ideas, 335 category trees, and 129 non-singleton category trees (Table~\ref{tab:idea_forest}).
-In general, category trees were shallow, with a maximum tree height of 5.
 The number of nodes within a category tree ranged from 1 to 66, with the distribution of values roughly following a Poisson distribution.
-The number of instances per node (i.e., the number of equivalent idea instances) also roughly follows a Poisson distribution, with number of instances ranging from 1 to 63.
+The number of instances per node (i.e., the number of equivalent idea instances) also roughly follows a Poisson distribution, with the number of instances ranging from 1 to 63.
 \begin{table}
 	\begin{tabular}[h!]{r | l l l l l l l}
-	\textbf{number condition} & 5 & 10 & 20 & 50 \\ \hline \hline
+	\textbf{number requested} & 5 & 10 & 20 & 50 \\ \hline \hline
 	HITs & 57 & 47 & 23 & 10\\
@@
 	non-singleton trees & 28 & 34 & 40 & 48 \\ \hline \hline
-	\textbf{number condition} & 75 & 100 & all \\ \hline \hline
+	\textbf{number requested} & 75 & 100 & all \\ \hline \hline
     HITs & 10 & 10 & 146\\
@@
 \subsection{Rates of Idea and Category Production}
-The idea forest allows us to model the rate at which new ideas and new categories of ideas are produced, testing hypotheses 1 and 2 (i.e., the notion that rates of new ideas and categories of ideas will diminish over time, and will do so as a function condition).
+Using the fully clustered data set, we can model the rate at which unique ideas and unique categories of ideas are produced, testing hypothesis 1 (i.e., the notion that rates of unique ideas and categories of ideas are non-linear over time).
 
-Figure~\ref{fig:cumulative_ideas} shows the cumulative count of new ideas as a function of the number of instances gathered in the experiment.
-For reference, a perfect ``1:1'' idea generation line is also plotted.
+Figure~\ref{fig:cumulative_ideas} shows the cumulative count of unique ideas per condition, as a function of the number of instances gathered in the experiment.
+The plots for each condition were obtained by considering each response instance in the order in which it was received, with its uniqueness assessed relative to all response instances received before it.
+For reference, an ideal ``1:1'' idea generation line is also plotted.
 
-We model this quantity using the following logarithmic model:
-\[n_{ideas} = b_0 + b_{cond} * \log (x_{scale} * x + x_{offset})\]
-where $n_{ideas}$ represents the number of new ideas, $b_0$, $x_{scale}$, and $x_{offset}$ are parameters shared between conditions that serve to position the curve, and $b_{cond}$ acts as a per-condition scaling factor.
-This model does not explicitly model variability due to different workers, a point we return to later.
+We model this quantity using the following exponential model:
+\[n_{ideas} = y_{scale} * x^{rate}\]
+where $n_{ideas}$ represents the number of unique ideas, $y_{scale}$ serves to scale the model, $x$ is the response number, and $rate$ models the rate of novel ideas.
+Importantly, this simplified model does not model either the condition or the effects of workers.
+In our analysis of the data, we found that individual workers introduced considerable variation, preventing us from separating the effects of the condition from the worker.
+Accordingly, we only consider this simplified model.
 
-We fit the parameters to the data using a Bayesian approach \cite{kruschke_doing_201}, using the STAN modeling software to perform the actual fit \cite{stan-software:2013}.
-In modeling the parameters, we use uninformative priors.
-The resulting posterior distributions allow us to determine whether the conditions are significantly different from one another by determining whether the highest density intervals (HDIs) of $b_{cond}$'s posterior distributions overlap or not.
-In our analyses, we use an HDI of 95\%.
-
-Analysis of the HDIs of $b_{cond}$ reveals no significant differences between the 5 and 10 conditions, with all other conditions significantly different from one another except for the 50 and 100 conditions.
-[TODO: Table of fitted values and their HDIs.] Notable in our data is the 75 condition, which rises above the 100 condition in this plot.
-Further analysis reveals that a single worker is responsible for that additional rise in number of new ideas.
-Removing that worker from the data reveals that the 75 condition is not different from either the 50 or 100 conditions.
-
-Taking the derivative of the models allows us to examine the same data, but as rates at which new ideas are produced.
-As expected, rates of new ideas quickly taper off as a function of the number of ideas received.
+Fitting this model with Stan, we obtain a $y_{scale}$ value of 1.49 (HDI 1.44-1.55) and a $rate$ value of 0.87 (HDI 0.86-0.88).
+Because $rate$'s HDI does not include 1, we find support for hypothesis 1: the number of unique ideas is non-linear as a function of the number of responses received.
+The fitted model is shown as the ``Best Fit'' line in Figure~\ref{fig:cumulative_ideas}.
 
@@
 
-We can perform a similar analysis for the number of new idea categories that arrive over time.
-As would be surmised from the previous graphs, we observe a similar pattern, which can also be modeled using a logarithmic model.
-In the fitted model, we see that the number of new idea categories produced increases more slowly than the number of new ideas.
-This finding suggests that although individuals may produce new, unique ideas over time, these new ideas are within the \emph{same} categories.
-In other words, they are ``riffing'' -- coming up with permutations on existing ideas.
+We use an identical model for modeling the rate at which new idea categories are introduced over time.
+Figure~\ref{fig:cumulative_categories} shows the data and fitted model for the cumulative number of novel categories over time.
+For this model, we obtain a $y_{scale}$ value of 2.1 (HDI 2.0-2.2), with a $rate$ of 0.67 (HDI 0.66-0.67).
+Again, we find very clear support that the rate of novel categories is non-linear as a function of the number of responses received.
 
-Collectively, these results largely affirm hypotheses 1 and 2: The rates of new idea and new idea categories production quickly taper off and tend toward zero over time.
-Furthermore, we see differences in rates between conditions.
-In our data, we see three clear groups of rates: the 5 and 10 conditions, the 20 condition, and the 50, 75, and 100 condition.
+The visibly lower rate of growth for novel categories growth suggests workers are more likely to produce new ideas within existing categories, rather than create both novel ideas that also represent novel categories.
+We more closely examine characteristics of individuals' brainstorming runs next.
 
@@
 \subsection{Characteristics of an Idea Within a Run}
-We now turn our attention to individual runs, and characterize the ideas produced throughout a run.
 
 \subsubsection{Most Popular Ideas}
-Our results suggest that individuals typically produce ideas from a common set idea categories early in the brainstorming process, then diverge and produce more original ideas.
-This phenomenon can be observed by examining the o-scores of ideas and idea categories in Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order}, which show the distribution of o-scores for individual ideas and idea categories, as a function of response number, across all conditions.
+Our data suggest that individuals typically produce ideas from a common set of idea categories early in the brainstorming process, then diverge and produce more original ideas.
+This phenomenon can be observed by examining the o-scores of ideas and idea categories.
+
+Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order} show the o-scores for response instances and idea categories, within a sliding window of size 10.
+Each point represents the mean o-score within that sliding window, with standard error bars provided for each point.
 
@@
 
-In both instances, we observe a steady increase in o-scores for the first 20-30 responses.
-After these first 20-30 responses, the o-scores essentially plateau, with no clear discernible pattern for either ideas or idea categories.
+In both cases, we observe a steady increase in o-scores for the first 20-30 responses.
+After these first 20-30 responses, the o-scores plateau, with no clear discernible pattern for either ideas or idea categories.
 
-These graphs lend support for hypothesis 3, the notion that there is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.
+These graphs lend support for hypothesis 2, the notion that there is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.
 
 To more explicitly confirm this hypothesis, we extracted the top 5\% most common ideas (66 in total).
-If our intuition is correct, these ideas should be more likely to appear in the first five instances of a run than they are to appear in any of the following instances.
+If individuals are more likely to initially generate the same set of ideas, the top 5\% of the ideas should be more likely to appear in the first five instances of a run than they are to appear in any of the following instances.
 We introduce two Bernoulli random variables to model this.
-The first, defined by parameter $\theta_f$ represents the probability that a response instance is in the top 5\% most common ideas, given it is found in the first five instances of a run ($P(\theta_f|top)$).
-The second, defined by parameter $\theta_l$, is the probability that an instance is in the common pool given it is one of the \emph{last} ($\textgreater5$) instances in a run ($P(\theta_l|\overline{top})$).
+The first, defined by parameter $\theta_f$ represents the probability that a response instance is in the top 5\% most common ideas, given it is found in the first five instances of a run (i.e., $P(\theta_f|top)$).
+The second, defined by parameter $\theta_l$, is the probability that an instance is in the top \% given it is one of the \emph{last} ($\textgreater5$) instances in a run (i.e., $P(\theta_l|\overline{top})$).
 
@@
 We chose this constraint so that no run could contribute evidence to $\theta_f$ without also contributing to $\theta_l$.
-The resulting parameters were $\theta_f = 0.564$ (95\% HDI 0.517-0.611) and $\theta_l = 0.323$ (95\% HDI 0.303-0.342).
-This difference is significant, supporing the hypothesis that common ideas are over-represented in the early part of a brainstorming run.
+Using an uninformative beta prior, the resulting parameters are $\theta_f = 0.56$ (95\% HDI 0.52-0.61) and $\theta_l = 0.32$ (95\% HDI 0.30-0.34).
+This difference is significant, supporting the hypothesis that common ideas are over-represented in the early part of a brainstorming run.
 
@@
 
-The increase in o-score over time in Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order}) also supports hypothesis 4, suggest by Parnes et al \cite{parnes_effects_1961}, that ideas generated in the second half a brainstorming session have higher o-scores than those in the first half.
-We test a stronger guarantee implied by our data: that ideas after the first 20 have higher o-scores than the first 20.
-Although we expected the o-scores to be normally distributed in each condition, they were not.
-Instead, we fit Beta distributions to o-score for each of the $\leq20$, $\textgreater20$ conditions.
-Table~\ref{tab:hyp4} and Figure~\ref{fig:idea_oscore_hyp4} show the resulting parameters and models.
+The increase in o-score over time in Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order} also supports hypothesis 3, the notion that ideas generated in the second half a brainstorming session have higher o-scores than those in the first half.
+We test a stronger guarantee of this hypothesis that is implied by our data: that ideas after the first 20 have higher o-scores than the first 20.
 
-\begin{table}
-	\begin{tabular}[h!]{r | l l }
-	    & $\alpha$ median & $\alpha$ 95\% HDI \\ \hline \hline
-        idea o-score (\textless=20) & 135.86 & 122.39-149.72 \\
-        idea o-score (\textgreater20) & 259.84 & 240.47-278.81 \\ \hline \hline
-	    & $\beta$ median & $\beta$ 95\% HDI \\ \hline \hline
-        category o-score (\textless=20) & 1.16 & 1.08-1.25 \\
-        category o-score (\textgreater20) & 0.82 & 0.77-0.86 \\
-	\end{tabular}
-    \caption{Hypothesis 4 Beta models}
-    \label{tab:hyp4}
-\end{table}
+Partitioning the data into instances $\leq20$ and $\textgreater20$, we found that these distributions of o-scores are best modeled using beta distributions.
+Figures~\ref{fig:idea_oscore_hyp4} and \ref{fig:cat_oscore_hyp4} show the data and fitted models.
+Examining the HDIs of the resultant model parameters shows significant differences between the models, confirming hypothesis 3.
+
+Collectively, these results provide a clear guideline for maximizing originality in brainstorming tasks on microtask marketplaces: it is more cost-effective to solicit $\textgreater20$ ideas from a smaller pool of participants than to ask for fewer ideas from more individuals.
+
+%The $\alpha$ parameter is significantly different in both the idea and category o-score comparison.
+The mean idea o-score is 0.997 in the $\textgreater20$ versus 0.995 in the $\leq20$ condition.
+Similarly, the $\leq20$ condition is favourable in category o-score with a mean of 0.982 vs 0.975.
+
+% \begin{table}
+% 	\begin{tabular}[h!]{r | l l }
+% 	    & $\alpha$ median & $\alpha$ 95\% HDI \\ \hline \hline
+%         idea o-score (\textless=20) & 135.86 & 122.39-149.72 \\
+%         idea o-score (\textgreater20) & 259.84 & 240.47-278.81 \\ \hline \hline
+% 	    & $\beta$ median & $\beta$ 95\% HDI \\ \hline \hline
+%         category o-score (\textless=20) & 1.16 & 1.08-1.25 \\
+%         category o-score (\textgreater20) & 0.82 & 0.77-0.86 \\
+% 	\end{tabular}
+%     \caption{Hypothesis 4 Beta models}
+%     \label{tab:hyp4}
+% \end{table}
 
@@
 
-The $\alpha$ parameter is significantly different in both the idea and category o-score comparison.
-The mean idea o-score is 0.997 in the $\textgreater20$ versus 0.995 in the $\leq20$ condition.
-Similarly, the $\leq20$ condition is favourable in category o-score with a mean of 0.982 vs 0.975.
+\subsection{Idea Generation Patterns}
+Given the clear differences in idea characteristics as a function of response number, we now consider idea generation patterns observable from the data.
+In this section, we examine these patterns in the context of hypotheses 4 and 5, the hypotheses that draw upon the SIAM model of brainstorming.
+This model predicts individuals will generate ideas within ``images'' or categories of ideas (i.e., ``riff'' on an idea or concept) before switching to a new category.
 
-This support the hypothesis, it provides a clear guideline for maximizing originality in brainstorming tasks on microtask marketplaces: it is more cost-effective to solicit $\textgreater20$ ideas from a smaller pool of participants than to ask for less from more.
+In our clustered data, we can identify riffing by examining whether a response instance belongs to the same category tree as the previous instance.
+We can then derive a number of statistics from these data.
 
-\subsection{Idea Generation Patterns}
-Given the clear differences in characteristics of ideas as a function of response number, we now consider idea generation patterns observable from the data.
-In this section, we examine these patterns in the context of hypotheses 5 and 6.
-Specifically, these hypotheses draw upon the SIAM model's conceptualization of brainstorming as a process in ideas are generated within one category image at a time.
+To contextualize this process, below is an excerpt of a particular participant's brainstorming run.
 
-In our data, we identify this pattern of creating new instances based on old ones as ``riffing'' (Table~\ref{tab:riffing}).
-The higher the number of instances requested, the more likely participants were to use old ideas as inspiration for new ideas (i.e., ``riff'' on previous ideas).
-However, these riff sequences are typically quite small in length, across conditions.
+\begin{quote}
+    \begin{enumerate}
+        \item recycle
+        \item take apart and refurbish
+        \item scrap it for any precious metals inside (some have gold contacts)
+    \end{enumerate}
+\end{quote}
+
+The participant first touches on the general idea of recycling, then becomes more specific by reducing the scope of recyling to specific parts.
+Finally, they limit the scope of parts to precious metals, particularly gold.
+At this point, the participant exhausts their working memory of the category and moves on.
+
+Table~\ref{tab:riffing} provides a summary of riffing statistics observed in our data set.
+We find that the higher the number of instances requested, the more likely participants are to use old ideas as inspiration for new ideas.
+However, we observe that riff sequences are typically quite small in length, across conditions.
 
@@
 
-As an example of riffing, below is an excerpt of a particular participant's brainstorming run.
-
-\begin{quote}
-    \begin{enumerate}
-        \item recycle
-        \item take apart and refurbish
-        \item scrap it for any precious metals inside (some have gold contacts)
-    \end{enumerate}
-\end{quote}
-
-The participant first touches on the general idea of recycling.
-Then, they become more specific by reducing the scope of recyling to specific parts.
-Finally, they limit the scope of parts to precious metals, particularly gold.
-At this point, the participant exhausts their working memory of the category an moves on.
-
-SIAM predict that individuals will generate ideas from an idea category until they exhaust that category, at which point they switch to another category (hypothesis 5).
-Furthermore, SIAM predicts that this category switching will be observable as taking more time to generate an idea in a new category than in the existing category (hypothesis 6).
+SIAM predict that individuals will generate ideas from an idea category until they exhaust that category, at which point they switch to another category (hypothesis 4).
+Furthermore, SIAM predicts that this category switching will be observable as taking more time to generate an idea in a new category than in the existing category (hypothesis 5).
+We examine each of these hypotheses in turn.
 
@@
 
-To determine category switching, we examine whether an idea was assigned to the same category tree as the previous, or whether it was assigned to a different category tree.
+To determine category switching, we examine whether an idea is assigned to the same category tree as the previous instance, or whether it was assigned to a different category tree.
 
-To explicitly test hypothesis 5, we model the probability of any idea category being followed by the same idea category with a Bernoulli random variable.
-Using a uniform beta prior, we find the probability of a response instance being in the same idea category as the previous to be 15.6\% (95\% HDI 14.3\% - 17.0\%).
+We model the probability of any idea category being followed by the same idea category with a Bernoulli random variable.
+Using a uniform beta prior, we find the probability of a response instance being in the same idea category as the previous to be 15.6\% (HDI 14.3\% - 17.0\%).
 To establish a baseline to understand whether this probability is greater than chance, the most popular category tree in our data set is \emph{music player}, representing 5.65\% of all instances.
-Thus, the greatest probability of one instance being in the same category as the previous instance, by chance, is 5.65\%, well below the 14.3\% lower bound of the calculated HDI.
+Thus, an upper bound on the greatest probability of an instance being in the same category as the previous instance, by chance, is 5.65\%, well below the 14.3\% lower bound of the calculated HDI.
 Thus, we accept the hypothesis that an instance follows a previous instance in the same category with greater probability than is explained by random chance.
@@
 
-To test hypothesis 6, we now examine the time it takes to generate a response instance. We calculate this as the time from the first activation of the form widget to the last.
+To calculate the time to generate a response instance, we examine the times logged for the first and last activation of a text entry widget.
 
-An examination of the distribution of data for time spent per instance suggests a log-normal distribution to the data.
+An examination of the timing data suggests a log-normal distribution.
 We fit two models: $t_w = \text{lognormal}(\mu_w, \sigma_w)$ for within-category idea generation and $t_b = \text{lognormal}(\mu_b, \sigma_b)$ for between-category idea generation, where the $t$ parameter in each model represents the time spent generating an instance.
@@
 
-The resulting fit models are shown in Figure~\ref{fig:hyp5_comparison}, overlapping the log-space histograms for observed values.
-The mean for the within-category condition was 9.5 in log space (13.4 seconds, 95\% HDI 9.4-9.6), while the mean for the between-categories condition was 10.2 (27.0 seconds, 95\% HDI 10.2-10.3).
-The variances were 1.1 (3 seconds) and 1.2 (3.3 seconds) respectively.
-These difference were both sigificant.
-Within-category instance generation took on average 12.6 seconds less than between-categories.
-This is consistent with the findings of Nijstad and Stoebe, who reported differences of 6-12 seconds between conditions.
-Thus, we find support for hypothesis 6 and also show the applicability of the SIAM model to brainstorming in microtask marketplaces.
+The resulting models fitted to the data are shown in Figure~\ref{fig:hyp5_comparison}.
+The mean for the within-category condition was 9.5 in log space (13.4 seconds, HDI 9.4-9.6), while the mean for the between-categories condition was 10.2 (27.0 seconds, HDI 10.2-10.3).
+The differences in mean are significant and confirm hypothesis 5.
+Notably, within-category instance generation took an average of 12.6 seconds less than between-categories instance generation, consistent with the findings of Nijstad and Stoebe, who reported differences of 6-12 seconds between these two idea generation processes.
 

### 2013-09-18 chi2014/abstract.tex
Edit abstract

@@
 We demonstrate that the rates of idea and category generation can be closely fit by a logarithmic model, and show that the first few responses of a brainstorming run are drawn from a pool of very common, obvious ideas.
-Our results suggest a cutoff for this strategy 20th idea.
+Our results suggest that ideas become more original after 20 ideas have been collected.
 Furthermore, we explore the phenomenon of \emph{riffing} in brainstorming tasks, and replicate two related findings from prior work.
-Our results provide recommendations for those leveraging the crowd for brainstorming, and comparable models of creative idea generation research community.
+Our results provide recommendations for those leveraging the crowd for brainstorming, and comparable models of creative idea generation for the research community.
 

### 2013-09-18 chi2014/conclusions.tex

--- 2013-09-16 chi2014/conclusions.tex
@@
 
-We set out to establish a baseline for brainstorming performance in crowd marketplaces, a new setting for idea generation in with spatial and temporal co-presence are not guaranteed.
+In this work, we developed models of idea generation, category geneartion, originality, and individual practice in brainstorming on microtask markets.
+These models provide a baseline for both structure and performance in this context.
+They allow us to provide recommendations for idea solicitors, and starting points and comparison statistics for researchers.
+
+ We summarize our findings:
+
+\begin{itemize}
+\item The rate of idea and category generation decays non-linearly over time. 
+\item Common ideas are more likely to be proposed early in a run. 
+\item Conversely, ideas beyond the first 20 are rarer and more original, as measured by o-scores. This relationship also holds for idea categories.
+\item Ideas are more likely to be followed by another idea from the same category than would be expected by random chance
+\item The time it takes to generate an idea when changing categories is longer than the time when generating within a category
+\end{itemize}
+
+\subsection{Implications}
+
+From these, we draw several implications. First, that there is a saturation point in number of unique ideas or idea categories received, beyond which results will stagnate.
+
+anyone seeking creative brainstorming results in microtask environments should ask for more than 20 ideas per participant.
+
+Although idea o-score goes up over time, category o-score increases and then reaches a point of high variability.
+This implies a change in the strategy of how categories are used.
+It may be that beyond 20 ideas, high idea o-scores are maintained via a mix of remixing within existing categories (low category o-score) and establishing entirely new images (high category o-score).
+
+%We set out to establish a baseline for brainstorming performance in crowd marketplaces, a new setting for idea generation in with spatial and temporal co-presence are not guaranteed.
 We did this by establishing a corpus of brainstorming responses, publically available at URL.
@@
 
-We found that originality increased as brainstorming participants were asked to generate more ideas individually.
+%We found that originality increased as brainstorming participants were asked to generate more ideas individually.
 Generality, however, decreased.
-In both cases, we found a "sweet spot" for this change at the 20 idea mark; participants' ideas reach their peak creativity and minimum generality after their 20th contribution.
+In both cases, we found a ``sweet spot" for this change at the 20 idea mark; participants' ideas reach their peak creativity and minimum generality after their 20th contribution.
 
-Furthermore, we verified that the findings and predictions of prior work in which participants were spatially and temporally copresent.
+%Furthermore, we verified that the findings and predictions of prior work in which participants were spatially and temporally copresent.
 We found that, in line with the predictions of Jigstad and Stroebe's SIAM model \cite{nijstad_how_2006}, ideas are more likely to follow an idea of the same category than would be randomly expected, and that the time spent to generate an idea within-category is lesser than that to generate ideas between categories.
@@
 
-We identified three qualitative phenomena in of brainstorming results.
+%We identified three qualitative phenomena in of brainstorming results.
 First, that brainstorming results are subject to the popular consciousness of the time they are gathered.
@@
 
-The most salient recommendation we can make is easily summarized: if you use the crowd for brainstorming and desire originality, it is most productive to ask few participants for more than 20 ideas each.
+%The most salient recommendation we can make is easily summarized: if you use the crowd for brainstorming and desire originality, it is most productive to ask few participants for more than 20 ideas each.

### 2013-09-18 chi2014/future.tex

@@
 
-\subsection{Interventions}
+The most immediate work that follows from this is to examine the same phenomena for multiple questions.
+A design space for brainstorming taks must be defined and tested to see how variations impact the rate of idea generation, originality, etc.
 
-\begin{itemize}
-\item multi-phase brainstorming: introducing common ideas first to force workers more quickly into the originality zone
-\end{itemize}
+The saturation point of unique ideas and categories presents an obstacle to large-scale brainstorming on microtask marketplaces.
+One could conceivable develop interventions to overcome and reset this stagnation behaviour.
+For example, ideas from previous responders could be interspersed in response fields of the task to promote riffing or combination.
 
-\subsection{Predicting Turker performance}
+We identified that ideas are more likely to be followed by one from the same category, and that changing categories takes time.
+Combining these models with others (such a natural language measures for text similarity), one could construct a posterior belief model for whether a participant is currently riffing.
 
-\subsection{Automatic Idea Forest Generation}
+The change in idea and category o-scores at the 20 instance point is reflected in other metrics not discussed in this paper.
+It is likely this represents a distinct strategy change.
+By modelling this strategy change explicitly, we may be able to identify more explicitly when this change happens and make predictions about later strategies based on those employed early in the brainstorming run.

### 2013-09-18 chi2014/future.tex
Future work

--- 2013-09-16 chi2014/future.tex
@@
 
-\subsection{Interventions}
+%The most immediate step that follows this work is to examine the same phenomena for multiple questions.
+A design space for brainstorming taks must be defined and tested to see how variations impact the rate of idea generation, originality, etc.
 
-\begin{itemize}
-\item multi-phase brainstorming: introducing common ideas first to force workers more quickly into the originality zone
-\end{itemize}
+%The saturation point of unique ideas and categories presents an obstacle to large-scale brainstorming on microtask marketplaces.
+One could conceivably develop interventions to overcome and reset this stagnation behaviour.
+For example, ideas from previous responders could be interspersed in response fields of the task to promote riffing or combination.
 
-\subsection{Predicting Turker performance}
+The next step following this work is to explore interventions to overcome the saturation point of unique ideas and categories presents an obstacle to large-scale brainstorming on microtask marketplaces.
+For example, ideas from previous responders could be strategically interspersed in response fields of the task to promote riffing or exploration of a design space in certain directions.
 
-\subsection{Automatic Idea Forest Generation}
+Our results show that ideas are more likely to be followed by one from the same category, and that changing categories takes time.
+Combining these models with others (such as natural language measures for text similarity), one could construct a posterior belief model to guess whether a participant is currently riffing.
+
+Once the point of riffing and the current category a participant is working on can be identified, we can come up with different mechanisms for crowd to explore a design space.
+One could imagine a brainstorming session where participants are encouraged to generate a lot of ideas within new or rare categories or discouraged from exhausting categories that are popular.
+
+%The change in idea and category o-scores at the 20 instance point is reflected in other metrics not discussed in this paper.
+It is likely this represents a distinct strategy change.
+By modelling this strategy change explicitly, we may be able to identify more explicitly when this change happens and make predictions about later strategies based on those employed early in the brainstorming run.

### 2013-09-18 chi2014/introduction.tex
Fix typo

@@
 \begin{itemize}
-\item The rates at which novel ideas and categories of ideas are produced follow logarithmic growth models. Thus, the rates of novel ideas and categories of ideas decrease exponentially.
+\item The rates at which novel ideas and categories of ideas are produced follow an exponential model with a decaying growth rate.
+The rates of novel ideas and categories of ideas decrease exponentially.
 %\item The rates for generating novel ideas changes with the number of responses requested per individual, even if the overall number of responses is held constant.
@@
 In our data, we see the prevalence of riffing steadily increase across the first 20 responses, after which its prevalence plateaus.
-\item We also replicate Nijstaf and Stroebe's findings with respect to the time it takes to generate a new idea \cite{nijstaf_how_2006}.
+\item We also replicate Nijstaf and Stroebe's findings with respect to the time it takes to generate a new idea \cite{nijstad_how_2006}.
 \item Requesting ``as many ideas as possible'' for a fixed price (fixed for the entire HIT, rather than per idea) is inadvisable, as it leads to a relatively low number of ideas produced (in our experiments, fewer than 15).

### 2013-09-18 chi2014/conclusions.tex
Erased future work

@@
-\section{Conclusion}
+\section{Discussion and Future Work}
 
-In this work, we developed models of idea generation, category geneartion, originality, and individual practice in brainstorming on microtask markets.
-These models provide a baseline for both structure and performance in this context.
-They allow us to provide recommendations for idea solicitors, and starting points and comparison statistics for researchers.
+In this paper, we have primarily focused on the uniqueness of ideas and idea categories in quantitative terms.
+In this section, we provide a more qualitative perspective on the data collected, then describe future work.
 
- We summarize our findings:
+\subsection{Variability of Worker Input}
+While we observed that individuals are likely to draw from a common pool of ideas early on in the brainstorming process, there were outliers in the data.
+For example, one individual in the 75 ideas condition consistently produced some of most original ideas for old MP3 players:
 
 \begin{itemize}
-\item The rate of idea and category generation decays non-linearly over time. 
-\item Common ideas are more likely to be proposed early in a run. 
-\item Conversely, ideas beyond the first 20 are rarer and more original, as measured by o-scores. This relationship also holds for idea categories.
-\item Ideas are more likely to be followed by another idea from the same category than would be expected by random chance
-\item The time it takes to generate an idea when changing categories is longer than the time when generating within a category
+\item With displays - could be attached to a small solar panel (like the USB chargers w/multi attachments, including solar and car, etc.) and all the power could be sent to the screen as a small, portable light, then sent to countries lacking power in rural areas.
+\item They could be loaded with coordinates or other info and sealed in containers, then used in ocean current studies - once the container was collected they could be plugged in and read, and the data could be sent to whomever.
+\item They could be plugged in at a base camp at the bottom of Everest, and then potential climbers could record their name, loved ones' address, and any last words before attempting the climb. 
 \end{itemize}
+
+This individual's responses were responsible for clearly raising the observed uniqueness of ideas in the 75 condition when plotted against the other conditions.
+
+Other individuals filled their quotas by creating uninspiring idea templates that they varied only slightly between responses. For example:
+
+\begin{itemize}
+\end{itemize}
+
+\subsection{Future Work}
+The most immediate work that follows from this is to examine the same phenomena for multiple questions.
+A design space for brainstorming taks must be defined and tested to see how variations impact the rate of idea generation, originality, etc.
+
+The saturation point of unique ideas and categories presents an obstacle to large-scale brainstorming on microtask marketplaces.
+One could conceivable develop interventions to overcome and reset this stagnation behaviour.
+For example, ideas from previous responders could be interspersed in response fields of the task to promote riffing or combination.
+
+We identified that ideas are more likely to be followed by one from the same category, and that changing categories takes time.
+Combining these models with others (such a natural language measures for text similarity), one could construct a posterior belief model for whether a participant is currently riffing.
+
+The change in idea and category o-scores at the 20 instance point is reflected in other metrics not discussed in this paper.
+It is likely this represents a distinct strategy change.
+By modelling this strategy change explicitly, we may be able to identify more explicitly when this change happens and make predictions about later strategies based on those employed early in the brainstorming run.
 
@@
 %The most salient recommendation we can make is easily summarized: if you use the crowd for brainstorming and desire originality, it is most productive to ask few participants for more than 20 ideas each.
+
+\section{Conclusion}
+
+In this work, we developed models of idea generation, category geneartion, originality, and individual practice in brainstorming on microtask markets.
+These models provide a baseline for both structure and performance in this context.
+They allow us to provide recommendations for idea solicitors, and starting points and comparison statistics for researchers.
+
+ We summarize our findings:
+
+\begin{itemize}
+\item The rate of idea and category generation decays non-linearly over time. 
+\item Common ideas are more likely to be proposed early in a run. 
+\item Conversely, ideas beyond the first 20 are rarer and more original, as measured by o-scores. This relationship also holds for idea categories.
+\item Ideas are more likely to be followed by another idea from the same category than would be expected by random chance.
+\item The time it takes to generate an idea when changing categories is longer than the time when generating within a category.
+\end{itemize}

### 2013-09-18 chi2014/introduction.tex
spellcheck

@@
 \begin{itemize}
-\item The rates at which novel ideas and categories of ideas are produced follow an exponential model with a decaying growth rate.
-The rates of novel ideas and categories of ideas decrease exponentially.
+\item The rates at which novel ideas and categories of ideas are produced follow an exponential model. The rates of novel ideas and categories of ideas decrease exponentially.
 %\item The rates for generating novel ideas changes with the number of responses requested per individual, even if the overall number of responses is held constant.
@@
 Given our results, we find there is little value in requesting fewer than 20 ideas from workers, since these early ideas are among the most common.
-\item Individuals' patterns of idea generation mimic those identified by Nijstaf and Stroebe's SIAM (search for ideas in associative memory) model \cite{nijstad_how_2006}: Workers produce a set of ideas from working memory, then switch strategies to generate ideas that are permutations of other ideas, a process we dub {\em riffing\/}.
-In our data, we see the prevalence of riffing steadily increase across the first 20 responses, after which its prevalence plateaus.
-\item We also replicate Nijstaf and Stroebe's findings with respect to the time it takes to generate a new idea \cite{nijstad_how_2006}.
+\item Individuals' patterns of idea generation mimic those identified by Nijstad and Stroebe's SIAM (search for ideas in associative memory) model \cite{nijstad_how_2006}: Workers produce a set of ideas from working memory, then switch strategies to generate ideas that are permutations of other ideas, a process we dub {\em riffing\/}.
+\item We also replicate Nijstad and Stroebe's findings with respect to the time it takes to generate a new idea \cite{nijstad_how_2006}.
 \item Requesting ``as many ideas as possible'' for a fixed price (fixed for the entire HIT, rather than per idea) is inadvisable, as it leads to a relatively low number of ideas produced (in our experiments, fewer than 15).

### 2013-09-18 chi2014/method.tex
spellcheck

@@
 \section{Method}
-In this section, we describe our process of recruiting and remunerating participants, the brainstorming question selected, and the user interface provided the worker.
-We then describe the data collected, and our method of organizing it into a hierarchical cluster.
+In this section, we first describe our process of recruiting and remunerating participants, the brainstorming question selected for our experiments, and the user interface provided to workers.
+We then describe the data collected, and our method of organizing it into a hierarchical cluster for analysis.
 
-\subsection{Recruitment and Remuneration}
+/subsection{Recruitment and Remuneration}
 To study brainstorming in a crowd environment, we published brainstorming HITs (human intelligence tasks) on Amazon's Mechanical Turk.
@@
 
-\subsection{Brainstorming Question}
-In conducting this research, we tested several brainstorming questions, including classic brainstorming questions (e.g., ``provide alternative uses for a mop''), a request to brainstorm factors that influence weight gain and weight loss, and more goal-oriented questions, such as one requesting people to brainstorm ways to raise funds for a charity.
+\subsection{Brainstorming Question: Uses for an Old MP3 Player}
+In conducting this research, we tested several brainstorming questions, including classic brainstorming questions (e.g., ``provide alternative uses for a mop''); a request to brainstorm factors that influence weight gain and weight loss; and more goal-oriented questions, such as one requesting people to brainstorm ways to raise funds for a charity.
 Ultimately, we found the goal-oriented brainstorming questions to be the most compelling since they reflect what we believe would be real-world uses of brainstorming on Mechanical Turk.
 
-The final brainstorming question we chose for this work is the following:
+The final brainstorming question we chose for this research is the following:
 \begin{quote}
@@
 
-We chose to focus on this problem because a) we expected workers to be familiar with problem area (old MP3 players), b) it represents a real-world problem for which innovative solutions are welcome (namely, what to do with old electronics that are still partially functional), and c) we anticipated there to be a wide range of creative solutions possible.
+We chose to focus on this problem because a) we expected workers to be familiar with problem area (old MP3 players), b) it represents a real-world problem (what to do with old electronics that are still partially functional) for which innovative solutions are welcome, and c) we anticipated there to be a wide range of creative solutions possible.
 All data presented in this paper refer only to responses received to this particular question.
@@
 Under more realistic, real-world conditions, workers would be expected to complete all responses to receive compensation.
-As we will report later, the incidence of workers abusing this policy was relatively low.
 
-The brainstorming task was presented in a standard web-based form that included an introduction, the specific brainstorming problem, a series of fields for entering the individual responses, and a larger text area where workers could enter additional ideas (Figure~\ref{fig:sample_task}).
+The brainstorming task was presented in a standard web-based form that included the following components: an introduction, the specific brainstorming problem, a series of fields for entering the individual responses, and a larger text area where workers could enter additional ideas (Figure~\ref{fig:sample_task}).
 
@@
 
-In addition to the visible input fields, HITs were instrumented to collect timing data.
+In addition to the visible text input fields, HITs were instrumented to collect timing data.
 Specifically, we recorded the time of the first activation and last de-activation of each form element, the time the HIT was accepted, and the time it was submitted.
@@
 
-Initially, we sought 10 HITs for each condition.
-However, we found that collecting only 10 HITs for the lower response conditions (5, 10, 20 responses) limited our ability to make meaningful comparisons between responses received from the larger response conditions (50, 75, 100).
-Accordingly, we sought to receive a minimum of 400 ideas for each condition.
-We found the 5 and 10 conditions to be particularly unpopular, and were not able to reach this minimum.
+Initially, we sought 10 HIT responses per condition.
+However, we found that collecting only 10 responses for the lower response conditions (5, 10, 20 responses) limited our ability to make meaningful comparisons between responses received from the larger response conditions (50, 75, 100).
+Accordingly, we sought to collect a minimum of 400 ideas for each condition.
+However, we found the 5 and 10 conditions to be particularly unpopular, and were not able to reach this minimum for these conditions.
 
@@
 
-If a worker completed multiple HITs (i.e., generated responses for multiple conditions), there is the risk their responses and idea generation processes evolved over time, making it difficult to compare their responses in the second HIT with responses from others who are performing the same HIT without any previous exposure to the problem.
+If a worker completed multiple HITs (i.e., generated responses for multiple conditions), there is the possibility that their responses and idea generation processes evolved over time, making it difficult to compare the responses to their second HIT with responses from others who are performing the same HIT for the first time.
 Accordingly, we retain only the first HIT submitted by each worker.
@@
 \subsection{Hierarchical Clustering}
+One of the primary goals of this research is to model the rate at which novel ideas and categories of ideas arrive.
+A significant challenge to creating these models is efficiently identifying what ideas and categories of ideas are similar to one another, given the volume of data collected.
+To address this problem, we organized the data into sets of hierarchical clusters, where the clusters encode similarity relationships between ideas.
+Specifically, parent nodes are more generalized versions of child nodes.
+With this information encoded in the clusters, we are then able to calculate quantities such as the rate of novel ideas or novel idea categories, as a function of responses received.
 
-To analyze the data, we organized it into sets hierarchical clusters, where the clusters encode similarity relationships between ideas.
-Specifically, parent nodes are more generalized versions of child nodes.
-This clustering enables a wide range of data analysis.
-We introduce specific terminology, motivate the technique, then describe the clusters and the process used to create them.
+In the rest of this section, we introduce terminology useful for describing these hierarchical clusters, then describe the clusters and the process used to create them.
 
 \subsubsection{Clustering: Terminology}
-An \emph{instance} is a single solution by a single participant to a brainstorming question.
+An \emph{instance} or \emph{response instance} is a single solution by a single participant to a brainstorming question.
 Most concretely, it is the text entered into a single field in the web form.
-For example, four \emph{instances} from the data set:
+For example, the following are four \emph{instances} from our data set:
 
@@
 Multiple instances of the same idea do not need to be identically phrased to refer to the same idea.
-For example, in Figure FIG, we consider ``Storage Container'' and ``Small Storage Box'' as two instances of the same \emph{idea}.
+For example, in Figure~\ref{fig:example_instances}, we consider ``Storage Container'' and ``Small Storage Box'' as two instances of the same \emph{idea}.
 However, we do not consider ``Coin Storage'' to be the same idea, since it provides a more specific use case than the previous two.
@@
 Despite the differences in specificity between ``Coin Storage'' and the more generic containers, it is useful to encode the fact that these ideas are related.
-We use hierarchical clustering to encode that relationship.
-We call a hierarchical cluster of ideas a \emph{category trees}.
+We use hierarchical clusters to encode these relationships.
+We call a hierarchical cluster of ideas a \emph{category tree}.
 
@@
 
-\subsubsection{Clustering: Motivation}
+% \subsubsection{Clustering: Motivation}
 
-Initially, a coding scheme was employed in which a master list of unique ideas was formed, and each response was assigned a categorical code from this list.
+% Initially, a coding scheme was employed in which a master list of unique ideas was formed, and each response was assigned a categorical code from this list.
 It proved difficult to determine and adhere to a constant granularity of distinction.
@@
 
-A category tree allows the encoding of generalization relationships such that useful metrics can be extracted without any assumptions about distance.
-For example, The number of descendants of an idea gives a measure of how broadly that node was explored. 
-The prevalance of similar responses can be approximated at multiple levels; the number of instances under a tree represents the popularity of a category, while the same for an individual node provides a weighted score of ideas that are exactly equivalance.
-An idea can be considered more closely related to its adjacent ideas than it is to all other ideas in the tree.
-Using this common structure to operationalize originality and uniqueness ensures consistency in analysis.
+% A category tree allows the encoding of generalization relationships such that useful metrics can be extracted without any assumptions about distance.
+% For example, The number of descendants of an idea gives a measure of how broadly that node was explored. 
+% The prevalance of similar responses can be approximated at multiple levels; the number of instances under a tree represents the popularity of a category, while the same for an individual node provides a weighted score of ideas that are exactly equivalance.
+% An idea can be considered more closely related to its adjacent ideas than it is to all other ideas in the tree.
+% Using this common structure to operationalize originality and uniqueness ensures consistency in analysis.
 Furthermore, the category forest relies primarily on generalization decisions for construction, which are more clearly defined than equivalence.
@@
 \subsubsection{Clustering: Process}
-To produce the idea forest, the primary author added instances to the idea forest one at a time using the following algorithm.
+To produce the idea forest, the primary author added instances to the idea forest one at a time using the following method.
 
@@
 In some cases, two ideas were clearly related, though no generalization of the two existed in the data set.
-In these cases, we introduced a node that is a generalization of the two.
+In these cases, we introduced a node that was a generalization of the two.
 For example, if we received only ``Coin Storage'' and ``Travel Jewelry Case'' as ideas, we would create a node labeled ``Storage Case'' to capture the fact that the two are related concepts.
@@
 
-Clustering was performed using a custom-built tool that assisted in finding instances that were similar to a given instance.
-Briefly, the tool represented each instance using a bag of words model which included synonyms for words.
-The words were stemmed, then individual instances were ranking according to a cosine similarity metric.
+Clustering was performed using a custom-built tool that assisted in finding similar instances.
+Briefly, the tool represented each instance using a bag-of-words model that included synonyms for instance's words.
+The words were stemmed, then individual instances were ranked according to a cosine similarity metric.
 
@@
 A researcher not involved in the project randomly sampled 40 category trees to assess whether ideas within the tree should be there, and whether parent-child relationships correctly encoded generalizations of concepts.
-10\% (HDI 6-14) of instances were marked as specific than their associated idea.
-0.7\% (HDI 1.3-2.8\%) of instances were marked as not generalized by their parents.
-No significant effect was found between conditions for probability of either error.
+10\% of the instances were marked as more specific than their associated idea, while 0.7\% of the instances were judged to be misclassified with their associated cluster.
 
 \subsection{Data Analysis}
-Throughout this paper, we will fit models to data and give confidence metrics for the parameters of those models.
-Our analysis techniques are based on Kruschke's \emph{Doing Bayesian Data Analysis: A Tutorial in R and Bugs} \cite{kruschke_doing_2010}.
-We express our confidence in parameters in terms of the 95\% HDI (highest density interval), or the interval which contains 95\% of the mass of our posterior belief distribution.
-
-To fit the models, we use the Bayesian inference software Stan \cite{stan-software:2013} via the package RStan.
+Throughout this paper, we employ Bayesian data analysis methods for modeling the data \cite{kruschke_doing_2010}.
+We express our results in terms of mean values and highest density intervals (HDI), or the interval that contains the most probable mass of the posterior distributions.
+In line with common practice, we choose an HDI of 95\%.
+To fit the models, we use the Bayesian inference software Stan \cite{stan-software:2013}, and choose uninformative priors for all modeled variables.
 

### 2013-09-18 chi2014/related.tex
spellcheck

@@
 Flexibility is often measured by the number of explored ``categories'' in a design space, or by subjective rating from domain experts \cite{lewis2011affective, marsh1996examples}.
-Category or feature-based quantities also yield other metrics such as within-catgories fluency and the length of train-of-thought when participants focus on a particular category \cite{nijstad_how_2006}.
+Category or feature-based quantities also yield other metrics such as within-categories fluency and the length of train-of-thought when participants focus on a particular category \cite{nijstad_how_2006}.
 Building in this past line of work, we consider both individual ideas and categories of ideas.
@@
 These low quality responses were not found in walk-in brainstorming studies.
-For exapmle, Bernstein et al.
+For example, Bernstein et al.
 found high variance of effort from Amazon Mechanical Turk workers \cite{soylent}.

### 2013-09-18 chi2014/results.tex
Erased future work
spellcheck

@@
 
-The participant first touches on the general idea of recycling, then becomes more specific by reducing the scope of recyling to specific parts.
+The participant first touches on the general idea of recycling, then becomes more specific by reducing the scope of recycling to specific parts.
 Finally, they limit the scope of parts to precious metals, particularly gold.
@@
 The differences in mean are significant and confirm hypothesis 5.
-Notably, within-category instance generation took an average of 12.6 seconds less than between-categories instance generation, consistent with the findings of Nijstad and Stoebe, who reported differences of 6-12 seconds between these two idea generation processes.
+Notably, within-category instance generation took an average of 12.6 seconds less than between-categories instance generation, consistent with the findings of Nijstad and Stroebe, who reported differences of 6-12 seconds between these two idea generation processes.
 

### 2013-09-18 chi2014/abstract.tex

@@
 We present a first characterization of brainstorming in this context, providing models for the rate of idea generation, the rate of category generation and the originality of generated ideas.
-We demonstrate that the rates of idea and category generation can be closely fit by a logarithmic model, and show that the first few responses of a brainstorming run are drawn from a pool of very common, obvious ideas.
+We demonstrate that the rates of idea and category generation can be closely fit by a exponential model, and show that the first few responses of a brainstorming run are drawn from a pool of very common, obvious ideas.
 Our results suggest that ideas become more original after 20 ideas have been collected.

### 2013-09-18 chi2014/conclusions.tex

@@
 
-Other individuals filled their quotas by creating uninspiring idea templates that they varied only slightly between responses. For example:
+Other individuals filled their quotas by creating uninspiring idea templates that they varied only slightly between responses.
+For example, this individual provides a list of essentially equivalent responses:
 
 \begin{itemize}
+\item a gift for a cousin
+\item a gift for a niece
+\item a gift for a nephew
+\item a gift for a second cousin
 \end{itemize}
 
+This variability in worker responses prevented us from modeling any potential effects that the number of questions requested might have on individuals' responses.
+Thus, while it is clear that there is little advantage to requesting 5 or 10 ideas from workers, it is not clear whether it is preferable to request 20, 50, 75, or 100 ideas.
+We suspect that requesting 50 or 75 is the ``sweet'' spot, but are not able to empirically support that recommendation at this time.
+
 \subsection{Future Work}
-The most immediate work that follows from this is to examine the same phenomena for multiple questions.
-A design space for brainstorming taks must be defined and tested to see how variations impact the rate of idea generation, originality, etc.
+The results of this research suggest several avenues for future work.
 
-The saturation point of unique ideas and categories presents an obstacle to large-scale brainstorming on microtask marketplaces.
-One could conceivable develop interventions to overcome and reset this stagnation behaviour.
-For example, ideas from previous responders could be interspersed in response fields of the task to promote riffing or combination.
+One of the most natural next steps is to explore interventions to overcome the tendency of workers to generate the most common and obvious ideas and idea categories.
+For example, ideas from previous workers could be strategically interspersed in response fields to promote riffing or exploration of a design space in certain directions.
 
-We identified that ideas are more likely to be followed by one from the same category, and that changing categories takes time.
-Combining these models with others (such a natural language measures for text similarity), one could construct a posterior belief model for whether a participant is currently riffing.
+Our results show that ideas are more likely to be followed by one from the same category, and that changing categories takes time.
+Combining these models with others (such as natural language measures for text similarity), one could imagine constructing a posterior belief model to infer whether a participant is currently riffing or not.
+If we can infer riffing and the general categories in which workers are riffing, one could imagine techniques to guide the crowd to explore different areas of the design space.
+For example, the system could encourage workers to generate ideas within new or rare categories.
 
-The change in idea and category o-scores at the 20 instance point is reflected in other metrics not discussed in this paper.
-It is likely this represents a distinct strategy change.
-By modelling this strategy change explicitly, we may be able to identify more explicitly when this change happens and make predictions about later strategies based on those employed early in the brainstorming run.
-
-\subsection{Implications}
-
-From these, we draw several implications. First, that there is a saturation point in number of unique ideas or idea categories received, beyond which results will stagnate.
-
-anyone seeking creative brainstorming results in microtask environments should ask for more than 20 ideas per participant.
-
-Although idea o-score goes up over time, category o-score increases and then reaches a point of high variability.
-This implies a change in the strategy of how categories are used.
-It may be that beyond 20 ideas, high idea o-scores are maintained via a mix of remixing within existing categories (low category o-score) and establishing entirely new images (high category o-score).
+In addition to studying potential interventions, we are also interested in whether there are differences in idea generation for different types of questions.
+For example, the results observed for the MP3 player condition may differ from questions for which the workers have more or less familiarity.
+For example, it would be interesting to see how the idea generation process differs if workers were asked for suggestions for improving Amazon's Mechanical Turk interface.
 
@@
 
-In this work, we developed models of idea generation, category geneartion, originality, and individual practice in brainstorming on microtask markets.
-These models provide a baseline for both structure and performance in this context.
-They allow us to provide recommendations for idea solicitors, and starting points and comparison statistics for researchers.
+In this work, we developed models of idea generation, category generation, originality, and individual practices for brainstorming on microtask markets.
+These models provide guidance and a baseline for both practitioners and researchers alike.
 
- We summarize our findings:
+To summarize, we found that:
 
@@
 \item Ideas are more likely to be followed by another idea from the same category than would be expected by random chance.
-\item The time it takes to generate an idea when changing categories is longer than the time when generating within a category.
+\item The time it takes to generate an idea when changing categories is longer than the time taken when generating ideas within a category.
 \end{itemize}
+
+The most obvious implication from these findings is that it is advisable to request more than 20 ideas from each participant; requesting 20 or fewer ideas is not cost effective.
+
+With respect to future research, this work suggests a need for interventions that can help people escape common areas of the design space.
+For example, systems that can help people transition out of familiar idea categories into less popular idea categories are likely to lead to more unique and novel ideas.

### 2013-09-18 chi2014/future.tex

@@
-\section{Future Work}
+%\section{Future Work}
 
@@
 
-The next step following this work is to explore interventions to overcome the saturation point of unique ideas and categories presents an obstacle to large-scale brainstorming on microtask marketplaces.
-For example, ideas from previous responders could be strategically interspersed in response fields of the task to promote riffing or exploration of a design space in certain directions.
-
-Our results show that ideas are more likely to be followed by one from the same category, and that changing categories takes time.
-Combining these models with others (such as natural language measures for text similarity), one could construct a posterior belief model to guess whether a participant is currently riffing.
-
-Once the point of riffing and the current category a participant is working on can be identified, we can come up with different mechanisms for crowd to explore a design space.
-One could imagine a brainstorming session where participants are encouraged to generate a lot of ideas within new or rare categories or discouraged from exhausting categories that are popular.
-
 %The change in idea and category o-scores at the 20 instance point is reflected in other metrics not discussed in this paper.

### 2013-09-18 chi2014/introduction.tex
small edit

@@
 \begin{itemize}
-\item The rates at which novel ideas and categories of ideas are produced follow an exponential model. The rates of novel ideas and categories of ideas decrease exponentially.
+\item The rates at which novel ideas and categories of ideas are produced follow an exponential model, decreasing exponentially over time.
 %\item The rates for generating novel ideas changes with the number of responses requested per individual, even if the overall number of responses is held constant.

### 2013-09-18 chi2014/abstract.tex
edits

--- 2013-09-18 chi2014/abstract.tex
@@
 Microtask marketplaces enable hybrid computational systems that use crowds to solve problems of creativity, such as brainstorming.
-We present a first characterization of brainstorming in this context, providing models for the rate of idea generation, the rate of category generation and the originality of generated ideas.
-We demonstrate that the rates of idea and category generation can be closely fit by a logarithmic model, and show that the first few responses of a brainstorming run are drawn from a pool of very common, obvious ideas.
-Our results suggest a cutoff for this strategy 20th idea.
-Furthermore, we explore the phenomenon of \emph{riffing} in brainstorming tasks, and replicate two related findings from prior work.
-Our results provide recommendations for those leveraging the crowd for brainstorming, and comparable models of creative idea generation research community.
+We present a first characterization of brainstorming in this context, providing models for the rate of idea generation, the rate of category generation, and the uniqueness of generated ideas.
+We demonstrate that the rates of idea and category generation can be closely fit with an exponential model.
+We also show that the first few responses of a brainstorming run are drawn from a pool of very common, obvious ideas, and become more original after 20 ideas have been produced by an individual.
+Our data also show support for the SIAM model (search for ideas in associative memory) \cite{nijstad_how_2006} in which workers draw ideas from a single category before moving on to a new category.
+Our results provide clear recommendations for practitioners and researchers alike who wish to leverage the crowd for brainstorming.
 

### 2013-09-19 chi2014/method.tex
Fixed reversed slash on subsection

@@
 
-/subsection{Recruitment and Remuneration}
+\subsection{Recruitment and Remuneration}
 To study brainstorming in a crowd environment, we published brainstorming HITs (human intelligence tasks) on Amazon's Mechanical Turk.

### 2013-09-30 chi2014/qualitative.tex
spellcheck
strategies section

@@
 
-I'm not going into any of these in detail because I'm not sure if they should go in.
+In addition to our quantitative analysis of the rate of idea generation, we also performed a qualitative analysis.
+We selected 30 brainstorming runs (5 from each condition).
+The primary author applied open coding to identify characteristics of responses and runs that are independent of the uniqueness and originality measures observed in the quantitative section of this paper.
+The characteristics primarily took the form of \emph{strategies} employed to different degrees by different participants.
+We constructed \emph{personas} around the use of these strategies.
 
-\subsection{The Miley Cyrus affect}
+\subsection{Strategies}
+
+A \emph{strategy} is an identifiable mechanism for idea generation that results in several ideas in a brainstorming run.
+A key component of a strategy is that it be present in some runs but not in others.
+
+One strategy employed was \emph{feature exploration}.
+An MP3 player may have many features that distinguish it from any other electronic device of the same physical proportions: audio output (and potentially input), data storage, a screen, general purpose computing abilities (apps), connectivity (WiFi, USB), a camera, a unique appearance, GPS.
+Many participants treated the MP3 player as a pocket-sized brick when generating ideas.
+In contrast, the feature use strategy led to responses such as:
 
 \begin{itemize}
-\item last batch of HITs had several references to Miley Cyrus
-\item previously, this pop culture reference never came up
-\item suggests the pool of available ideas is time-sensitive
+    \item use them in museums to give information on various installations (audio output)
+    \item use in a scavenger hunt, device must be plugged in at another location to see clue
 \end{itemize}
 
-\subsection{Reward vs Relative Reward in Completion Time}
+Runs utilizing the feature exploration strategy produce responses that seem intuitively better than those that do not.
+It is satisfying to have a use for a device that "fits" the device as closely as possible, so that none of the functionality is "wasted".
+Despite this, the proportion of runs utilizing feature exploration seemed roughly constant across all other strategies, and the strategy does not form a basis for discrimination between any personas.
+
+The second strategy identified is \emph{riffing}.
+While riffing was examined earlier in relation to time spent on idea generation, we found in our qualitative analysis that it was manifested in many ways.
+\emph{Specificity riffing} occurs when a participant generates two or more ideas and one is a generalization of the other.
+For example, two consecutive ideas given by a participant:
 
 \begin{itemize}
-\item HITs with more overall reward perform faster, despite same payout per idea
-\item design implications
+    \item brick
+    \item building material
 \end{itemize}
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.9\columnwidth]{batch_time}
-    \caption{Time to complete batch of HITs}
-\end{figure}
-
-\subsection{Personas}
+Another, more common type of riffing is \emph{exhaustive mutation}.
+Under this strategy, a participant repeats a response nearly verbatim many times, each time replacing a single word or concept in such a way that no new ideas are formed.
+For example:
 
 \begin{itemize}
-\item Comment on the effect of single participants - the zoomed in 75'er effect.
+    \item wrap yarn around it
+    \item wrap twine around it
+    \item wrap string around it
 \end{itemize}
+
+\emph{Exploratory riffing} is a strategy in which participants hold at least one element of a previous response constant when generating a new response, such that a new idea is formed.
+This can be chained many times in a kind of idea "telephone game" such that the final ideas no longer resemble the original source.
+For example:
+
+\begin{itemize}
+    \item Jukebox music selector in bars
+    \item Commercials in bar bathrooms (retains bar element)
+    \item Tapper handles for beer (retain alcohol element)
+\end{itemize}
+
+Finally, \emph{Follow-up riffing} occurs when a participants response assumes that the solution in one of their previous responses has already been implemented. For example:
+
+\begin{itemize}
+    \item I suppose you could just grind them down into a sand
+    \item You could take the sand... and put it in an hourglass
+\end{itemize}
+
+Follow-up riffing, while interesting, is incredibly rare in the data and we will not discuss it further.
+
+There are two special variants of riffing that can occur across all riffing strategies.
+Some participants \emph{pair} their riffs, such that they often riff on old ideas but never produce more than two versions of each idea.
+
+Another special case is is \emph{linguistic riffing}. Some participants seem to use common sentence structure as queues to help them generate ideas. For example:
+
+\begin{itemize}
+    \item Make sunglasses out of them
+    \item Make shoes out of them
+    \item Make a handbag out of them
+\end{itemize}
+
+Note that linguistic riffing and exhaustive mutation have significant overlap. A participant that is employing exhaustive mutation is always employing linguistic riffing, but not vice-versa.
+
+Finally, \emph{reaching back} is a variant of riffing employed by almost all participants who provided more than 10 responses.
+Reaching back entails riffing on an idea that occured further back in the run than the previous response.
+The presence of reaching back speaks to the exaustion of the participant's creativity, as they produce new variants of old ideas rather than new ideas altogether.
+
+Almost all participants employ some form of riffing.
+This is not surprising, as it is ingrained in the rules of brainstorming given at the beginning of the task: "combinations of ideas count as new ideas".
+However, the different sub-strategies of riffing have very different impacts on the qualitative assessment of a run.
+Exhaustive mutation is often unsatisfying, as it feels as though the participant is trying to wring every scrap of task completion out of a single idea.
+On the other hand, exploratory riffing can produce answers that are incredibly unexpected but still of high quality.
+Riffing strategies paint a significant picture of the persona of the participant.
+
+Two time-saving strategies were also identified.
+The first, perhaps overly obvious strategy, is \emph{brevity}.
+This is simply the act of keeping responses as concise as possible, presumably in the interest of finishing quickly.
+The second time-saving strategy was \emph{cop-outs}.
+Cop-outs occur when participants choose to resort to vague and unsatisfying responses that do not actually provide any solution to the problem.
+For example, one participant's response to the MP3 player problem was to "Remove glass screen to make something".
+Cop-out responses may provide one or more steps towards a solution, but no solution.

### 2013-09-30 chi2014/qualitative.tex
personas

@@
 The second time-saving strategy was \emph{cop-outs}.
-Cop-outs occur when participants choose to resort to vague and unsatisfying responses that do not actually provide any solution to the problem.
+Cop-outs occur when participants choose to resort to vague or iunrelated responses that do not actually provide any solution to the problem.
 For example, one participant's response to the MP3 player problem was to "Remove glass screen to make something".
 Cop-out responses may provide one or more steps towards a solution, but no solution.
+
+\subsection{Personas}
+
+We use personas to capture collections of strategy use we see repeated across participants.
+While some strategies proved discriminatory, others were used more or less equally by all personas.
+Drill-down riffing, pairing and language riffing, while interesting strategies, are either minimal in the makeup of a run or have no impact of the subjective quality of the resulting ideas.
+
+Six personas were identified: quitters, lazy creatives, diligent struggles, lazy strugglers, hard workers, and over-deliverers.
+
+\emph{Quitters} are the easiest to summarize. These participants do not provide all the requested responses, and often resort to cop-outs or exhaustive mutation in the ideas that they do provide.
+
+\emph{Lazy creatives} provide primarily uncommon responses, but do not put any particular effort into the responses.
+They employ the brevity, reach back and exhaustive mutation strategies to a great extent.
+These participants seem like prime targets for future interventions, as their ideas are of interest but they focus their strategies towards rapid completion.
+Lazy creatives employ minimal feature utilization.
+
+\emph{Diligent strugglers} are the opposite of lazy creatives.
+These participants do not employ the brevity or exhaustive mutation strategies, but nonetheless produce primarily common, unoriginal ideas.
+These participants also seem like potential targets for interventions, as they apply themselves to the task but may need external assistance to move beyond their comfort zone.
+Strategies employed include exploratory riffing and feature utilization.
+
+\emph{Lazy strugglers} are simply those who produce no creative ideas and are predisposed towards unproductive strategies (primarily exhaustive mutation and brevity).
+
+\emph{Hard workers} are notable for their minimal use of riffing strategies.
+The riffing strategies they do use tend to exclude exhaustive mutation.
+While brevity is occasionally employed for short periods, it is generally only where a short description is sufficient to communicate an idea.
+
+\emph{Over-deliverers} are notable for their lack of strategy in general. Their ideas are verbose, distinct and feature extremely minimal riffing.

### 2013-10-02 chi2014/qualitative.tex
started revision

@@
 
-In addition to our quantitative analysis of the rate of idea generation, we also performed a qualitative analysis.
+In addition to modelling data and confirming our hypotheses, we performed a qualitative analysis.
 We selected 30 brainstorming runs (5 from each condition).
-The primary author applied open coding to identify characteristics of responses and runs that are independent of the uniqueness and originality measures observed in the quantitative section of this paper.
-The characteristics primarily took the form of \emph{strategies} employed to different degrees by different participants.
-We constructed \emph{personas} around the use of these strategies.
+The primary author applied open coding to identify characteristics of responses and runs that are independent of the uniqueness and originality measures observed above.
+In particular, we identified groups of ideas within runs that shared characteristics.
+These characteristics, in addition to temporal order, form the basis for what we call \emph{strategies}: repeatedly observed classes of ideas that are generated by some participants but not others.
+Furthermore, we use the presence or lack of presence of strategies to define \emph{personas}.
 
@@
 
-A \emph{strategy} is an identifiable mechanism for idea generation that results in several ideas in a brainstorming run.
-A key component of a strategy is that it be present in some runs but not in others.
+A \emph{strategy} is a class of ideas repeatedly observed in groups, though not neccessarily consecutively, within brainstorming runs.
+These ideas are defined by some common characteristics and may or may not be further defined by temporal order.
+Not all participants utilized all strategies in their brainstorming runs.
 
-One strategy employed was \emph{feature exploration}.
-An MP3 player may have many features that distinguish it from any other electronic device of the same physical proportions: audio output (and potentially input), data storage, a screen, general purpose computing abilities (apps), connectivity (WiFi, USB), a camera, a unique appearance, GPS.
-Many participants treated the MP3 player as a pocket-sized brick when generating ideas.
-In contrast, the feature use strategy led to responses such as:
+We identified three categories of strategy.
+The first, \emph{feature exploration}, is specific to the MP3 player problem presented in the paper.
+This category entails focussing on specific features a device possesses and leveraging them as the basis for generating new ideas.
+The second category expands on the \emph{riffing} phenomenon introduced earlier, identifying six(?) different types of riffing exployed.
+Finally, \emph{time-saving} strategies are those that do not neccesarily impact the kinds of ideas generated, but allow the participant to generate at a faster rate.
+
+\subsubsection{Feature exploration}
+
+One category of strategies employed was \emph{feature exploration}.
+An MP3 player has many features that distinguish it from any other electronic device of the same physical proportions: audio output (and potentially input), data storage, a screen, general purpose computing abilities (apps), connectivity (WiFi, USB), a camera, a unique appearance, GPS.
+Each of these features forms the basis for a strategy of idea generation.
+Participants would choose a feature and generate ideas that leveraged that feature in various ways. For example, here are several answers from a participant who explored audio input and output:
 
 \begin{itemize}
-    \item use them in museums to give information on various installations (audio output)
-    \item use in a scavenger hunt, device must be plugged in at another location to see clue
+    \item have them as a resource on public transportation -- people must supply their own headphones
+    \item use them in museums to give information on various installations
+    \item have cities install them in tourist areas, so people can listen about where they are
 \end{itemize}
 
-Runs utilizing the feature exploration strategy produce responses that seem intuitively better than those that do not.
+While it might be expected that all participants would employ this strategy to some degree, many treated the MP3 player as a generic electronic device of equivalent size.
+Runs were commonly composed of answers that could apply to almost any un-used household item:
+
+\begin{itemize}
+    \item doorstop
+    \item use in abstract art
+    \item recycle
+\end{itemize}
+
+Runs that used a feature exploration strategy produce responses that seem intuitively better than those that do not.
 It is satisfying to have a use for a device that "fits" the device as closely as possible, so that none of the functionality is "wasted".
-Despite this, the proportion of runs utilizing feature exploration seemed roughly constant across all other strategies, and the strategy does not form a basis for discrimination between any personas.
 
-The second strategy identified is \emph{riffing}.
-While riffing was examined earlier in relation to time spent on idea generation, we found in our qualitative analysis that it was manifested in many ways.
-\emph{Specificity riffing} occurs when a participant generates two or more ideas and one is a generalization of the other.
+\subsubsection{Riffing}
+
+The second category of strategies is \emph{riffing}.
+We found riffing manifested itself many ways, and identified six distinct riffing strategies used by participants: \emph{Specificity riffing} occurs when a participant generates two or more ideas and one is a generalization of the other.
 For example, two consecutive ideas given by a participant:

### 2013-10-02 chi2014/qualitative.tex
table

@@
 The second category of strategies is \emph{riffing}.
-We found riffing manifested itself many ways, and identified six distinct riffing strategies used by participants: \emph{Specificity riffing} occurs when a participant generates two or more ideas and one is a generalization of the other.
-For example, two consecutive ideas given by a participant:
+We found riffing manifested itself many ways, and identified four distinct riffing strategies used by participants: specificity riffing, exhaustive mutation, exploratory riffing, and follow-up riffing.
+We describe each of these in detail, and then touch on three common phenomena found under riffing strategies.
+
+\emph{Specificity riffing} occurs when a participant generates two or more ideas and one is a generalization of the other. For example, two consecutive ideas given by a participant:
 
@@
 
-Finally, \emph{Follow-up riffing} occurs when a participants response assumes that the solution in one of their previous responses has already been implemented. For example:
+Finally, \emph{Follow-up riffing} occurs when a participant's response assumes that the solution in one of their previous responses has already been implemented. For example:
 
@@
 
-Follow-up riffing, while interesting, is incredibly rare in the data and we will not discuss it further.
+Within and across these strategies, there are a few repeated phenomena.
+Some participants \emph{pair} their riffs.
+These participants  often riff on old ideas but never produce more than two versions of each idea.
+This may represent a desire to generate more ideas to meet the task expectations, but a trepitation that repeated ideas are less satisfying than completely unique ones.
 
-There are two special variants of riffing that can occur across all riffing strategies.
-Some participants \emph{pair} their riffs, such that they often riff on old ideas but never produce more than two versions of each idea.
-
-Another special case is is \emph{linguistic riffing}. Some participants seem to use common sentence structure as queues to help them generate ideas. For example:
+Another phenomenon is \emph{linguistic riffing}. Some participants seem to use common sentence structure as queues when generating ideas. For example:
 
 \begin{itemize}
-    \item Make sunglasses out of them
-    \item Make shoes out of them
-    \item Make a handbag out of them
+    \item We could use the hard-drives inside for different electronics.
+    \item We could use them in place of rocks (to throw at things, to use in pavement.)
+    \item We could use them to back up our music collections on our current devices.
 \end{itemize}
 
-Note that linguistic riffing and exhaustive mutation have significant overlap. A participant that is employing exhaustive mutation is always employing linguistic riffing, but not vice-versa.
+Note the distinction from exhaustive mutation. In the exhaustive mutation example above, instances using the same language were interchangeable.
+In this example, while language is re-used, each instance is a new idea.
 
@@
 Reaching back entails riffing on an idea that occured further back in the run than the previous response.
-The presence of reaching back speaks to the exaustion of the participant's creativity, as they produce new variants of old ideas rather than new ideas altogether.
+The presence of reaching back speaks to the exhaustion of the participant's creativity, as they produce new variants of old ideas rather than new ideas altogether.
 
@@
 This is not surprising, as it is ingrained in the rules of brainstorming given at the beginning of the task: "combinations of ideas count as new ideas".
-However, the different sub-strategies of riffing have very different impacts on the qualitative assessment of a run.
+However, the different strategies of riffing have very different impacts on the qualitative assessment of a run.
 Exhaustive mutation is often unsatisfying, as it feels as though the participant is trying to wring every scrap of task completion out of a single idea.
-On the other hand, exploratory riffing can produce answers that are incredibly unexpected but still of high quality.
-Riffing strategies paint a significant picture of the persona of the participant.
+On the other hand, exploratory riffing can produce answers that are unexpected and of high quality.
+
+\subsubsection{Time-saving}
 
@@
 Six personas were identified: quitters, lazy creatives, diligent struggles, lazy strugglers, hard workers, and over-deliverers.
+These personas are described below.
+A summary, in terms of strategies used, is in Table TAB.
 
@@
 \emph{Over-deliverers} are notable for their lack of strategy in general. Their ideas are verbose, distinct and feature extremely minimal riffing.
+
+\begin{table*}
+    \begin{tabular}{|l|l|l|}
+        \hline
+        \textbf{Persona} & \textbf{Common strategies} & \textbf{Uncommon strategies} \\
+        \hline
+        Quitters & cop-outs, exhaustive mutation & \\
+        Lazy creatives & & \\
+        Diligent strugglers & & \\
+        Lazy struggles & & \\
+        Hard workers & & \\
+        Over-deliverers & & \\
+        \hline
+    \end{tabular}
+\end{table*}

### 2013-10-02 chi2014/qualitative.tex
revision

@@
 
-We identified three categories of strategy.
+We identified four categories of strategy.
 The first, \emph{feature exploration}, is specific to the MP3 player problem presented in the paper.
@@
 The second category expands on the \emph{riffing} phenomenon introduced earlier, identifying six(?) different types of riffing exployed.
-Finally, \emph{time-saving} strategies are those that do not neccesarily impact the kinds of ideas generated, but allow the participant to generate at a faster rate.
+\emph{Time-saving} strategies are those that allow the participant to generate at a faster rate.
+Finally, the \emph{humour} strategy emphasizes generating ideas that are amusing rather than practical.
 
@@
 
+Time-saving strategies tend to negatively impact the quality of responses.
+While there are succinct, creative responses, many brief responses are often a sign that the ideas within the run are "low-hanging fruit".
+Cop-out ideas are poor quality by definition.
+
+\subsubsection{Humour}
+
+Some participants used \emph{humour} as a method of generating ideas.
+For example, one response given was "glue it to the side walk and let the fun ensue".
+While ideas generated under this strategy are impractical individually, this strategy is most often employed by participants who otherwise take the idea generation task seriously and perform successfully.
+
 \subsection{Personas}
@@
 We use personas to capture collections of strategy use we see repeated across participants. 
-While some strategies proved discriminatory, others were used more or less equally by all personas.
-Drill-down riffing, pairing and language riffing, while interesting strategies, are either minimal in the makeup of a run or have no impact of the subjective quality of the resulting ideas.
-
 Six personas were identified: quitters, lazy creatives, diligent struggles, lazy strugglers, hard workers, and over-deliverers.
@@
 
-\emph{Lazy creatives} provide primarily uncommon responses, but do not put any particular effort into the responses.
-They employ the brevity, reach back and exhaustive mutation strategies to a great extent.
+\emph{Lazy creatives} provide primarily uncommon responses, but do not put effort into the responses.
+They employ brevity, reach back and exhaustive mutation strategies to a great extent.
 These participants seem like prime targets for future interventions, as their ideas are of interest but they focus their strategies towards rapid completion.
@@
 \begin{table*}
-    \begin{tabular}{|l|l|l|}
+    \begin{tabular}{|l|p{0.35\linewidth}|l|}
         \hline
-        \textbf{Persona} & \textbf{Common strategies} & \textbf{Uncommon strategies} \\
+        \textbf{Persona} & \textbf{Common strategies} & \textbf{Rare strategies} \\
         \hline
-        Quitters & cop-outs, exhaustive mutation & \\
-        Lazy creatives & & \\
-        Diligent strugglers & & \\
-        Lazy struggles & & \\
-        Hard workers & & \\
-        Over-deliverers & & \\
+        Quitters & cop-outs, exhaustive mutation & follow-up riffing \\
+        \hline
+        Lazy creatives & exploratory riffing, specificity riffing, \newline
+        exhaustive mutation, brevity, reach-backs  & follow-up riffing  \\
+        \hline
+        Diligent strugglers & exploratory riffing & exhaustive mutation, brevity, \newline
+        follow-up riffing \\
+        \hline
+        Lazy strugglers & exhaustive mutation, exploratory riffing, \newline
+        cop-outs, brevity  & follow-up riffing \\
+        \hline
+        Hard workers & feature use, humour, \newline
+        exploratory riffing, specificity riffing  & follow-up riffing, exhaustive mutation \\
+        \hline
+        Over-deliverers & feature use & riffing, reach-backs, brevity \\
         \hline
     \end{tabular}
+    \caption{Strategies used by personas}
 \end{table*}

### 2013-10-04 chi2014/qualitative.tex
more revisions in progress

@@
 In addition to modelling data and confirming our hypotheses, we performed a qualitative analysis.
-We selected 30 brainstorming runs (5 from each condition).
-The primary author applied open coding to identify characteristics of responses and runs that are independent of the uniqueness and originality measures observed above.
-In particular, we identified groups of ideas within runs that shared characteristics.
-These characteristics, in addition to temporal order, form the basis for what we call \emph{strategies}: repeatedly observed classes of ideas that are generated by some participants but not others.
+The primary goal was to identify trends  that were repeated between participants.
+These trends can then be used to inform and justify interventions to change the types of ideas produced.
+
+We selected 30 brainstorming runs (5 randomly from each condition).
+The primary author applied an open coding strategy, flagging ideas within runs that shared common characteristics, such as length, wording, or core problem solution.
+
+In particular, we identify classes of ideas within runs, often temporally close, that hold common characteristics.
+For each class that was observed across multiple participants, we infer the existence of a \emph{strategy}: a plan or rule that aids in generating ideas.
+We describe the observed strategies in detail below.
 Furthermore, we use the presence or lack of presence of strategies to define \emph{personas}.
@@
 
-A \emph{strategy} is a class of ideas repeatedly observed in groups, though not neccessarily consecutively, within brainstorming runs.
-These ideas are defined by some common characteristics and may or may not be further defined by temporal order.
+A \emph{strategy} is a plan, rule, or idea that is used by a participant to generate brainstorming responses.
+Without debriefing participants, we cannot infer the goal of a strategy, but define it in terms of its observable effects.
+These effects are repeated, cross-participant presence of classes of similar ideas.
 Not all participants utilized all strategies in their brainstorming runs.
 
-We identified four categories of strategy.
-The first, \emph{feature exploration}, is specific to the MP3 player problem presented in the paper.
-This category entails focussing on specific features a device possesses and leveraging them as the basis for generating new ideas.
-The second category expands on the \emph{riffing} phenomenon introduced earlier, identifying six(?) different types of riffing exployed.
-\emph{Time-saving} strategies are those that allow the participant to generate at a faster rate.
-Finally, the \emph{humour} strategy emphasizes generating ideas that are amusing rather than practical.
+We identified X \emph{orthogonal} strategy categories - each category of strategies can be employed by a participant with or without any of the other categories.
+The first, \emph{problem scoping}, is a process in which a participant chooses a limited implication or component of the problem and provides multiple specific solutions.
+The second category expands on the \emph{riffing} phenomenon introduced earlier, identifying X different types of riffing exployed.
+Finally, \emph{time-saving} strategies increase the rate of idea generation.
+%Finally, the \emph{humour} strategy emphasizes generating ideas that are amusing rather than practical.
 
-\subsubsection{Feature exploration}
+\subsubsection{Problem scoping}
 
-One category of strategies employed was \emph{feature exploration}.
-An MP3 player has many features that distinguish it from any other electronic device of the same physical proportions: audio output (and potentially input), data storage, a screen, general purpose computing abilities (apps), connectivity (WiFi, USB), a camera, a unique appearance, GPS.
-Each of these features forms the basis for a strategy of idea generation.
-Participants would choose a feature and generate ideas that leveraged that feature in various ways. For example, here are several answers from a participant who explored audio input and output:
+To complete the brainstorming task, participants had to generate some number of uses for a partially broken MP3 player.
+Under the problem scoping strategy, this problem can reduced in scope to one of many sub-problems: brainstorm uses for \emph{an audio output device}; brainstorm uses for \emph{an electronic screen}; brainstorm uses for \emph{a hard drive}; and so on.
+The participant chooses one sub-problem at a time and generates solutions, each of which is also a solution to the original problem.
+For example, the following are generated by a participant focussing on the audio output capabilities of an MP3 player:
 
@@
 
-While it might be expected that all participants would employ this strategy to some degree, many treated the MP3 player as a generic electronic device of equivalent size.
-Runs were commonly composed of answers that could apply to almost any un-used household item:
+The previous examples of problem scoping are examples of emphasizing specific details of the problem.
+Problem scoping can also entail neglecting details.
+For example, the following ideas from a participant apply to almost any physical object, not just the MP3 player described:
 
@@
 
-Runs that used a feature exploration strategy produce responses that seem intuitively better than those that do not.
-It is satisfying to have a use for a device that "fits" the device as closely as possible, so that none of the functionality is "wasted".
+We refer to the former, detail-oriented kind of problem scoping as \emph{scoping inward} and the alternative as \emph{scoping outward}.
+Problem scoping is employed by all participants, but the type of scoping (inwards vs outward) is a useful discriminant.
 
@@
 
-\emph{Specificity riffing} occurs when a participant generates two or more ideas and one is a generalization of the other. For example, two consecutive ideas given by a participant:
+\emph{Generalization riffing} occurs when a participant generates two or more ideas and one is a generalization of the other. For example, two consecutive ideas given by a participant:
 
@@
 
-Another, more common type of riffing is \emph{exhaustive mutation}.
+Another type of riffing is \emph{exhaustive mutation}.
 Under this strategy, a participant repeats a response nearly verbatim many times, each time replacing a single word or concept in such a way that no new ideas are formed.
@@
 
-\emph{Exploratory riffing} is a strategy in which participants hold at least one element of a previous response constant when generating a new response, such that a new idea is formed.
-This can be chained many times in a kind of idea "telephone game" such that the final ideas no longer resemble the original source.
+\emph{Hold riffing} is a strategy in which participants hold at least one element of a previous response constant when generating a new response, such that a new idea is formed.
+This can be thought of as analogous to the game of \emph{draw poker}, in which players hold some cards while receiving new ones, with the intent of building a winning hand complementing the held cards.
 For example:
@@
     \item Jukebox music selector in bars
-    \item Commercials in bar bathrooms (retains bar element)
-    \item Tapper handles for beer (retain alcohol element)
+    \item Commercials in bar bathrooms (hold the bar element)
+    \item Tapper handles for beer (hold the alcohol element)
 \end{itemize}
 
-Finally, \emph{Follow-up riffing} occurs when a participant's response assumes that the solution in one of their previous responses has already been implemented. For example:
+Finally, \emph{Continuation riffing} is the strategy of creating another idea that assumes the previous idea has already been implemented.
+The new ideas relies on the implications or consequences of the previous.
+For example:
 

### 2013-10-05 chi2014/qualitative.tex
revised first half

@@
 
-Another type of riffing is \emph{exhaustive mutation}.
-Under this strategy, a participant repeats a response nearly verbatim many times, each time replacing a single word or concept in such a way that no new ideas are formed.
-For example:
+Another type of riffing is \emph{repeat riffing}. Under this strategy, a participant produces multiple reponses with only slight differences between them. For example:
 
 \begin{itemize}
-    \item wrap yarn around it
-    \item wrap twine around it
-    \item wrap string around it
+    \item recycle to make better MP3/ipod players
+    \item use the circuits to for the ipod/MP3 to power other things that are more recent
 \end{itemize}
+
+In this case, the participant is providing the same idea (namely, creating new electronic devices from the parts) twice.
 
@@
 Within and across these strategies, there are a few repeated phenomena.
-Some participants \emph{pair} their riffs.
-These participants  often riff on old ideas but never produce more than two versions of each idea.
-This may represent a desire to generate more ideas to meet the task expectations, but a trepitation that repeated ideas are less satisfying than completely unique ones.
+Some participants \emph{pair} their riffs. These participants  often riff on old ideas but never produce more than two versions of each idea. For example, paired exhaustive riffing from a participant:
 
-Another phenomenon is \emph{linguistic riffing}. Some participants seem to use common sentence structure as queues when generating ideas. For example:
+\begin{itemize}
+\item TODO
+\item TODO
+\end{itemize}
+
+Another phenomenon is \emph{language repitition}. Some participants seem to use common sentence structure as queues when generating ideas. For example:
 
@@
 
-Note the distinction from exhaustive mutation. In the exhaustive mutation example above, instances using the same language were interchangeable.
 In this example, while language is re-used, each instance is a new idea.
+However, language repitition is especially common in repeat riffing, as this example shows:
+
+\begin{itemize}
+    \item wrap yarn around it
+    \item wrap twine around it
+    \item wrap string around it
+\end{itemize}
 

### 2013-10-08 chi2014/qualitative.tex
updates

@@
 The second category expands on the \emph{riffing} phenomenon introduced earlier, identifying X different types of riffing exployed.
-Finally, \emph{time-saving} strategies increase the rate of idea generation.
+Finally, \emph{language use} strategies... TODO
 %Finally, the \emph{humour} strategy emphasizes generating ideas that are amusing rather than practical.
@@
 
-Another phenomenon is \emph{language repitition}. Some participants seem to use common sentence structure as queues when generating ideas. For example:
+Finally, \emph{reaching back} is a variant of riffing employed by almost all participants who provided more than 10 responses.
+Reaching back entails riffing on an idea that occured further back in the run than the previous response.
+The presence of reaching back speaks to the exhaustion of the participant's creativity, as they produce new variants of old ideas rather than new ideas altogether.
+
+Almost all participants employ some form of riffing.
+This is not surprising, as it is ingrained in the rules of brainstorming given at the beginning of the task: "combinations of ideas count as new ideas".
+However, the different strategies of riffing have very different impacts on the qualitative assessment of a run.
+Exhaustive mutation is often unsatisfying, as it feels as though the participant is trying to wring every scrap of task completion out of a single idea.
+On the other hand, exploratory riffing can produce answers that are unexpected and of high quality.
+
+\subsubsection{Language use}
+
+Participants varied in the language they used to express their ideas. Three strategies were observed: language repitition, brevity, and ambiguity.
+
+Some participants re-use sentence structure or phrasings accross multiple responses. We call this strategy \emph{language repitition}. For example:
 
@@
 
-Finally, \emph{reaching back} is a variant of riffing employed by almost all participants who provided more than 10 responses.
-Reaching back entails riffing on an idea that occured further back in the run than the previous response.
-The presence of reaching back speaks to the exhaustion of the participant's creativity, as they produce new variants of old ideas rather than new ideas altogether.
+\emph{Brevity} is the simple strategy of producing short responses.
+Answers under this strategy are not neccessarily less specific than longer answers.
+For example "textbook audio reader" is a concise description of a very specific use for an MP3 player.
 
-Almost all participants employ some form of riffing.
-This is not surprising, as it is ingrained in the rules of brainstorming given at the beginning of the task: "combinations of ideas count as new ideas".
-However, the different strategies of riffing have very different impacts on the qualitative assessment of a run.
-Exhaustive mutation is often unsatisfying, as it feels as though the participant is trying to wring every scrap of task completion out of a single idea.
-On the other hand, exploratory riffing can produce answers that are unexpected and of high quality.
+\subsubsection{Partial solutions}
 
-\subsubsection{Time-saving}
+Under this category of strategies, responses provide some elements of a solution to the problem, but further idea generation is required to implement the solution.
+Some responses provide a reduced scope, but do not explore within that scope.
+Others would provide a goal for a solution to obtain but not provide a plan for achieving that goal.
+Finally, some responses passed the problem to a third party.
 
-Two time-saving strategies were also identified.
-The first, perhaps overly obvious strategy, is \emph{brevity}.
-This is simply the act of keeping responses as concise as possible, presumably in the interest of finishing quickly.
-The second time-saving strategy was \emph{cop-outs}.
-Cop-outs occur when participants choose to resort to vague or iunrelated responses that do not actually provide any solution to the problem.
+Above, we described the problem scoping strategy.
+Some participants performed \emph{problem scoping without solution}, a variant of this strategy in which problem scoping is the only component of the response.
 For example, one participant's response to the MP3 player problem was to "Remove glass screen to make something".
-Cop-out responses may provide one or more steps towards a solution, but no solution.
+In this case, the participant has provided a scoping of the problem (focus on the glass component) but has not actually provided a solution to the scoped problem.
 
-Time-saving strategies tend to negatively impact the quality of responses.
-While there are succinct, creative responses, many brief responses are often a sign that the ideas within the run are "low-hanging fruit".
-Cop-out ideas are poor quality by definition.
+When a solution is provided, it is still occasionally lacking in sufficient information to act upon.
+We call this strategy \emph{goal without plan}.
+For example, a common response to the problem was "use as art".
+In this case, an end goal has been established, but any implementation of the goal requires further idea generation.
 
-\subsubsection{Humour}
-
-Some participants used \emph{humour} as a method of generating ideas.
-For example, one response given was "glue it to the side walk and let the fun ensue".
-While ideas generated under this strategy are impractical individually, this strategy is most often employed by participants who otherwise take the idea generation task seriously and perform successfully.
+Finally, responses in the \emph{pass the problem} strategy simply relocate the need for idea generation to a third party.
+Common responses under this strategy are "give as gift" or "donate to cancer research".
+These solve the original problem, but simultaneously re-open the problem under slightly changed conditions.
+In this case, some cancer researchers must determine what to do with broken MP3 players.
 

### 2013-10-08 chi2014/qualitative.tex
new personas

@@
 The second category of strategies is \emph{riffing}.
-We found riffing manifested itself many ways, and identified four distinct riffing strategies used by participants: specificity riffing, exhaustive mutation, exploratory riffing, and follow-up riffing.
+We found riffing manifested itself many ways, and identified four distinct riffing strategies used by participants: generalization riffing, repeat riffing, hold riffing, and continuation riffing.
 We describe each of these in detail, and then touch on three common phenomena found under riffing strategies.
@@
 
-Finally, \emph{Continuation riffing} is the strategy of creating another idea that assumes the previous idea has already been implemented.
+Hold riffing can also involve holding language terms constant even when the meaning of the term changes between responses, as in free association. For example:
+
+\begin{itemize}
+    \item beer coaster
+    \item drink coaster (generalization riffing)
+    \item roller coaster (hold riffing)
+\end{itemize}
+
+\emph{Continuation riffing} is the strategy of creating another idea that assumes the previous idea has already been implemented.
 The new ideas relies on the implications or consequences of the previous.
@@
 
-Within and across these strategies, there are a few repeated phenomena.
-Some participants \emph{pair} their riffs. These participants  often riff on old ideas but never produce more than two versions of each idea. For example, paired exhaustive riffing from a participant:
+Within and across these strategies, we identified two sub-strategies or tweaked behaviours.
+Some participants \emph{pair} their riffs. These participants  often riff on old ideas but never produce more than two versions of each idea. For example, paired hold riffing from a participant:
 
@@
 This is not surprising, as it is ingrained in the rules of brainstorming given at the beginning of the task: "combinations of ideas count as new ideas".
-However, the different strategies of riffing have very different impacts on the qualitative assessment of a run.
-Exhaustive mutation is often unsatisfying, as it feels as though the participant is trying to wring every scrap of task completion out of a single idea.
-On the other hand, exploratory riffing can produce answers that are unexpected and of high quality.
 
@@
 In this case, an end goal has been established, but any implementation of the goal requires further idea generation.
+The inverse of this strategy is \emph{plan without goal}, as in the response "change the settings".
+In this case, one could certainly perform the operation on the MP3 player, but it doesn't provide any obvious benefit.
+
+In many cases, responses that do not have an explicit goal or plan nonetheless imply one.
+For example, "throw it in the ocean" is an action performed with the MP3 player with no stated goal, but the implied goal is to get rid of the broken device.
+In these cases, as judged by the coder, reponses were not considered part of a partial solution strategy.
 
@@
 
-We use personas to capture collections of strategy use we see repeated across participants. 
-Six personas were identified: quitters, lazy creatives, diligent struggles, lazy strugglers, hard workers, and over-deliverers.
-These personas are described below.
-A summary, in terms of strategies used, is in Table TAB.
+A persona can be thought of as a toolbox of strategies that we see repeated across participants.
+Participants of a given persona have a subset of strategies that are used often, and a subset that are rarely used.
+In a process analogous to affinity diagramming, brainstorming runs were clustered based on similarity of strategy use.
+This resulted in six clusters of repeated behaviour, each of which is a basis of a persona.
+These personas are summarized in Table TAB.
 
-\emph{Quitters} are the easiest to summarize. These participants do not provide all the requested responses, and often resort to cop-outs or exhaustive mutation in the ideas that they do provide.
+The \emph{high concept} persona consists of participants who view the MP3 player as an abstract object and provide many distinct, if often vague, responses.
+Commonly employed strategies are scoping outward, hold riffing, and partial responses.
 
-\emph{Lazy creatives} provide primarily uncommon responses, but do not put effort into the responses.
-They employ brevity, reach back and exhaustive mutation strategies to a great extent.
-These participants seem like prime targets for future interventions, as their ideas are of interest but they focus their strategies towards rapid completion.
-Lazy creatives employ minimal feature utilization.
+\emph{Outside the box} personas are similar to high concept, except that their responses are more well-defined.
+Common strategies are scoping outward, hold riffing and brevity.
+Partial responses are rare.
 
-\emph{Diligent strugglers} are the opposite of lazy creatives.
-These participants do not employ the brevity or exhaustive mutation strategies, but nonetheless produce primarily common, unoriginal ideas.
-These participants also seem like potential targets for interventions, as they apply themselves to the task but may need external assistance to move beyond their comfort zone.
-Strategies employed include exploratory riffing and feature utilization.
+and so on...
 
-\emph{Lazy strugglers} are simply those who produce no creative ideas and are predisposed towards unproductive strategies (primarily exhaustive mutation and brevity).
+\subsubsection{Outliers}
 
-\emph{Hard workers} are notable for their minimal use of riffing strategies.
-The riffing strategies they do use tend to exclude exhaustive mutation.
-While brevity is occasionally employed for short periods, it is generally only where a short description is sufficient to communicate an idea.
-
-\emph{Over-deliverers} are notable for their lack of strategy in general. Their ideas are verbose, distinct and feature extremely minimal riffing.
+In addition to the observed personas, there were several outliers.
+One participant employed no strategies to a signficant degree (in the case of problem scoping, they scoped inward and outward with similar frequency).
+A second outlier responded with questions, such as "What can the average person do to better the environment?".
 
@@
         \hline
-        Quitters & cop-outs, exhaustive mutation & follow-up riffing \\
+        High concept & scoping outward, hold riffing, partial responses & brevity, reach-backs \\
         \hline
-        Lazy creatives & exploratory riffing, specificity riffing, \newline
-        exhaustive mutation, brevity, reach-backs  & follow-up riffing  \\
+        Outside the box & scoping outward, hold riffing, brevity & partial responses, reach-backs \\
         \hline
-        Diligent strugglers & exploratory riffing & exhaustive mutation, brevity, \newline
-        follow-up riffing \\
+        broad maximizers & repeat riffing, brevity, \newline
+        partial responses  & reach-backs \\ 
         \hline
-        Lazy strugglers & exhaustive mutation, exploratory riffing, \newline
-        cop-outs, brevity  & follow-up riffing \\
+        broad explorers & hold riffing, partial responses & \\
         \hline
-        Hard workers & feature use, humour, \newline
-        exploratory riffing, specificity riffing  & follow-up riffing, exhaustive mutation \\
+        kitchen sink & scoping outward, repeat riffing, brevity, \newline
+        partial responses, reach-backs & \\ 
         \hline
-        Over-deliverers & feature use & riffing, reach-backs, brevity \\
+        detail explorers & scoping inward, hold riffing & brevity, partial responses, reach-backs\\
         \hline

### 2013-10-16 chi2014/qualitative.tex
revision with stats

@@
 
-In addition to modelling data and confirming our hypotheses, we performed a qualitative analysis.
+In addition to modeling data and confirming our hypotheses, we performed a qualitative analysis.
 The primary goal was to identify trends  that were repeated between participants.
@@
 Without debriefing participants, we cannot infer the goal of a strategy, but define it in terms of its observable effects.
-These effects are repeated, cross-participant presence of classes of similar ideas.
+These effects are observed as repeated, cross-participant classes of similar ideas.
 Not all participants utilized all strategies in their brainstorming runs.
 
-We identified X \emph{orthogonal} strategy categories - each category of strategies can be employed by a participant with or without any of the other categories.
+We identified 3 \emph{orthogonal} strategy categories - each category of strategies can be employed by a participant with or without any of the other categories.
 The first, \emph{problem scoping}, is a process in which a participant chooses a limited implication or component of the problem and provides multiple specific solutions.
-The second category expands on the \emph{riffing} phenomenon introduced earlier, identifying X different types of riffing exployed.
-Finally, \emph{language use} strategies... TODO
+The second category expands on the \emph{riffing} phenomenon introduced earlier, identifying 4 different types of riffing explored.
+Finally, \emph{language use} strategies relate not to the meaning of the ideas generated, but to the way they are expressed.
 %Finally, the \emph{humour} strategy emphasizes generating ideas that are amusing rather than practical.
@@
 The participant chooses one sub-problem at a time and generates solutions, each of which is also a solution to the original problem.
-For example, the following are generated by a participant focussing on the audio output capabilities of an MP3 player:
+For example, the following are generated by PXXX focusing on the audio output capabilities of an MP3 player:
 
@@
 Problem scoping can also entail neglecting details.
-For example, the following ideas from a participant apply to almost any physical object, not just the MP3 player described:
+For example, the following ideas from PXXX apply to almost any physical object, not just the MP3 player described:
 
@@
 We refer to the former, detail-oriented kind of problem scoping as \emph{scoping inward} and the alternative as \emph{scoping outward}.
-Problem scoping is employed by all participants, but the type of scoping (inwards vs outward) is a useful discriminant.
+These two types of scoping cover almost all responses.
+However, some responses employ no scoping whatsoever.
+These responses utilize the device exactly as it was intended, as a portable MP3 player.
+For example, "get it fixed and give it to a needy kid".
+
+Problem scoping is employed by all participants, but the type of scoping is a useful discriminant.
 
@@
 
-\emph{Generalization riffing} occurs when a participant generates two or more ideas and one is a generalization of the other. For example, two consecutive ideas given by a participant:
+\emph{Generalization riffing} occurs when a participant generates two or more ideas and one is a generalization of the other. For example, two consecutive ideas given by PXXX:
 
@@
 
-Another type of riffing is \emph{repeat riffing}. Under this strategy, a participant produces multiple reponses with only slight differences between them. For example:
+Another type of riffing is \emph{repeat riffing}. Under this strategy, a participant produces multiple responses with only slight differences between them. For example (PXXX):
 
 \begin{itemize}
-    \item recycle to make better MP3/ipod players
-    \item use the circuits to for the ipod/MP3 to power other things that are more recent
+    \item recycle to make better MP3/iPod players
+    \item use the circuits to for the iPod/MP3 to power other things that are more recent
 \end{itemize}
@@
 
-\emph{Hold riffing} is a strategy in which participants hold at least one element of a previous response constant when generating a new response, such that a new idea is formed.
-This can be thought of as analogous to the game of \emph{draw poker}, in which players hold some cards while receiving new ones, with the intent of building a winning hand complementing the held cards.
-For example:
+\emph{Hold riffing} is a strategy in which participants hold at least one element of a previous response constant when generating a new response, such that a new idea is formed. For example (PXXX):
 
@@
 
-Hold riffing can also involve holding language terms constant even when the meaning of the term changes between responses, as in free association. For example:
+Hold riffing can also involve holding language terms constant even when the meaning of the term changes between responses, as in free association. For example (PXXX):
 
@@
 \emph{Continuation riffing} is the strategy of creating another idea that assumes the previous idea has already been implemented.
-The new ideas relies on the implications or consequences of the previous.
-For example:
+The new idea on the implications or consequences of the previous.
+For example (PXXX):
 
@@
 
-Within and across these strategies, we identified two sub-strategies or tweaked behaviours.
-Some participants \emph{pair} their riffs. These participants  often riff on old ideas but never produce more than two versions of each idea. For example, paired hold riffing from a participant:
+\paragraph{Spatial separation}
+
+The strategies above describe techniques for taking one idea and transforming it into another. In this section, we briefly describe patterns of \emph{where} participants implemented these strategies.
+
+The most natural separation between a riffed idea and its source is none - the riffed idea directly follows the source.
+We call this \emph{linear riffing}.
+Some participants \emph{pair} their riffs.
+Paired riffing is a special case on linear riffing in which only a single riffed idea is produced, and no further riffs are produced on that source for the remainder of the run.
+For example, paired repeat riffing from PXXX:
 
 \begin{itemize}
-\item TODO
-\item TODO
+\item cut up and use to decorate shoes
+\item cut up and use to decorate vase
 \end{itemize}
 
-Finally, \emph{reaching back} is a variant of riffing employed by almost all participants who provided more than 10 responses.
-Reaching back entails riffing on an idea that occured further back in the run than the previous response.
-The presence of reaching back speaks to the exhaustion of the participant's creativity, as they produce new variants of old ideas rather than new ideas altogether.
-
-Almost all participants employ some form of riffing.
-This is not surprising, as it is ingrained in the rules of brainstorming given at the beginning of the task: "combinations of ideas count as new ideas".
+Finally, \emph{reach backs} comprise all other riffs. reaching back entails riffing on an idea that occurred further back in the run than the previous response.
 
@@
 
-Participants varied in the language they used to express their ideas. Three strategies were observed: language repitition, brevity, and ambiguity.
-
-Some participants re-use sentence structure or phrasings accross multiple responses. We call this strategy \emph{language repitition}. For example:
+Some participants re-use sentence structure or phrasings across multiple responses. We call this strategy \emph{language repetition}. For example (PXXX):
 
@@
 
-In this example, while language is re-used, each instance is a new idea.
-However, language repitition is especially common in repeat riffing, as this example shows:
+In this example, while language is re-used, each instance is also a new idea unrelated to the previous.
+However, language repetition is especially common in repeat riffing, as this example (PXXX) shows:
 
@@
 \end{itemize}
-
-\emph{Brevity} is the simple strategy of producing short responses.
-Answers under this strategy are not neccessarily less specific than longer answers.
-For example "textbook audio reader" is a concise description of a very specific use for an MP3 player.
 
@@
 Some participants performed \emph{problem scoping without solution}, a variant of this strategy in which problem scoping is the only component of the response.
-For example, one participant's response to the MP3 player problem was to "Remove glass screen to make something".
+For example, on of PXXX's responses to the MP3 player problem was to "Remove glass screen to make something".
 In this case, the participant has provided a scoping of the problem (focus on the glass component) but has not actually provided a solution to the scoped problem.
 
-When a solution is provided, it is still occasionally lacking in sufficient information to act upon.
+In some responses, an end goal has been established, but any implementation of the goal requires further idea generation.
 We call this strategy \emph{goal without plan}.
-For example, a common response to the problem was "use as art".
-In this case, an end goal has been established, but any implementation of the goal requires further idea generation.
-The inverse of this strategy is \emph{plan without goal}, as in the response "change the settings".
-In this case, one could certainly perform the operation on the MP3 player, but it doesn't provide any obvious benefit.
+For example, a selection of responses from PXXX:
+
+\begin{itemize}
+    \item use old parts to make a new device
+    \item old parts can use to make something
+    \item create a new device
+    \item reuse all working parts
+\end{itemize}
+
+In this case, the goal of creating a new electronic device is given, but it is not clear what will be created or how the materials will be re-used.
+
+The inverse of this strategy is \emph{plan without goal}, as demonstrated in these responses (also from PXXX):
+
+\begin{itemize}
+    \item melt down old parts
+    \item see what old parts are usable
+\end{itemize}
+
+In this case, the operations of melting down the device or checking for working parts can be performed, but don't provide any obvious benefit.
 
 In many cases, responses that do not have an explicit goal or plan nonetheless imply one.
-For example, "throw it in the ocean" is an action performed with the MP3 player with no stated goal, but the implied goal is to get rid of the broken device.
-In these cases, as judged by the coder, reponses were not considered part of a partial solution strategy.
+For example, "throw it in the ocean" (PXXX) is an action performed with the MP3 player with no stated goal, but the implied goal is to get rid of the broken device.
+In these cases, as judged by the coder, responses were not assumed to employ a partial solution strategy.
 
-Finally, responses in the \emph{pass the problem} strategy simply relocate the need for idea generation to a third party.
-Common responses under this strategy are "give as gift" or "donate to cancer research".
+Finally, responses in the \emph{pass the problem} strategy simply relocate the need for idea generation to a third party. These responses from PXXX provide an example:
+
+\begin{itemize}
+    \item donate to less fortunate that are trying to assist the homeless
+    \item give to university
+    \item give to non-profit that can benefit
+    \item give to an organization that has the knowledge to use these devices
+\end{itemize}
+
 These solve the original problem, but simultaneously re-open the problem under slightly changed conditions.
-In this case, some cancer researchers must determine what to do with broken MP3 players.
+In this case, a university/non-profit/other body must determine what to do with broken MP3 players.
 
-\subsection{Personas}
+\subsection{Strategy use}
 
-A persona can be thought of as a toolbox of strategies that we see repeated across participants.
-Participants of a given persona have a subset of strategies that are used often, and a subset that are rarely used.
-In a process analogous to affinity diagramming, brainstorming runs were clustered based on similarity of strategy use.
-This resulted in six clusters of repeated behaviour, each of which is a basis of a persona.
-These personas are summarized in Table TAB.
+In this section, we briefly describe how strategies were employed by participants.
+The primary author again coded 30 randomly selected runs (5 from each condition, 1150 instances total) for the strategies identified above.
+Note that examples of riffing identified in this section are distinct from the algorithmic/quantitative identification of riffing described in the method section above.
+The prevalence of each strategy as a function of total number of instances is given in Table TAB.
 
-The \emph{high concept} persona consists of participants who view the MP3 player as an abstract object and provide many distinct, if often vague, responses.
-Commonly employed strategies are scoping outward, hold riffing, and partial responses.
+\begin{table}
+    \begin{tabular}{r | l l}
+        \textbf{strategy} & \textbf{\# instances} & \textbf{\% of instances} \\
+        inward scoping & 267 & 23.2 \\
+        outward scoping & 828 & 72 \\
+        no scoping & 55 & 4.8 \\
+        \hline \hline
+        hold riffing & 278 & 24.2 \\
+        continuation riffing & 3 & 0.3 \\
+        repeat riffing & 119 & 10.3 \\
+        generalization riffing & 48 & 4.2 \\
+        \hline
+        any riffing & 448 & 39.0 \\
+        pair & 108 & 9.4 \\
+        reach back & 129 & 11.2 \\
+        \hline \hline
+        scoping without solution & 3 & 0.3 \\
+        passing the problem & 69 & 6 \\
+        plan without goal & 58 & 5.0 \\
+        goal without plan & 58 & 5.0 \\ 
+        \hline
+        any partial solution & 188 & 16.3 \\
+    \end{tabular}
+\end{table}
 
-\emph{Outside the box} personas are similar to high concept, except that their responses are more well-defined.
-Common strategies are scoping outward, hold riffing and brevity.
-Partial responses are rare.
+Outward scoping is by far the most commons strategy, occurring in 72\% of instances.
+Riffing is also very common, with 40\% of instances employing some form of riffing.
+Riffs are most often consecutive (linear riffing), and hold riffing is the most common type.
 
-and so on...
+Partial solutions are fairly uncommon (16.3\%), and most often involve passing the problem (6\%).
+Continuation riffing and scoping without solution are the least common strategies, each with only 3 instances.
 
-\subsubsection{Outliers}
+\subsubsection{Strategy location}
 
-In addition to the observed personas, there were several outliers.
-One participant employed no strategies to a signficant degree (in the case of problem scoping, they scoped inward and outward with similar frequency).
-A second outlier responded with questions, such as "What can the average person do to better the environment?".
+We also consider where in a brainstorming run the strategies are more likely to occur. The results are summarized in Table TAB.
 
-\begin{table*}
-    \begin{tabular}{|l|p{0.35\linewidth}|l|}
+\begin{table}
+    \begin{tabular}{r | l}
+        \textbf{strategy} & \textbf{mean position in run} \\
+        inward scoping & 26.2 \\
+        outward scoping & 34.8 \\
+        no scoping & 18.1 \\
+        \hline \hline
+        hold riffing & 34.9 \\
+        continuation riffing & 16.3 \\
+        repeat riffing & 40.3 \\
+        generalization riffing & 26.4 \\
         \hline
-        \textbf{Persona} & \textbf{Common strategies} & \textbf{Rare strategies} \\
+        pair & 31.9 \\
+        reach back & 33.1 \\
+        \hline \hline
+        scoping without solution & 39.7 \\
+        passing the problem & 40 \\
+        plan without goal & 29.6 \\
+        goal without plan & 20.0 \\ 
+    \end{tabular}
+\end{table}
+
+Most of the differences run position between strategies are minimal.
+Most discriminatory are the differences in scoping.
+Lack of scoping is, on average, very early in the run, while scoping inward towards some detail of the problem occurs later.
+Outward scoping, or dropping details of the problem, occurs latest of all.
+This gradual increase hints that more "outside the box" ideas may not show up until later in runs, after participants have finished exploring specific details of the problem.
+
+\subsubsection{Strategy co-occurrence}
+
+Of final interest was the question of which strategies occurred together.
+For each run, a count was taken of the number of occurrences of each strategy, and the Pearson r of these counts was taken for each pairing of strategies to give a rough measure of how the strategies were grouped in runs.
+Strong correlations (greater than 0.6) are given in Table TAB.
+
+\begin{table}
+    \begin{tabular}{l l l}
+        \textbf{strategy 1} & \textbf{strategy 2} & \textbf{Pearson r} \\
         \hline
-        High concept & scoping outward, hold riffing, partial responses & brevity, reach-backs \\
-        \hline
-        Outside the box & scoping outward, hold riffing, brevity & partial responses, reach-backs \\
-        \hline
-        broad maximizers & repeat riffing, brevity, \newline
-        partial responses  & reach-backs \\ 
-        \hline
-        broad explorers & hold riffing, partial responses & \\
-        \hline
-        kitchen sink & scoping outward, repeat riffing, brevity, \newline
-        partial responses, reach-backs & \\ 
-        \hline
-        detail explorers & scoping inward, hold riffing & brevity, partial responses, reach-backs\\
-        \hline
+        passing the problem & repeat riffing & 0.89 \\
+        hold riffing & outward scoping & 0.85 \\
+        pairing & scoping without solution & 0.64 \\
+        reaching back & hold riffing & 0.64 \\
+        pairing & outward scoping & 0.63 \\
+        reaching back & outward scoping & 0.62 \\
+        plan without goal & hold riffing & 0.61 \\
     \end{tabular}
-    \caption{Strategies used by personas}
-\end{table*}
+\end{table}
+
+Passing the problem and repeat riffing were highly correlated (0.89), indicating that responses utilizing pass the problem were likely to be repeated many times to generate more responses.
+For each of the other correlations, the potential implications less obvious.
+Further analysis is required to tease out the implications of strategy co-occurrence.

### 2013-10-16 chi2014/qualitative.tex
2/8 mike feedback

@@
 
-To complete the brainstorming task, participants had to generate some number of uses for a partially broken MP3 player.
-Under the problem scoping strategy, this problem can reduced in scope to one of many sub-problems: brainstorm uses for \emph{an audio output device}; brainstorm uses for \emph{an electronic screen}; brainstorm uses for \emph{a hard drive}; and so on.
-The participant chooses one sub-problem at a time and generates solutions, each of which is also a solution to the original problem.
-For example, the following are generated by PXXX focusing on the audio output capabilities of an MP3 player:
+Under the problem scoping strategy, the original problem is transformed to a new problem of reduced scope.
+An MP3 player has many functions which are difficult to leverage simultaneously.
+Thus, many transformations are possible, for example: brainstorm uses for \emph{an audio output device}; brainstorm uses for a small box-shaped object; brainstorm uses for \emph{a hard drive}; and so on.
+The participant chooses a problem transformation and generates solutions, each of which is also a solution to the original problem.
+For example (PXXX):
 
@@
 
-The previous examples of problem scoping are examples of emphasizing specific details of the problem.
-Problem scoping can also entail neglecting details.
-For example, the following ideas from PXXX apply to almost any physical object, not just the MP3 player described:
+The participant's responses satisfy the transformed question "brainstorming uses for an audio output device".
+The resulting ideas no not rely on or acknowledge other aspects of the problem, such as portability, the screen, the ability to store files, or the appearance or size of the device.
+Nonetheless, each response is also a valid response to the original problem.
+
+We distinguish between kinds of problem scoping.
+\emph{Inward scoping} transformations reduce scope by limiting consideration to a single detail of the problem.
+The previous responses, which focus on audio output, are an example of inward scoping.
+
+\emph{Outward scoping} transformations reduce complexity by stripping all meaningful details.
+In the case of the MP3 player problem, this can be understood as essentially reducing the problem to "brainstorm uses for a thing".
+For example (PXXX):
 
@@
 
-We refer to the former, detail-oriented kind of problem scoping as \emph{scoping inward} and the alternative as \emph{scoping outward}.
+The inward scoping solutions were applicable to the transformed problem and the original problem, but not many others.
+In contrast, the outward scoping solutions could be applied to a wide body of problems unrelated to the original, such as "brainstorm uses for a broom".
+Thus, a useful rule for distinguishing the types of scoping is to consider whether the \emph{set of problems solved by the solutions} is small, or large.
+
 These two types of scoping cover almost all responses.
@@
 For example, "get it fixed and give it to a needy kid".
-
-Problem scoping is employed by all participants, but the type of scoping is a useful discriminant.
 

### 2013-10-16 chi2014/qualitative.tex

@@
 
-Another type of riffing is \emph{repeat riffing}. Under this strategy, a participant produces multiple responses with only slight differences between them. For example (PXXX):
+Another type of riffing is \emph{repeat riffing}.
+Under this strategy, a participant produces multiple responses with only slight differences between them, such that the group of ideas could be easily summarized as one idea without significant loss of information.
+For example (PXXX):
 
 \begin{itemize}
-    \item recycle to make better MP3/iPod players
-    \item use the circuits to for the iPod/MP3 to power other things that are more recent
+    \item extract usage data
+    \item extract texts
+    \item extract gps data
 \end{itemize}
 
-In this case, the participant is providing the same idea (namely, creating new electronic devices from the parts) twice.
+In this case, all three ideas could be summarized with "extract the data on the device".
 
@@
     \item Jukebox music selector in bars
-    \item Commercials in bar bathrooms (hold the bar element)
-    \item Tapper handles for beer (hold the alcohol element)
+    \item Commercials in bar bathrooms 
+    \item Tapper handles for beer 
 \end{itemize}
 
-Hold riffing can also involve holding language terms constant even when the meaning of the term changes between responses, as in free association. For example (PXXX):
+In this example, the setting of the application is held constant across riffs: a bar.
+
+Another realization of hold riffing holds language term constant between unrelated ideas. For example (PXXX):
 
 \begin{itemize}
-    \item beer coaster
-    \item drink coaster (generalization riffing)
-    \item roller coaster (hold riffing)
+    \item We could use the hard-drives inside for different electronics.
+    \item We could use them in place of rocks (to throw at things, to use in pavement.)
+    \item We could use them to back up our music collections on our current devices.
 \end{itemize}
 
-\emph{Continuation riffing} is the strategy of creating another idea that assumes the previous idea has already been implemented.
-The new idea on the implications or consequences of the previous.
+\emph{Continuation riffing} is the strategy of creating another idea that relies on the implications or consequences of a previous idea.
+The new idea cannot be understood without the context of the previous.
 For example (PXXX):
@@
 
-\subsubsection{Language use}
-
-Some participants re-use sentence structure or phrasings across multiple responses. We call this strategy \emph{language repetition}. For example (PXXX):
-
-\begin{itemize}
-    \item We could use the hard-drives inside for different electronics.
-    \item We could use them in place of rocks (to throw at things, to use in pavement.)
-    \item We could use them to back up our music collections on our current devices.
-\end{itemize}
-
-In this example, while language is re-used, each instance is also a new idea unrelated to the previous.
-However, language repetition is especially common in repeat riffing, as this example (PXXX) shows:
-
-\begin{itemize}
-    \item wrap yarn around it
-    \item wrap twine around it
-    \item wrap string around it
-\end{itemize}
-
 \subsubsection{Partial solutions}
@@
 Under this category of strategies, responses provide some elements of a solution to the problem, but further idea generation is required to implement the solution.
-Some responses provide a reduced scope, but do not explore within that scope.
-Others would provide a goal for a solution to obtain but not provide a plan for achieving that goal.
-Finally, some responses passed the problem to a third party.
+The following ideas (PXXX) are all examples of partial solutions:
+
+\begin{itemize}
+    \item use old parts to make a new device
+    \item old parts can use to make something
+    \item create a new device
+    \item reuse all working parts
+\end{itemize}
+
+In these responses, there is a clear recommendation that we recycle components of the MP3 player to produce new electronics, but it is unclear what we would make or how we would use the parts.
+Essentially, we are left with new brainstorming problems to solve.
+
+As with riffing and problem scoping, there are four sub-strategies of partial solutions: problem scoping without solution, goal without plan, plan without goal, and passing the problem.
 
@@
 We call this strategy \emph{goal without plan}.
-For example, a selection of responses from PXXX:
-
-\begin{itemize}
-    \item use old parts to make a new device
-    \item old parts can use to make something
-    \item create a new device
-    \item reuse all working parts
-\end{itemize}
-
-In this case, the goal of creating a new electronic device is given, but it is not clear what will be created or how the materials will be re-used.
+The list of responses at the beginning of this section are examples of the goal without plan strategy.
+The goal of creating a new electronic device is given, but it is not clear what will be created or how the materials will be re-used.
 

### 2013-10-16 chi2014/qualitative.tex
8/8 mike recs

@@
 Note that examples of riffing identified in this section are distinct from the algorithmic/quantitative identification of riffing described in the method section above.
-The prevalence of each strategy as a function of total number of instances is given in Table TAB.
+The prevalence of each strategy as a function of total number of instances is given in Table TAB, and a comparative plot in Figure FIG.
 
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{strategy_counts}
+    \caption{presence of strategies}
+    \label{fig:strategy_counts}
+\end{figure}
+
 Outward scoping is by far the most commons strategy, occurring in 72\% of instances.
@@
 \begin{table}
-    \begin{tabular}{r | l}
-        \textbf{strategy} & \textbf{mean position in run} \\
-        inward scoping & 26.2 \\
-        outward scoping & 34.8 \\
-        no scoping & 18.1 \\
+    \begin{tabular}{r | l l l l l l}
+        \textbf{strategy} & \textbf{5} & \textbf{10} & \textbf{25} & \textbf{50} & \textbf{75} & \textbf{100} \\
+        inward scoping & 2 & 5 & 11 & 23 & 29 & 44 \\
+        outward scoping & 3 & 3 & 8.5 & 27.5 & 39.5 & 41 \\
+        no scoping & 3 & 5 & 9 & 18 & 22.5 & 20.5 \\
         \hline \hline
-        hold riffing & 34.9 \\
-        continuation riffing & 16.3 \\
-        repeat riffing & 40.3 \\
-        generalization riffing & 26.4 \\
+        hold riffing & 3 & 8 & 12.5 & 24 & 44.5 & 57.5 \\
+        continuation riffing & - & - & 18 & 4 & 27 & - \\
+        repeat riffing & 4 & 7 & 2 & 25 & 49.5 & 45 \\
+        generalization riffing & 4 & 4 & 11 & 29.5 & 34 & 30 \\
         \hline
-        pair & 31.9 \\
-        reach back & 33.1 \\
+        pair & 3.5 & 7.5 & 12.5 & 25.5 & 40 & 33.5 \\
+        reach back & 3.5 & 6 & 11 &29.5 & 40 & 47.5 \\
         \hline \hline
-        scoping without solution & 39.7 \\
-        passing the problem & 40 \\
-        plan without goal & 29.6 \\
-        goal without plan & 20.0 \\ 
+        scoping without solution & - & - & - & 13 & - & 53 \\
+        passing the problem & 1.5 & - & 14.5 & 40 & 35.5 & 60.5  \\
+        plan without goal & - & - & 6 & 18 & 38.5 & 30 \\
+        goal without plan & - & 3 & 2 & 15 & 11 & 29 \\ 
     \end{tabular}
+    \caption{Median position of strategy in run by condition}
 \end{table}
 
-Most of the differences run position between strategies are minimal.
-Most discriminatory are the differences in scoping.
-Lack of scoping is, on average, very early in the run, while scoping inward towards some detail of the problem occurs later.
-Outward scoping, or dropping details of the problem, occurs latest of all.
-This gradual increase hints that more "outside the box" ideas may not show up until later in runs, after participants have finished exploring specific details of the problem.
+The distributions of strategy occurence across position in run tend towards uniform when there are more than a few examples.
+The notable exception to this is problem scoping.
+There are fewer occurences of the inward scoping strategy later in brainstorming runs.
+
+The number of reach-backs increase with position. This may simply be because there are more ideas to draw from for inspiration later in a run.
 

### 2013-10-17 chi2014/abstract.tex
10 page revision first attempt

@@
 Our data also show support for the SIAM model (search for ideas in associative memory) \cite{nijstad_how_2006} in which workers draw ideas from a single category before moving on to a new category.
-Our results provide clear recommendations for practitioners and researchers alike who wish to leverage the crowd for brainstorming.
+Our results provide clear recommendations for practitioners and researchers alike who wish to leverage the crowd for brainstorming, and a qualitative understanding of brainstorming strategies employed by crowd workers.
 

### 2013-10-17 chi2014/conclusions.tex
10 page revision first attempt

@@
-\section{Discussion and Future Work}
+\section{Future Work}
 
-In this paper, we have primarily focused on the uniqueness of ideas and idea categories in quantitative terms.
-In this section, we provide a more qualitative perspective on the data collected, then describe future work.
-
-\subsection{Variability of Worker Input}
-While we observed that individuals are likely to draw from a common pool of ideas early on in the brainstorming process, there were outliers in the data.
-For example, one individual in the 75 ideas condition consistently produced some of most original ideas for old MP3 players:
-
-\begin{itemize}
-\item With displays - could be attached to a small solar panel (like the USB chargers w/multi attachments, including solar and car, etc.) and all the power could be sent to the screen as a small, portable light, then sent to countries lacking power in rural areas.
-\item They could be loaded with coordinates or other info and sealed in containers, then used in ocean current studies - once the container was collected they could be plugged in and read, and the data could be sent to whomever.
-\item They could be plugged in at a base camp at the bottom of Everest, and then potential climbers could record their name, loved ones' address, and any last words before attempting the climb. 
-\end{itemize}
-
-This individual's responses were responsible for clearly raising the observed uniqueness of ideas in the 75 condition when plotted against the other conditions.
-
-Other individuals filled their quotas by creating uninspiring idea templates that they varied only slightly between responses.
-For example, this individual provides a list of essentially equivalent responses:
-
-\begin{itemize}
-\item a gift for a cousin
-\item a gift for a niece
-\item a gift for a nephew
-\item a gift for a second cousin
-\end{itemize}
-
-This variability in worker responses prevented us from modeling any potential effects that the number of questions requested might have on individuals' responses.
-Thus, while it is clear that there is little advantage to requesting 5 or 10 ideas from workers, it is not clear whether it is preferable to request 20, 50, 75, or 100 ideas.
-We suspect that requesting 50 or 75 is the ``sweet'' spot, but are not able to empirically support that recommendation at this time.
-
-\subsection{Future Work}
 The results of this research suggest several avenues for future work.
@@
 \item The time it takes to generate an idea when changing categories is longer than the time taken when generating ideas within a category.
+\item There are repeatedly observable strategies employed by workers brainstorming in microtask marketplaces
 \end{itemize}

### 2013-10-17 chi2014/introduction.tex
10 page revision first attempt

@@
 
-Microtask marketplaces such as Amazon's Mechanical Turk have proven to be useful platforms for outsourcing a wide range of tasks, from proofreading documents \cite{soylent} to estimating the caloric content of food \cite{platemate}.
+Microtask marketplaces such as Amazon's Mechanical Turk have proven to be useful platforms for outsourcing of tasks, from proofreading documents \cite{soylent} to estimating the caloric content of food \cite{platemate}.
 These platforms make possible human computation systems \cite{quinn_2011}, hybrid systems that leverage humans' abilities to solve problems that currently cannot be solved by machines alone.
@@
 A particularly compelling use of microtask marketplaces is to tap into the collective creativity of crowds.
-This enables one to quickly sample a wide range of ideas from a diverse population, which is useful for tasks such as brainstorming or design.
-Aaron Koblin's art piece ``The Sheep Market'' \cite{koblin_2009} was an early demonstration of this potential.
+This enables one to quickly sample a wide range of ideas from a diverse population for tasks such as brainstorming or design.
+Aaron Koblin's art piece ``The Sheep Market'' \cite{koblin_2009} was an early demonstration of this application.
 In this project, he asked 10,000 individuals to draw a ``sheep facing left,'' yielding a wide range of unique results.
@@
 In this paper, we focus on crowd-based brainstorming.
-While brainstorming has been an implicit component of many creative endeavors in microtask marketplaces, we are unaware of any research that specifically describes and models the brainstorming process in this environment.
+While brainstorming has been an implicit component of many creative endeavors in microtask marketplaces, we are unaware of any research that describes and models the brainstorming process in this environment.
 Furthermore, while there is a rich literature describing brainstorming processes of individuals and groups (e.g., \cite{osborn_applied_1957, pinsonneault_electronic_1999, nijstad_how_2006,brown1998modeling}), there are important qualitative differences in microtask marketplaces.
 For example, crowd workers exhibit high variance of effort in completing a task \cite{soylent}.
-Brainstorming participants in a microtask marketplace are also paid for their participation, and some might be more eager than others to complete a task as fast as possible to maximize earnings \cite{kittur2008crowdsourcing}.
+Brainstorming participants in a microtask marketplace are paid for their participation, and some might be more eager than others to complete a task as fast as possible to maximize earnings \cite{kittur2008crowdsourcing}.
 Furthermore, participants in microtask marketplaces cannot be expected to be spatially or temporally co-located.
@@
 Across all conditions, 154 workers submitted responses to 170 HITs (Human Intelligence Tasks), yielding 3292 ideas.
-From these ideas, we present a series of models characterizing features of crowd brainstorming, such as the rate of novel idea production, the rate at which novel {\em categories\/} of ideas are produced (where a category represents a set of conceptually similar ideas that are distinct from other categories), qualities of the ideas (for example, whether they are common or not), and characteristics of the brainstorming process itself.
+From these ideas, we present a series of models characterizing features of crowd brainstorming: the rate of novel idea production, the rate at which novel {\em categories\/} of ideas are produced, qualities of the ideas (for example, whether they are common or not), and characteristics of the brainstorming process itself.
 
@@
 \item Requesting ``as many ideas as possible'' for a fixed price (fixed for the entire HIT, rather than per idea) is inadvisable, as it leads to a relatively low number of ideas produced (in our experiments, fewer than 15).
+\item we identify and describe a set of strategies employed by crowd workers that result in distinct kinds of produced ideas
 \end{itemize}
@@
 Our results provide baseline models useful for two audiences: 1) those who wish to leverage crowds for brainstorming, and 2) researchers interested in developing techniques to optimize the crowd brainstorming processes.
-For the former, our models provide clear guidance and recommendations for various parameters of brainstorming in this environment (such as the number of questions to request from workers).
+For the former, our models provide recommendations for parameters of brainstorming in this environment (such as the number of questions to request from workers).
 For the latter group, our models provide a baseline for comparing novel techniques intended to increase the quality of brainstorming output by crowds.
@@
 
-In the rest of this paper, we review prior work, then describe our experimental design, results from our experiments, and implications for practitioners and researchers.
+In the rest of this paper, we review prior work, then describe our experimental design, results from our experiments, implications for practitioners and researchers, and a qualitative understanding of idea generation strategies.

### 2013-10-17 chi2014/qualitative.tex
10 page revision first attempt

@@
-\section{Qualitative findings}
-
-In addition to modeling data and confirming our hypotheses, we performed a qualitative analysis.
-The primary goal was to identify trends  that were repeated between participants.
-These trends can then be used to inform and justify interventions to change the types of ideas produced.
+\section{Qualitative analysis}
 
 We selected 30 brainstorming runs (5 randomly from each condition).
-The primary author applied an open coding strategy, flagging ideas within runs that shared common characteristics, such as length, wording, or core problem solution.
-
-In particular, we identify classes of ideas within runs, often temporally close, that hold common characteristics.
+The primary author applied an open coding strategy, identifying classes of ideas within runs that hold common characteristics.
 For each class that was observed across multiple participants, we infer the existence of a \emph{strategy}: a plan or rule that aids in generating ideas.
 We describe the observed strategies in detail below.
-Furthermore, we use the presence or lack of presence of strategies to define \emph{personas}.
+We describe the distribution of these strategies across runs and comment on variability between participants.
 
@@
 
-A \emph{strategy} is a plan, rule, or idea that is used by a participant to generate brainstorming responses.
-Without debriefing participants, we cannot infer the goal of a strategy, but define it in terms of its observable effects.
-These effects are observed as repeated, cross-participant classes of similar ideas.
-Not all participants utilized all strategies in their brainstorming runs.
-
+A \emph{strategy} is a plan or rule, that is used by a participant to generate brainstorming responses.
 We identified 3 \emph{orthogonal} strategy categories - each category of strategies can be employed by a participant with or without any of the other categories.
 The first, \emph{problem scoping}, is a process in which a participant chooses a limited implication or component of the problem and provides multiple specific solutions.
-The second category expands on the \emph{riffing} phenomenon introduced earlier, identifying 4 different types of riffing explored.
-Finally, \emph{language use} strategies relate not to the meaning of the ideas generated, but to the way they are expressed.
+The second category expands on the \emph{riffing} phenomenon introduced earlier, identifying 4 different types of riffing.
 %Finally, the \emph{humour} strategy emphasizes generating ideas that are amusing rather than practical.
+Finally, the \emph{partial solution} strategy entails creating an idea that requires further idea generation to solve the problem.
 
@@
 Under the problem scoping strategy, the original problem is transformed to a new problem of reduced scope.
-An MP3 player has many functions which are difficult to leverage simultaneously.
-Thus, many transformations are possible, for example: brainstorm uses for \emph{an audio output device}; brainstorm uses for a small box-shaped object; brainstorm uses for \emph{a hard drive}; and so on.
+Many transformations of the MP3 problem are possible, for example: brainstorm uses for \emph{an audio output device}; brainstorm uses for a small box-shaped object; brainstorm uses for \emph{a hard drive}; and so on.
 The participant chooses a problem transformation and generates solutions, each of which is also a solution to the original problem.
@@
 
-The participant's responses satisfy the transformed question "brainstorming uses for an audio output device".
-The resulting ideas no not rely on or acknowledge other aspects of the problem, such as portability, the screen, the ability to store files, or the appearance or size of the device.
-Nonetheless, each response is also a valid response to the original problem.
+The participant's responses satisfy the transformed question "brainstorm uses for an audio output device". Resulting ideas no not rely on other aspects of the problem.
 
-We distinguish between kinds of problem scoping.
+We distinguish between two types of problem scoping.
 \emph{Inward scoping} transformations reduce scope by limiting consideration to a single detail of the problem.
@@
 
-\emph{Outward scoping} transformations reduce complexity by stripping all meaningful details.
-In the case of the MP3 player problem, this can be understood as essentially reducing the problem to "brainstorm uses for a thing".
-For example (PXXX):
+\emph{Outward scoping} transformations reduce complexity by stripping all meaningful details. For example (PXXX):
 
@@
     \item use in abstract art
-    \item recycle
 \end{itemize}
 
-The inward scoping solutions were applicable to the transformed problem and the original problem, but not many others.
-In contrast, the outward scoping solutions could be applied to a wide body of problems unrelated to the original, such as "brainstorm uses for a broom".
+These solutions could be applied to a wide body of problems unrelated to the original, such as "brainstorm uses for a broom".
+In contrast, inward scoping solutions are applicable to the transformed problem and the original problem, but not many others.
 Thus, a useful rule for distinguishing the types of scoping is to consider whether the \emph{set of problems solved by the solutions} is small, or large.
 
-These two types of scoping cover almost all responses.
-However, some responses employ no scoping whatsoever.
-These responses utilize the device exactly as it was intended, as a portable MP3 player.
-For example, "get it fixed and give it to a needy kid".
+Some responses employ no scoping whatsoever. These responses utilize the device exactly as it was intended, as a portable MP3 player. For example, "get it fixed and give it to a needy kid" (PXXX).
 
@@
 We found riffing manifested itself many ways, and identified four distinct riffing strategies used by participants: generalization riffing, repeat riffing, hold riffing, and continuation riffing.
-We describe each of these in detail, and then touch on three common phenomena found under riffing strategies.
 
@@
     \item We could use them in place of rocks (to throw at things, to use in pavement.)
-    \item We could use them to back up our music collections on our current devices.
 \end{itemize}
@@
 
-The strategies above describe techniques for taking one idea and transforming it into another. In this section, we briefly describe patterns of \emph{where} participants implemented these strategies.
+The strategies above describe techniques for taking one idea and transforming it into another. In this section, we describe patterns of \emph{where} participants implemented these strategies.
 
-The most natural separation between a riffed idea and its source is none - the riffed idea directly follows the source.
-We call this \emph{linear riffing}.
-Some participants \emph{pair} their riffs.
+The most natural separation between a riffed idea and its source is none - the riffed idea directly follows the source. We call this \emph{linear riffing}.
+\emph{Reach backs} entail riffing on an idea that occurred further back in the run than the previous response.
+
+Finally, some participants \emph{pair} their riffs.
 Paired riffing is a special case on linear riffing in which only a single riffed idea is produced, and no further riffs are produced on that source for the remainder of the run.
@@
 \end{itemize}
-
-Finally, \emph{reach backs} comprise all other riffs. reaching back entails riffing on an idea that occurred further back in the run than the previous response.
 
@@
     \item create a new device
-    \item reuse all working parts
 \end{itemize}
@@
 In these responses, there is a clear recommendation that we recycle components of the MP3 player to produce new electronics, but it is unclear what we would make or how we would use the parts.
-Essentially, we are left with new brainstorming problems to solve.
 
-As with riffing and problem scoping, there are four sub-strategies of partial solutions: problem scoping without solution, goal without plan, plan without goal, and passing the problem.
-
-Above, we described the problem scoping strategy.
-Some participants performed \emph{problem scoping without solution}, a variant of this strategy in which problem scoping is the only component of the response.
-For example, on of PXXX's responses to the MP3 player problem was to "Remove glass screen to make something".
-In this case, the participant has provided a scoping of the problem (focus on the glass component) but has not actually provided a solution to the scoped problem.
-
+There are four sub-strategies of partial solutions: problem scoping without solution, goal without plan, plan without goal, and passing the problem. 
 In some responses, an end goal has been established, but any implementation of the goal requires further idea generation.
 We call this strategy \emph{goal without plan}.
-The list of responses at the beginning of this section are examples of the goal without plan strategy.
+The above responses are examples of the goal without plan strategy.
 The goal of creating a new electronic device is given, but it is not clear what will be created or how the materials will be re-used.
@@
 The inverse of this strategy is \emph{plan without goal}, as demonstrated in these responses (also from PXXX):
-
 \begin{itemize}
@@
 \end{itemize}
-
 In this case, the operations of melting down the device or checking for working parts can be performed, but don't provide any obvious benefit.
 
-In many cases, responses that do not have an explicit goal or plan nonetheless imply one.
-For example, "throw it in the ocean" (PXXX) is an action performed with the MP3 player with no stated goal, but the implied goal is to get rid of the broken device.
-In these cases, as judged by the coder, responses were not assumed to employ a partial solution strategy.
+Above, we described the problem scoping strategy. Some participants performed \emph{problem scoping without solution}, in which problem scoping is the only component of the response.
+For example, one of PXXX's responses to the MP3 player problem was to "Remove glass screen to make something".
+In this case, the participant has provided a scoping of the problem (focus on the glass component) but has not actually provided a solution to the scoped problem.
 
-Finally, responses in the \emph{pass the problem} strategy simply relocate the need for idea generation to a third party. These responses from PXXX provide an example:
+Responses that do not have an explicit goal or plan may nonetheless imply one.
+For example, "throw it in the ocean" (PXXX) has the implied goal of disposing of the broken device.
+In these cases, responses do not employ a partial solution strategy.
+
+Finally, responses in the \emph{pass the problem} strategy relocate the need for idea generation to a third party. These responses from PXXX provide an example:
 
 \begin{itemize}
-    \item donate to less fortunate that are trying to assist the homeless
-    \item give to university
     \item give to non-profit that can benefit
@@
 
-These solve the original problem, but simultaneously re-open the problem under slightly changed conditions.
-In this case, a university/non-profit/other body must determine what to do with broken MP3 players.
+These solve the original problem, but simultaneously re-open the problem under slightly changed conditions. In this case, a third party must determine what to do with broken MP3 players.
 
@@
 
-In this section, we briefly describe how strategies were employed by participants.
+In this section, we DESCRIBE how strategies were employed by participants.
 The primary author again coded 30 randomly selected runs (5 from each condition, 1150 instances total) for the strategies identified above.
-Note that examples of riffing identified in this section are distinct from the algorithmic/quantitative identification of riffing described in the method section above.
+Note that examples of riffing identified in this section are distinct from the category tree identification of riffing described in the method section above.
 The prevalence of each strategy as a function of total number of instances is given in Table TAB, and a comparative plot in Figure FIG.
@@
 
-\subsubsection{Strategy co-occurrence}
+%\subsubsection{Strategy co-occurrence}
 
-Of final interest was the question of which strategies occurred together.
+%Of final interest was the question of which strategies occurred together.
 For each run, a count was taken of the number of occurrences of each strategy, and the Pearson r of these counts was taken for each pairing of strategies to give a rough measure of how the strategies were grouped in runs.
@@
 
-\begin{table}
-    \begin{tabular}{l l l}
-        \textbf{strategy 1} & \textbf{strategy 2} & \textbf{Pearson r} \\
-        \hline
-        passing the problem & repeat riffing & 0.89 \\
-        hold riffing & outward scoping & 0.85 \\
-        pairing & scoping without solution & 0.64 \\
-        reaching back & hold riffing & 0.64 \\
-        pairing & outward scoping & 0.63 \\
-        reaching back & outward scoping & 0.62 \\
-        plan without goal & hold riffing & 0.61 \\
-    \end{tabular}
-\end{table}
+%\begin{table}
+%    \begin{tabular}{l l l}
+%        \textbf{strategy 1} & \textbf{strategy 2} & \textbf{Pearson r} \\
+%        \hline
+%        passing the problem & repeat riffing & 0.89 \\
+%        hold riffing & outward scoping & 0.85 \\
+%        pairing & scoping without solution & 0.64 \\
+%        reaching back & hold riffing & 0.64 \\
+%        pairing & outward scoping & 0.63 \\
+%        reaching back & outward scoping & 0.62 \\
+%        plan without goal & hold riffing & 0.61 \\
+%    \end{tabular}
+%\end{table}
 
-Passing the problem and repeat riffing were highly correlated (0.89), indicating that responses utilizing pass the problem were likely to be repeated many times to generate more responses.
+%Passing the problem and repeat riffing were highly correlated (0.89), indicating that responses utilizing pass the problem were likely to be repeated many times to generate more responses.
 For each of the other correlations, the potential implications less obvious.
 Further analysis is required to tease out the implications of strategy co-occurrence.
+
+\subsection{Variability of Worker Input}
+While we observed that individuals are likely to draw from a common pool of ideas early on in the brainstorming process, there were outliers in the data.
+For example, one individual in the 75 ideas condition consistently produced some of most original ideas for old MP3 players:
+
+\begin{itemize}
+\item With displays - could be attached to a small solar panel (like the USB chargers w/multi attachments, including solar and car, etc.) and all the power could be sent to the screen as a small, portable light, then sent to countries lacking power in rural areas.
+\item They could be loaded with coordinates or other info and sealed in containers, then used in ocean current studies - once the container was collected they could be plugged in and read, and the data could be sent to whomever.
+\item They could be plugged in at a base camp at the bottom of Everest, and then potential climbers could record their name, loved ones' address, and any last words before attempting the climb. 
+\end{itemize}
+
+This individual's responses were responsible for clearly raising the observed uniqueness of ideas in the 75 condition when plotted against the other conditions.
+In contrast, other individuals filled their quotas with excessive repeat riffing chains.
+
+This variability in worker responses prevented us from modeling any potential effects that the number of questions requested might have on individuals' responses.
+Thus, while it is clear that there is little advantage to requesting 5 or 10 ideas from workers, it is not clear whether it is preferable to request 20, 50, 75, or 100 ideas.
+We suspect that requesting 50 or 75 is the ``sweet'' spot, but are not able to empirically support that recommendation at this time.

### 2013-10-17 chi2014/related.tex
10 page revision first attempt

@@
 Following Osborne's work, Taylor et al.\ established that brainstorming by \emph{nominal groups}, or individuals working in isolation from one another, yields better results than collocated groups in terms of number of ideas generated \cite{taylor_does_1958}.
-These effects can be partially explained by the absence of counterproductive social forces, such as production blocking (the inability for multiple people to speak at once) and the fear of judgment.
+These effects can be partially explained by the absence of counterproductive social forces, such as fear of judgment.
 Bouchard and Hare \cite{bouchard_jr_size_1970} later found that nominal brainstorming groups are able to generate quantities of ideas linear in the number of members in the group.
@@
 Electronic brainstorming is a variant in which individuals brainstorm at separate computer terminals, with the exchange of ideas performed over the network \cite{gallupe_electronic_1992}.
-Participants can enter ideas while simultaneously viewing the ideas of all other ideas, which are anonymously presented on-screen.
+Participants enter ideas while others' ideas are anonymously presented on-screen.
 As with nominal brainstorming, Gallupe \cite{gallupe_electronic_1992} observed that electronic brainstorming can reduce counterproductive social effects.
 However, Pinsonneault et al.\ \cite{pinsonneault_electronic_1999} identify productivity impediments introduced by group electronic brainstorming, such as being distracted by other ideas appearing, and individuals witholding ideas they feel are derivative of others.
-Given these drawbacks, their results suggest there is little evidence to support the proposition that electronic brainstorming outperforms nominal brainstorming.
+They find little evidence to support the proposition that electronic brainstorming outperforms nominal brainstorming.
 
-Finally, Diehl and Stroebe found that groups that generate good ideas also generate \emph{many} ideas \cite{diehl_productivity_1987}, a finding that has been verified many times \cite{briggs1997quality, parnes1959effects, parnes_effects_1961, shah2003metrics, cross1996creativity}.
+Finally, Diehl and Stroebe found that groups that generate many ideas also generate \emph{good} ideas \cite{diehl_productivity_1987}, a finding that has been verified many times \cite{briggs1997quality, parnes1959effects, parnes_effects_1961, shah2003metrics, cross1996creativity}.
 
@@
 SIAM defines two stages of idea generation: the generation of ideas based on working memory, and idea production within a single activated ``image'' (where an image can be thought of as a category of ideas).
-This model makes predictions applicable to all brainstorming contexts, with one of the most important predictions being that an idea is more likely to be followed by an idea in the same category than would be expected by chance.
+This model makes predictions, one of the most important being that an idea is more likely to be followed by an idea in the same category than would be expected by chance.
 Their model also predicts that new ideas in the same category as the previous idea are generated faster than ideas between categories.
@@
 The most common approach to measure novelty and appropriateness is to ask domain experts to subjectively make scale-rated assessment of each \cite{lewis2011affective, amabile_1983}.
-Depending on the types of artifacts and researchers' interest, these two measures are sometimes combined into one measure of quality \cite{little2010exploring}.
-
-Instead of subjective ratings, some researchers opt for a measure that can be derived objectively.
+These two measures are sometimes combined into one measure of quality \cite{little2010exploring}.
+Some researchers opt for a measure that can be derived objectively.
 For example, Jansson et al.
-used `o' scores, derived from the frequency with which an idea occurs, to evaluate products novelty \cite{jansson_design_1991}.
+used `o' scores, derived from the frequency with which an idea occurs, to evaluate novelty \cite{jansson_design_1991}.
 We adopt this use of o-score in our later analyses.
@@
 Flexibility is often measured by the number of explored ``categories'' in a design space, or by subjective rating from domain experts \cite{lewis2011affective, marsh1996examples}.
-Category or feature-based quantities also yield other metrics such as within-categories fluency and the length of train-of-thought when participants focus on a particular category \cite{nijstad_how_2006}.
+Category or feature-based quantities yield metrics such as within-categories fluency and length of train-of-thought when participants focus on a particular category \cite{nijstad_how_2006}.
 Building in this past line of work, we consider both individual ideas and categories of ideas.
@@
 The simplest imaginable instantiation of brainstorming in a microtask marketplace mirrors that of nominal brainstorming.
-More sophisticated brainstorming environments are possible, such as those that provide awareness of ideas that others have generated (electronic brainstorming), or those that rely on the crowd to organize or rank the ideas produced by others, as in evolutionary techniques \cite{yu_cooks_2011}.
+More sophisticated environments are possible, such as those that forward ideas that others have generated (electronic brainstorming).
 %Examples for these environments are platforms for generating social innovation ideas or product concepts like OpenIDEO, Quirky and Threadless that employ structured process where participants take on different responsibility from generating ideas to evaluating others' ideas.
-In this paper, we constrain our focus to nominal brainstorming in this environment with the goal of establishing a set of baselines.
+We constrain our focus to nominal brainstorming in this environment with the goal of establishing baselines.
 
@@
 For example, Yu and Nickerson created an evolutionary crowd algorithm to generate creative chair designs \cite{yu_cooks_2011}.
-Their work demonstrates the creative potential of the crowd, but does not specifically address brainstorming or analogous processes in which participants generate many ideas.
+Their work demonstrates the creative potential of the crowd, but does not specifically address brainstorming or processes in which participants generate many ideas.
 
@@
 Participants in the parallel condition worked in isolation without seeing other participants' ideas.
-Subjective ratings of the quality of ideas reveal that while the ideas from the iterative condition were rated higher on average, the best ideas from the parallel condition were rated higher than the best ideas from the iterative conditions.
-Thus, the parallel process was found to be more likely to generate names with high ratings, advocating the case for nominal brainstorming over traditional electronic brainstorming.
-However, the study did not mention categorization of generated ideas, an important measure on flexibility of idea generation \cite{lewis2011affective, nijstad_how_2006, finke1992creative, shah2003metrics}.
+The parallel process was found to be more likely to generate names with high ratings, advocating the case for nominal brainstorming.
+The study did not mention categorization of generated ideas, an important measure on flexibility of idea generation \cite{lewis2011affective, nijstad_how_2006, finke1992creative, shah2003metrics}.
 Each participant also generated only five ideas, leaving the number of ideas requested per participant an open dimension to be explored.
 
-The nature of the microtask marketplace may also require alternative ways of structuring the brainstorming activity in this environment.
+The nature of the microtask marketplace may also require alternative ways of structuring the brainstorming activity.
 % calls for an alternative structure for brainstorming tasks in this setting.
@@
 found high variance of effort from Amazon Mechanical Turk workers \cite{soylent}.
-They characterized two worker personas at opposite ends of the effort spectrum: the {\em Eager Beaver\/} who goes beyond task requirement and the {\em Lazy Turker\/} who does as little work as necessary to get paid.
+They characterized two worker personas at opposite ends of the effort spectrum: the {\em Eager Beaver\/} who goes beyond task requirement and the {\em Lazy Turker\/} who does as little work as necessary.
 Given this finding, a brainstorming task that asks participants to generate as many ideas as they can in a fixed time period might not be cost-effective, since a {\em Lazy Turker\/} can simply fill in a few ideas and wait for the time to run out instead of exerting full effort.
@@
 
-In summary, while brainstorming has been examined in traditional and electronic settings, its application in crowd marketplaces has received limited attention, leaving many open questions about how this process occurs in this environment.
+In summary, while brainstorming has been examined in traditional and electronic settings, its application in crowd marketplaces has received limited attention, leaving open questions.

### 2013-10-17 chi2014/results.tex
10 page revision first attempt

@@
 where $n_{ideas}$ represents the number of unique ideas, $y_{scale}$ serves to scale the model, $x$ is the response number, and $rate$ models the rate of novel ideas.
-Importantly, this simplified model does not model either the condition or the effects of workers.
+Importantly, this model does not model either the condition or the effects of workers.
 In our analysis of the data, we found that individual workers introduced considerable variation, preventing us from separating the effects of the condition from the worker.
@@
 
-The visibly lower rate of growth for novel categories growth suggests workers are more likely to produce new ideas within existing categories, rather than create both novel ideas that also represent novel categories.
+The visibly lower rate of growth for novel categories suggests workers are more likely to produce new ideas within existing categories, rather than create both novel ideas that also represent novel categories.
 We more closely examine characteristics of individuals' brainstorming runs next.
@@
 \subsubsection{Most Popular Ideas}
-Our data suggest that individuals typically produce ideas from a common set of idea categories early in the brainstorming process, then diverge and produce more original ideas.
-This phenomenon can be observed by examining the o-scores of ideas and idea categories.
 
 Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order} show the o-scores for response instances and idea categories, within a sliding window of size 10.
-Each point represents the mean o-score within that sliding window, with standard error bars provided for each point.
+Each point represents the mean o-score within that sliding window, with standard error bars.
 
@@
 
-These graphs lend support for hypothesis 2, the notion that there is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.
+This lends support for hypothesis 2, the notion that there is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.
 
-To more explicitly confirm this hypothesis, we extracted the top 5\% most common ideas (66 in total).
+To confirm this hypothesis, we extracted the top 5\% most common ideas (66 in total).
 If individuals are more likely to initially generate the same set of ideas, the top 5\% of the ideas should be more likely to appear in the first five instances of a run than they are to appear in any of the following instances.
 We introduce two Bernoulli random variables to model this.
-The first, defined by parameter $\theta_f$ represents the probability that a response instance is in the top 5\% most common ideas, given it is found in the first five instances of a run (i.e., $P(\theta_f|top)$).
-The second, defined by parameter $\theta_l$, is the probability that an instance is in the top \% given it is one of the \emph{last} ($\textgreater5$) instances in a run (i.e., $P(\theta_l|\overline{top})$).
+The first, defined by parameter $\theta_f$ represents the probability that a response instance is in the top 5\% most common ideas, given it is found in the first five instances of a run (i.e., $P(\theta_f|first)$).
+The second, defined by parameter $\theta_l$, is the probability that an instance is in the top \% given it is one of the \emph{last} ($\textgreater5$) instances in a run (i.e., $P(\theta_l|\overline{first})$).
 
@@
 We chose this constraint so that no run could contribute evidence to $\theta_f$ without also contributing to $\theta_l$.
-Using an uninformative beta prior, the resulting parameters are $\theta_f = 0.56$ (95\% HDI 0.52-0.61) and $\theta_l = 0.32$ (95\% HDI 0.30-0.34).
+Using an uninformative beta prior, the resulting parameters are $\theta_f = 0.56$ (HDI 0.52-0.61) and $\theta_l = 0.32$ (HDI 0.30-0.34).
 This difference is significant, supporting the hypothesis that common ideas are over-represented in the early part of a brainstorming run.
@@
 Figures~\ref{fig:idea_oscore_hyp4} and \ref{fig:cat_oscore_hyp4} show the data and fitted models.
-Examining the HDIs of the resultant model parameters shows significant differences between the models, confirming hypothesis 3.
+We found significant differences between model parameters, confirming hypothesis 3.
 
-Collectively, these results provide a clear guideline for maximizing originality in brainstorming tasks on microtask marketplaces: it is more cost-effective to solicit $\textgreater20$ ideas from a smaller pool of participants than to ask for fewer ideas from more individuals.
+Collectively, these results provide a guideline for maximizing originality in brainstorming tasks on microtask marketplaces: it is more effective to solicit $\textgreater20$ ideas from a smaller pool of participants than to ask for fewer ideas from more individuals.
 
@@
 \subsection{Idea Generation Patterns}
-Given the clear differences in idea characteristics as a function of response number, we now consider idea generation patterns observable from the data.
-In this section, we examine these patterns in the context of hypotheses 4 and 5, the hypotheses that draw upon the SIAM model of brainstorming.
+Given the differences in originality and rarity as a function of response number, we now consider idea generation patterns observable from the data.
+We examine these patterns in the context of hypotheses 4 and 5, which draw upon the SIAM model of brainstorming.
 This model predicts individuals will generate ideas within ``images'' or categories of ideas (i.e., ``riff'' on an idea or concept) before switching to a new category.
@@
 Using a uniform beta prior, we find the probability of a response instance being in the same idea category as the previous to be 15.6\% (HDI 14.3\% - 17.0\%).
-To establish a baseline to understand whether this probability is greater than chance, the most popular category tree in our data set is \emph{music player}, representing 5.65\% of all instances.
-Thus, an upper bound on the greatest probability of an instance being in the same category as the previous instance, by chance, is 5.65\%, well below the 14.3\% lower bound of the calculated HDI.
+To understand whether this probability is greater than chance, we use the most popular category tree in our data set (\emph{music player}), representing 5.65\% of all instances as an upper bound.
+This well below the 14.3\% lower bound of the calculated HDI.
 Thus, we accept the hypothesis that an instance follows a previous instance in the same category with greater probability than is explained by random chance.

### 2013-10-18 chi2014/method.tex

@@
 
-Initially, workers were randomly paired with a question from a pool of questions.
-However, we eventually dropped all other questions, so workers would only receive the MP3 player question.
-%When testing multiple questions, we attempted to prevent workers from answering the same question twice (for example, if they chose the HIT requesting 10 responses, then the HIT requesting 20 responses) .
-However, we were not always successful in preventing this from happening.
-Thus, we only use a worker's first set of responses if they answered the question multiple times.
-
 %Because workers could choose which \emph{number of ideas} condition to participate in, self-selection bias is a reasonable concern.
@@
 Upon accepting a HIT, participants were asked to give consent for participating in the study, and informed that they could leave the study at any point without financial consequences.
-Under more realistic, real-world conditions, workers would be expected to complete all responses to receive compensation.
+Under real-world conditions, workers would be expected to complete all responses to receive compensation.
 
@@
 
-In total, 154 workers submitted 170 HITs, for a total of 3292 responses.
+154 workers submitted 170 HITs, for a total of 3292 responses.
 As we received data in the unlimited condition, it became obvious that this was a poor strategy to employ -- the number of responses received was less than 15 in all cases.
@@
 
-If a worker completed multiple HITs (i.e., generated responses for multiple conditions), there is the possibility that their responses and idea generation processes evolved over time, making it difficult to compare the responses to their second HIT with responses from others who are performing the same HIT for the first time.
+If a worker completed multiple HITs (i.e., generated responses for multiple conditions), there is the possibility that their responses and idea generation processes evolved over time.
 Accordingly, we retain only the first HIT submitted by each worker.
@@
 One of the primary goals of this research is to model the rate at which novel ideas and categories of ideas arrive.
-A significant challenge to creating these models is efficiently identifying what ideas and categories of ideas are similar to one another, given the volume of data collected.
-To address this problem, we organized the data into sets of hierarchical clusters, where the clusters encode similarity relationships between ideas.
-Specifically, parent nodes are more generalized versions of child nodes.
-With this information encoded in the clusters, we are then able to calculate quantities such as the rate of novel ideas or novel idea categories, as a function of responses received.
+A significant challenge to creating these models is efficiently identifying what ideas and categories of ideas are similar to one another.
+To address this problem, we organized the data into sets of hierarchical clusters, where parent nodes are more generalized versions of child nodes.
+With this relationship, we are then able to calculate quantities such as the rate of novel ideas or novel idea categories, as a function of responses received.
 
@@
 A researcher not involved in the project randomly sampled 40 category trees to assess whether ideas within the tree should be there, and whether parent-child relationships correctly encoded generalizations of concepts.
-10\% of the instances were marked as more specific than their associated idea, while 0.7\% of the instances were judged to be misclassified with their associated cluster.
+10\% of the instances were marked as more specific than their associated idea, while 0.7\% of the instances were judged to be misclassified with their associated cluster tree.
 

### 2013-10-18 chi2014/conclusions.tex
down to 11 pages...

@@
 Our results show that ideas are more likely to be followed by one from the same category, and that changing categories takes time.
-Combining these models with others (such as natural language measures for text similarity), one could imagine constructing a posterior belief model to infer whether a participant is currently riffing or not.
-If we can infer riffing and the general categories in which workers are riffing, one could imagine techniques to guide the crowd to explore different areas of the design space.
-For example, the system could encourage workers to generate ideas within new or rare categories.
+One could imagine constructing a posterior belief model to infer whether a participant is currently riffing or not, guide the crowd to explore ideas within new or rare categories.
 
@@
 \item Common ideas are more likely to be proposed early in a run. 
-\item Conversely, ideas beyond the first 20 are rarer and more original, as measured by o-scores. This relationship also holds for idea categories.
+\item Conversely, ideas beyond the first 20 are rarer and more original, as measured by o-scores. %This relationship also holds for idea categories.
 \item Ideas are more likely to be followed by another idea from the same category than would be expected by random chance.

### 2013-10-18 chi2014/goals.tex
10 page revision first attempt
down to 11 pages...

@@
 
-In modeling these quantities, we wish to explore the following hypotheses:
+In modeling these quantities, we explore the following hypotheses:
 
@@
 
-\textbf{Hypothesis 3:} O-scores generated in the second half of a brainstorming session will be higher than ideas generated in the first half.
+%\textbf{Hypothesis 3:} O-scores generated in the second half of a brainstorming session will be higher than ideas generated in the first half.
 This hypothesis provides a complementary perspective to the former.
@@
 
-\textbf{Hypothesis 4:} An idea from one semantic category of ideas should more often be followed by an idea from the same category than expected by random chance.
+\textbf{Hypothesis 3:} An idea from one semantic category of ideas should more often be followed by an idea from the same category than expected by random chance.
 
-\textbf{Hypothesis 5:} The time it takes to generate an idea should be longer when changing semantic categories than when generating ideas within a category.
+\textbf{Hypothesis 4:} The time it takes to generate an idea should be longer when changing semantic categories than when generating ideas within a category.

### 2013-10-18 chi2014/introduction.tex
down to 11 pages...

@@
 In this project, he asked 10,000 individuals to draw a ``sheep facing left,'' yielding a wide range of unique results.
-More recently, there has been interest in systems that support crowd-based design.
-For example, Yu and Nickerson developed a system that combines genetic algorithms with design input from crowd workers to explore a design space \cite{yu_cooks_2011}.
 
@@
 \item Individuals' patterns of idea generation mimic those identified by Nijstad and Stroebe's SIAM (search for ideas in associative memory) model \cite{nijstad_how_2006}: Workers produce a set of ideas from working memory, then switch strategies to generate ideas that are permutations of other ideas, a process we dub {\em riffing\/}.
-\item We also replicate Nijstad and Stroebe's findings with respect to the time it takes to generate a new idea \cite{nijstad_how_2006}.
-\item Requesting ``as many ideas as possible'' for a fixed price (fixed for the entire HIT, rather than per idea) is inadvisable, as it leads to a relatively low number of ideas produced (in our experiments, fewer than 15).
+We also replicate Nijstad and Stroebe's findings with respect to the time it takes to generate a new idea \cite{nijstad_how_2006}.
 \item we identify and describe a set of strategies employed by crowd workers that result in distinct kinds of produced ideas

### 2013-10-18 chi2014/method.tex
down to 11 pages...

@@
 \subsection{Brainstorming Question: Uses for an Old MP3 Player}
-In conducting this research, we tested several brainstorming questions, including classic brainstorming questions (e.g., ``provide alternative uses for a mop''); a request to brainstorm factors that influence weight gain and weight loss; and more goal-oriented questions, such as one requesting people to brainstorm ways to raise funds for a charity.
+In conducting this research, we tested several brainstorming questions, including classic brainstorming questions from prior work (e.g., ``provide alternative uses for a mop''); a request to brainstorm factors that influence weight gain and weight loss; and more goal-oriented questions, such as one requesting people to brainstorm ways to raise funds for a charity.
 Ultimately, we found the goal-oriented brainstorming questions to be the most compelling since they reflect what we believe would be real-world uses of brainstorming on Mechanical Turk.
@@
 
-In addition to the visible text input fields, HITs were instrumented to collect timing data.
-Specifically, we recorded the time of the first activation and last de-activation of each form element, the time the HIT was accepted, and the time it was submitted.
+In addition to the visible text input fields, HITs were instrumented to collect the time of the first activation and last de-activation of each form element.
 
@@
 
-If a worker completed multiple HITs (i.e., generated responses for multiple conditions), there is the possibility that their responses and idea generation processes evolved over time.
-Accordingly, we retain only the first HIT submitted by each worker.
+If a worker completed multiple HITs, there is the possibility that their responses and idea generation processes evolved over time. Accordingly, we retain only the first HIT submitted by each worker.
 
@@
 
-\begin{table}[h!]
-\begin{tabular}{r | l l l l l l l }
-    & 5 & 10 & 20 & 50 & 75 & 100 & all \\ \hline
-    participants & 57 & 39 & 21 & 9 & 10 & 10 & 146 \\
-    solutions & 283 & 372 & 413 & 450 & 635 & 855 & 3007
-\end{tabular}
-\caption{Data after removing unlimited condition and repeats}
-\label{tab:data_collected}
-\end{table}
+%\begin{table}[h!]
+%\begin{tabular}{r | l l l l l l l }
+%    & 5 & 10 & 20 & 50 & 75 & 100 & all \\ \hline
+%    participants & 57 & 39 & 21 & 9 & 10 & 10 & 146 \\
+%    solutions & 283 & 372 & 413 & 450 & 635 & 855 & 3007
+%\end{tabular}
+%\caption{Data after removing unlimited condition and repeats}
+%\label{tab:data_collected}
+%\end{table}
 
@@
 \subsection{Hierarchical Clustering}
-One of the primary goals of this research is to model the rate at which novel ideas and categories of ideas arrive.
-A significant challenge to creating these models is efficiently identifying what ideas and categories of ideas are similar to one another.
+A significant challenge to creating model of brainstorming is efficiently identifying what ideas and categories of ideas are similar to one another.
 To address this problem, we organized the data into sets of hierarchical clusters, where parent nodes are more generalized versions of child nodes.
-With this relationship, we are then able to calculate quantities such as the rate of novel ideas or novel idea categories, as a function of responses received.
+%With this relationship, we are then able to calculate quantities such as the rate of novel ideas or novel idea categories, as a function of responses received.
 
@@
 An \emph{idea} is a unique solution to the brainstorming problem.
-An instance is associated with exactly one \emph{idea}, though multiple instances can be associated with the same idea (i.e., when multiple workers come up with the same idea).
+An instance is associated with exactly one \emph{idea}, though multiple instances can be associated with the same idea.
 Multiple instances of the same idea do not need to be identically phrased to refer to the same idea.
@@
 
-Despite the differences in specificity between ``Coin Storage'' and the more generic containers, it is useful to encode the fact that these ideas are related.
+Despite the differences in specificity between ``Coin Storage'' and the more generic containers, it is useful to encode their relationship.
 We use hierarchical clusters to encode these relationships.

### 2013-10-18 chi2014/qualitative.tex
down to 11 pages...

@@
 We selected 30 brainstorming runs (5 randomly from each condition).
-The primary author applied an open coding strategy, identifying classes of ideas within runs that hold common characteristics.
-For each class that was observed across multiple participants, we infer the existence of a \emph{strategy}: a plan or rule that aids in generating ideas.
+The primary author applied an open coding strategy, identifying \emph{strategies}: plans or rules that aid in generating ideas.
 We describe the observed strategies in detail below.
@@
 A \emph{strategy} is a plan or rule, that is used by a participant to generate brainstorming responses.
-We identified 3 \emph{orthogonal} strategy categories - each category of strategies can be employed by a participant with or without any of the other categories.
-The first, \emph{problem scoping}, is a process in which a participant chooses a limited implication or component of the problem and provides multiple specific solutions.
-The second category expands on the \emph{riffing} phenomenon introduced earlier, identifying 4 different types of riffing.
+We infer the existence of strategies through similar classes of ideas observed between participants.
+We identified 3 strategy categories, each of which can be employed with or without any of the other categories.
+The first, \emph{problem scoping}, is a process in which a participant chooses a limited implication or component of the problem and provides multiple solutions.
+The second category expands on the \emph{riffing} phenomenon introduced earlier.
 %Finally, the \emph{humour} strategy emphasizes generating ideas that are amusing rather than practical.
@@
 Under the problem scoping strategy, the original problem is transformed to a new problem of reduced scope.
-Many transformations of the MP3 problem are possible, for example: brainstorm uses for \emph{an audio output device}; brainstorm uses for a small box-shaped object; brainstorm uses for \emph{a hard drive}; and so on.
+Many transformations of the MP3 problem are possible, for example: brainstorm uses for \emph{an audio output device}; brainstorm uses for \emph{a hard drive}; and so on.
 The participant chooses a problem transformation and generates solutions, each of which is also a solution to the original problem.
@@
     \item use them in museums to give information on various installations
-    \item have cities install them in tourist areas, so people can listen about where they are
+    %\item have cities install them in tourist areas, so people can listen about where they are
 \end{itemize}
@@
 
-\emph{Outward scoping} transformations reduce complexity by stripping all meaningful details. For example (PXXX):
+\emph{Outward scoping} transformations reduce complexity by stripping meaningful details. For example (PXXX):
 
@@
 Another type of riffing is \emph{repeat riffing}.
-Under this strategy, a participant produces multiple responses with only slight differences between them, such that the group of ideas could be easily summarized as one idea without significant loss of information.
+Under this strategy, a participant produces multiple responses, such that the group of ideas could be easily summarized as one idea without significant loss of information.
 For example (PXXX):
@@
 
-\emph{Hold riffing} is a strategy in which participants hold at least one element of a previous response constant when generating a new response, such that a new idea is formed. For example (PXXX):
+\emph{Hold riffing} is a strategy in which participants hold at least one element of a previous response constant when generating a new response. For example (PXXX):
 
@@
 
-\emph{Continuation riffing} is the strategy of creating another idea that relies on the implications or consequences of a previous idea.
-The new idea cannot be understood without the context of the previous.
-For example (PXXX):
+\emph{Continuation riffing} is the strategy of creating another idea that cannot be understood without the context of the previous. For example (PXXX):
 
@@
 
-The strategies above describe techniques for taking one idea and transforming it into another. In this section, we describe patterns of \emph{where} participants implemented these strategies.
+%The strategies above describe techniques for taking one idea and transforming it into another.
+In this section, we describe patterns of \emph{where} participants implemented strategies.
 
@@
 
-Finally, some participants \emph{pair} their riffs.
-Paired riffing is a special case on linear riffing in which only a single riffed idea is produced, and no further riffs are produced on that source for the remainder of the run.
-For example, paired repeat riffing from PXXX:
+Finally, some participants \emph{pair} their riffs. Paired riffing is a special case on linear riffing in which only a single riffed idea is produced. For example, paired repeat riffing from PXXX:
 
@@
     \item old parts can use to make something
-    \item create a new device
+    %\item create a new device
 \end{itemize}
 
-In these responses, there is a clear recommendation that we recycle components of the MP3 player to produce new electronics, but it is unclear what we would make or how we would use the parts.
+In these responses, there is a recommendation that we produce new electronics, but it is unclear what we would make.
 
 There are four sub-strategies of partial solutions: problem scoping without solution, goal without plan, plan without goal, and passing the problem.
-In some responses, an end goal has been established, but any implementation of the goal requires further idea generation.
+
+In some responses, an end goal has been established, but executing the goal requires additional information.
 We call this strategy \emph{goal without plan}.
 The above responses are examples of the goal without plan strategy.
-The goal of creating a new electronic device is given, but it is not clear what will be created or how the materials will be re-used.
 
@@
 
-Above, we described the problem scoping strategy. Some participants performed \emph{problem scoping without solution}, in which problem scoping is the only component of the response.
-For example, one of PXXX's responses to the MP3 player problem was to "Remove glass screen to make something".
-In this case, the participant has provided a scoping of the problem (focus on the glass component) but has not actually provided a solution to the scoped problem.
-
 Responses that do not have an explicit goal or plan may nonetheless imply one.
@@
 In these cases, responses do not employ a partial solution strategy.
+
+Above, we described the problem scoping strategy. Some participants performed \emph{problem scoping without solution}, in which problem scoping is the only component of the response.
+For example (PXXX): "Remove glass screen to make something".
+In this case, the participant has provided a scoping of the problem (focus on the glass component) but has not actually provided a solution to the scoped problem.
 
@@
 
-These solve the original problem, but simultaneously re-open the problem under slightly changed conditions. In this case, a third party must determine what to do with broken MP3 players.
+In this case, a third party must determine what to do with broken MP3 players.
 
@@
 Note that examples of riffing identified in this section are distinct from the category tree identification of riffing described in the method section above.
-The prevalence of each strategy as a function of total number of instances is given in Table TAB, and a comparative plot in Figure FIG.
+The prevalence of each strategy is given in Figure FIG.
 
-\begin{table}
-    \begin{tabular}{r | l l}
-        \textbf{strategy} & \textbf{\# instances} & \textbf{\% of instances} \\
-        inward scoping & 267 & 23.2 \\
-        outward scoping & 828 & 72 \\
-        no scoping & 55 & 4.8 \\
-        \hline \hline
-        hold riffing & 278 & 24.2 \\
-        continuation riffing & 3 & 0.3 \\
-        repeat riffing & 119 & 10.3 \\
-        generalization riffing & 48 & 4.2 \\
-        \hline
-        any riffing & 448 & 39.0 \\
-        pair & 108 & 9.4 \\
-        reach back & 129 & 11.2 \\
-        \hline \hline
-        scoping without solution & 3 & 0.3 \\
-        passing the problem & 69 & 6 \\
-        plan without goal & 58 & 5.0 \\
-        goal without plan & 58 & 5.0 \\ 
-        \hline
-        any partial solution & 188 & 16.3 \\
-    \end{tabular}
-\end{table}
+%\begin{table}
+%    \begin{tabular}{r | l l}
+%        \textbf{strategy} & \textbf{\# instances} & \textbf{\% of instances} \\
+%        inward scoping & 267 & 23.2 \\
+%        outward scoping & 828 & 72 \\
+%        no scoping & 55 & 4.8 \\
+%        \hline \hline
+%        hold riffing & 278 & 24.2 \\
+%        continuation riffing & 3 & 0.3 \\
+%        repeat riffing & 119 & 10.3 \\
+%        generalization riffing & 48 & 4.2 \\
+%        \hline
+%        any riffing & 448 & 39.0 \\
+%        pair & 108 & 9.4 \\
+%        reach back & 129 & 11.2 \\
+%        \hline \hline
+%        scoping without solution & 3 & 0.3 \\
+%        passing the problem & 69 & 6 \\
+%        plan without goal & 58 & 5.0 \\
+%        goal without plan & 58 & 5.0 \\ 
+%        \hline
+%        any partial solution & 188 & 16.3 \\
+%    \end{tabular}
+%\end{table}
 
@@
 
-Outward scoping is by far the most commons strategy, occurring in 72\% of instances.
+Outward scoping is by far the most common strategy, occurring in 72\% of instances.
 Riffing is also very common, with 40\% of instances employing some form of riffing.
@@
 
-The distributions of strategy occurence across position in run tend towards uniform when there are more than a few examples.
-The notable exception to this is problem scoping.
+The distributions of strategy occurence across position in run tend towards uniform.
+The notable exceptions are problem scoping and reach-backs.
 There are fewer occurences of the inward scoping strategy later in brainstorming runs.
-
-The number of reach-backs increase with position. This may simply be because there are more ideas to draw from for inspiration later in a run.
+The number of reach-backs increase with position.
 
@@
 \item They could be loaded with coordinates or other info and sealed in containers, then used in ocean current studies - once the container was collected they could be plugged in and read, and the data could be sent to whomever.
-\item They could be plugged in at a base camp at the bottom of Everest, and then potential climbers could record their name, loved ones' address, and any last words before attempting the climb. 
+%\item They could be plugged in at a base camp at the bottom of Everest, and then potential climbers could record their name, loved ones' address, and any last words before attempting the climb. 
 \end{itemize}
 
-This individual's responses were responsible for clearly raising the observed uniqueness of ideas in the 75 condition when plotted against the other conditions.
+This individual's responses raised the observed uniqueness of ideas in the 75 condition. 
 In contrast, other individuals filled their quotas with excessive repeat riffing chains.

### 2013-10-18 chi2014/related.tex
down to 11 pages...

@@
 \subsection{Crowdsourced Brainstorming}
-The simplest imaginable instantiation of brainstorming in a microtask marketplace mirrors that of nominal brainstorming.
+%The simplest imaginable instantiation of brainstorming in a microtask marketplace mirrors that of nominal brainstorming.
 More sophisticated environments are possible, such as those that forward ideas that others have generated (electronic brainstorming).
@@
 The parallel process was found to be more likely to generate names with high ratings, advocating the case for nominal brainstorming.
-The study did not mention categorization of generated ideas, an important measure on flexibility of idea generation \cite{lewis2011affective, nijstad_how_2006, finke1992creative, shah2003metrics}.
-Each participant also generated only five ideas, leaving the number of ideas requested per participant an open dimension to be explored.
+The study did not mention categorization of generated ideas, an important measure on flexibility of idea generation \cite{lewis2011affective, nijstad_how_2006, finke1992creative, shah2003metrics}, nor did it explore requesting more than five ideas.
 
@@
 They characterized two worker personas at opposite ends of the effort spectrum: the {\em Eager Beaver\/} who goes beyond task requirement and the {\em Lazy Turker\/} who does as little work as necessary.
-Given this finding, a brainstorming task that asks participants to generate as many ideas as they can in a fixed time period might not be cost-effective, since a {\em Lazy Turker\/} can simply fill in a few ideas and wait for the time to run out instead of exerting full effort.
 

### 2013-10-18 chi2014/results.tex
down to 11 pages...

@@
 
-\subsubsection{Originality}
+The above plots also suggest that there is a pivot point around 20 ideas after which o-scores are higher.
 
-The increase in o-score over time in Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order} also supports hypothesis 3, the notion that ideas generated in the second half a brainstorming session have higher o-scores than those in the first half.
+%The increase in o-score over time in Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order} also supports hypothesis, the notion that ideas generated in the second half a brainstorming session have higher o-scores than those in the first half.
 We test a stronger guarantee of this hypothesis that is implied by our data: that ideas after the first 20 have higher o-scores than the first 20.
@@
 Partitioning the data into instances $\leq20$ and $\textgreater20$, we found that these distributions of o-scores are best modeled using beta distributions.
-Figures~\ref{fig:idea_oscore_hyp4} and \ref{fig:cat_oscore_hyp4} show the data and fitted models.
-We found significant differences between model parameters, confirming hypothesis 3.
-
+Figure~\ref{fig:idea_oscore_hyp4} shows the data and fitted models.
+We found significant differences between model parameters.
 Collectively, these results provide a guideline for maximizing originality in brainstorming tasks on microtask marketplaces: it is more effective to solicit $\textgreater20$ ideas from a smaller pool of participants than to ask for fewer ideas from more individuals.
@@
     \label{fig:idea_oscore_hyp4}
-    \includegraphics[width=0.9\columnwidth]{hyp4_cats}
-    \caption{Beta models for category o-score}
-    \label{fig:cat_oscore_hyp4}
+    %\includegraphics[width=0.9\columnwidth]{hyp4_cats}
+    %\caption{Beta models for category o-score}
+    %\label{fig:cat_oscore_hyp4}
 \end{figure}
@@
 
-In our clustered data, we can identify riffing by examining whether a response instance belongs to the same category tree as the previous instance.
+We can identify riffing by examining whether a response instance belongs to the same category tree as the previous instance.
 We can then derive a number of statistics from these data.
-
-To contextualize this process, below is an excerpt of a particular participant's brainstorming run.
+Below is an excerpt of a particular participant's brainstorming run.
 
@@
 
-The participant first touches on the general idea of recycling, then becomes more specific by reducing the scope of recycling to specific parts.
-Finally, they limit the scope of parts to precious metals, particularly gold.
-At this point, the participant exhausts their working memory of the category and moves on.
+The participant first touches on the general idea of recycling, then becomes more specific by reducing the scope of recycling to specific parts, and then to precious metals.
 
@@
 
-SIAM predict that individuals will generate ideas from an idea category until they exhaust that category, at which point they switch to another category (hypothesis 4).
-Furthermore, SIAM predicts that this category switching will be observable as taking more time to generate an idea in a new category than in the existing category (hypothesis 5).
+SIAM predict that individuals will generate ideas from an idea category until they exhaust that category, at which point they switch to another category (hypothesis 3).
+Furthermore, SIAM predicts that this category switching will be observable as taking more time to generate an idea in a new category than in the existing category (hypothesis 4).
 We examine each of these hypotheses in turn.
@@
 The mean for the within-category condition was 9.5 in log space (13.4 seconds, HDI 9.4-9.6), while the mean for the between-categories condition was 10.2 (27.0 seconds, HDI 10.2-10.3).
-The differences in mean are significant and confirm hypothesis 5.
-Notably, within-category instance generation took an average of 12.6 seconds less than between-categories instance generation, consistent with the findings of Nijstad and Stroebe, who reported differences of 6-12 seconds between these two idea generation processes.
+The differences in mean are significant and confirm hypothesis 4.
+Notably, within-category instance generation took an average of 12.6 seconds less than between-categories instance generation, consistent with the findings of Nijstad and Stroebe, who reported differences of 6-12 seconds.
 

### 2013-10-18 chi2014/conclusions.tex
so close

@@
 \item Common ideas are more likely to be proposed early in a run.
-\item Conversely, ideas beyond the first 20 are rarer and more original, as measured by o-scores. %This relationship also holds for idea categories.
-\item Ideas are more likely to be followed by another idea from the same category than would be expected by random chance.
-\item The time it takes to generate an idea when changing categories is longer than the time taken when generating ideas within a category.
+Conversely, ideas beyond the first 20 are rarer and more original, as measured by o-scores.
+%This relationship also holds for idea categories.
+\item Ideas are more likely to be followed by another idea from the same category than would be expected by random chance. The time it takes to generate an idea is longer when changing categories.
 \item There are repeatedly observable strategies employed by workers brainstorming in microtask marketplaces

### 2013-10-18 chi2014/introduction.tex
so close

@@
 A particularly compelling use of microtask marketplaces is to tap into the collective creativity of crowds.
-This enables one to quickly sample a wide range of ideas from a diverse population for tasks such as brainstorming or design.
 Aaron Koblin's art piece ``The Sheep Market'' \cite{koblin_2009} was an early demonstration of this application.
-In this project, he asked 10,000 individuals to draw a ``sheep facing left,'' yielding a wide range of unique results.
+%In this project, he asked 10,000 individuals to draw a ``sheep facing left,'' yielding a wide range of unique results.
 
-In this paper, we focus on crowd-based brainstorming.
+In this paper, we model brainstorming in microtask marketplaces (specifically, Amazon's Mechanical Turk). 
 While brainstorming has been an implicit component of many creative endeavors in microtask marketplaces, we are unaware of any research that describes and models the brainstorming process in this environment.
@@
 
-This paper presents the results of a series of experiments intended to model brainstorming in microtask marketplaces (specifically, Amazon's Mechanical Turk).
-In our experiments, we asked individuals to brainstorm ways to use an MP3 player whose batteries were dead.
+We asked individuals to brainstorm ways to use an MP3 player whose batteries were dead.
 We varied the number of ideas requested (5, 10, 20, 50, 75, 100, or unlimited), paying a fixed amount per idea (3.5 cents).
@@
 \item Individuals' patterns of idea generation mimic those identified by Nijstad and Stroebe's SIAM (search for ideas in associative memory) model \cite{nijstad_how_2006}: Workers produce a set of ideas from working memory, then switch strategies to generate ideas that are permutations of other ideas, a process we dub {\em riffing\/}.
-We also replicate Nijstad and Stroebe's findings with respect to the time it takes to generate a new idea \cite{nijstad_how_2006}.
+%We also replicate Nijstad and Stroebe's findings with respect to the time it takes to generate a new idea \cite{nijstad_how_2006}.
 \item we identify and describe a set of strategies employed by crowd workers that result in distinct kinds of produced ideas
@@
 Our results provide baseline models useful for two audiences: 1) those who wish to leverage crowds for brainstorming, and 2) researchers interested in developing techniques to optimize the crowd brainstorming processes.
-For the former, our models provide recommendations for parameters of brainstorming in this environment (such as the number of questions to request from workers).
-For the latter group, our models provide a baseline for comparing novel techniques intended to increase the quality of brainstorming output by crowds.
+For the former, our models provide recommendations for number of questions to request from workers.
+For the latter group, our models provide a baseline for comparing novel techniques intended to increase the quality of brainstorming output.
 

### 2013-10-18 chi2014/qualitative.tex
so close

@@
 We describe the observed strategies in detail below.
-We describe the distribution of these strategies across runs and comment on variability between participants.
 
@@
 Under the problem scoping strategy, the original problem is transformed to a new problem of reduced scope.
-Many transformations of the MP3 problem are possible, for example: brainstorm uses for \emph{an audio output device}; brainstorm uses for \emph{a hard drive}; and so on.
-The participant chooses a problem transformation and generates solutions, each of which is also a solution to the original problem.
-For example (PXXX):
+Many transformations of the MP3 problem are possible, for example: brainstorm uses for \emph{an audio output device}; brainstorm uses for \emph{a hard drive}; and so on. For example (PXXX):
 
@@
 
-The participant's responses satisfy the transformed question "brainstorm uses for an audio output device". Resulting ideas no not rely on other aspects of the problem.
+The participant's responses satisfy the transformed question "brainstorm uses for an audio output device", and satisfy the original problem.
 
@@
 The previous responses, which focus on audio output, are an example of inward scoping.
+Response generated by inward scoping are applicabable to the transformed problem and the original problem, but few others.
 
@@
 These solutions could be applied to a wide body of problems unrelated to the original, such as "brainstorm uses for a broom".
-In contrast, inward scoping solutions are applicable to the transformed problem and the original problem, but not many others.
 Thus, a useful rule for distinguishing the types of scoping is to consider whether the \emph{set of problems solved by the solutions} is small, or large.
@@
 %The strategies above describe techniques for taking one idea and transforming it into another.
-In this section, we describe patterns of \emph{where} participants implemented strategies.
+%In this section, we describe patterns of \emph{where} participants implemented strategies.
 
@@
 
-\subsection{Variability of Worker Input}
-While we observed that individuals are likely to draw from a common pool of ideas early on in the brainstorming process, there were outliers in the data.
+%\subsection{Variability of Worker Input}
+%While we observed that individuals are likely to draw from a common pool of ideas early on in the brainstorming process, there were outliers in the data.
 For example, one individual in the 75 ideas condition consistently produced some of most original ideas for old MP3 players:
 
-\begin{itemize}
-\item With displays - could be attached to a small solar panel (like the USB chargers w/multi attachments, including solar and car, etc.) and all the power could be sent to the screen as a small, portable light, then sent to countries lacking power in rural areas.
-\item They could be loaded with coordinates or other info and sealed in containers, then used in ocean current studies - once the container was collected they could be plugged in and read, and the data could be sent to whomever.
-%\item They could be plugged in at a base camp at the bottom of Everest, and then potential climbers could record their name, loved ones' address, and any last words before attempting the climb. 
-\end{itemize}
-
-This individual's responses raised the observed uniqueness of ideas in the 75 condition. 
-In contrast, other individuals filled their quotas with excessive repeat riffing chains.
-
-This variability in worker responses prevented us from modeling any potential effects that the number of questions requested might have on individuals' responses.
+%\begin{itemize}
+%\item With displays - could be attached to a small solar panel (like the USB chargers w/multi attachments, including solar and car, etc.) and all the power could be sent to the screen as a small, portable light, then sent to countries lacking power in rural areas.
+%\item They could be loaded with coordinates or other info and sealed in containers, then used in ocean current studies - once the container was collected they could be plugged in and read, and the data could be sent to whomever.
+%%\item They could be plugged in at a base camp at the bottom of Everest, and then potential climbers could record their name, loved ones' address, and any last words before attempting the climb. 
+%\end{itemize}
+%
+%This individual's responses raised the observed uniqueness of ideas in the 75 condition. 
+%In contrast, other individuals filled their quotas with excessive repeat riffing chains. 
+%
+%This variability in worker responses prevented us from modeling any potential effects that the number of questions requested might have on individuals' responses.
 Thus, while it is clear that there is little advantage to requesting 5 or 10 ideas from workers, it is not clear whether it is preferable to request 20, 50, 75, or 100 ideas.

### 2013-10-19 chi2014/introduction.tex
10 pages

@@
 
-We asked individuals to brainstorm ways to use an MP3 player whose batteries were dead.
-We varied the number of ideas requested (5, 10, 20, 50, 75, 100, or unlimited), paying a fixed amount per idea (3.5 cents).
-Across all conditions, 154 workers submitted responses to 170 HITs (Human Intelligence Tasks), yielding 3292 ideas.
-From these ideas, we present a series of models characterizing features of crowd brainstorming: the rate of novel idea production, the rate at which novel {\em categories\/} of ideas are produced, qualities of the ideas (for example, whether they are common or not), and characteristics of the brainstorming process itself.
+We asked individuals to brainstorm ways (5, 10, 20, 50, 75, 100, or as many as possible) to use an MP3 player whose batteries were dead, paying a fixed amount per idea (3.5 cents).
+154 workers submitted responses to 170 HITs (Human Intelligence Tasks), yielding 3292 ideas.
+We create a series of models characterizing features of crowd brainstorming: the rate of novel idea production, the rate at which novel {\em categories\/} of ideas are produced, qualities of the ideas (for example, whether they are common or not), and characteristics of the brainstorming process itself.
 
@@
 \item People are likely to come up with the same overall set of ideas early in their brainstorming process, but then diverge from this set of common ideas as they generate more ideas.
-Given our results, we find there is little value in requesting fewer than 20 ideas from workers, since these early ideas are among the most common.
+We find there is little value in requesting fewer than 20 ideas from workers.
 \item Individuals' patterns of idea generation mimic those identified by Nijstad and Stroebe's SIAM (search for ideas in associative memory) model \cite{nijstad_how_2006}: Workers produce a set of ideas from working memory, then switch strategies to generate ideas that are permutations of other ideas, a process we dub {\em riffing\/}.
 %We also replicate Nijstad and Stroebe's findings with respect to the time it takes to generate a new idea \cite{nijstad_how_2006}.
-\item we identify and describe a set of strategies employed by crowd workers that result in distinct kinds of produced ideas
+\item we describe a set of strategies employed by crowd workers that result in distinct kinds of produced ideas
 \end{itemize}

### 2013-10-19 chi2014/method.tex
10 pages

@@
 \subsection{Recruitment and Remuneration}
-To study brainstorming in a crowd environment, we published brainstorming HITs (human intelligence tasks) on Amazon's Mechanical Turk.
-We restricted our worker pool to residents of the United States to increase the chances of English language comprehension and a shared cultural background.
-No other screening was performed on workers.
+We published brainstorming HITs (human intelligence tasks) on Amazon's Mechanical Turk.
+Workers were residents of the United States to increase the chances of English language comprehension and a shared cultural background.
 
-HITs were placed on the marketplace asking for 5, 10, 20, 50, 75, 100, or an unlimited number of ideas for a brainstorming problem.
-We paid a fixed amount per HIT, computed at a rate of approximately 3.5 cents per idea requested, yielding reward amounts of 18 cents (5), 35 cents (10), 70 cents (20), \$1.75 (50), \$2.65 (75), and \$3.50 (100).
+HITs were placed on the marketplace asking for 5, 10, 20, 50, 75, 100, or as many possible ideas.
+HITs paid approximately 3.5 cents per idea requested, yielding reward amounts of 18 cents (5), 35 cents (10), 70 cents (20), \$1.75 (50), \$2.65 (75), and \$3.50 (100).
 For the unlimited condition, we offered \$1.75, pricing it the same as the 50 ideas condition.
@@
 In conducting this research, we tested several brainstorming questions, including classic brainstorming questions from prior work (e.g., ``provide alternative uses for a mop''); a request to brainstorm factors that influence weight gain and weight loss; and more goal-oriented questions, such as one requesting people to brainstorm ways to raise funds for a charity.
-Ultimately, we found the goal-oriented brainstorming questions to be the most compelling since they reflect what we believe would be real-world uses of brainstorming on Mechanical Turk.
+Ultimately, we chose the goal-oriented brainstorming questions since they reflect what we believe would be real-world uses of brainstorming.
 
-The final brainstorming question we chose for this research is the following:
+The final brainstorming question we chose is the following:
 \begin{quote}
@@
 
-We chose to focus on this problem because a) we expected workers to be familiar with problem area (old MP3 players), b) it represents a real-world problem (what to do with old electronics that are still partially functional) for which innovative solutions are welcome, and c) we anticipated there to be a wide range of creative solutions possible.
+We focus on this problem because a) we expected workers to be familiar with problem area (old MP3 players), b) it represents a real-world problem (what to do with old electronics that are still partially functional), and c) we anticipated there to be a wide range of creative solutions possible.
 All data presented in this paper refer only to responses received to this particular question.
@@
 
-The introduction listed Osborne's four rules of brainstorming \cite{osborn_applied_1957}, adapted to this context:
-
-\begin{enumerate}
-\item There are no bad ideas. Don't criticise your choices.
-\item Wild ideas and building off of old ideas are okay.
-\item Quantity of ideas is prioritized.
-\item Combinations of ideas count as new ideas.
-\end{enumerate}
-
+The introduction listed Osborne's four rules of brainstorming \cite{osborn_applied_1957}, adapted to this context.
+%\begin{enumerate}
+%\item There are no bad ideas. Don't criticise your choices.
+%\item Wild ideas and building off of old ideas are okay.
+%\item Quantity of ideas is prioritized.
+%\item Combinations of ideas count as new ideas.
+%\end{enumerate}
 In addition to the visible text input fields, HITs were instrumented to collect the time of the first activation and last de-activation of each form element.
@@
 As we received data in the unlimited condition, it became obvious that this was a poor strategy to employ -- the number of responses received was less than 15 in all cases.
-Accordingly, we remove that data from our analysis.
+We remove that data from our analysis.
 
-If a worker completed multiple HITs, there is the possibility that their responses and idea generation processes evolved over time. Accordingly, we retain only the first HIT submitted by each worker.
+If a worker completed multiple HITs, there is the possibility that their responses and idea generation processes evolved over time. Only the first HIT submitted by each worker is retained.
 
-After removing these data, we are left with 146 distinct participants for a total of 146 HITs, yielding 3007 responses.
-A breakdown of responses across conditions is given in Table~\ref{tab:data_collected}.
+We are left with 146 distinct participants for a total of 146 HITs, yielding 3007 responses. %A breakdown of responses across conditions is given in Table~\ref{tab:data_collected}.
 
@@
 
-A \emph{brainstorming run}, or \emph{run}, is the temporally-ordered series of instances that were received in a single HIT by a single worker.
+A \emph{brainstorming run}, or \emph{run}, is the temporally-ordered series of instances that were received in a single HIT.
 
@@
 An instance is associated with exactly one \emph{idea}, though multiple instances can be associated with the same idea.
-Multiple instances of the same idea do not need to be identically phrased to refer to the same idea.
 For example, in Figure~\ref{fig:example_instances}, we consider ``Storage Container'' and ``Small Storage Box'' as two instances of the same \emph{idea}.
 However, we do not consider ``Coin Storage'' to be the same idea, since it provides a more specific use case than the previous two.
-
-Despite the differences in specificity between ``Coin Storage'' and the more generic containers, it is useful to encode their relationship.
 We use hierarchical clusters to encode these relationships.
@@
 \subsubsection{Clustering: Process}
-To produce the idea forest, the primary author added instances to the idea forest one at a time using the following method.
+To produce the idea forest, the primary author added instances to the idea forest using the following method.
 
@@
 Clustering was performed using a custom-built tool that assisted in finding similar instances.
-Briefly, the tool represented each instance using a bag-of-words model that included synonyms for instance's words.
-The words were stemmed, then individual instances were ranked according to a cosine similarity metric.
+The tool represented each instance using a query-expanded bag-of-words model ranked according to a cosine similarity metric.
 

### 2013-10-19 chi2014/related.tex
10 pages

@@
 Electronic brainstorming is a variant in which individuals brainstorm at separate computer terminals, with the exchange of ideas performed over the network \cite{gallupe_electronic_1992}.
-Participants enter ideas while others' ideas are anonymously presented on-screen.
 As with nominal brainstorming, Gallupe \cite{gallupe_electronic_1992} observed that electronic brainstorming can reduce counterproductive social effects.
-However, Pinsonneault et al.\ \cite{pinsonneault_electronic_1999} identify productivity impediments introduced by group electronic brainstorming, such as being distracted by other ideas appearing, and individuals witholding ideas they feel are derivative of others.
+However, Pinsonneault et al.\ \cite{pinsonneault_electronic_1999} identify productivity impediments introduced by group electronic brainstorming, such as being distracted by other ideas appearing, and individuals witholding ideas they feel are derivative.
 They find little evidence to support the proposition that electronic brainstorming outperforms nominal brainstorming.
@@
 SIAM defines two stages of idea generation: the generation of ideas based on working memory, and idea production within a single activated ``image'' (where an image can be thought of as a category of ideas).
-This model makes predictions, one of the most important being that an idea is more likely to be followed by an idea in the same category than would be expected by chance.
-Their model also predicts that new ideas in the same category as the previous idea are generated faster than ideas between categories.
+This model predicts that an idea is more likely to be followed by an idea in the same category than would be expected by chance.
+SIAM also predicts that new ideas in the same category as the previous idea are generated faster than ideas between categories.
 Nijstad and Stroebe verified these predictions in the context of nominal brainstorming.
 
-To assess the brainstorming process and its outputs, a number of measures have been proposed.
-Typically, the measures of concern are the novelty of generated ideas, the practicality/appropriateness of ideas, the variety of ideas (flexibility), and the quantity of ideas generated.
-\cite{finke1992creative, shah2003metrics}.
-
-The most common approach to measure novelty and appropriateness is to ask domain experts to subjectively make scale-rated assessment of each \cite{lewis2011affective, amabile_1983}.
+To evaluate brainstorming, a number of measures have been proposed: novelty , practicality/appropriateness, variety, and quantity. \cite{finke1992creative, shah2003metrics}.
+The most common approach to measure novelty and appropriateness is to ask domain experts to make scale-rated assessment of each \cite{lewis2011affective, amabile_1983}.
 These two measures are sometimes combined into one measure of quality \cite{little2010exploring}.
@@
 %Examples for these environments are platforms for generating social innovation ideas or product concepts like OpenIDEO, Quirky and Threadless that employ structured process where participants take on different responsibility from generating ideas to evaluating others' ideas.
-We constrain our focus to nominal brainstorming in this environment with the goal of establishing baselines.
+%We constrain our focus to nominal brainstorming in this environment with the goal of establishing baselines.
 

### 2013-10-19 chi2014/results.tex
10 pages

@@
 
-We use an identical model for modeling the rate at which new idea categories are introduced over time.
+We use an identical model for the rate at which new idea categories are introduced over time.
 Figure~\ref{fig:cumulative_categories} shows the data and fitted model for the cumulative number of novel categories over time.
@@
 
-In both cases, we observe a steady increase in o-scores for the first 20-30 responses.
-After these first 20-30 responses, the o-scores plateau, with no clear discernible pattern for either ideas or idea categories.
-
+In both cases, we observe a steady increase in o-scores for the first 20-30 responses, followed by a plateau.
 This lends support for hypothesis 2, the notion that there is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.
@@
 \subsection{Idea Generation Patterns}
-Given the differences in originality and rarity as a function of response number, we now consider idea generation patterns observable from the data.
-We examine these patterns in the context of hypotheses 4 and 5, which draw upon the SIAM model of brainstorming.
-This model predicts individuals will generate ideas within ``images'' or categories of ideas (i.e., ``riff'' on an idea or concept) before switching to a new category.
+Given the differences in originality and rarity as a function of response number, we now consider idea generation patterns in the context of hypotheses 4 and 5.
+The SIAM model of brainstorming predicts individuals will generate ideas within ``images'' or categories of ideas (i.e., ``riff'' on an idea or concept) before switching to a new category.
 
-We can identify riffing by examining whether a response instance belongs to the same category tree as the previous instance.
+We identify riffing by examining whether a response instance belongs to the same category tree as the previous instance.
 We can then derive a number of statistics from these data.
-Below is an excerpt of a particular participant's brainstorming run.
+Below is an excerpt of a brainstorming run (PXXX):
 
@@
 Table~\ref{tab:riffing} provides a summary of riffing statistics observed in our data set.
-We find that the higher the number of instances requested, the more likely participants are to use old ideas as inspiration for new ideas.
-However, we observe that riff sequences are typically quite small in length, across conditions.
+The higher the number of instances requested, the more likely participants are to use old ideas as inspiration.
+Riff sequences are typically quite small in length.
 
@@
 Furthermore, SIAM predicts that this category switching will be observable as taking more time to generate an idea in a new category than in the existing category (hypothesis 4).
-We examine each of these hypotheses in turn.
+%We examine each of these hypotheses in turn.
 
@@
 
-To determine category switching, we examine whether an idea is assigned to the same category tree as the previous instance, or whether it was assigned to a different category tree.
+To determine category switching, we examine whether an idea is assigned to the same category tree as the previous instance.%, or whether it was assigned to a different category tree.
 
@@
 
-To calculate the time to generate a response instance, we examine the times logged for the first and last activation of a text entry widget.
-
-An examination of the timing data suggests a log-normal distribution.
+To calculate the time to generate a response instance, we take the difference between the first and last activation of a text entry widget.
 We fit two models: $t_w = \text{lognormal}(\mu_w, \sigma_w)$ for within-category idea generation and $t_b = \text{lognormal}(\mu_b, \sigma_b)$ for between-category idea generation, where the $t$ parameter in each model represents the time spent generating an instance.
@@
 The mean for the within-category condition was 9.5 in log space (13.4 seconds, HDI 9.4-9.6), while the mean for the between-categories condition was 10.2 (27.0 seconds, HDI 10.2-10.3).
-The differences in mean are significant and confirm hypothesis 4.
+The differences in mean are significant.
 Notably, within-category instance generation took an average of 12.6 seconds less than between-categories instance generation, consistent with the findings of Nijstad and Stroebe, who reported differences of 6-12 seconds.

### 2013-10-19 bootstrap.sh
fixed running notebook start to finish

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 apt-get update
-apt-get install -y python-pip python-dev python-requests build-essential sqlite3 libfreetype6-dev libpng-dev python-scipy python-matplotlib python-tornado python-zmq python-nltk pandoc python-pygments python-sphinx pyside-tools
+apt-get install -y python-pip python-dev python-requests build-essential sqlite3 libfreetype6-dev libpng-dev python-scipy python-matplotlib python-tornado python-zmq python-nltk pandoc python-pygments python-sphinx pyside-tools r-base
 pip install --upgrade ipython
 pip install turkflow numpy pyyaml pandas
 

### 2013-11-17 exps/format_data.py
some hypotheses recreated

@@ -300,12 +300,6 @@ def do_format_data(processed_data_folder, filter_instances = None):
 
 
     # ========================================================
-    # CLUSTER TOPOLOGY
-    # ========================================================
-
-    cluster_forests = {qc: cluster_forest(cluster_tree_csvs[qc]) for qc in cluster_tree_csvs.keys()}
-
-    # ========================================================
     # INSTANCE DATA
     # ========================================================
 
@@ -360,6 +354,13 @@ def do_format_data(processed_data_folder, filter_instances = None):
     # ======================================================== 
     if filter_instances:
         df = filter_instances(df)
+    
+    # ========================================================
+    # CLUSTER TOPOLOGY
+    # ========================================================
+
+    cluster_forests = {qc: cluster_forest(cluster_tree_csvs[qc]) for qc in cluster_tree_csvs.keys()
+            if len(df[df['question_code'] == qc]) > 0}
 
 
     # ========================================================
@@ -390,6 +391,9 @@ def do_format_data(processed_data_folder, filter_instances = None):
 
     for qc in qc_conds:
         sub_df = df[df['question_code'] == qc]
+        if len(sub_df) == 0:
+            continue
+
         #print len(sub_df)
         #print sum(sub_df['is_repeat_worker'])
         

### 2013-11-18 chi2014/rebuttal_notes.md
first pass of rebuttal constructions. Lots to go

# Main rebuttal content

## M1

> The results of the work are not generalizable because the data came
from only a single task and a single run of that task and there were no
test of generalizability on unseen data. The models presented in the
paper are appreciated and quite interesting, but a significant weakness
is that the authors built the models using only a single brainstorming
question and are trying to make generalizable claims. For example, the
authors state (p. 7), “there is a set of general, common ideas that
make up the first several responses of every crowd brainstorming
session.” How can this claim be made when there was only a single
brainstorming question posed? Technically, the claims in the paper can
only be made for that specific brainstorming question (how to reuse old
mp-3 players). The paper would be far more convincing had the authors
showed that the models fit the data generated from multiple questions.
Alternatively, it could have been acceptable if the authors conducted
additional runs of the same question and showed that the models remained
the same or similar. Claiming generalizability of models trained on only
a single data set is a major limitation of the work. It does not mean the
work is bad, only that a larger data set would greatly strengthen the
work. Even R1, the strongest advocate for the paper, observes that there
is no validation of the model and questions its validity, which is the
main contribution of the work.

It was not our intention to suggest the findings were highly generalizable, but rather to provide evidence in support of the stated hypotheses as well as present models that can be applied to generalized datasets. Examining phenomena in the context of a single problem is not unusual in creativity literature, and the paper cites several examples in psychology (3, 4, 17, 22) and crowd creativity research specifically (28, 29).

In addition, this is the first work to examine crowd creativity in the large scale. Even in the context of a single question, this work explores the diversity of a massive sampling.

REV: change quoted line to refer to limit application to study data explicitly

## M2

> The motivation for the work is very weak. It is not clear why it is
important or interesting to build models of brainstorming of workers on
Mechanical Turk. For example, in prior work, such as on the SIAM model
developed by Diehl and Strobe, it is clear that the research was directed
at testing a particular model of human memory as it relates to
brainstorming. In the current paper, however, it is difficult to
understand for what purpose the authors are constructing the models. Are
you testing a particular theory of human memory? Are you testing a
particular theory of group brainstorming? How do the authors intend for
the models to be used, e.g., in what work contexts and for what purpose?
As implied by R2, it also does not appear that there was anything
specific about the use of mechanical turk for this work and therefore the
authors should offer a deeper comparison to brainstorming in group
contexts.

Crowds are uniquely suited to creativity generation tasks. Brainstorming has been independently introduced as either a component of a task or a major task itself in prior crowd creativity work [16, 29]. In cases where idea generation is a prerequisite to creative task completion, it is desirable to have a set of models and guidelines to inform task design. Finally, it cannot be assumed that findings in group or electronic brainstorming are generalizable to a microtask marketplace context.

## M3

> There were a number of methodology confounds. One serious confound is
that workers were producing different numbers of ideas. For example, in
the “5” condition, 57 people produced 287 ideas; whereas in the
“50” condition, 9 people produced 450 ideas. Given that prior results
on brainstorming show that better ideas come later in the idea generation
phase for an individual, the fact that this was not accounted for in the
modeling is a serious oversight.  An alternative might have been to ask
every worker to produce the same number of ideas, but manipulate the
number of workers asked. This may remove the confound yet still allow for
modeling and interesting outcomes.

The models treat all instances that occur in the same position in a brianstorming run equivalently, across condition. We tested for differences in originality scores between the first 5, 10, 25, etc responses as a function of condition, and found no two conditions significantly different.

Note, it may be more valuable to just run our analysis separately for each condition and see if the hypotheses hold.

REV: add this to the analysis section

# M4 (+ other reviewers)

> How were the off-topic (non-sense) responses filtered from the data
set? The problem with originality is that it rewards off-topic responses.
As noted by R3, the more “original” ideas may also have been the more
off-topic ideas and it is likely that these types of ideas start to
appear after the easy ones run out. Asking one person to come up with 100
sensible responses for using an old MP-3 player might be a stretch. Can
the authors offer evidence that the ideas after the first 20 were both
on-topic and original?

> Listing 100 ideas at one time seems to be a challenging task. Prior work
shows that when a task becomes difficult, workers attempt to give random
answers. In the analysis, the authors only removed the repeated responses
from the workers, but did not remove the off-topic responses. The
analysis shows that requesting more than 20 ideas from each worker can
improve the “originality”. However, another explanation is that when
a worker is required to generate more 20 ideas, s/he is more likely to
produce random and irrelevant responses.

> The researchers described a few example cases where participants had
been extremely creative or made systematic manipulations of the same
idea, which prevented the researchers identifying differences between
their conditions.  How common where these cases?  Why not remove these
subjects from your dataset?

Nonsense responses were flagged under the condition that their application to the iPod problem could not be visualized by the coder. This was not mentioned in the analysis due to their low prevalance (less than 12 instances total). We re-ran all analyses with these responses removed and all hypothesis tests held (with very slight differences in posterior parameters). They were initially left in as representative of "real world" results.
REV: describe nonsense answers in data description and coding process

## M5 (+ other reviewers)

> The authors need to better describe how the o-scores were computed. In
the Jannson and Smith paper, the o-scores were computed from pictorial
representations of the designs. Here, the ideas are in the form of text.
So how were the o-scores computed and how can you interpret the o-score
from the reader’s perspective? Also, as noted by R3, the o-scores
appear to be way off. How can all of the o-scores be between .994 and
.999 (Figure 5).

> The authors very clearly describe early on in the paper how
originality will be measured e.g. o-scores.  However, this leaves some
confusion around how categories of ideas will be measure --- note that
the authors do explain this in the paper, but question if a reference to
how categories of ideas are measured earlier in the paper would help
readership.

> The o-score in the worst condition is still more than 0.994 (see Figure
5), which is very close to the maximum value of o-score. This means that
most of the ideas generated in the conditions are highly original and
originality is not the main issue for this particular brainstorming
question (the reuse of MP3 players). The authors should redesign the
brainstorming question to test the hypotheses.

O-scores were computed as follows. A count of all instances in the idea node (for idea o-score) or category tree (for category o-score) was taken. This was divided by the total number of instances and subtracted from one to arrive at the o-score. While the scores seems unusually clumped, the o-score is a relative measure for distinguishing between ideas dominated by the total number of unique responses. The "squishing" of scores into the .994-.999 range is an artifact of the volume of the data set and representative of the space of possible responses.
REV: clarify the o-score measure and explain the meaning of the response range

## M6 (+ other reviewers)

> The clustering technique did not follow standard protocol, yet the
validity of the models depend on how the clustering was done.  Both R1
and R3 raised this concern. For example, R3 notes that “the clustering
results are only produced by one primary author. The authors need to
follow the standard procedure in qualitative research and the second
researcher should produce the clustering results independently (not just
review the results produced the primary author).”

> I slightly question the process and validity of the inter-judge
reliability test.  It seems as if the second rater was given the
categories and asked to validate them.  A better approach would have been
to give the second rater the ideas belonging to those randomly sampled
category tree, form category trees themselves and then look at
differences.  As the authors describe, the rating does not seem to
validate different instances of the same idea e.g. "Storage Container"
and "Small Storage Box".

> The validity of the findings highly depends on the clustering results.
But the clustering results are only produced by one primary author. The
authors need to follow the standard procedure in qualitative research and
the second researcher should produce the clustering results independently
(not just review the results produced the primary author). Another
possible way is to use clustering methods to get the results
automatically without the interference of the primary author.

Mike: I need your help here.

## M7

> The hypotheses in the paper have no framing and may be better
characterized as research questions. For instance, the first hypothesis
states that the probability that an idea is novel decreases non-linearly
as a function of the number of responses received. Where did this
hypothesis come from? How was it derived? How does it relate to prior
work? Perhaps the hypotheses could be better framed as a set of
interesting research questions, which would be fine by me. The first
question, for example, might be better phrased as “how does the
probability that an idea is novel relate to the number of ideas
received?” This allows a bit more freedom into exploring the various
models as opposed to locking into a specific set of models (e.g.
non-linear models).

We agree that the first and second hypotheses would benefit by being expressed as research questions. The other three hypotheses are informed by prior work.
REV: As suggested.

## M-Extra

> Why are you providing Osborne’s rules of brainstorming to individual
users? As far as I know, those rules were targeted at group brainstorms,
not for individuals.

The rules of brainstorming are often given in an individual context, including in very early comparisons of group vs nominal brainstorming [27].

> The authors should focus on idea quality, rather than quantity or
originality. The problem with originality is that highly ‘original’
responses could be utter non-sense. The authors should report how the
ideas were filtered to remove any irrelevant results and possibly
consider idea quality in future work.

We chose to emphasize uniqueness as a common component of creativity. Number of unique ideas also correlates with quality overall [7, 4, 22, 21, 25, 6]. We agree that an explicit analysis of quality is imperative future work.
REV: explicitly state choice to emphasize uniqueness, add quality to future work

# Other reviewers

> I was interested as to why the researchers dropped all other
questions than the MP3 question.  Either don't mention it or provide an
explanation.

Not sure what to say here. "We didn't have time!"

> The researchers modeled their data using the full dataset.  Usually,
when creating a model you split your dataset into three.  One used to
create your model, one used to test your model and another used to
validate your model.  As such, there is no validation of this model,
which is one of the main contributions of this paper.  Furthermore, given
the high variance described, I would further question the validity of
this model.

This paper leverages its models specifically to test hypotheses. We do not evaluate the fit or predictive power of the models.

> The authors do not really discuss how this strategy for brainstorming
differs from, or is similar to, group or individual brainstorming not
online. This deserves a substantial discussion and proposal of future
work.

> The future work and conclusion sections are the weakest part of the
paper. The authors should carefully consider how their results can be
used by other researchers and what clear (and generalizable) findings can
be made. In addition, I challenge the authors to more thoughtfully
consider the future work to drive the future of HCI research in this
field.

TODO: future work revisions

> To compare the originality in condition <= 20 and condition >20 (section
“Originality”, page 7), the authors first used models to fit the data
and then compared parameters of these models. However, some of the models
obviously did not fit the data well (see Figure 8 left bottom). This made
the comparison results inaccurate. Since it is non-normal distributed
data, the authors could perform a non-parametric test (e.g. Mann-Whitney)
to directly compare o-scores in these two conditions. Similar issues are
also present in the section “Most popular Ideas”.

TODO: run this test

> In the section “Most popular Ideas”, the authors compared the
originality in the condition “first 5 instances” and the condition
“last 5 instances”. In the “first 5 instances” condition, the
ideas are the mixture of ideas in all the conditions including condition
5, 10, 20, 50, 75, and 100, while the ideas in the “last 5 instances”
condition are only from the condition 100. I’m not sure this is a fair
comparison. It would be better if the authors performed the analysis
separately in each condition. Also, if a worker needs to submit 100
ideas, it is reasonable to consider that the first 5 ideas are generated
in the early phase. But if a worker is aware of that he only needs to
submit 5 ideas, these 5 ideas are generated from the entire brainstorming
session (NOT the “early” phase of the brainstorming session).

With respect to the test of hypothesis 3, we feel the description of the test was poorly given and led to misunderstanding among reviewers. We did not compare the first 5 instances and last 5 instances for the presence of common ideas, but rather the first 5 instances with the *remaining* instances (all those with position greater than 5). In addition, this was done only for runs with 10 or more responses (conditions 10, 25, 50, 75, 100), to ensure that no participant could contribute to one distribution without contributing to the other.
REV: Rephrase the description of the hypothesis test to be more descriptive

# Maybe Ignore

> The authors have clearly stated the methodology of the paper and have
presented testable hypothesis. Their contribution to the HCI community is
clearly stated (It should be noted that I greatly appreciate the detailed
contribution statements and the bullets at the end of the introduction-
this provides a very clear outline for the paper!) and I feel that it
does add to the community knowledge. I do wish that the authors had
compared their findings from Mechanical turk to a non- distributed team
and compare findings. This type of study would help bridge the gap and
identify differences in these two communities and also aid researchers in
understanding which study findings generalize to both audiences. I think
this deserves some discussion in the paper.

> The authors only had participants textually describe their ideas -
however they spend little to know time describing why. This is important
as prior research has explored differences in textual and pictorial
brainstorming. Please clarify and provide relevant source work.

> The authors mention the popular shah and vargas-hernandez method for
rating idea novelty but make little reference to it later on. It should
be better noted why this method was not used to test originality and
popular ideas... Because this is the 'go-to' method, it makes the
findings and comparison of these results to other studies more difficult.

> It would be great if the authors reported more information on the
collected ideas (e.g. the length of the collected ideas). It seems that
the length of example ideas in the paper is much shorter than the median
length of ideas in previous work (e.g. 60 words).

# Textual changes

> The authors provide a lot of details in the data collection section on
why they didn't do certain things, e.g. removed data from the unlimited
condition etc. However, this additional information makes this section
hard to comprehend. It may be a better strategy to just simply say what
you did and why you did it - revise.

> Title headers are floating at the bottom of page 4, 5, and 6.

> Details of clustering was too brief on page 5 to comprehend how this
was accomplished, "Clustering was performed using a custom-built tool".
Either provide more details or a reference for this strategy.

> on page 7 the list is split across the two pages. Please fix.

> The first sentence of the abstract is difficult to understand and should
be re-written.

> There are a few places that need to be edited, such as the floating
sentence on page 2, "Finally, Diehl and Stroebe found that groups that
generate good ideas also generate many ideas [7], a ﬁnding that has
been veriﬁed many times [4, 22, 21, 25, 6]." In fact, there are several
weird 2 sentence 'paragraphs' throughout that should be edited. In
addition, the authors mention several places throughout, "we are unaware
of..." This should be edited to be sound more professional and accurate.
When there is something like this written it makes me question if the
authors have completed a substantial review of the literature.

# M8 + others: Related work additions

> Both R1 and R2 list a series of citations that should be included in a
revision of the paper but none of them appear detrimental to the
contribution of the present work.

> The literature review includes a good mix of work on creativity from
the psychology community and later work on Group Decision Support
Systems.  Some recent research the authors have missed or not included
is:

> Warr, A. and O'Neill, E. (2005) The Effect of Operational Mechanisms on
Creativity in Design. Proc. Interact 2005, Rome, Italy. Springer LNCS
3585, 629-642.

> Warr, A. and O'Neill, E. (2006). The Effect of Group Composition on
Divergent Thinking in an Interaction Design Activity. DIS 2006, 26-28
June 2006, Penn State, USA.

> The above research presents an experiment that examines the effect of
group compositions on creativity in terms of number of generated ideas
and number of categories of ideas, respectively.

> According to the literature, a creative idea is an idea that is both
novel and appropriate.  Where novelty is an indicator of uniqueness and
appropriateness is what makes an idea creative.

> Amabile, T. M. (1983) The Social Psychology of Creativity.
Springer-Verlag, NewvYork.

> Hence, while measuring novelty is a measure of uniqueness, the measure of
appropriateness is what differentiates creative ideas from the bizarre
and odd.

> Other references recommended for review and citation:

> Furnham, A., & Yazdanpanahi, T. (1995). Personality differences and group
versus individual 
brainstorming. Personality and Individual Differences(19), 73-80.

> Nijstad, B. A., Stroebe, W., & Lodewijkz, H. F. M. (2002). Cognitive
Stimulation and interference in 
groups: Exposure effects in an idea generation task. Journal of
Experimental Psychology, 
38(2002), 535-544.

> Nijstad, B.A., Stroebe, W. and Lodewijkx, H.F., 2003, “Production
blocking and idea generation:
Does blocking interfere with cognitive processes”, Journal of
Experimental Social Psychology,
39(6): 531-548.

> Xu, A. and B.P. Bailey. A Crowdsourcing Model for Receiving Design
Critique. Proceedings of the ACM Conference on Human Factors in Computing
Systems, Extended Abstracts, 2011, 1183-1188

# Re-usable praise

> This paper makes a number of contributions:
Demonstrates the application of microtask marketplaces to
brainstorming
Provides models for rates of idea generation, divergence and
uniqueness
Show that the first 20 ideas are less original than those generated
after

> Overall, I would argue for accepting this paper.  IMO, it is one of the
best papers I have read on the study of creativity recently and is
furthermore, very scientific in its application --- something very
difficult in the domain of creativity.

> The researcher conclude by making actionable recommendations based on
data that are applicable to researchers and practitioners.

> I am truly excited by the scope of this research in the future.  For
example, explaining differences between demographics, personality types,
etc... As well as validating the research with different brainstorming
tasks.

> Through a series of experiments, the authors provide a characterization
of brainstorming on a microtask marketplace (Amazon Mechanical Turk), and
provide a model for the rate of idea generation, rate of category
generation and uniqueness of generated ideas through a series of
experiments.  The authors have also identified two clear audiences for
their findings including both individuals who wish to leverage the crowd
for brainstorming and those interested in developing methods for
optimizing the brainstorming process. I feel that these items clearly
contribute to the HCI community but the article needs to be revised
before it is acceptable for publication.

> This paper reports on an investigation of brainstorming in micro-task
marketplace. The topic is very interesting and research hypotheses are
original. I hope this work is published eventually. But the experimental
design and data analysis were flawed.

# Boilerplate

> This paper presents a series of experiments aimed at modeling
brainstorming behavior in crowds (as defined by Mechanical Turk).
Specifically, the authors modeled the rate of idea production, including
originality, to a brainstorming task and offered data-driven insights
into how to best leverage the use of crowds for brainstorming.

> The reviewers were very much split on this paper. R1 strongly advocates
for acceptance and states that “this one of the best papers I have read
on the study of creativity recently and is furthermore, very scientific
in its application --- something very difficult in the domain of
creativity.” I agree the authors are pursuing an original line of
scientific inquiry and appreciate R1's enthusiasm for the work. R2 also
proposes to accept the paper and believes that the paper adds to the
community knowledge. R3, however, is not sold on the work and offered a
very detailed analysis of the statistical results. From my own reading of
paper, I think this paper could be outstanding but currently has a lot of
weaknesses, some serious, that might require additional work.

> The paper will likely receive an additional review and discussion at the
program committee meeting. So the authors are kindly asked to direct
their rebuttal at these points:

> There are several points in the paper that need clarification:

### 2013-11-19 chi2014/rebuttal_notes.md
kfold

@@
+# Text of rebuttal
+
+Below are responses and revisions (REV) to address the 7 points identified by the AC, follow by a general discussion.
+
+1) Generalizability
+We did not intend to suggest the findings were generalizable, but rather to provide evidence in support of the hypotheses as well as present models that can be applied to generalized datasets.
+Exploring a single problem is not unusual in creativity literature, and the paper cites several examples in psychology (3, 4, 17, 22) and crowd creativity research that do similarly (28, 29).
+REV: change language to limit generalizability to study data
+
+2) Motivation
+As stated in the introduction, brainstorming has been independently introduced as either a component of a task or a major task itself in prior crowd creativity work [16, 29].
+It is desirable to have a set of models and guidelines to inform task design in these cases.
+It cannot be assumed that findings in group or electronic brainstorming are generalizable to the microtask marketplace context.
+
+3) Number of ideas produced
+The models assume all instances that occur in the same ordinal (run) position have similar characteristics, across condition.
+We tested for differences in ordinal originality scores between condition, and found no significantly differences.
+In addition, the Bayesian methods applied do not require equivalent sample sizes in compared model fits - adding more samples to either model simply tightens the posterior parameter distributions for that model.
+REV: state assumption and supporting evidence explicitly
+
+4) Off-topic answers
+Off-topic answers were flagged in the coding process.
+The criteria was that the use case for the iPod could not be visualized.
+There were 12 off-topic instances total, showing that participants took the task very seriously.
+All hypothesis tests held with or without off-topic answers included.
+The AC and R3 both felt that asking for 100 ideas would result in noise.
+This work is a necessary exploration of this kind of task design by intuition, and participants in fact generated primarily non-random ideas.
+REV: a brief treatment of off-topic answers in analysis
+
+5) O-scores
+O-scores were computed as follows.
+A count of all instances in the idea node (for idea o-score) or category tree (for category o-score) was taken.
+This was divided by the total number of instances and subtracted from one to arrive at the o-score.
+The o-score is a relative measure dominated by the total number of unique responses.
+The "squishing" of scores into the .994-.999 range is an artifact of the volume of the data set and representative of the diversity of responses.
+REV: clarify o-score measure and response range
+
+6) Clustering validity
+Our analyses were primarily concerned with originality, computed as a function of number of instances in an idea node.
+The clustering validity test found 0.7% of the instances were falsely identified as belonging to a cluster to which they had no relation.
+TODO: discovered a problem here with our upper bound theory, need another angle of attack
+
+7) Hypotheses
+We agree that the first and second hypotheses would benefit by being expressed as research questions. The other three hypotheses are informed by prior work.
+REV: As suggested.
+
+This is the first work to examine crowd creativity with such a large volume of ideas (3292 reponses).
+As stated by R1, the approach taken is "very scientific in its application --- something very difficult in the domain of creativity".
+Even in the context of a single question and a single output of interest (uniqueness/originality), our methods and models contribute an approach for understanding idea generation at a massive scale.
+
+The primary criticisms are of methodology, which we would like to further address:
+The AC questioned giving the rules of brainstorming are often given in an individual context. This is often done, including in comparisons of group vs nominal brainstorming [27].
+The AC further suggested that appropriateness should have been examined.
+We chose to emphasize uniqueness as a core component in all creativity measures.
+Number of unique ideas also correlates with quality overall [7, 4, 22, 21, 25, 6], and can be evaluated domain-independently, which appropriateness cannot.
+Cross-domain applicability is key to a comparable baseline.
+
+R1 questioned the lack of model validation.
+This paper leverages its models to test hypotheses, not for predictive power.
+In addition the fit of a Bayesian model is determined by its posterior distribution over parmameters, which we report.
+Regardless, we evaluated the idea rate model in a 10-fold cross validation, and found a mean squared error of 341.58 (about 18 ideas).
+
+R3 suggested a Mann-Whitney U test to better compare originality conditions. We performed this test and found a difference (p < 0.01) for both ideas and categories.
+R3 also expressed a slight misunderstanding of the test for common ideas (page 7).
+We tested the difference between the first 5 instances and the remaining 5-95 instances (not the last 5 instances) in a run.
+This test was performed only for runs with >=10 instances, so runs in all conditions contributed samples to both distributions.
+REV: revise description of hypothesis test to be more clear
+
+Finally, we appreciate the related work and formatting error revisions, and gladly accept them.
+
 # Main rebuttal content
@@
 work. Even R1, the strongest advocate for the paper, observes that there
-is no validation of the model and questions its validity, which is the
+is no validation of the model and quesfrom the reader’s perspective? Also, as noted by R3, the o-scores
+tions its validity, which is the
 main contribution of the work.

### 2014-02-11 src/clusterer/IdeaListModel.py
latest clusterer

@@ -1,0 +1,154 @@
+from PySide import QtCore
+import sqlite3 as db
+
+class IdeaListModel(QtCore.QAbstractListModel):
+  def __init__(self, db_file_name, parent=None):
+    super(IdeaListModel, self).__init__(parent)
+    self.conn = db.connect(db_file_name)
+    self.cur_question_code = ""
+    self.cur_ideas = [] # Each entry: (id, idea)
+    self.last_label = None # hack
+    self._init_from_db()
+  def _init_from_db(self):
+    self.cur_question_code = self.get_question_codes()[0]
+    self.read_in_ideas()
+  def read_in_ideas(self):
+    self.cur_ideas = []
+    cursor = self.conn.cursor()
+    cursor.execute("SELECT id, idea FROM ideas WHERE question_code = ? AND id NOT IN (SELECT idea_id FROM used_ideas) ORDER BY id ASC", (self.cur_question_code,))
+    for idea_id, idea in cursor.fetchall():
+      self.cur_ideas.append((idea_id, idea))
+    cursor.close()
+  
+  def get_ideas_for_indices(self, indices):
+    return [self.cur_ideas[i] for i in indices]
+
+  def mean_similarity(self, group1_ids, group2_ids):
+    cursor = self.conn.cursor()
+    tups = [(x, y) for x in group1_ids for y in group2_ids]
+
+    scores = []
+
+    for id1, id2 in tups:
+      cursor.execute("""SELECT score FROM similarity_scores
+                        WHERE idea1 = ? AND idea2 = ?""", (id1, id2))
+      res = cursor.fetchone()
+      if res:
+        scores.append(res[0])
+      else:
+        print("Missing similarity score", id1, id2)
+
+    return sum(scores) / len(scores)
+
+  def get_ids_for_text(self, text):
+    l = text
+    id_loc = l.find('(_')
+    ids = []
+    if id_loc > 0:
+        old_id = l[id_loc+2:-1]
+        l_p = l[0:id_loc - 2]
+        l_p = l_p.strip('-')
+
+        cursor = self.conn.cursor()   
+        cursor.execute("SELECT id FROM ideas WHERE question_code = ? AND idea = ?", (self.cur_question_code, l_p,))
+        for (idea_id,) in cursor.fetchall():
+            ids.append(idea_id)
+
+        return l_p, ids
+    else:
+      print("Couldn't find ID", l)
+      return None
+
+  def resolve(self, idea_tree_model):
[…]

### 2014-02-11 src/clusterer/IdeaTreeModel.py

@@ -1,0 +1,393 @@
+import datetime, csv
+import sqlite3 as db
+import pickle
+from IdeaTreeNode import IdeaTreeNode
+from PySide import QtCore
+
+def get_current_time():
+  return datetime.datetime.now()
+
+class IdeaTreeModel(QtCore.QAbstractItemModel):
+  def __init__(self, question_code, db_file_name, parent=None, root = None):
+    super(IdeaTreeModel, self).__init__(parent)
+    self.root = IdeaTreeNode([], None, question_code) if not root else root
+    self.conn = db.connect(db_file_name)
+    self.question_code = question_code
+
+    if root is None:
+      self.load()
+
+  def columnCount(self, parent = None):
+    return 1
+
+  def data(self, index, role):
+    item = index.internalPointer()
+    if role == QtCore.Qt.DisplayRole:
+      return item.label()
+    elif role == QtCore.Qt.ToolTipRole:
+      return item.tooltip()
+    return None
+
+  def index(self, row, column, parent):
+    p = parent.internalPointer() if parent.isValid() else self.root
+    c = p.child(row)
+    if c is None:
+      return None
+    return self.createIndex(row, column, c)
+
+  def parent(self, index):
+    c = index.internalPointer()
+    p = c.get_parent()
+    if p == self.root:
+      return QtCore.QModelIndex()
+
+    return self.createIndex(p.row(), 0, p)
+
+  def rowCount(self, parent=QtCore.QModelIndex()):
+    p = parent.internalPointer() if parent.isValid() else self.root
+    return p.child_count()
+
+  def get_next_depth(self, dq):
+    for l in dq:
+      if len(l) == 0:
+        continue
+      return get_line_indent(l)
+
+  # Please note that all this code is awful
+  def _get_line_node(self, l, current_node, used_ids, idea_model):
+    ids = extract_idea_ids_from_text(l)
+    assert len(ids) <= 1
+
+    if not ids:
+      # Check if this is a label
+      if len(l) > 0:
+        return IdeaTreeNode([], current_node, l.strip('-'))
+      # Otherwise this is a newline, close clusters up to level of depth
+      else:
+        return None
+
+    idea, ids = idea_model.get_ids_for_text(l)
+    new_ids = [i for i in ids if not i in used_ids]
+    if len(new_ids) == 0:
+        print("Couldn't find idea:", l)
+        return None
+    new_id = new_ids[0]
+    used_ids.append(new_id)
[…]

### 2014-02-11 src/clusterer/IdeaTreeNode.py

@@ -1,7 +1,3 @@
-import pickle
-import sys
-import random
-
 class IdeaTreeNode(object):
   def __init__(self, ideas, parent, label = None):
     assert not isinstance(ideas, str)
@@ -9,7 +5,7 @@ class IdeaTreeNode(object):
       self._label = label
     else:
       self._label = None
-    self.ideas = ideas # list of (text, id) tuples
+    self.ideas = ideas # list of (text, id, time) tuples
     self.parent = parent
     self.children = []
 
@@ -18,7 +14,7 @@ class IdeaTreeNode(object):
 
     if self._label and len(self._label) > 0:
       outlines.append(self._label)
-    for text, i in self.ideas:
+    for text, i, time in self.ideas:
       outlines.append(text + '  (_id:' + str(i) + ')')
 
     for i, c in enumerate(self.children):
@@ -31,8 +27,8 @@ class IdeaTreeNode(object):
     return outlines
 
   def add_ideas(self, ideas):
-    present_ids = [iid for idea, iid in self.ideas]
-    ideas = [(idea, iid) for (idea, iid) in ideas if not iid in present_ids]
+    present_ids = [iid for idea, iid, time in self.ideas]
+    ideas = [(idea, iid, time) for (idea, iid, time) in ideas if not iid in present_ids]
     self.ideas.extend(ideas)
 
   def merge(self, other_node):
@@ -42,6 +38,8 @@ class IdeaTreeNode(object):
     self.children = sorted(self.children, key=lambda x: x.label().lower())
 
   def append_child(self, child_node):
+    if not child_node.parent is None:
+        child_node.parent.remove_child(child_node)
     self.children.append(child_node)
     child_node.parent = self
     self.sort_children()
@@ -102,7 +100,6 @@ class IdeaTreeNode(object):
   def get_parent(self):
     return self.parent
 
-  def get_ids(self):
     ids = [i[1] for i in self.ideas]
 
     for c in self.children:
@@ -118,18 +115,4 @@ class IdeaTreeNode(object):
 
     return ret
 
-root = None
-with open("out/_iPod_IdeaTreeNode.pickle", 'rb') as f:
-    root = pickle.load(f)
 
-if root is None:
-    print("error")
-    sys.exit(1)
-
-all_trees = root.children
-
-sampled_trees = random.sample(all_trees, 40)
-
-for i in sampled_trees:
-    print('\n'.join(i.as_text()))
-    print("\n\n\n\n\n\n\n\n\n\n")

### 2014-02-11 src/clusterer/IdeaTreeNodeListModel.py

@@ -1,0 +1,20 @@
+from PySide import QtCore
+
+class IdeaTreeNodeListModel(QtCore.QAbstractListModel):
+  def __init__(self, itn, parent=None):
+    super(IdeaTreeNodeListModel, self).__init__(parent)
+    self.node = itn 
+ 
+# QAbstractListModel required methods
+  def rowCount(self, parent):
+    return len(self.node.ideas)
+  def data(self, index, role=QtCore.Qt.DisplayRole):
+    row = index.row()
+    if role == QtCore.Qt.DisplayRole or role == QtCore.Qt.ToolTipRole:
+      return self.node.ideas[row][0]
+#    if role == QtCore.Qt.TextAlignmentRole and self.alignment_data:
+#      return self.alignment_data[col]
+    return None
+# End QAbstractListModel required methods
+
+

### 2014-02-11 src/clusterer/change_parent.py

@@ -1,0 +1,38 @@
+# -*- coding: utf-8 -*-
+
+# Form implementation generated from reading ui file 'parent_chooser.ui'
+#
+# Created: Thu Jan 23 17:10:51 2014
+#      by: pyside-uic 0.2.14 running on PySide 1.2.1
+#
+# WARNING! All changes made in this file will be lost!
+
+from PySide import QtCore, QtGui
+
+class Ui_Dialog(object):
+    def setupUi(self, Dialog):
+        Dialog.setObjectName("Dialog")
+        Dialog.resize(438, 469)
+        sizePolicy = QtGui.QSizePolicy(QtGui.QSizePolicy.Preferred, QtGui.QSizePolicy.Preferred)
+        sizePolicy.setHorizontalStretch(0)
+        sizePolicy.setVerticalStretch(0)
+        sizePolicy.setHeightForWidth(Dialog.sizePolicy().hasHeightForWidth())
+        Dialog.setSizePolicy(sizePolicy)
+        self.btn_change_parent = QtGui.QPushButton(Dialog)
+        self.btn_change_parent.setGeometry(QtCore.QRect(10, 420, 211, 41))
+        self.btn_change_parent.setObjectName("btn_change_parent")
+        self.btn_cancel = QtGui.QPushButton(Dialog)
+        self.btn_cancel.setGeometry(QtCore.QRect(230, 420, 201, 41))
+        self.btn_cancel.setObjectName("btn_cancel")
+        self.tree_options = QtGui.QTreeView(Dialog)
+        self.tree_options.setGeometry(QtCore.QRect(10, 10, 421, 401))
+        self.tree_options.setObjectName("tree_options")
+
+        self.retranslateUi(Dialog)
+        QtCore.QMetaObject.connectSlotsByName(Dialog)
+
+    def retranslateUi(self, Dialog):
+        Dialog.setWindowTitle(QtGui.QApplication.translate("Dialog", "Change Parent", None, QtGui.QApplication.UnicodeUTF8))
+        self.btn_change_parent.setText(QtGui.QApplication.translate("Dialog", "Change Parent", None, QtGui.QApplication.UnicodeUTF8))
+        self.btn_cancel.setText(QtGui.QApplication.translate("Dialog", "Cancel", None, QtGui.QApplication.UnicodeUTF8))
+

### 2014-02-11 src/clusterer/idea_remover.py

@@ -1,0 +1,39 @@
+# -*- coding: utf-8 -*-
+
+# Form implementation generated from reading ui file 'idea_remover.ui'
+#
+# Created: Thu Jan 30 11:47:38 2014
+#      by: pyside-uic 0.2.14 running on PySide 1.2.1
+#
+# WARNING! All changes made in this file will be lost!
+
+from PySide import QtCore, QtGui
+
+class Ui_Dialog(object):
+    def setupUi(self, Dialog):
+        Dialog.setObjectName("Dialog")
+        Dialog.resize(438, 469)
+        sizePolicy = QtGui.QSizePolicy(QtGui.QSizePolicy.Preferred, QtGui.QSizePolicy.Preferred)
+        sizePolicy.setHorizontalStretch(0)
+        sizePolicy.setVerticalStretch(0)
+        sizePolicy.setHeightForWidth(Dialog.sizePolicy().hasHeightForWidth())
+        Dialog.setSizePolicy(sizePolicy)
+        self.btn_remove_ideas = QtGui.QPushButton(Dialog)
+        self.btn_remove_ideas.setGeometry(QtCore.QRect(10, 420, 211, 41))
+        self.btn_remove_ideas.setObjectName("btn_remove_ideas")
+        self.btn_cancel = QtGui.QPushButton(Dialog)
+        self.btn_cancel.setGeometry(QtCore.QRect(230, 420, 201, 41))
+        self.btn_cancel.setObjectName("btn_cancel")
+        self.lst_node_ideas = QtGui.QListView(Dialog)
+        self.lst_node_ideas.setGeometry(QtCore.QRect(10, 10, 421, 401))
+        self.lst_node_ideas.setSelectionMode(QtGui.QAbstractItemView.ExtendedSelection)
+        self.lst_node_ideas.setObjectName("lst_node_ideas")
+
+        self.retranslateUi(Dialog)
+        QtCore.QMetaObject.connectSlotsByName(Dialog)
+
+    def retranslateUi(self, Dialog):
+        Dialog.setWindowTitle(QtGui.QApplication.translate("Dialog", "Change Parent", None, QtGui.QApplication.UnicodeUTF8))
+        self.btn_remove_ideas.setText(QtGui.QApplication.translate("Dialog", "Remove Selections", None, QtGui.QApplication.UnicodeUTF8))
+        self.btn_cancel.setText(QtGui.QApplication.translate("Dialog", "Cancel", None, QtGui.QApplication.UnicodeUTF8))
+

### 2014-02-11 src/clusterer/qt_app.py

@@ -2,593 +2,163 @@
 import os
 import sys
 import json
-import pickle
 import re
 import string
+from collections import deque, defaultdict
+
+# UI imports
 from PySide import QtCore, QtGui
-import sqlite3 as db
 import clustering_app
 import import_cluster_text
 import similarity_selector
 import change_parent
+import idea_remover
 import coverage
 import title
-import csv
-import datetime
-from collections import deque, defaultdict
+import IdeaTreeModel, IdeaListModel, IdeaTreeNodeListModel
 
 db_file_name = '/home/crowdbrainstorm/Desktop/clusters.db'
 
-
-def get_current_time():
-  return datetime.datetime.now()
-
-def get_line_indent(l):
-  depth_count = 0
-  l = l.strip()
-  front = l
-  while front and front[0] == '-':
-    depth_count = depth_count + 1
-    front = front[1:]
-  return depth_count
-
-def extract_idea_ids_from_text(s):
-  matches = []
-  for match in re.findall('\(_id:[0-9]+\)', s):
-    matches.append(int(re.findall('[0-9]+', match)[0]))
-  return matches
-
-class IdeaTreeNode(object):
-  def __init__(self, ideas, parent, label = None):
-    assert not isinstance(ideas, str)
-    if label is not None and len(label) > 0:
-      self._label = label
-    else:
-      self._label = None
-    self.ideas = ideas # list of (text, id, time) tuples
-    self.parent = parent
-    self.children = []
-
-  def as_text(self):
-    outlines = []
-
-    if self._label and len(self._label) > 0:
-      outlines.append(self._label)
-    for text, i, time in self.ideas:
-      outlines.append(text + '  (_id:' + str(i) + ')')
-
-    for i, c in enumerate(self.children):
-      recur = c.as_text()
-      outlines += ['' if l == '' else ('-' + l)
-                        for l in recur]
-
-      outlines.append('')
-
-    return outlines
-
-  def add_ideas(self, ideas):
-    present_ids = [iid for idea, iid, time in self.ideas]
-    ideas = [(idea, iid, time) for (idea, iid, time) in ideas if not iid in present_ids]
-    self.ideas.extend(ideas)
-
-  def merge(self, other_node):
-    self.add_ideas(other_node.ideas)
-
-  def sort_children(self):
-    self.children = sorted(self.children, key=lambda x: x.label().lower())
-
-  def append_child(self, child_node):
-    if not child_node.parent is None:
-        child_node.parent.remove_child(child_node)
[…]

### 2014-02-11 src/clusterer/text_fns.py

@@ -1,0 +1,16 @@
+def get_line_indent(l):
+  depth_count = 0
+  l = l.strip()
+  front = l
+  while front and front[0] == '-':
+    depth_count = depth_count + 1
+    front = front[1:]
+  return depth_count
+
+def extract_idea_ids_from_text(s):
+  matches = []
+  for match in re.findall('\(_id:[0-9]+\)', s):
+    matches.append(int(re.findall('[0-9]+', match)[0]))
+  return matches
+
+

### 2014-02-11 chi2014/rebuttal.txt
long overdue rebuttal commit

We thank the reviewers for their thoughtful comments. We first respond to the AC's numbered points, then to other comments.

1,2: Currently, if a practitioner wishes to solicit brainstorming ideas from a microtask marketplace, there are no models to provide guidance wrt the number of ideas to request, the number of workers to ask, or expected output; yet brainstorming is a critical component of crowd creativity work [16,29]. Similarly, if a researcher wishes to produce a tool to improve crowd brainstorming, there is no baseline for comparison. One cannot assume past work in brainstorming is applicable in this environment (pp 1-2 explicitly make these points). Our work addresses these gaps, and is the first to examine brainstorming at this scale (>100 participants, >3000 ideas).

While any model can be improved by testing under more conditions, we note that in creativity research, it is not uncommon to provide a single task (e.g., [3,4,17,22] in the psych literature and [28,29] in the crowdsourcing literature). Nonetheless, our results replicate findings from past brainstorming research (e.g., SIAM), suggesting they are not idiosyncratic, while usefully extending the literature on this subject. We agree with the AC's suggestion for qualifying the language on p7 (and in related areas).

We posted the HIT at several different times (separated by days or weeks), so it was not a single run.

3: We do not believe our models suffer the confounds the AC suggests. The Bayesian methods used do not require equivalent sample sizes across conditions to compare posterior parameter distributions, and it is unclear how manipulating the number of workers employed while keeping the number of ideas requested would be useful. It was necessary to request varying numbers of ideas to examine how workers would respond to the various conditions. In doing so, we found useful results that challenge expectations (see 4 below). Finally, while not reported, we also tested ordinal originality scores between conditions, and found no significant differences.

4: Off-topic answers were flagged in the coding process. There were only 12 off-topic instances in total, showing that the task was taken seriously. While the AC and R3 felt asking for 100 ideas would result in numerous off-topic responses, we have no evidence supporting this. The examples on p8 from a 75 condition indicate that workers could produce original, on-topic responses later in a run (those examples were ideas 1,27,32; their last 10 ideas included a vegetable peeler, tool for darning socks, and tool for building sand castles).

5: To compute o-scores, we count all instances in the idea node (for idea o-score) or category tree (for category o-score), divide by the total number of instances, and subtract from one. The squashing of scores into the .994-.999 range is an artifact of the diversity of ideas received and size of the data set.

6: The volume of data (>3000 ideas) necessitates that we sample the validity of the clustering, rather than fully replicate it. However, we note that clustering this data is more concrete than other types of qualitative coding as it relies on only two judgements: all equivalent ideas should be placed in the same node, and a more general version of an idea should be a parent to the more specific idea. While there is still room for subjectivity and error, these criteria are easier to follow and test. Our custom tool also assisted in this process: For every unclustered idea, it produced a list of idea clusters ordered by similarity to that idea, with final cluster assignment done manually. The extremely low incidence of misclassified ideas (0.7%) suggests the validity of cluster membership in our data, which was used to derive measures such as o-scores and rates of unique idea generation. Errors in parent-child relationships would not affect the statistical tests, only riffing statistics. We'd be happy to add these points to the paper.

7: We agree and would be happy to frame the hypotheses as research questions.

R1: Because we used uninformative priors, the posterior distributions for model parameters should capture all of our uncertainty in the model. However, in line with R1's suggestion, we performed a 10-fold validation of the exponential model, predicting the number of unique ideas as a function of total number of responses. We found a mean squared error of 18.48 unique ideas.

R3: We performed a Mann-Whitney U test and found a statistical difference (p < 0.00001) for both ideas and categories, confirming our earlier tests.

R3: In the test for popular ideas (p7), we tested the difference between the first 5 instances and all remaining instances (not the last 5 instances) in a run. This test was performed only for runs with >=10 instances, so runs in all conditions contributed samples to both distributions.

We're happy to make these changes, add references suggested by R1+R2. and correct the formatting errors identified.

### 2014-02-12 exps/rate_counting_model.py
some models

@@ -1,0 +1,127 @@
+import pandas as pd
+import scipy as sp
+import numpy as np
+import csv, os, re, itertools, random, math, json, pystan, format_data, pickle
+import scipy.stats as stats
+import networkx as nx
+from imp import reload
+import matplotlib.pyplot as plt
+import stats_fns as mystats
+from collections import defaultdict, OrderedDict
+
+
+model_string = """
+data {
+    int N; // number of instances
+    int novel[N]; // whether there was a novel idea at this point
+    int x[N]; // ordinal position of the instance in its condition
+}
+
+parameters {
+    real <upper=0> rate;
+}
+
+model {
+    real mu[N];
+    real theta;
+    int t;
+    for (i in 1:N) {
+        theta <- exp(rate * x[i]);
+        novel[i] ~ bernoulli(theta);
+    }
+}
+"""
+
+def compute_model(max_x, rate):
+    ys = [0]
+    for x in range(1, max_x):
+        ys.append(ys[-1] + math.exp(rate * x))
+    return ys
+
+def gen_uniques_counts(adf, field):
+    adf = adf.sort(columns=['submit_datetime', 'answer_num'], ascending=[1, 1])
+    uniques = set()
+    
+    counts = []
+    for thing in adf[field]:
+        uniques.add(thing)
+        counts.append(len(uniques))
+    return counts
+
+def gen_data(df, field):
+    dat = defaultdict(list)
+
+    for nr in set(df['num_requested']):
+        nrdf = df[df['num_requested'] == nr]
+        for qc in set(df['question_code']):
+            qcdf = nrdf[nrdf['question_code'] == qc]
+            uniques_counts = gen_uniques_counts(qcdf, field)
+            uniques_diffs = np.diff(uniques_counts)
+            
+            dat['novel'].extend(uniques_diffs)
+            dat['x'].extend(range(len(uniques_diffs)))
+                            
+    return {'x': dat['x'],
+            'novel': dat['novel'],
+            'N': len(dat['x'])}
+
+def view_fit(df, field, fit):
+    la = fit.extract(permuted=True)
+    rate = mystats.mean_and_hpd(la['rate'], 0.95)
+
+    plot_rate_model(rate, df, field)
+
+def plot_rate_model(rate, df, field):
+    fig = plt.figure(figsize=(10, 10))
+    ax = fig.add_subplot(111)
[…]

### 2014-02-12 exps/rate_counting_types_model.py

@@ -1,0 +1,137 @@
+import pandas as pd
+import scipy as sp
+import numpy as np
+import csv, os, re, itertools, random, math, json, pystan, format_data, pickle
+import scipy.stats as stats
+import networkx as nx
+from imp import reload
+import matplotlib.pyplot as plt
+import stats_fns as mystats
+from collections import defaultdict, OrderedDict
+
+model_string = """
+data {
+    int<lower=0> N; // number of instances
+    int<lower=0,upper=1> novel[N]; // whether there was a novel idea at this point
+    int<lower=0,upper=N> x[N]; // ordinal position of the instance in its condition
+    int<lower=0,upper=N> M; // number of turkers
+    int<lower=0,upper=M> t[N]; // which turker this is
+}
+
+parameters {
+    real <lower=-10, upper=0> rate[M];
+}
+
+model {
+    for (i in 1:N) {
+        novel[i] ~ bernoulli( exp(rate[t[i]] *  x[i]) );
+    }
+}
+"""
+
+def compute_model(turkers, turker_rates):
+    ys = [0]
+    for t in turkers:
+        rate = turker_rates[t]
+        ys.append(ys[-1] + math.exp(rate * x))
+    return ys
+
+def gen_uniques_counts(adf, field):
+    adf = adf.sort(columns=['submit_datetime', 'answer_num'], ascending=[1, 1])
+    uniques = set()
+    
+    counts = []
+    turkers = []
+    for i in adf.index:
+        thing = adf[field][i]
+        uniques.add(thing)
+        counts.append(len(uniques))
+        turkers.append(int(adf['worker_int'][i]))
+    return counts, turkers
+
+def gen_data(df, field):
+    dat = defaultdict(list)
+
+    for nr in set(df['num_requested']):
+        nrdf = df[df['num_requested'] == nr]
+        for qc in set(df['question_code']):
+            qcdf = nrdf[nrdf['question_code'] == qc]
+            uniques_counts, turkers = gen_uniques_counts(qcdf, field)
+            temp = [0] + uniques_counts
+            uniques_diffs = np.diff(temp)
+            
+            dat['novel'].extend(uniques_diffs)
+            dat['x'].extend(range(len(uniques_diffs)))
+            dat['t'].extend(turkers)
+                            
+    return {'M': len(set(dat['t'])),
+            'x': dat['x'],
[…]

### 2014-02-12 exps/rate_model.py

@@ -1,0 +1,139 @@
+import pandas as pd
+import scipy as sp
+import numpy as np
+import csv, os, re, itertools, random, math, json, pystan, format_data, pickle
+import scipy.stats as stats
+import networkx as nx
+from imp import reload
+import matplotlib.pyplot as plt
+import stats_fns as mystats
+from collections import defaultdict, OrderedDict
+
+
+rate_model_string = """
+data {
+    int N; // number of instances
+    real y[N]; // the number of ideas or categories receieved up to and including instance n
+    int x[N]; // ordinal position of the instance in its condition
+}
+
+parameters {
+    real<lower=0, upper=1> rate;
+    real<lower=0> y_scale;
+    real<lower=0> sigma;
+}
+
+model {
+    real mu[N];
+    for (i in 1:N) {
+        mu[i] <- y_scale * pow(x[i], rate);
+        y[i] ~ normal(mu[i], sigma);
+    }
+}
+"""
+
+def compute_rate_model(x, y_scale, rate):
+    return y_scale * (x ** rate)
+
+def gen_uniques_counts(adf, field):
+    adf = adf.sort(columns=['submit_datetime', 'answer_num'], ascending=[1, 1])
+    uniques = set()
+    
+    counts = []
+    for thing in adf[field]:
+        uniques.add(thing)
+        counts.append(len(uniques))
+    return counts
+
+def gen_rate_model_data(df, field):
+    dat = defaultdict(list)
+
+    for nr in set(df['num_requested']):
+        nrdf = df[df['num_requested'] == nr]
+        for qc in set(df['question_code']):
+            qcdf = nrdf[nrdf['question_code'] == qc]
+            uniques_counts = gen_uniques_counts(qcdf, field)
+            dat['y'].extend(uniques_counts)
+            dat['x'].extend(range(1, len(uniques_counts) + 1))
+                            
+    assert(len(dat['x']) == len(dat['y']))
+
+    return {'x': dat['x'],
+            'y': dat['y'],
+            'N': len(dat['x'])}
+
+def view_rate_model_fit(df, field, fit):
+    la = fit.extract(permuted=True)
+    rates = la['rate']
+    y_scale = np.mean( la['y_scale'])
+    
+    left, right = mystats.hpd(rates, 0.95)
+    rate_mean = np.mean(rates)
+
[…]

### 2014-02-12 exps/split_originality_blend_model.py

@@ -1,0 +1,120 @@
+import pandas as pd
+import scipy as sp
+import numpy as np
+import csv, os, re, itertools, random, math, json, pystan, format_data, pickle
+import scipy.stats as stats
+import networkx as nx
+from imp import reload
+import matplotlib.pyplot as plt
+import stats_fns as mystats
+from collections import defaultdict, OrderedDict
+
+model_string = """
+data {
+    int N; // number of instances
+    int order[N]; // order of instance in brainstorming run
+    real oscore[N]; // oscore of instance at position
+}
+
+parameters {
+    real<lower=1, upper=100> split;
+    real s1_m;
+    real s1_b;
+    real s2_b;
+    real sigma;
+}
+
+model {
+    real s1_score[N];
+    real s2_score[N];
+    real mu[N];
+    for (i in 1:N) {
+        s1_score[i] <- s1_m * order[N] + s1_b;
+        s2_score[i] <- s2_b;
+        mu[i] <- (order[N] <= split) * s1_score[i] + (order[N] > split) * s2_score[i];
+        oscore[i] ~ normal(mu[i], sigma);
+    }
+}
+"""
+def gen_dat(df, field):
+    return {'order': df['answer_num'],
+            'oscore': df[field],
+            'N': len(df) }
+
+def view_fit(dat, fit):
+    la = fit.extract(permuted=True)
+
+    s1_m = mystats.mean_and_hpd(la['s1_m'], 0.95)
+    s1_b = mystats.mean_and_hpd(la['s1_b'], 0.95)
+    s2_b = mystats.mean_and_hpd(la['s2_b'], 0.95)
+    switch = mystats.mean_and_hpd(la['split'], 0.95)
+    sigma = mystats.mean_and_hpd(la['sigma'], 0.95)
+
+    print("Estimated normal variance:", sigma[0])
+
+    plot_fit(dat, s1_m, s1_b, s2_b, switch)
+ 
+def plot_linear_hpd(ax, m, b, xs):
+    bottom_ys = [min(m) * x + min(b) for x in xs]
+    top_ys = [max(m) * x + max(b) for x in xs]
+    ax.fill_between(xs, bottom_ys, top_ys, color='g', alpha=0.10)
+
+def plot_fit(dat, s1_m, s1_b, s2_b, switch):
+    fig = plt.figure(figsize=(10, 10))
+    ax = fig.add_subplot(111)
+    ax.set_xlabel("answer number")
+    ax.set_ylabel("oscore")
[…]

### 2014-02-12 exps/split_originality_model.py

@@ -1,0 +1,146 @@
+import pandas as pd
+import scipy as sp
+import numpy as np
+import csv, os, re, itertools, random, math, json, pystan, format_data, pickle
+import scipy.stats as stats
+import networkx as nx
+from imp import reload
+import matplotlib.pyplot as plt
+import stats_fns as mystats
+from collections import defaultdict, OrderedDict
+
+originality_log_model_string = """
+data {
+    int N;
+    real y[N];
+}
+
+parameters {
+    real<lower=0,upper=1> phi;
+    real<lower=0.1> lambda;
+}
+
+transformed parameters {
+    real<lower=0> alpha;
+    real<lower=0> beta;
+    alpha <- lambda * phi;
+    beta <- lambda * (1 - phi);
+}
+
+model {
+    phi ~ beta(1,1);
+    lambda ~ pareto(0.1,1.5);
+    for (i in 1:N) {
+        log(y[i]) ~ beta(alpha, beta);
+    }
+}
+"""
+def gen_originality_model_dat(df, field):
+    left_df = df[df['answer_num'] < 20]
+    right_df = df[df['answer_num'] >= 20]
+    
+    left_dat = { 'N': len(left_df),
+                 'y': left_df[field] }
+    right_dat = { 'N': len(right_df),
+                  'y': right_df[field] }
+
+
+    return left_dat, right_dat
+
+def view_originality_model_fit(left_fit, right_fit, df, field):
+    for name, fit in [('LEFT', left_fit), ('RIGHT', right_fit)]:
+        la = fit.extract(permuted=True)
+        alpha = la['alpha']
+        beta = la['beta']
+        
+        print(name)
+        hl, hr = hpd(alpha, 0.95)
+        print("alpha:", hl, hr)
+        hl, hr = hpd(beta, 0.95)
+        print("beta:", hl, hr)
+ 
+def plot_originality_model(df, field):
+    fig = plt.figure(figsize=(10, 10))
+    ax = fig.add_subplot(111)
+    ax.set_xlabel("idea_oscore")
+    ax.set_ylabel("number of instances")
+
+    min_x = min(df[field])
+    #xs = range(min_)
+    #ax.set_xlim(0, max_x)
+    #ax.set_ylim(0, max_x)
+
+    # plot the hpd area
+    #bottom_ys = [compute_rate_model(x, y_scale, rate_hpd[0]) for x in xs]
[…]

### 2014-02-12 exps/format_data.py
sljdfaljsd

@@ -6,6 +6,11 @@ import numpy as np
 from numpy import uint8, uint16, uint32, uint64, datetime64, int64, int32, float64
 from collections import defaultdict, OrderedDict
 
+def filter_repeats(df):
+    new_df = df[df['is_repeat_worker'] == 0]
+    return new_df
+
+
 def metrics_folder(pd_folder, x):
         return '/%s/pilot18_metrics/%s' % (pd_folder, x)
 
@@ -233,7 +238,6 @@ def mk_repeat_workers_series(adf):
     is_repeat = pd.Series([0 for i in adf.index], index=adf.index)
     adf = adf.sort(['submit_datetime'])
     runs = adf.groupby(['worker_id', 'question_code', 'num_requested', 'submit_datetime'])
-    print("repeat worker check count", len(list(runs)))
 
     seen_keys = set()
     for rid, ((wid, qc, nr, sdt), run) in enumerate(runs):

### 2014-02-12 exps/stats_fns.py

@@ -1,0 +1,119 @@
+import subprocess
+import numpy as np
+
+pystan_model_cache = dict()
+
+_hdi_stats_cache = {}
+# Use this one instead
+def get_hdi(a, b, confidence_interval):
+  if (a,b,confidence_interval) in _hdi_stats_cache:
+    return _hdi_stats_cache[(a,b,confidence_interval)]
+  #print "get_hdi:", a, b, confidence_interval
+  if a == 1 or b == 1:
+    if a == 1 and b == 1:
+      lower_bound = (1.0-confidence_interval) / 2
+      upper_bound = 1.0 - lower_bound
+    else:
+      #print "special qbeta:", a, b, confidence_interval
+      results = subprocess.check_output(["R", "-q", "-e", 'qbeta(%f, 1, %d)' % (confidence_interval, max(a,b))])
+      results = str(results)
+      results = results.split('\\n')[1]
+      upper_bound = float(results[results.index(' ')+1:])
+      if a < b:
+        lower_bound = 0
+      else:
+        lower_bound = 1.0 - upper_bound
+        upper_bound = 1.0
+  else:
+    results = subprocess.check_output(["R", "-q", "-e", 
+        'get_hdi = function(a, b, level=0.95) { density_diff = function(lower, a, b) { p_lower = pbeta(lower, a, b); p_upper = min(1.0, p_lower + level); upper = qbeta(p_upper, a, b); return(dbeta(lower, a, b) - dbeta(upper, a, b)); }; lower = uniroot(density_diff, c(0, qbeta(1.0-level, a, b)), a=a, b=b)$root; upper = qbeta(pbeta(lower, a, b) + level, a, b); return(c(lower, upper)); }; get_hdi(%d, %d, %f)' % (a, b, confidence_interval)])
+    results = str(results)
+    results = results.split('\\n')[1]
+    results = results[results.index(' ')+1:]
+    lower_bound, upper_bound = [float(x) for x in results.split(' ')]
+  _hdi_stats_cache[(a,b,confidence_interval)] = (lower_bound, upper_bound)
+  return lower_bound, upper_bound
+
+def beta_bernoulli_posterior(num_successes, total_num):
+  a_prior = 1
+  b_prior = 1
+  a_post = a_prior + num_successes
+  b_post = b_prior + (total_num - num_successes)
+  posterior_mean = a_post / float(total_num + a_prior + b_prior)
[…]

### 2014-02-13 exps/split_originality_blend_model.py
working model assuming normal distribution of samples

@@ -18,21 +18,20 @@ data {
 
 parameters {
     real<lower=1, upper=100> split;
-    real s1_m;
-    real s1_b;
-    real s2_b;
-    real sigma;
+    real<lower=0, upper=1> s1_b;
+    real<lower=0, upper=1> s2_b;
+    real<lower=0, upper=1> s1_m;
+    real<lower=0, upper=1> sigma1;
+    real<lower=0, upper=1> sigma2;
 }
 
 model {
-    real s1_score[N];
-    real s2_score[N];
-    real mu[N];
     for (i in 1:N) {
-        s1_score[i] <- s1_m * order[N] + s1_b;
-        s2_score[i] <- s2_b;
-        mu[i] <- (order[N] <= split) * s1_score[i] + (order[N] > split) * s2_score[i];
-        oscore[i] ~ normal(mu[i], sigma);
+        if ( order[i] <= split ) {
+            oscore[i] ~ normal(s1_m * order[i] + s1_b, sigma1);
+        } else {
+            oscore[i] ~ normal(s2_b, sigma2);
+        }
     }
 }
 """
@@ -48,9 +47,9 @@ def view_fit(dat, fit):
     s1_b = mystats.mean_and_hpd(la['s1_b'], 0.95)
     s2_b = mystats.mean_and_hpd(la['s2_b'], 0.95)
     switch = mystats.mean_and_hpd(la['split'], 0.95)
-    sigma = mystats.mean_and_hpd(la['sigma'], 0.95)
+    #sigma = mystats.mean_and_hpd(la['sigma'], 0.95)
 
-    print("Estimated normal variance:", sigma[0])
+    #print("Estimated normal variance:", sigma[0])
 
     plot_fit(dat, s1_m, s1_b, s2_b, switch)
  
@@ -67,6 +66,7 @@ def plot_fit(dat, s1_m, s1_b, s2_b, switch):
 
     ax.set_xlim(0, 100)
     split_point = int(switch[0])
+    #split_point = 20
     m1_xs = range(0, split_point + 1)
     m2_xs = range(split_point + 1, 100)
 
@@ -81,7 +81,7 @@ def plot_fit(dat, s1_m, s1_b, s2_b, switch):
     # plot the model
     m1_ys = [s1_m[0] * x + s1_b[0] for x in m1_xs]
     m2_ys = [s2_b[0] for x in m2_xs]
-    print(s2_b[0])
+    print("intercept for right model:", s2_b[0])
     ax.plot(m1_xs, m1_ys, '--', color='k')
[…]

### 2014-02-13 exps/rate_counting_model.py
uuurrghhhh zombie

@@ -56,12 +56,14 @@ def gen_data(df, field):
         for qc in set(df['question_code']):
             qcdf = nrdf[nrdf['question_code'] == qc]
             uniques_counts = gen_uniques_counts(qcdf, field)
-            uniques_diffs = np.diff(uniques_counts)
+            temp = [0] + uniques_counts
+            uniques_diffs = np.diff(temp)
             
             dat['novel'].extend(uniques_diffs)
             dat['x'].extend(range(len(uniques_diffs)))
                             
-    return {'x': dat['x'],
+    return {'M': len(set(dat['t'])),
+            'x': dat['x'],
             'novel': dat['novel'],
             'N': len(dat['x'])}
 

### 2014-02-13 exps/rate_counting_types_model.py

@@ -12,10 +12,10 @@ from collections import defaultdict, OrderedDict
 model_string = """
 data {
     int<lower=0> N; // number of instances
-    int<lower=0,upper=1> novel[N]; // whether there was a novel idea at this point
+    int<lower=0,upper=1> success[N]; // whether there was a novel idea at this point
     int<lower=0,upper=N> x[N]; // ordinal position of the instance in its condition
     int<lower=0,upper=N> M; // number of turkers
-    int<lower=0,upper=M> t[N]; // which turker this is
+    int<lower=1,upper=M> t[N]; // which turker this is
 }
 
 parameters {
@@ -24,7 +24,7 @@ parameters {
 
 model {
     for (i in 1:N) {
-        novel[i] ~ bernoulli( exp(rate[t[i]] *  x[i]) );
+        success[i] ~ bernoulli( exp(rate[t[i]] *  x[i]) );
     }
 }
 """
@@ -60,15 +60,20 @@ def gen_data(df, field):
             temp = [0] + uniques_counts
             uniques_diffs = np.diff(temp)
             
-            dat['novel'].extend(uniques_diffs)
+            dat['success'].extend(uniques_diffs)
             dat['x'].extend(range(len(uniques_diffs)))
             dat['t'].extend(turkers)
+
+    # Assertion based on potential error identified by Bob Carpenter on the
+    # Stan mailing list (success should be 1 at t=0 with P = 1)
+    for x, s in zip(dat['x'], dat['success']):
+        assert(not (x == 0 and s == 0))
                             
     return {'M': len(set(dat['t'])),
-            'x': dat['x'],
-            'novel': dat['novel'],
+            'x': [int(x) for x in dat['x']],
+            'success': [int(s) for s in dat['success']],
             'N': len(dat['x']),
-            't': dat['t']}
+            't': [int(t) for t in dat['t']]}
 
 def view_fit(df, field, fit):
     la = fit.extract(permuted=True)
@@ -122,6 +127,7 @@ if __name__ == '__main__':
     idf['worker_int'] += 1
 
     dat = gen_data(idf, 'idea')
+    json.dump(dat, open('cache/rate_counting_types_model_stan_dat.json', 'w'))
 
     model_file = 'cache/rate_counting_types_model_stan'
     if os.path.isfile(model_file): 

### 2014-02-13 exps/split_originality_blend_model.py

@@ -11,54 +11,68 @@ from collections import defaultdict, OrderedDict
 
 model_string = """
 data {
-    int N; // number of instances
-    int order[N]; // order of instance in brainstorming run
-    real oscore[N]; // oscore of instance at position
+    int<lower=0> N; // number of instances
+    int<lower=1, upper=100> order[N]; // order of instance in brainstorming run
+    real<lower=0, upper=1> oscore[N]; // oscore of instance at position
 }
 
 parameters {
     real<lower=1, upper=100> split;
-    real<lower=0, upper=1> s1_b;
-    real<lower=0, upper=1> s2_b;
-    real<lower=0, upper=1> s1_m;
-    real<lower=0, upper=1> sigma1;
-    real<lower=0, upper=1> sigma2;
+
+    real<lower=0> alpha1;
+    real<lower=0> alpha2;
+    real<lower=0> beta1;
+    real<lower=0> beta2;
 }
 
 model {
+    real mix[N];
     for (i in 1:N) {
-        if ( order[i] <= split ) {
-            oscore[i] ~ normal(s1_m * order[i] + s1_b, sigma1);
-        } else {
-            oscore[i] ~ normal(s2_b, sigma2);
-        }
+        if (order[i] <= split)
+            mix[i] <- order[i] / split;
+        else
+            mix[i] <- 1;
+
+        oscore[i] ~ beta((1 - mix[i]) * alpha1 + mix[i] * alpha2,
+                         (1 - mix[i]) * beta1 + mix[i] * beta2);
     }
+
 }
 """
 def gen_dat(df, field):
-    return {'order': df['answer_num'],
+    return {'order': df['answer_num'] + 1,
             'oscore': df[field],
             'N': len(df) }
 
 def view_fit(dat, fit):
     la = fit.extract(permuted=True)
 
-    s1_m = mystats.mean_and_hpd(la['s1_m'], 0.95)
-    s1_b = mystats.mean_and_hpd(la['s1_b'], 0.95)
-    s2_b = mystats.mean_and_hpd(la['s2_b'], 0.95)
+    alpha1 = mystats.mean_and_hpd(la['alpha1'], 0.95)
+    alpha2 = mystats.mean_and_hpd(la['alpha2'], 0.95)
[…]

### 2014-02-14 exps/format_data.py
added charity data

@@ -371,10 +371,10 @@ def do_format_data(processed_data_folder, filter_instances = None):
             [11, 12, 13, 14, 16, 17, 18])
 
     idea_cluster_csvs = {qc: metrics_folder(processed_data_folder, "_%s.csv" % qc) for qc in \
-                         ['iPod', 'turk']}
+                         ['iPod', 'turk', 'charity']}
                          #['charity', 'iPod', 'forgot_name', 'turk']}
     cluster_tree_csvs = {qc: metrics_folder(processed_data_folder, "_%s_clusters.csv" % qc) for qc in \
-                         ['iPod', 'turk']}
+                         ['iPod', 'turk', 'charity']}
                          #['charity', 'iPod', 'forgot_name', 'turk']}
         
     
@@ -425,5 +425,6 @@ def do_format_data(processed_data_folder, filter_instances = None):
 
 if __name__ == "__main__":
     idf, cfs = do_format_data("/home/fil/enc_projects/crowbrain/processed_data")
+    df, rmdf, clusters_df, ann_cfs = mk_redundant(idf, cfs)
     
-    print(idf)
+    print(df)

### 2014-02-14 exps/old/rate_counting_types_model_2.py
some old models

@@ -1,0 +1,129 @@
+import pandas as pd
+import scipy as sp
+import numpy as np
+import csv, os, re, itertools, random, math, json, pystan, format_data, pickle
+import scipy.stats as stats
+import networkx as nx
+from imp import reload
+import matplotlib.pyplot as plt
+import stats_fns as mystats
+from collections import defaultdict, OrderedDict
+
+model_string = """
+data {
+    int<lower=0> N; // number of instances
+    int<lower=1, upper=100> order[N]; // order of instance in brainstorming run
+    real<lower=0, upper=1> oscore[N]; // oscore of instance at position
+}
+
+parameters {
+    real<lower=1, upper=100> split;
+    real<lower=0, upper=1> s1_b;
+    real<lower=0, upper=1> s2_b;
+    real<lower=0, upper=1> s1_m;
+    real<lower=0, upper=1> sigma;
+
+    real<lower=0> alpha[100];
+    real<lower=0> beta[100];
+}
+
+model {
+    # hyperparameters on alpha, beta
+    real mu[100];
+    for (i in 1:100) {
+        mu[i] <- alpha[i] / (alpha[i] + beta[i]);
+        if (i < split) {
+            mu[i] ~ normal(s1_m * i + s1_b, sigma);
+        } else {
+            mu[i] ~ normal(s2_b, sigma);
+        }
+    }
+
+    for (i in 1:N) {
+        oscore[i] ~ beta(alpha[order[i]], beta[order[i]]);
+    }
+}
+"""
+def gen_dat(df, field):
+    return {'order': df['answer_num'] + 1,
+            'oscore': df[field],
+            'N': len(df) }
+
+def view_fit(dat, fit):
+    la = fit.extract(permuted=True)
+
+    s1_m = mystats.mean_and_hpd(la['s1_m'], 0.95)
+    s1_b = mystats.mean_and_hpd(la['s1_b'], 0.95)
+    s2_b = mystats.mean_and_hpd(la['s2_b'], 0.95)
+    switch = mystats.mean_and_hpd(la['split'], 0.95)
+    #sigma = mystats.mean_and_hpd(la['sigma'], 0.95)
+
+    #print("Estimated normal variance:", sigma[0])
+
+    plot_fit(dat, s1_m, s1_b, s2_b, switch)
+ 
+def plot_linear_hpd(ax, m, b, xs):
+    bottom_ys = [min(m) * x + min(b) for x in xs]
[…]

### 2014-02-14 exps/old/split_originality_blend_model_3.py

@@ -1,0 +1,128 @@
+import pandas as pd
+import scipy as sp
+import numpy as np
+import csv, os, re, itertools, random, math, json, pystan, format_data, pickle
+import scipy.stats as stats
+import networkx as nx
+from imp import reload
+import matplotlib.pyplot as plt
+import stats_fns as mystats
+from collections import defaultdict, OrderedDict
+
+model_string = """
+data {
+    int<lower=0> N; // number of instances
+    int<lower=1, upper=100> order[N]; // order of instance in brainstorming run
+    real<lower=0, upper=1> oscore[N]; // oscore of instance at position
+}
+
+parameters {
+    real<lower=1, upper=100> split;
+    real<lower=0, upper=1> s1_b;
+    real<lower=0, upper=1> s2_b;
+    real<lower=0, upper=1> s1_m;
+    real<lower=0, upper=1> sigma;
+
+    real<lower=0, upper=1> mu[100];
+    real<lower=0> v;
+}
+
+model {
+    # hyperparameters on alpha, beta
+    for (i in 1:100) {
+        if (i < split) {
+            mu[i] ~ normal(s1_m * i + s1_b, sigma);
+        } else {
+            mu[i] ~ normal(s2_b, sigma);
+        }
+    }
+
+    for (i in 1:N) {
+        oscore[i] ~ beta(mu[order[i]] * v,
+                         (1 - mu[order[i]]) * v);
+    }
+}
+"""
+def gen_dat(df, field):
+    return {'order': df['answer_num'] + 1,
+            'oscore': df[field],
+            'N': len(df) }
+
+def view_fit(dat, fit):
+    la = fit.extract(permuted=True)
+
+    s1_m = mystats.mean_and_hpd(la['s1_m'], 0.95)
+    s1_b = mystats.mean_and_hpd(la['s1_b'], 0.95)
+    s2_b = mystats.mean_and_hpd(la['s2_b'], 0.95)
+    switch = mystats.mean_and_hpd(la['split'], 0.95)
+    #sigma = mystats.mean_and_hpd(la['sigma'], 0.95)
+
+    #print("Estimated normal variance:", sigma[0])
+
+    plot_fit(dat, s1_m, s1_b, s2_b, switch)
+ 
+def plot_linear_hpd(ax, m, b, xs):
+    bottom_ys = [min(m) * x + min(b) for x in xs]
+    top_ys = [max(m) * x + max(b) for x in xs]
[…]
